### 2025-03-24 docs/paper/paper.tex

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={A Configurable Library for Generating and Manipulating
Maze Datasets},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
Spies, Rusheb Shah, Brandon Knutson, Cecilia Diniz Behn, Samy Wu Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={N/A},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alex F. Spies, Rusheb Shah, Brandon Knutson, Cecilia
Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\usepackage{fixltx2e} % provides \textsubscript

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
\renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{joss/logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
et al. (1970). A Configurable Library for Generating and Manipulating
Maze Datasets. \emph{Journal of Open Source Software},
\emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

\definecolor{c53baa1}{RGB}{83,186,161}
\definecolor{c202826}{RGB}{32,40,38}
\def \rorglobalscale {0.1}
\newcommand{\rorlogo}{%
\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364).. controls (3.9643, 1.4576) and (4.0508, 1.5081) .. (4.1098, 1.5878).. controls (4.169, 1.6674) and (4.1984, 1.7642) .. (4.1984, 1.8777).. controls (4.1984, 1.9719) and (4.182, 2.0503) .. (4.1495, 2.1132).. controls (4.1169, 2.1762) and (4.0727, 2.2262) .. (4.0174, 2.2635).. controls (3.9621, 2.3006) and (3.8976, 2.3273) .. (3.824, 2.3432).. controls (3.7505, 2.359) and (3.6727, 2.367) .. (3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688).. controls (2.9625, 1.8833) and (2.9572, 1.8976) .. (2.9514, 1.9119).. controls (2.9083, 2.0164) and (2.848, 2.1056) .. (2.7705, 2.1791).. controls (2.6929, 2.2527) and (2.6014, 2.3093) .. (2.495, 2.3487).. controls (2.3889, 2.3881) and (2.2728, 2.408) .. (2.1468, 2.408).. controls (2.0209, 2.408) and (1.905, 2.3881) .. (1.7986, 2.3487).. controls (1.6925, 2.3093) and (1.6007, 2.2527) .. (1.5232, 2.1791).. controls (1.4539, 2.1132) and (1.3983, 2.0346) .. (1.3565, 1.9436).. controls (1.3504, 2.009) and (1.3351, 2.0656) .. (1.3105, 2.1132).. controls (1.2779, 2.1762) and (1.2338, 2.2262) .. (1.1785, 2.2635).. controls (1.1232, 2.3006) and (1.0586, 2.3273) .. (0.985, 2.3432).. controls (0.9115, 2.359) and (0.8337, 2.367) .. (0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364).. controls (1.1255, 1.4576) and (1.2121, 1.5081) .. (1.2711, 1.5878).. controls (1.2737, 1.5915) and (1.2761, 1.5954) .. (1.2787, 1.5991).. controls (1.2782, 1.5867) and (1.2779, 1.5743) .. (1.2779, 1.5616).. controls (1.2779, 1.4327) and (1.2996, 1.3158) .. (1.3428, 1.2113).. controls (1.3859, 1.1068) and (1.4462, 1.0176) .. (1.5237, 0.944).. controls (1.601, 0.8705) and (1.6928, 0.8139) .. (1.7992, 0.7744).. controls (1.9053, 0.735) and (2.0214, 0.7152) .. (2.1474, 0.7152).. controls (2.2733, 0.7152) and (2.3892, 0.735) .. (2.4956, 0.7744).. controls (2.6016, 0.8139) and (2.6935, 0.8705) .. (2.771, 0.944).. controls (2.8482, 1.0176) and (2.9086, 1.1068) .. (2.952, 1.2113).. controls (2.9578, 1.2253) and (2.9631, 1.2398) .. (2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777).. controls (0.9438, 1.7534) and (0.92, 1.7357) .. (0.8911, 1.7243).. controls (0.8623, 1.7129) and (0.83, 1.706) .. (0.7945, 1.7039).. controls (0.7588, 1.7015) and (0.7252, 1.7005) .. (0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667).. controls (0.7477, 2.0667) and (0.7805, 2.0643) .. (0.8139, 2.0598).. controls (0.8472, 2.0553) and (0.8768, 2.0466) .. (0.9025, 2.0336).. controls (0.9282, 2.0206) and (0.9496, 2.0021) .. (0.9663, 1.9778).. controls (0.9829, 1.9534) and (0.9914, 1.9209) .. (0.9914, 1.8799).. controls (0.9914, 1.8362) and (0.9819, 1.8021) .. (0.9628, 1.7777) -- cycle(2.6125, 1.3533).. controls (2.5889, 1.2904) and (2.5553, 1.2359) .. (2.5112, 1.1896).. controls (2.4672, 1.1433) and (2.4146, 1.1073) .. (2.3529, 1.0814).. controls (2.2916, 1.0554) and (2.2228, 1.0427) .. (2.1471, 1.0427).. controls (2.0712, 1.0427) and (2.0026, 1.0557) .. (1.9412, 1.0814).. controls (1.8799, 1.107) and (1.8272, 1.1433) .. (1.783, 1.1896).. controls (1.7391, 1.2359) and (1.7052, 1.2904) .. (1.6817, 1.3533).. controls (1.6581, 1.4163) and (1.6465, 1.4856) .. (1.6465, 1.5616).. controls (1.6465, 1.6359) and (1.6581, 1.705) .. (1.6817, 1.7687).. controls (1.7052, 1.8325) and (1.7388, 1.8873) .. (1.783, 1.9336).. controls (1.8269, 1.9799) and (1.8796, 2.0159) .. (1.9412, 2.0418).. controls (2.0026, 2.0675) and (2.0712, 2.0804) .. (2.1471, 2.0804).. controls (2.223, 2.0804) and (2.2916, 2.0675) .. (2.3529, 2.0418).. controls (2.4143, 2.0161) and (2.467, 1.9799) .. (2.5112, 1.9336).. controls (2.5551, 1.8873) and (2.5889, 1.8322) .. (2.6125, 1.7687).. controls (2.636, 1.705) and (2.6477, 1.6359) .. (2.6477, 1.5616).. controls (2.6477, 1.4856) and (2.636, 1.4163) .. (2.6125, 1.3533) -- cycle(3.8015, 1.7777).. controls (3.7825, 1.7534) and (3.7587, 1.7357) .. (3.7298, 1.7243).. controls (3.701, 1.7129) and (3.6687, 1.706) .. (3.6333, 1.7039).. controls (3.5975, 1.7015) and (3.5639, 1.7005) .. (3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667).. controls (3.5864, 2.0667) and (3.6192, 2.0643) .. (3.6526, 2.0598).. controls (3.6859, 2.0553) and (3.7155, 2.0466) .. (3.7412, 2.0336).. controls (3.7669, 2.0206) and (3.7883, 2.0021) .. (3.805, 1.9778).. controls (3.8216, 1.9534) and (3.8301, 1.9209) .. (3.8301, 1.8799).. controls (3.8301, 1.8362) and (3.8206, 1.8021) .. (3.8015, 1.7777) -- cycle;
  \end{scope}
\end{tikzpicture}
}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\renewcommand{\affil}[2][]%
   {\newaffiltrue\let\AB@blk@and\AB@pand
      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
        \setcounter{Maxaffil}{0}\fi
        \begingroup
        \let\href=\href@Orig
        \let\protect\@unexpandable@protect
        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
        \@temptokena=\expandafter{\AB@authors}%
        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
         \gdef\AB@las{}\gdef\AB@au@str{}%
        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
        \@temptokena=\expandafter{\AB@affillist}%
        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
      \endgroup
       \let\AB@affilsep\AB@affilsepx
}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{fontspec}

  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\sffamily]{Ligatures=TeX}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

%% Font settings
\usepackage{fontsetup} % Lazy way to get proper Greek lowercase glyphs

% Use Hack https://sourcefoundry.org/hack/
\setmonofont{Hack}

\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\urlstyle{same}  % don't use monospace font for urls
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{A Configurable Library for Generating and Manipulating Maze
Datasets}

\author[1%
%
\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
  \,\orcidlink{0000-0002-4213-4993}\,%
}
\author[%
%
]{Aaron Sandoval%
}
\author[2%
%
]{Alex F. Spies%
  \,\orcidlink{0000-0002-8708-1530}\,%
}
\author[%
%
]{Rusheb Shah%
  \,\orcidlink{0009-0008-3058-0217}\,%
}
\author[%
%
]{Brandon Knutson%
}
\author[1%
%
]{Cecilia Diniz Behn%
}
\author[1%
%
]{Samy Wu Fung%
}

\affil[1]{Colorado School of Mines, Department of Applied Mathematics
and Statistics%
}
\affil[2]{Imperial College London%
}
\affil[$\mathparagraph$]{Corresponding author}
\date{\vspace{-2.5ex}}

\begin{document}
\maketitle

\marginpar{

  \begin{flushleft}
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/N/A}{\color{linky}{N/A}}

  \vspace{2mm}
    {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{https://github.com/openjournals}{\color{linky}{Review}} \ExternalLink
    \item \href{https://github.com/openjournals}{\color{linky}{Repository}} \ExternalLink
    \item \href{https://doi.org/10.5281}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

    \par\noindent\hrulefill\par

  \vspace{2mm}

  {\bfseries Editor:} \href{https://joss.theoj.org}{Open
Journals} \ExternalLink
   \\
  \vspace{1mm}
    {\bfseries Reviewers:}
  \begin{itemize}
  \setlength\itemsep{0em}
    \item \href{https://github.com/openjournals}{@openjournals}
    \end{itemize}
    \vspace{2mm}

    {\bfseries Submitted:} 01 January 1970\\
    {\bfseries Published:} 01 January 1970

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{https://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).

  \end{flushleft}
}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

Understanding how machine learning models respond to distributional
shifts and different input representations is a key research challenge.
Mazes serve as an excellent testbed for this problem due to the degree
of control we have over quantifiable distributional shifts. To enable
systematic investigations of model behavior on maze tasks, we present
\texttt{maze-dataset}, a comprehensive library for generating,
processing, visualizing, and tokenizing datasets consisting of
maze-solving tasks. With this library, researchers can easily create
datasets and have extensive control over the generation algorithm used,
the parameters fed to the algorithm of choice, and the filters that
generated mazes must satisfy. Furthermore, this library supports
multiple output formats, including rasterized and a large number of
text-based formats, catering to convolutional neural networks and
autoregressive transformer models. These formats, along with tools for
visualizing and converting between them, ensure versatility and
adaptability in research applications.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Out-of-distribution generalization is a critical challenge in modern
machine learning (ML) research. For interpretability and behavioral
research in this area, training on algorithmic tasks offers benefits by
allowing systematic data generation and task decomposition, as well as
simplifying the process of circuit discovery
(\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
Although mazes are well suited for these investigations, we have found
that existing maze generation packages
(\protect\hyperlink{ref-cobbe2019procgen}{Cobbe et al., 2019};
\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
\protect\hyperlink{ref-harriesMazeExplorerCustomisable3D2019}{Harries et
al., n.d.}; \protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}) do not support flexible maze generation
algorithms that provide fine-grained control of generation parameters
and the ability to easily transform between multiple representations of
the mazes (Images, Textual, Tokenized) for training and testing models.

This work aims to facilitate deeper research into generalization and
interpretability by addressing these limitations. We introduce
\href{https://github.com/understanding-search/maze-dataset}{\texttt{maze-dataset}},
an accessible Python package
(\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et al.,
2023a}). This package offers flexible configuration options for maze
dataset generation, allowing users to select from a range of algorithms
and adjust corresponding parameters (Section
\protect\hyperlink{generation}{Generation}). Furthermore, this package
supports various output formats tailored to different ML architectures
(Section \protect\hyperlink{visual-output-formats}{Visual Output
Formats}, Section \protect\hyperlink{tokenized-output-formats}{Tokenized
Output Formats}).

TODO: links to docs, not just github

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth,height=\textheight]{figures/demo-dataset.pdf}
\caption{Example mazes from various algorithms. Left to right:
randomized depth-first search (RDFS), RDFS without forks, constrained
RDFS, Wilson's (\protect\hyperlink{ref-wilson}{Wilson, 1996}), RDFS with
percolation (p=0.1), RDFS with percolation (p=0.4), random stack RDFS.}
\end{figure}

\hypertarget{generation}{%
\section{Generation and Usage}\label{generation}}

Our package can be installed from
\href{https://pypi.org/project/maze-dataset/}{PyPi} via
\texttt{pip\ install\ maze-dataset}, or directly from the
\href{https://github.com/understanding-search/maze-dataset}{git
repository} (\protect\hyperlink{ref-maze-dataset-github}{M. I.
Ivanitskiy et al., 2023a}).

To create a dataset, we first create a \texttt{MazeDatasetConfig}
configuration object, which specifies the seed, number, and size of
mazes, as well as the generation algorithm and its corresponding
parameters. This object is passed to a \texttt{MazeDataset} class to
create a dataset. Crucially, this \texttt{MazeDataset} inherits from a
PyTorch (\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
\texttt{Dataset}, and can thus be easily incorporated into existing data
pre-processing and training pipelines, e.g., through the use of a
\texttt{Dataloader} class.

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }
\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{, }
\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }
\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs,}
\NormalTok{)}
\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(cfg)}
\end{Highlighting}
\end{Shaded}

When initializing mazes, further configuration options can be specified
through the \texttt{from\_config()} factory method as necessary. Options
include 1) whether to generate the dataset during runtime or load an
existing dataset, 2) if and how to parallelize generation, and 3) where
to store the generated dataset. Full documentation of configuration
options is available in our repository
(\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et al.,
2023a}). Available maze generation algorithms are static methods of the
\texttt{LatticeMazeGenerators} class and include the following:

\begin{itemize}
\tightlist
\item
  \texttt{gen\_dfs} \textbf{(randomized depth-first search)}: Parameters
  can be passed to constrain the number of accessible cells, the number
  of forks in the maze, and the maximum tree depth. Creates a spanning
  tree by default or a partially spanning tree if constrained.
\item
  \texttt{gen\_wilson} \textbf{(Wilson's algorithm)}: Generates a random
  spanning tree via loop-erased random walk
  (\protect\hyperlink{ref-wilson}{Wilson, 1996}).
\item
  \texttt{gen\_percolation} \textbf{(percolation)}: Starting with no
  connections, every possible lattice connection is set to either true
  or false with some probability \(p\), independently of all other
  connections. For the kinds of graphs that this process generates, we
  refer to existing work
  (\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
  \protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam,
  2004}).
\item
  \texttt{gen\_dfs\_percolation} \textbf{(randomized depth-first search
  with percolation)}: A connection exists if it exists in a maze
  generated via \texttt{gen\_dfs\ OR\ gen\_percolation}. Useful for
  generating mazes that are not acyclic graphs.
\end{itemize}

Furthermore, a dataset of mazes can be filtered to satisfy certain
properties:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dataset\_filtered: MazeDataset }\OperatorTok{=}\NormalTok{ dataset.filter\_by.path\_length(min\_length}\OperatorTok{=}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Custom filters can be specified, and several filters are included:

\begin{itemize}
\tightlist
\item
  \texttt{path\_length(min\_length:\ int)}: shortest length from the
  origin to target should be at least \texttt{min\_length}.
\item
  \texttt{start\_end\_distance(min\_distance:\ int)}: Manhattan distance
  between start and end should be at least \texttt{min\_distance},
  ignoring walls.
\item
  \texttt{remove\_duplicates(...)}: remove mazes which are similar to
  others in the dataset, measured via Hamming distance.
\item
  \texttt{remove\_duplicates\_fast()}: remove mazes which are exactly
  identical to others in the dataset.
\end{itemize}

All implemented maze generation algorithms are stochastic by nature. For
reproducibility, the \texttt{seed} parameter of
\texttt{MazeDatasetConfig} may be set. In practice, we do not find that
exact duplicates of mazes are generated with any meaningful frequency,
even when generating large datasets.

\hypertarget{visual-output-formats}{%
\section{Visual Output Formats}\label{visual-output-formats}}

Internally, mazes are \texttt{SolvedMaze} objects, which have path
information, and a connection list optimized for storing sub-graphs of a
lattice. These objects can be converted to and from several formats.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3250}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3250}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\texttt{as\_ascii()}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\texttt{as\_pixels()}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\texttt{MazePlot()}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Simple text format for displaying mazes, useful for debugging in a
terminal environment. & \texttt{numpy} array of \texttt{dtype=uint8} and
shape \texttt{(height,\ width,\ 3)}. The last dimension is RGB color. &
feature-rich plotting utility with support for multiple paths, heatmaps
over positions, and more. \\
\end{longtable}

\begin{figure}
\centering
\includegraphics{figures/output-fmts.pdf}
\caption{Various output formats. Top row (left to right): ASCII diagram,
rasterized pixel grid, and advanced display.}
\end{figure}

\hypertarget{training}{%
\subsection{Visual Outputs for Training and Evaluation}\label{training}}

In previous work, maze tasks have been used with Recurrent Convolutional
Neural Network (RCNN) derived architectures
(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
Huang, et al., 2021}). To facilitate the use of our package in this
context, we replicate the format of
(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}) and provide the \texttt{RasterizedMazeDataset}
class which returns rasterized pairs of (input, target) mazes as shown
in Figure below.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
\caption{Input is the rasterized maze without the path marked (left),
and provide as a target the maze with all but the correct path removed.
Configuration options exist to adjust whether endpoints are included and
if empty cells should be filled in.}
\end{figure}

\hypertarget{tokenized-output-formats}{%
\section{Tokenized Output Formats}\label{tokenized-output-formats}}

To train autoregressive text models such as transformers, we convert
mazes to token sequences in two steps. First, the maze is stringified
using \texttt{as\_tokens()}. The \texttt{MazeTokenizerModular} class
provides a powerful interface for configuring maze stringification
behavior. Second, the sequence of strings is tokenized into integers
using \texttt{encode()}. Tokenization uses a fixed vocabulary for
simplicity. Mazes up to 50x50 are supported using unique tokens, and up
to 128x128 when using coordinate tuple tokens.

\hypertarget{mtm}{%
\subsection{\texorpdfstring{Stringification Options and
\texttt{MazeTokenizerModular}}{Stringification Options and MazeTokenizerModular}}\label{mtm}}

There are many algorithms by which one might tokenize a 2D maze into a
1D format usable by autoregressive text models. Training multiple models
on the encodings output from each of these algorithms may produce very
different internal representations, learned solution algorithms, and
levels of performance. To explore how different maze tokenization
algorithms affect these models, the \texttt{MazeTokenizerModular} class
contains a rich set of options to customize how mazes are stringified.
This class contains 19 discrete parameters, resulting in 5.9 million
unique tokenizers. But wait, there's more! There are 6 additional
parameters available in the library which are untested but further
expand the the number of tokenizers by a factor of \(44/3\) to 86
million.

All output sequences consist of four token regions representing
different features of the maze. These regions are distinguished by color
in Figure below.

\begin{itemize}
\tightlist
\item
  {Adjacency list}: A text representation of the lattice graph
\item
  {Origin}: Starting coordinate
\item
  {Target}: Ending coordinate
\item
  {Path}: Maze solution sequence from the start to the end
\end{itemize}

\begin{figure}
\centering
\includegraphics{figures/outputs-tokens-colored.tex}
\caption{Example text output format with token regions highlighted.}
\end{figure}

Each \texttt{MazeTokenizerModular} is constructed from a set of several
\texttt{\_TokenizerElement} objects, each of which specifies how
different token regions or other elements of the stringification are
produced.

\begin{figure}
\centering
\includegraphics{figures/TokenizerElement_structure.pdf}
\caption{Nested internal structure of \texttt{\_TokenizerElement}
objects inside a typical \texttt{MazeTokenizerModular} object.}
\end{figure}

Optional delimiter tokens may be added in many places in the output.
Delimiter options are all configured using the parameters named
\texttt{pre}, \texttt{intra}, and \texttt{post} in various
\texttt{\_TokenizerElement} classes. Each option controls a unique
delimiter token. Here we describe each \texttt{\_TokenizerElement} and
the behaviors they support. We also discuss some of the model behaviors
and properties that may be investigated using these options.

\hypertarget{coordtokenizer}{%
\subsubsection{Coordinates}\label{coordtokenizer}}

The \texttt{\_CoordTokenizer} object controls how coordinates in the
lattice are represented in across all token regions. Options include:

\begin{itemize}
\tightlist
\item
  \textbf{Unique tokens}: Each coordinate is represented as a single
  unique token \texttt{"(i,j)"}
\item
  \textbf{Coordinate tuple tokens}: Each coordinate is represented as a
  sequence of 2 tokens, respectively encoding the row and column
  positions: \texttt{{[}"i",\ ",",\ "j"{]}}
\end{itemize}

\hypertarget{adjlisttokenizer}{%
\subsubsection{Adjacency List}\label{adjlisttokenizer}}

The \texttt{\_AdjListTokenizer} object controls this token region. All
tokenizations represent the maze connectivity as a sequence of
connections or walls between pairs of adjacent coordinates in the
lattice.

\begin{itemize}
\tightlist
\item
  \texttt{\_EdgeSubset}: Specifies the subset of lattice edges to be
  tokenized

  \begin{itemize}
  \tightlist
  \item
    \textbf{All edges}: Every edge in the lattice
  \item
    \textbf{Connections}: Only edges which contain a connection
  \item
    \textbf{Walls}: Only edges which contain a wall
  \end{itemize}
\item
  \texttt{\_EdgePermuter}: Specifies how to sequence the two coordinates
  in each lattice edge

  \begin{itemize}
  \tightlist
  \item
    \textbf{Random}
  \item
    \textbf{Sorted}: The smaller coordinate always comes first
  \item
    \textbf{Both permutations}: Each edge is represented twice, once
    with each permutation. This option attempts to represent connections
    in a more directionally symmetric manner. Including only one
    permutation of each edge may affect models' internal representations
    of edges, treating a path traversing the edge differently depending
    on if the coordinate sequence in the path matches the sequence in
    the adjacency list.
  \end{itemize}
\item
  \texttt{shuffle\_d0}: Whether to shuffle the edges randomly or sort
  them in the output by their first coordinate
\item
  \texttt{connection\_token\_ordinal}: Location in the sequence of the
  token representing whether the edge is a connection or a wall
\end{itemize}

\hypertarget{pathtokenizer}{%
\subsubsection{Path}\label{pathtokenizer}}

The \texttt{\_PathTokenizer} object controls this token region. Paths
are all represented as a sequence of steps moving from the start to the
end position.

\begin{itemize}
\tightlist
\item
  \texttt{\_StepSize}: Specifies the size of each step

  \begin{itemize}
  \tightlist
  \item
    \textbf{Singles}: Every coordinate traversed between start and end
    is directly represented
  \item
    \textbf{Forks}: Only coordinates at forking points in the maze are
    represented. The paths between forking points are implicit. Using
    this option might train models more directly to represent forking
    points differently from coordinates where the maze connectivity
    implies an obvious next step in the path.
  \end{itemize}
\item
  \texttt{\_StepTokenizer}: Specifies how an individual step is
  represented

  \begin{itemize}
  \tightlist
  \item
    \textbf{Coordinate}: The coordinates of each step are directly
    tokenized using a \texttt{\_CoordTokenizer}
  \item
    \textbf{Cardinal direction}: A single token corresponding to the
    cardinal direction taken at the starting position of that step.
    E.g., \texttt{NORTH}, \texttt{SOUTH}. If using a \texttt{\_StepSize}
    other than \textbf{Singles}, this direction may not correspond to
    the final direction traveled to arrive at the end position of the
    step.
  \item
    \textbf{Relative direction}: A single token corresponding to the
    first-person perspective relative direction taken at the starting
    position of that step. E.g., \texttt{RIGHT}, \texttt{LEFT}.
  \item
    \textbf{Distance}: A single token corresponding to the number of
    coordinate positions traversed in that step. E.g., using a
    \texttt{\_StepSize} of \textbf{Singles}, the \textbf{Distance} token
    would be the same for each step, corresponding to a distance of 1
    coordinate. This option is only of interest in combination with a
    \texttt{\_StepSize} other than \textbf{Singles}.
  \end{itemize}
\end{itemize}

A \texttt{\_PathTokenizer} contains a sequence of one or more unique
\texttt{\_StepTokenizer} objects. Different step representations may be
mixed and permuted, allowing for investigation of model representations
of multiple aspects of a maze solution at once.

\hypertarget{token-training}{%
\subsection{Tokenized Outputs for Training and
Evaluation}\label{token-training}}

During deployment we provide only the prompt up to the
\texttt{\textless{}PATH\_START\textgreater{}} token.

Examples of usage of this dataset to train autoregressive transformers
can be found in our \texttt{maze-transformer} library
(\protect\hyperlink{ref-maze-transformer-github}{M. I. Ivanitskiy et
al., 2023b}). Other tokenization and vocabulary schemes are also
included, such as representing each coordinate as a pair of \(i,j\)
index tokens.

\hypertarget{extensibility}{%
\subsection{Extensibility}\label{extensibility}}

The tokenizer architecture is purposefully designed such that adding and
testing a wide variety of new tokenization algorithms is fast and
minimizes disturbances to functioning code. This is enabled by the
modular architecture and the automatic inclusion of any new tokenizers
in integration tests. To create a new tokenizer, developers forking the
library may simply create their own \texttt{\_TokenizerElement} subclass
and implement the abstract methods. If the behavior change is
sufficiently small, simply adding a parameter to an existing
\texttt{\_TokenizerElement} subclass and updating its implementation
will suffice. For small additions, simply adding new cases to existing
unit tests will suffice.

The breadth of tokenizers is also easily scaled in the opposite
direction. Due to the exponential scaling of parameter combinations,
adding a small number of new features can significantly slow certain
procedures which rely on constructing all possible tokenizers, such as
integration tests. If any existing subclass contains features which
aren't needed, a developer tool decorator is provided which can be
applied to the unneeded \texttt{\_TokenizerElement} subclasses to prune
those features and compact the available space of tokenizers.

\hypertarget{benchmarks}{%
\section{Benchmarks of Generation Speed}\label{benchmarks}}

We provide approximate benchmarks for relative generation time across
various algorithms, parameter choices, maze sizes, and dataset sizes.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1680}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2080}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1360}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1520}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1360}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Method \& Parameters
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Average time per maze (ms)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Generation algorithm & Generation parameters & all sizes & small
(\(g \leq 10\)) & medium (\(10 < g \leq 32\)) & large (\(g > 32\)) \\
:===============: & :===============: & :===============: &
:===============: & :===============: & :===============: \\
gen\_dfs & accessible\_cells=20 & 2.4 & 2.4 & 2.6 & 2.4 \\
gen\_dfs & do\_forks=False & 3.0 & 2.4 & 3.7 & 3.8 \\
gen\_dfs & max\_tree\_depth=0.5 & 4.5 & 2.2 & 4.9 & 11.6 \\
gen\_dfs & -- & 31.1 & 2.8 & 28.0 & 136.5 \\
gen\_dfs\_percolation & p=0.1 & 53.9 & 3.6 & 42.5 & 252.9 \\
gen\_dfs\_percolation & p=0.4 & 58.8 & 3.7 & 44.7 & 280.2 \\
gen\_percolation & -- & 59.1 & 3.3 & 43.6 & 285.2 \\
gen\_wilson & -- & 767.9 & 10.1 & 212.9 & 4530.4 \\
:===============: & :===============: & :===============: &
:===============: & :===============: & :===============: \\
\textbf{median (all runs)} & & 10.8 & 6.0 & 44.4 & 367.7 \\
\textbf{mean (all runs)} & & 490.0 & 11.7 & 187.2 & 2769.6 \\
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth,height=\textheight]{figures/gridsize-vs-gentime.pdf}
\caption{Plots of maze generation time. Generation time scales
exponentially with maze size for all algorithms (left). Generation time
does not depend on the number of mazes being generated, and there is
minimal overhead to initializing the generation process for a small
dataset (right). Wilson's algorithm is notably less efficient than
others and has high variance. Note that for both plots, values are
averaged across all parameter sets for that algorithm, and
parallelization is disabled.}
\end{figure}

\hypertarget{implementation}{%
\section{Implementation}\label{implementation}}

We refer to our GitHub repository
(\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et al.,
2023a}) for documentation and up-to-date implementation details.

This package utilizes a simple, efficient representation of mazes. Using
an adjacency list to represent mazes would lead to a poor lookup time of
whether any given connection exists, whilst using a dense adjacency
matrix would waste memory by failing to exploit the structure (e.g.,
only 4 of the diagonals would be filled in). Instead, we describe mazes
with the following simple representation: for a \(d\)-dimensional
lattice with \(r\) rows and \(c\) columns, we initialize a boolean array
\(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
a \texttt{connection\_list}. The value at \(A[0,i,j]\) determines
whether a downward connection exists from node \([i,j]\) to
\([i+1, j]\). Likewise, the value at \(A[1,i,j]\) determines whether a
rightwards connection to \([i, j+1]\) exists. Thus, we avoid duplication
of data about the existence of connections, at the cost of requiring
additional care with indexing when looking for a connection upwards or
to the left. Note that this setup allows for a periodic lattice.

To produce solutions to mazes, two points are selected uniformly at
random without replacement from the connected component of the maze, and
the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
1968}) is applied to find the shortest path between them.

Parallelization is implemented via the \texttt{multiprocessing} module
in the Python standard library, and parallel generation can be
controlled via keyword arguments to the
\texttt{MazeDataset.from\_config()} function.

\hypertarget{relation-work}{%
\subsection{Relation to Existing Works}\label{relation-work}}

As mentioned in the introduction, a multitude of public and open-source
software packages exist for generating mazes
(\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}). However, our package provides more flexibility
and efficiency in the following ways:

\begin{itemize}
\tightlist
\item
  For rigorous investigations of the response of a model to various
  distributional shifts, preserving metadata about the generation
  algorithm with the dataset itself is essential. To this end, our
  package efficiently stores the dataset along with its metadata in a
  single human-readable file (\protect\hyperlink{ref-zanj}{M.
  Ivanitskiy, n.d.}). This metadata is loaded when the dataset is
  retrieved from disk and reduces the complexity of discerning the
  parameters under which a dataset was created.
\item
  Prior works provide maze datasets in only a rasterized format, which
  is not suitable for training autoregressive text-based transformer
  models. As discussed in Section
  \protect\hyperlink{visual-output-formats}{Visual Output Formats} and
  Section \protect\hyperlink{tokenized-output-formats}{Tokenized Output
  Formats}, our package provides these different formats natively.
\item
  Our package provides a selection of maze generation algorithms, which
  all write to a single unified format. All output formats are
  reversible, and operate to and from this unified format.
\end{itemize}

As mentioned in Section \protect\hyperlink{training}{Training}, we also
include the \texttt{RasterizedMazeDataset} class in our codebase, which
can exactly mimic the outputs provided in
\texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
Borgnia, Gupta, Bansal, et al., 2021}). Our \texttt{as\_ascii()} method
provides a format similar to that used in
(\protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023}). The text format
provided by \texttt{as\_tokens()} is similar to that of
(\protect\hyperlink{ref-eval-LLM-graphs}{Liu \& Wu, 2023}), but provides
a custom tokenization scheme.

\hypertarget{limitations}{%
\subsection{\texorpdfstring{Limitations of
\texttt{maze-dataset}}{Limitations of maze-dataset}}\label{limitations}}

For simplicity, the package primarily supports mazes that are sub-graphs
of a 2-dimensional rectangular lattice. Some support for
higher-dimensional lattices is present, but not all output formats are
adapted for higher dimensional mazes.
\protect\hyperlink{implementation-implementation}{Implementation}
\protect\hyperlink{implementation-implementation}{Implementation}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The
\href{https://github.com/understanding-search/maze-dataset}{\texttt{maze-dataset}}
library (\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et
al., 2023a}) introduced in this paper provides a flexible and extensible
toolkit for generating, processing, and analyzing maze datasets. By
supporting various procedural generation algorithms and conversion
utilities, it enables the creation of mazes with customizable properties
to suit diverse research needs. Planned improvements to the
\texttt{maze-dataset} include adding more generation algorithms (such as
Prim's algorithm (\protect\hyperlink{ref-dijkstra-prim}{Dijkstra, 1959};
\protect\hyperlink{ref-jarnik-prim}{Jarnık, 1930};
\protect\hyperlink{ref-prim}{Prim, 1957}) and Kruskal's algorithm
(\protect\hyperlink{ref-kruskal}{Kruskal, 1956}), among others
(\protect\hyperlink{ref-mazegen_analysis}{Gabrovšek, 2019})), adding the
ability to augment a maze with an adjacency list to add ``shortcuts'' to
the maze, and resolving certain limitations detailed in Section
\protect\hyperlink{limitations}{Limitations}. Future work will make
extensive use of this library to study interpretability and
out-of-distribution generalization in autoregressive transformers
(\protect\hyperlink{ref-maze-transformer-github}{M. I. Ivanitskiy et
al., 2023b}), recurrent convolutional neural networks
(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
Huang, et al., 2021}), and implicit networks
(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022};
\protect\hyperlink{ref-mckenzie2023faster}{McKenzie et al., 2023}).

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This work was partially supported by and many of the authors were
brought together by AI Safety Camp and AI Safety Support. This work was
partially funded by National Science Foundation awards DMS-2110745 and
DMS-2309810. We thank the Mines Optimization and Deep Learning group
(MODL) for fruitful discussions. We also thank Michael Rosenberg for
advice on Finite State Transducers.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0.5}
\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
procedural generation to benchmark reinforcement learning. \emph{arXiv
Preprint arXiv:1912.01588}.

\leavevmode\vadjust pre{\hypertarget{ref-dijkstra-prim}{}}%
Dijkstra, E. W. (1959). \emph{A note on two problems in connexion with
graphs:({Numerische Mathematik}, 1 (1959), p 269-271)}.

\leavevmode\vadjust pre{\hypertarget{ref-percolation}{}}%
Duminil-Copin, H. (2017). \emph{Sixty years of percolation} (No.
arXiv:1712.04651). {arXiv}. \url{http://arxiv.org/abs/1712.04651}

\leavevmode\vadjust pre{\hypertarget{ref-gh_Ehsan_2022}{}}%
Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}

\leavevmode\vadjust pre{\hypertarget{ref-percolation-clustersize}{}}%
Fisher, M. E., \& Essam, J. W. (2004). Some {Cluster Size} and
{Percolation Problems}. \emph{Journal of Mathematical Physics},
\emph{2}(4), 609--619. \url{https://doi.org/10.1063/1.1703745}

\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
(2022). Jfb: Jacobian-free backpropagation for implicit networks.
\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
\emph{36}, 6648--6656.

\leavevmode\vadjust pre{\hypertarget{ref-mazegen_analysis}{}}%
Gabrovšek. (2019). Analysis of maze generating algorithms. \emph{{IPSI
Transactions} on {Internet Research}}, \emph{15.1}, 23--30.
\url{http://www.ipsitransactions.org/journals/papers/tir/2019jan/p5.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
{Games CoG}}, 1--4.

\leavevmode\vadjust pre{\hypertarget{ref-A_star}{}}%
Hart, P. E., Nilsson, N. J., \& Raphael, B. (1968). A {Formal Basis} for
the {Heuristic Determination} of {Minimum Cost Paths}. \emph{IEEE
Transactions on Systems Science and Cybernetics}, \emph{4}(2), 100--107.
\url{https://doi.org/10.1109/TSSC.1968.300136}

\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
Ivanitskiy, M. (n.d.). \emph{ZANJ}.
\url{https://github.com/mivanit/ZANJ}

\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
Ivanitskiy, M. I., Shah, R., Spies, A. F., Räuker, T., Valentine, D.,
Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a). \emph{Maze
dataset}. \url{https://github.com/understanding-search/maze-dataset}

\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
Ivanitskiy, M. I., Shah, R., Spies, A. F., Räuker, T., Valentine, D.,
Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b). \emph{Maze
transformer interpretability}.
\url{https://github.com/understanding-search/maze-transformer}

\leavevmode\vadjust pre{\hypertarget{ref-jarnik-prim}{}}%
Jarnık, V. (1930). About a certain minimal problem. \emph{Práce Moravské
Prırodovedecké Spolecnosti}, \emph{6}, 57--63.

\leavevmode\vadjust pre{\hypertarget{ref-kruskal}{}}%
Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
the traveling salesman problem. \emph{Proceedings of the American
Mathematical Society}, \emph{7}(1), 48--50.
\url{https://doi.org/10.1090/S0002-9939-1956-0078686-7}

\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
Performance insights and comparative analysis. \emph{arXiv Preprint
arXiv:2308.11224}.

\leavevmode\vadjust pre{\hypertarget{ref-mckenzie2023faster}{}}%
McKenzie, D., Fung, S. W., \& Heaton, H. (2023). Faster
predict-and-optimize with three-operator splitting. \emph{arXiv Preprint
arXiv:2301.13395}.

\leavevmode\vadjust pre{\hypertarget{ref-gh_Nemeth_2019}{}}%
Németh, F. (2019). \emph{Maze-generation-algorithms}.
\url{https://github.com/ferenc-nemeth/maze-generation-algorithms}

\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G.,
Killeen, T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., Kopf,
A., Yang, E., DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S.,
Steiner, B., Fang, L., \ldots{} Chintala, S. (2019). {PyTorch: An
Imperative Style, High-Performance Deep Learning Library}. In H.
Wallach, H. Larochelle, A. Beygelzimer, F. d'Alché-Buc, E. Fox, \& R.
Garnett (Eds.), \emph{Advances in neural information processing systems
32} (pp. 8024--8035). Curran Associates, Inc.
\url{http://papers.neurips.cc/paper/9015-pytorch-an-imperative-style-high-performance-deep-learning-library.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-prim}{}}%
Prim, R. C. (1957). Shortest connection networks and some
generalizations. \emph{The Bell System Technical Journal}, \emph{36}(6),
1389--1401.

\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
transparent ai: A survey on interpreting the inner structures of deep
neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
Machine Learning (SaTML)}, 464--483.

\leavevmode\vadjust pre{\hypertarget{ref-easy_to_hard}{}}%
Schwarzschild, A., Borgnia, E., Gupta, A., Bansal, A., Emam, Z., Huang,
F., Goldblum, M., \& Goldstein, T. (2021). \emph{Datasets for {Studying
Generalization} from {Easy} to {Hard Examples}} (No. arXiv:2108.06011).
{arXiv}. \url{https://doi.org/10.48550/arXiv.2108.06011}

\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
Generalizing from easy to hard problems with recurrent networks.
\emph{Advances in Neural Information Processing Systems}, \emph{34},
6695--6706.

\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
\emph{arXiv Preprint arXiv:2308.02522}.

\leavevmode\vadjust pre{\hypertarget{ref-wilson}{}}%
Wilson, D. B. (1996). Generating random spanning trees more quickly than
the cover time. \emph{Proceedings of the Twenty-Eighth Annual {ACM}
Symposium on {Theory} of Computing - {STOC} '96}, 296--303.
\url{https://doi.org/10.1145/237814.237880}

\end{CSLReferences}

\end{document}

### 2025-03-24 maze_dataset/generation/__init__.py
some docs

@@ -1,7 +1,6 @@
 """generation functions have signature `(grid_shape: Coord, **kwargs) -> LatticeMaze` and are methods in `LatticeMazeGenerators`
 
 `DEFAULT_GENERATORS` is a list of generator name, generator kwargs pairs used in tests and demos
-
 """
 
 from maze_dataset.generation.generators import (

### 2025-03-25 docs/paper/paper.tex
modify paper, recompile

@@
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Rusheb Shah, Brandon Knutson, Cecilia Diniz Behn, Samy Wu Fung},
+Spies, Rusheb Shah, Brandon Knutson, , Cecilia Diniz Behn, Samy Wu
+Fung},
     pdfpublication={Journal of Open Source Software},
@@
     pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Rusheb Shah, Brandon Knutson, Cecilia
+Aaron Sandoval, Alex F. Spies, Rusheb Shah, Brandon Knutson, , Cecilia
 Diniz Behn, Samy Wu Fung},
@@
 %
+]{%
+  \,\orcidlink{0009-0004-8413-0239}\,%
+}
+\author[1%
+%
 ]{Cecilia Diniz Behn%
@@
 
-Understanding how machine learning models respond to distributional
-shifts and different input representations is a key research challenge.
-Mazes serve as an excellent testbed for this problem due to the degree
-of control we have over quantifiable distributional shifts. To enable
-systematic investigations of model behavior on maze tasks, we present
-\texttt{maze-dataset}, a comprehensive library for generating,
-processing, visualizing, and tokenizing datasets consisting of
-maze-solving tasks. With this library, researchers can easily create
-datasets and have extensive control over the generation algorithm used,
-the parameters fed to the algorithm of choice, and the filters that
-generated mazes must satisfy. Furthermore, this library supports
-multiple output formats, including rasterized and a large number of
-text-based formats, catering to convolutional neural networks and
-autoregressive transformer models. These formats, along with tools for
-visualizing and converting between them, ensure versatility and
-adaptability in research applications.
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Out-of-distribution generalization is a critical challenge in modern
-machine learning (ML) research. For interpretability and behavioral
-research in this area, training on algorithmic tasks offers benefits by
-allowing systematic data generation and task decomposition, as well as
-simplifying the process of circuit discovery
+Solving mazes is a classic problem in computer science and artificial
+intelligence, and humans have been constructing mazes for thousands of
+years. Although finding the shortest path through a maze is a solved
+problem, this makes it an excellent testbed for studying how machine
+learning algorithms solve problems and represent spatial information. In
+this paper, we introduce \texttt{maze-dataset}, a Python library for
+generating, processing, and visualizing datasets of mazes. This library
+supports a variety of maze generation algorithms providing both mazes
+with loops and ``perfect'' mazes without them. These generation
+algorithms can be configured with various parameters, and the resulting
+mazes can be filtered to satisfy a desired property. Also provided are
+tools for converting mazes to and from various formats, such as
+rasterized images and tokenized text sequences, making it suitable for
+training or evaluating convolutional neural networks and autoregressive
+transformer models, as well as various visualization tools. As well as
+providing a simple interface for generating, storing, and loading these
+datasets, \texttt{maze-dataset} is extensively tested, type hinted,
+benchmarked, and documented.
+
+\textbf{TODO: mention the examples, put in a figure
+https://understanding-search.github.io/maze-dataset/examples/maze\_examples.html}
+
+\hypertarget{statement-of-need}{%
+\section{Statement of Need}\label{statement-of-need}}
+
+The generation of mazes with a given algorithm is not inherently a
+complex task, but the ability to seamlessly switch out algorithms,
+modify algorithm parameters, or filter by desired properties all while
+preserving the ability to convert between different representations of
+the maze is not trivial. This library aims to greatly streamline the
+process of generating and working with datasets of mazes that can be
+described as subgraphs of an \(n \times n\) lattice with boolean
+connections and, optionally, start and end points that are nodes in the
+graph. Furthermore, we place emphasis on a wide variety of possible text
+output formats aimed at evaluating the spatial reasoning capabilities of
+Large Language Models and other text-based transformer models.
+
+For interpretability and behavioral research, algorithmic tasks offer
+benefits by allowing systematic data generation and task decomposition,
+as well as simplifying the process of circuit discovery
 (\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
@@
 
-This work aims to facilitate deeper research into generalization and
-interpretability by addressing these limitations. We introduce
-\href{https://github.com/understanding-search/maze-dataset}{\texttt{maze-dataset}},
-an accessible Python package
-(\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et al.,
-2023a}). This package offers flexible configuration options for maze
-dataset generation, allowing users to select from a range of algorithms
-and adjust corresponding parameters (Section
-\protect\hyperlink{generation}{Generation}). Furthermore, this package
-supports various output formats tailored to different ML architectures
-(Section \protect\hyperlink{visual-output-formats}{Visual Output
-Formats}, Section \protect\hyperlink{tokenized-output-formats}{Tokenized
-Output Formats}).
-
-TODO: links to docs, not just github
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\textwidth,height=\textheight]{figures/demo-dataset.pdf}
-\caption{Example mazes from various algorithms. Left to right:
-randomized depth-first search (RDFS), RDFS without forks, constrained
-RDFS, Wilson's (\protect\hyperlink{ref-wilson}{Wilson, 1996}), RDFS with
-percolation (p=0.1), RDFS with percolation (p=0.4), random stack RDFS.}
-\end{figure}
+\hypertarget{related-works}{%
+\subsection{Related Works}\label{related-works}}
+
+A multitude of public and open-source software packages exist for
+generating mazes (\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
+\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
+\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
+Bansal, et al., 2021}). However, nearly all of these packages generate
+and store mazes in a form that is not optimized for storage space or,
+more importantly, computer readability. The mazes produces by other
+packages are usually rasterized or in some form of image, rather than
+the underlying graph structure, and this makes it difficult to work with
+these datasets.
+
+\begin{itemize}
+\item
+  Most prior works provide mazes in some kind of image or raster format,
+  which is not suitable for training autoregressive text-based
+  transformer models -- a key usage case this work seeks to enable.
+  However, we still provide a variety of similar output formats:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    we also include the \texttt{RasterizedMazeDataset} class, utilizing
+    \texttt{as\_pixels()}, in our codebase, which can exactly mimic the
+    outputs provided in
+    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
+    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
+    similar to the outputs of
+    (\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019}).
+  \item
+    Our \texttt{as\_ascii()} method provides a format similar to that
+    used in Oppenheim
+    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{2018}).
+  \item
+    Our \texttt{MazePlot} class provides a feature-rich plotting utility
+    with support for multiple paths, heatmaps over positions, and more.
+    This is similar to the outputs of Ehsan
+    (\protect\hyperlink{ref-gh_Ehsan_2022}{2022})
+  \end{itemize}
+\item
+  The text format provided by \texttt{SolvedMaze(...).as\_tokens()} is
+  similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \& Wu,
+  2023}), but provides over 5.8 million unique formats for converting
+  mazes to a text stream. We maintain a single underlying format,
+  meaning that the same maze can be turned into a variety of text
+  streams to assess how the precise format of the text stream affects
+  the model.
+\item
+  For rigorous investigations of the response of a model to various
+  distributional shifts, preserving metadata about the generation
+  algorithm with the dataset itself is essential. To this end, our
+  package efficiently stores the dataset along with its metadata in a
+  single human-readable file (\protect\hyperlink{ref-zanj}{M.
+  Ivanitskiy, n.d.}). This metadata is loaded when the dataset is
+  retrieved from disk and makes it simple to understand how exactly each
+  maze was generated. As far as we are aware, no existing packages do
+  this reliably.
+\item
+  Storing mazes as images is not only difficult to work with, but also
+  inefficient. Directly storing adjacency matrices is also inefficient
+  as subgraphs of the lattice are sparse. Storing adjacency lists can be
+  efficient, but comes with a higher lookup cost and possible high
+  comparison cost. We use a simple, efficient representation of mazes
+  that is optimized for subgraphs of a \(d\)-dimensional finite lattice
+  that we do not believe is used in any existing maze generation
+  package.
+\item
+  Our package is easily installable with source code freely available.
+  It is extensively tested, type hinted, benchmarked, and documented.
+  Many other maze generation packages lack this level of rigor and
+  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
+  2008}) appear to simply no longer be accessible.
+\end{itemize}
+
+\hypertarget{features-and-implementation}{%
+\section{Features and
+Implementation}\label{features-and-implementation}}
 
 \hypertarget{generation}{%
-\section{Generation and Usage}\label{generation}}
+\subsection{Generation and Usage}\label{generation}}
+
+\textbf{TODO: JOSS says don't document the API itself. we should talk
+about the capabilities, but just rewrite this to link to the
+documentation.}
 
@@
 \href{https://github.com/understanding-search/maze-dataset}{git
-repository} (\protect\hyperlink{ref-maze-dataset-github}{M. I.
+repository} (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
 Ivanitskiy et al., 2023a}).
@@
 options is available in our repository
-(\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et al.,
-2023a}). Available maze generation algorithms are static methods of the
-\texttt{LatticeMazeGenerators} class and include the following:
+(\protect\hyperlink{ref-maze-dataset-github}{Michael I. Ivanitskiy et
+al., 2023a}). Available maze generation algorithms are static methods of
+the \texttt{LatticeMazeGenerators} class and include the following:
 
@@
 can be found in our \texttt{maze-transformer} library
-(\protect\hyperlink{ref-maze-transformer-github}{M. I. Ivanitskiy et
-al., 2023b}). Other tokenization and vocabulary schemes are also
+(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
+et al., 2023b}). Other tokenization and vocabulary schemes are also
 included, such as representing each coordinate as a pair of \(i,j\)
@@
 We refer to our GitHub repository
-(\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et al.,
-2023a}) for documentation and up-to-date implementation details.
+(\protect\hyperlink{ref-maze-dataset-github}{Michael I. Ivanitskiy et
+al., 2023a}) for documentation and up-to-date implementation details.
 
@@
 
-\hypertarget{relation-work}{%
-\subsection{Relation to Existing Works}\label{relation-work}}
-
-As mentioned in the introduction, a multitude of public and open-source
-software packages exist for generating mazes
-(\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). However, our package provides more flexibility
-and efficiency in the following ways:
-
-\begin{itemize}
-\tightlist
-\item
-  For rigorous investigations of the response of a model to various
-  distributional shifts, preserving metadata about the generation
-  algorithm with the dataset itself is essential. To this end, our
-  package efficiently stores the dataset along with its metadata in a
-  single human-readable file (\protect\hyperlink{ref-zanj}{M.
-  Ivanitskiy, n.d.}). This metadata is loaded when the dataset is
-  retrieved from disk and reduces the complexity of discerning the
-  parameters under which a dataset was created.
-\item
-  Prior works provide maze datasets in only a rasterized format, which
-  is not suitable for training autoregressive text-based transformer
-  models. As discussed in Section
-  \protect\hyperlink{visual-output-formats}{Visual Output Formats} and
-  Section \protect\hyperlink{tokenized-output-formats}{Tokenized Output
-  Formats}, our package provides these different formats natively.
-\item
-  Our package provides a selection of maze generation algorithms, which
-  all write to a single unified format. All output formats are
-  reversible, and operate to and from this unified format.
-\end{itemize}
-
-As mentioned in Section \protect\hyperlink{training}{Training}, we also
-include the \texttt{RasterizedMazeDataset} class in our codebase, which
-can exactly mimic the outputs provided in
-\texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
-Borgnia, Gupta, Bansal, et al., 2021}). Our \texttt{as\_ascii()} method
-provides a format similar to that used in
-(\protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023}). The text format
-provided by \texttt{as\_tokens()} is similar to that of
-(\protect\hyperlink{ref-eval-LLM-graphs}{Liu \& Wu, 2023}), but provides
-a custom tokenization scheme.
-
 \hypertarget{limitations}{%
@@
 
+\hypertarget{usage-in-research}{%
+\section{Usage in Research}\label{usage-in-research}}
+
+This package was originally built for the needs of the
+(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
+et al., 2023b}) project, which aims to investigate spatial planning and
+world models in autoregressive transformer models trained on mazes Spies
+et al. (\protect\hyperlink{ref-spies2024causalworldmodels}{2024}). This
+project has also adapted itself to be useful for work on understanding
+the mechanisms by which recurrent convolutional and implicit networks
+(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
+given a rasterized view
+(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
+2024}), and for this we match the output format of
+(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
+Bansal, et al., 2021}).
+
+This package has also been utilized in work by other groups:
+
+\begin{itemize}
+\item
+  (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) use
+  \texttt{maze-dataset} to compare the effectiveness of transformers
+  trained with the MLM-\(\mathcal{U}\)
+  (\protect\hyperlink{ref-MLMU-kitouni2024factorization}{Kitouni et al.,
+  2024}) multistep prediction objective against standard autoregressive
+  training for multi-step planning on our maze task.
+\item
+  (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
+  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) use
+  \texttt{maze-dataset} to study the effectiveness of imperative
+  learning
+\end{itemize}
+
 \hypertarget{conclusion}{%
@@
 \href{https://github.com/understanding-search/maze-dataset}{\texttt{maze-dataset}}
-library (\protect\hyperlink{ref-maze-dataset-github}{M. I. Ivanitskiy et
-al., 2023a}) introduced in this paper provides a flexible and extensible
-toolkit for generating, processing, and analyzing maze datasets. By
-supporting various procedural generation algorithms and conversion
-utilities, it enables the creation of mazes with customizable properties
-to suit diverse research needs. Planned improvements to the
+library (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
+Ivanitskiy et al., 2023a}) introduced in this paper provides a flexible
+and extensible toolkit for generating, processing, and analyzing maze
+datasets. By supporting various procedural generation algorithms and
+conversion utilities, it enables the creation of mazes with customizable
+properties to suit diverse research needs. Planned improvements to the
 \texttt{maze-dataset} include adding more generation algorithms (such as
@@
 the maze, and resolving certain limitations detailed in Section
-\protect\hyperlink{limitations}{Limitations}. Future work will make
-extensive use of this library to study interpretability and
-out-of-distribution generalization in autoregressive transformers
-(\protect\hyperlink{ref-maze-transformer-github}{M. I. Ivanitskiy et
-al., 2023b}), recurrent convolutional neural networks
-(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
-Huang, et al., 2021}), and implicit networks
-(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022};
-\protect\hyperlink{ref-mckenzie2023faster}{McKenzie et al., 2023}).
+\protect\hyperlink{limitations}{Limitations}.
 
@@
 partially funded by National Science Foundation awards DMS-2110745 and
-DMS-2309810. We thank the Mines Optimization and Deep Learning group
-(MODL) for fruitful discussions. We also thank Michael Rosenberg for
-advice on Finite State Transducers.
+DMS-2309810. We are also grateful to LTFF and FAR Labs for hosting three
+of the authors for a Residency Visit, and to various members of FAR's
+technical staff for their advice. We thank the Mines Optimization and
+Deep Learning group (MODL) for fruitful discussions. We also thank
+Michael Rosenberg for recommending the usage of Finite State Transducers
+for storing tokenizer validation information.
 
@@
 \begin{CSLReferences}{1}{0.5}
+\leavevmode\vadjust pre{\hypertarget{ref-mazegenerator-net}{}}%
+Alance AB. (2019). \emph{Maze generator}.
+\url{http://www.mazegenerator.net}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
+Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
+1.0: A complete set of tools to prepare, present, and analyze
+navigational and spatial cognitive neuroscience experiments.
+\emph{Behavior Research Methods}, \emph{40}, 353--359.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
+Chen, X., Yang, F., \& Wang, C. (2024). iA \(\^{}*\): Imperative
+learning-based a \(\^{}*\) search for pathfinding. \emph{arXiv Preprint
+arXiv:2403.15870}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
+Guo, C., Barthelet, L., \& Morris, R. (2011). \emph{Maze generator and
+solver}. Wolfram Demonstrations Project,
+\url{https://demonstrations.wolfram.com/MazeGeneratorAndSolver/}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
-Ivanitskiy, M. I., Shah, R., Spies, A. F., Räuker, T., Valentine, D.,
-Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a). \emph{Maze
-dataset}. \url{https://github.com/understanding-search/maze-dataset}
+Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
+D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
+\emph{Maze dataset}.
+\url{https://github.com/understanding-search/maze-dataset}
 
 \leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
-Ivanitskiy, M. I., Shah, R., Spies, A. F., Räuker, T., Valentine, D.,
-Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b). \emph{Maze
-transformer interpretability}.
+Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
+D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
+\emph{Maze transformer interpretability}.
 \url{https://github.com/understanding-search/maze-transformer}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
+Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
+Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
+D., \& others. (2023). Structured world representations in maze-solving
+transformers. \emph{arXiv Preprint arXiv:2312.02566}.
 
@@
 Prırodovedecké Spolecnosti}, \emph{6}, 57--63.
+
+\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
+Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
+Ibrahim, M. (2024). The factorization curse: Which tokens you predict
+underlie the reversal curse and more. \emph{Advances in Neural
+Information Processing Systems}, \emph{37}, 112329--112355.
+
+\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
+Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
+Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
+extrapolation for mazes with recurrent and implicit networks.
+\emph{arXiv Preprint arXiv:2410.03020}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mckenzie2023faster}{}}%
-McKenzie, D., Fung, S. W., \& Heaton, H. (2023). Faster
-predict-and-optimize with three-operator splitting. \emph{arXiv Preprint
-arXiv:2301.13395}.
+\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
+Nag, A. (2020). MDL suite: A language, generator and compiler for
+describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
+1815.
 
@@
 \url{https://github.com/ferenc-nemeth/maze-generation-algorithms}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
+Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
+Transformers can navigate mazes with multi-step prediction. \emph{arXiv
+Preprint arXiv:2412.05117}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gh-oppenheimj2018maze}{}}%
+Oppenheim, J. (2018). \emph{Maze-generator: Generate a random maze
+represented as a 2D array using depth-first search}.
+\url{https://github.com/oppenheimj/maze-generator/}; GitHub.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
+Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
+Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
+world models in maze-solving tasks. \emph{arXiv Preprint
+arXiv:2412.11867}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
+Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
+Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
+self-supervised neural-symbolic learning framework for robot autonomy.
+\emph{arXiv Preprint arXiv:2406.16087}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-wilson}{}}%

### 2025-03-28 maze_dataset/utils.py
separate dep extra for tokenization, move stuff around
rust-fst doesn't work on macOS, so we put that and frozendict into the `tokenization` extras
we also move all_instances and FiniteValued and related code out of utils.py to allow this

@@ -198,241 +198,3 @@ def adj_list_to_nested_set(adj_list: list) -> set:
 		for start_coord, end_coord in adj_list
 	}
 
-
-FiniteValued = TypeVar("FiniteValued", bound=bool | IsDataclass | enum.Enum)
-"""
-# `FiniteValued`
-The details of this type are not possible to fully define via the Python 3.10 typing library.
-This custom generic type is a generic domain of many types which have a finite, discrete, and well-defined range space.
-`FiniteValued` defines the domain of supported types for the `all_instances` function, since that function relies heavily on static typing.
-These types may be nested in an arbitrarily deep tree via Container Types and Superclass Types (see below).
-The leaves of the tree must always be Primitive Types.
-
-# `FiniteValued` Subtypes
-*: Indicates that this subtype is not yet supported by `all_instances`
-
-## Non-`FiniteValued` (Unbounded) Types
-These are NOT valid subtypes, and are listed for illustrative purposes only.
-This list is not comprehensive.
-While the finite and discrete nature of digital computers means that the cardinality of these types is technically finite,
-they are considered unbounded types in this context.
-- No Container subtype may contain any of these unbounded subtypes.
-- `int`
-- `float`
-- `str`
-- `list`
-- `set`: Set types without a `FiniteValued` argument are unbounded
-- `tuple`: Tuple types without a fixed length are unbounded
-
-## Primitive Types
-Primitive types are non-nested types which resolve directly to a concrete range of values
-- `bool`: has 2 possible values
-- *`enum.Enum`: The range of a concrete `Enum` subclass is its set of enum members
-- `typing.Literal`: Every type constructed using `Literal` has a finite set of possible literal values in its definition.
-This is the preferred way to include limited ranges of non-`FiniteValued` types such as `int` or `str` in a `FiniteValued` hierarchy.
-
-## Container Types
-Container types are types which contain zero or more fields of `FiniteValued` type.
-The range of a container type is the cartesian product of their field types, except for `set[FiniteValued]`.
-- `tuple[FiniteValued]`: Tuples of fixed length whose elements are each `FiniteValued`.
-- `IsDataclass`: Concrete dataclasses whose fields are `FiniteValued`.
[…]

### 2025-03-28 maze_dataset/tokenization/modular/all_instances.py
mock behavior for when rust_fst not loaded
see https://github.com/understanding-search/maze-dataset/issues/57

@@ -18,7 +18,12 @@ from typing import (
 	overload,
 )
 
-import frozendict
+try:
+	import frozendict
+except ImportError as e:
+	raise ImportError(
+		"You need to install the `frozendict` package to use `all_instances` -- try installing `maze_dataset[tokenization]`"
+	) from e
 import numpy as np
 from jaxtyping import Bool, Int, Int8
 from muutils.misc import IsDataclass, flatten, is_abstract

### 2025-03-28 maze_dataset/tokenization/modular/fst_load.py

@@ -13,14 +13,33 @@ thanks to https://github.com/rozbb for suggesting doing this instead of storing 
 
 from functools import cache
 from pathlib import Path
+import warnings
 
-from rust_fst import Set as FstSet  # type: ignore[import-untyped]
+_RUST_FST_LOADED: bool = False
+"""if the rust_fst module was loaded successfully"""
+
+_RUST_FST_ERR_MSG: str = (
+	"you need the `rust_fst` package to use `maze_dataset.tokenization.modular` properly. installing `maze-dataset[tokenization]` will install it\n"
+	"Note that rust-fst doesn't work on mac, see https://github.com/understanding-search/maze-dataset/issues/57\n"
+	"and this makes modular tokenizers not checkable on mac. Things should still work, but you will have no guarantee that a tokenizer is tested.\n"
+	"If you can find away around this, please let us know!\n"
+)
+
+class RustFstNotLoadedWarning(UserWarning):
+	"""warning for when `rust_fst` is not loaded"""
+
+try:
+	from rust_fst import Set as FstSet  # type: ignore[import-untyped]
+	_RUST_FST_LOADED = True
+except ImportError as e:
+	warnings.warn(_RUST_FST_ERR_MSG + str(e), RustFstNotLoadedWarning)
+	_RUST_FST_LOADED = False
 
 MMT_FST_PATH: Path = Path(__file__).parent / "MazeTokenizerModular_tested.fst"
 
 
 @cache
-def get_tokenizers_fst() -> FstSet:
+def get_tokenizers_fst() -> "FstSet":
 	"""(cached) load the tokenizers fst set from `MMT_FST_PATH`"""
 	return FstSet(MMT_FST_PATH.as_posix())
 
@@ -52,3 +71,22 @@ def check_tokenizer_in_fst(tokenizer_name: str, do_except: bool = False) -> bool
 		raise ValueError(err_msg)
 
 	return in_fst
+
+
+def _check_tokenizer_in_fst_mock(tokenizer_name: str, do_except: bool = False) -> bool: # noqa: ARG001
+	"""mock function for `check_tokenizer_in_fst`
+
+	runs when we cant import `rust_fst` which sets `_RUST_FST_LOADED` to `False`
+	"""
+	warnings.warn(
+		_RUST_FST_ERR_MSG
+		+ "you are seeing this warning probably because you tried to run"
+		"`MazeTokenizerModular(...).is_tested_tokenizer()` on a mac or without `rust_fst` installed"
[…]

### 2025-03-28 maze_dataset/tokenization/modular/all_instances.py
format

@@ -2,7 +2,6 @@
 
 import enum
 import itertools
-import math
 import typing
 from dataclasses import Field  # noqa: TC003
 from functools import cache, wraps
@@ -15,7 +14,6 @@ from typing import (
 	TypeVar,
 	get_args,
 	get_origin,
-	overload,
 )
 
 try:
@@ -24,8 +22,6 @@ except ImportError as e:
 	raise ImportError(
 		"You need to install the `frozendict` package to use `all_instances` -- try installing `maze_dataset[tokenization]`"
 	) from e
-import numpy as np
-from jaxtyping import Bool, Int, Int8
 from muutils.misc import IsDataclass, flatten, is_abstract
 
 FiniteValued = TypeVar("FiniteValued", bound=bool | IsDataclass | enum.Enum)

### 2025-03-28 maze_dataset/tokenization/modular/fst_load.py

@@ -11,9 +11,9 @@ thanks to https://github.com/rozbb for suggesting doing this instead of storing 
 
 """
 
+import warnings
 from functools import cache
 from pathlib import Path
-import warnings
 
 _RUST_FST_LOADED: bool = False
 """if the rust_fst module was loaded successfully"""
@@ -25,11 +25,14 @@ _RUST_FST_ERR_MSG: str = (
 	"If you can find away around this, please let us know!\n"
 )
 
+
 class RustFstNotLoadedWarning(UserWarning):
 	"""warning for when `rust_fst` is not loaded"""
 
+
 try:
 	from rust_fst import Set as FstSet  # type: ignore[import-untyped]
+
 	_RUST_FST_LOADED = True
 except ImportError as e:
 	warnings.warn(_RUST_FST_ERR_MSG + str(e), RustFstNotLoadedWarning)
@@ -73,7 +76,7 @@ def check_tokenizer_in_fst(tokenizer_name: str, do_except: bool = False) -> bool
 	return in_fst
 
 
-def _check_tokenizer_in_fst_mock(tokenizer_name: str, do_except: bool = False) -> bool: # noqa: ARG001
+def _check_tokenizer_in_fst_mock(tokenizer_name: str, do_except: bool = False) -> bool:  # noqa: ARG001
 	"""mock function for `check_tokenizer_in_fst`
 
 	runs when we cant import `rust_fst` which sets `_RUST_FST_LOADED` to `False`

### 2025-03-28 maze_dataset/utils.py

@@ -1,27 +1,12 @@
 "misc utilities for the `maze_dataset` package"
 
-import enum
-import itertools
 import math
-import typing
-from dataclasses import Field  # noqa: TC003
-from functools import cache, wraps
-from types import UnionType
 from typing import (
-	Callable,
-	Generator,
-	Iterable,
-	Literal,
-	TypeVar,
-	get_args,
-	get_origin,
 	overload,
 )
 
-import frozendict
 import numpy as np
 from jaxtyping import Bool, Int, Int8
-from muutils.misc import IsDataclass, flatten, is_abstract
 
 
 def bool_array_from_string(
@@ -197,4 +182,3 @@ def adj_list_to_nested_set(adj_list: list) -> set:
 		frozenset([tuple(start_coord), tuple(end_coord)])
 		for start_coord, end_coord in adj_list
 	}
-

### 2025-03-28 maze_dataset/utils.py
fix some typing and tests

@@ -44,9 +44,7 @@ def bool_array_from_string(
 	expected_symbol_count = math.prod(shape)
 	symbol_count = len(stripped)
 	if len(stripped) != expected_symbol_count:
-		err_msg: str = (
-			f"Connection List contains the wrong number of symbols. Expected {expected_symbol_count}. Found {symbol_count} in {stripped}.",
-		)
+		err_msg: str = f"Connection List contains the wrong number of symbols. Expected {expected_symbol_count}. Found {symbol_count} in {stripped}."
 		raise ValueError(err_msg)
 
 	bools = [(symbol == true_symbol) for symbol in stripped]
@@ -94,8 +92,10 @@ return indices[sorted_order]
 def manhattan_distance(
 	edges: Int[np.ndarray, "edges coord=2 row_col=2"],
 ) -> Int8[np.ndarray, " edges"]: ...
+# TYPING: error: Overloaded function signature 2 will never be matched: signature 1's parameter type(s) are the same or broader  [overload-cannot-match]
+# this is because mypy doesn't play nice with jaxtyping
 @overload
-def manhattan_distance(
+def manhattan_distance(  # type: ignore[overload-cannot-match]
 	edges: Int[np.ndarray, "coord=2 row_col=2"],
 ) -> int: ...
 def manhattan_distance(

### 2025-03-28 tests/unit/tokenization/test_all_instances.py

@@ -5,7 +5,7 @@ from typing import Callable, Iterable, Literal
 import pytest
 from muutils.misc import IsDataclass, dataclass_set_equals
 
-from maze_dataset.utils import FiniteValued, all_instances
+from maze_dataset.tokenization.modular.all_instances import FiniteValued, all_instances
 
 
 # Test classes

### 2025-03-29 docs/paper/paper.tex
wip

@@
     pdfaconformance=B,
-    pdftitle={A Configurable Library for Generating and Manipulating
-Maze Datasets},
+    pdftitle={maze-dataset},
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
@@
 \fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). A Configurable Library for Generating and Manipulating
-Maze Datasets. \emph{Journal of Open Source Software},
+et al. (1970). maze-dataset. \emph{Journal of Open Source Software},
 \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
@@
 
-\title{A Configurable Library for Generating and Manipulating Maze
-Datasets}
+\title{maze-dataset}
 
@@
 ]{Aaron Sandoval%
+  \,\orcidlink{0009-0002-8380-6140}\,%
 }
@@
 ]{Cecilia Diniz Behn%
+  \,\orcidlink{0000-0002-8078-5105}\,%
 }
@@
 ]{Samy Wu Fung%
+  \,\orcidlink{0000-0002-2926-4582}\,%
 }
@@
 https://understanding-search.github.io/maze-dataset/examples/maze\_examples.html}
+
+\colorbox[RGB]{ 176,152,232 }{ \texttt{ <ADJLIST\_START> (0,1) <--> (0,2) ; (1,0) <--> (1,1) ; (0,1) <--> (0,0) ; } }
+\colorbox[RGB]{ 176,152,232 }{ \texttt{ (4,0) <--> (3,0) ; (0,2) <--> (0,3) ; (0,4) <--> (1,4) ; (2,2) <--> (3,2) ; } }
+\colorbox[RGB]{ 176,152,232 }{ \texttt{ (3,4) <--> (2,4) ; (2,0) <--> (3,0) ; (3,4) <--> (4,4) ; (3,2) <--> (3,3) ; } }
+\colorbox[RGB]{ 176,152,232 }{ \texttt{ (2,4) <--> (2,3) ; (1,3) <--> (0,3) ; (1,3) <--> (1,4) ; (1,0) <--> (0,0) ; } }
+\colorbox[RGB]{ 176,152,232 }{ \texttt{ (2,1) <--> (2,0) ; (4,2) <--> (4,3) ; (3,2) <--> (3,1) ; (1,2) <--> (0,2) ; } }
+\colorbox[RGB]{ 176,152,232 }{ \texttt{ (4,1) <--> (4,0) ; (4,2) <--> (4,1) ; (4,3) <--> (4,4) ; (1,2) <--> (2,2) ; } }
+\colorbox[RGB]{ 176,152,232 }{ \texttt{ (2,1) <--> (2,2) ; <ADJLIST\_END> } }
+\newline
+\colorbox[RGB]{ 154,239,123 }{ \texttt{ <ORIGIN\_START> (1,3) <ORIGIN\_END> } }
+\colorbox[RGB]{ 246,136,136 }{ \texttt{ <TARGET\_START> (2,3) <TARGET\_END> } }
+\colorbox[RGB]{ 111,187,254 }{ \texttt{ <PATH\_START> (1,3) (0,3) (0,2) (1,2) (2,2) (2,1) (2,0) (3,0) (4,0) (4,1) } }
+\colorbox[RGB]{ 111,187,254 }{ \texttt{ (4,2) (4,3) (4,4) (3,4) (2,4) (2,3) <PATH\_END> } }
 

### 2025-03-29 docs/paper/diagram/test.tex
wip pdf export for figure
links dont work unless you use a chromium browser to export to pdf?

\documentclass{article}

% Basic packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{xcolor}  % Required for colors
\usepackage{transparent}  % Required for transparency
\usepackage{import}  % For importing the PDF_TEX file

% Page margins
\usepackage[margin=1in]{geometry}

% Title information
\title{Maze Visualization}
\author{Your Name}
\date{\today}

\begin{document}

\maketitle

\section{Maze Diagram}

Below is the maze representation generated from Inkscape:

\begin{figure}[h]
    \centering
    % Set the desired width of the SVG image
    \def\svgwidth{0.8\textwidth}
    % Import the PDF_TEX file
    \input{diagram-4.pdf_tex}
    \caption{Maze representation with different visualization methods}
    \label{fig:maze}
\end{figure}

\section{Notes on the Diagram}

This diagram shows various ways to visualize and represent a maze:

\begin{itemize}
    \item The top-left shows the configuration setup with \texttt{MazeDatasetConfig}
    \item The center shows how to access a specific maze with \texttt{ds[0]}
    \item The right side shows ASCII representation with \texttt{m.as\_ascii()}
    \item The bottom sections show other visualization methods:
    \begin{itemize}
        \item \texttt{m.as\_pixels()} for pixel representation
        \item \texttt{MazePlot(m)} for a graphical plot
        \item \texttt{m.as\_tokens(...)} for a tokenized representation
    \end{itemize}
\end{itemize}

\section{Using the Maze Dataset}

The diagram illustrates the typical workflow for working with the maze dataset:

\begin{enumerate}
    \item Configure the maze parameters using \texttt{MazeDatasetConfig}
    \item Create the dataset with \texttt{MazeDataset.from\_config(cfg)}
    \item Access individual maze objects from the dataset
    \item Visualize or process mazes using various methods
\end{enumerate}

\end{document}

### 2025-03-29 docs/paper/paper.tex
trying to get these links to work. madness.

--- 2025-03-25 docs/paper/paper.tex
@@
 \fi
+\usepackage{graphicx}
+\usepackage{tikz}
+\usepackage[luatex, pdfversion=1.7]{hyperref}
+\usetikzlibrary{calc}
+\tikzset{ % Define a TikZ style for an external hyperlink node.
+  hyperlink node url/.style={
+    alias=sourcenode,
+    append after command={
+      let \p1 = (sourcenode.north west),
+          \p2 = (sourcenode.south east),
+          \n1 = {\x2-\x1},
+          \n2 = {\y1-\y2} in
+      node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
+          {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
+    }
+  }
+}
+\providecommand{\XeTeXLinkBox}[1]{#1}
 \ifLuaTeX
@@
 
+\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
+
 \hypertarget{summary}{%
@@
 
-Solving mazes is a classic problem in computer science and artificial
-intelligence, and humans have been constructing mazes for thousands of
-years. Although finding the shortest path through a maze is a solved
-problem, this makes it an excellent testbed for studying how machine
-learning algorithms solve problems and represent spatial information. In
-this paper, we introduce \texttt{maze-dataset}, a Python library for
-generating, processing, and visualizing datasets of mazes. This library
-supports a variety of maze generation algorithms providing both mazes
-with loops and ``perfect'' mazes without them. These generation
-algorithms can be configured with various parameters, and the resulting
-mazes can be filtered to satisfy a desired property. Also provided are
-tools for converting mazes to and from various formats, such as
-rasterized images and tokenized text sequences, making it suitable for
-training or evaluating convolutional neural networks and autoregressive
-transformer models, as well as various visualization tools. As well as
-providing a simple interface for generating, storing, and loading these
-datasets, \texttt{maze-dataset} is extensively tested, type hinted,
-benchmarked, and documented.
-
 \textbf{TODO: mention the examples, put in a figure
 https://understanding-search.github.io/maze-dataset/examples/maze\_examples.html}
+
+\begin{figure} 
+\begin{tikzpicture}[remember picture]
+  % Place a clickable blue box at the top left corner (offset down by 1cm)
+  \node[draw=blue, fill=blue!20, fill opacity=0.3,
+        hyperlink node url={https://github.com},
+        minimum width=2cm, minimum height=1cm,
+        anchor=north west]
+    at ($(current page.north west)+(0,-1cm)$) {};
+\end{tikzpicture}
+\caption{testinggggggggggggggggg}
+\end{figure}
+
+\newpage
 

### 2025-03-29 docs/paper/paper.tex
diagram in pdf working
wip paper

@@
 
-\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
-
 \hypertarget{summary}{%
@@
 
-\textbf{TODO: mention the examples, put in a figure
-https://understanding-search.github.io/maze-dataset/examples/maze\_examples.html}
+Solving mazes is a classic problem in computer science and artificial
+intelligence, and humans have been constructing mazes for thousands of
+years. Although finding the shortest path through a maze is a solved
+problem, this very fact makes it an excellent testbed for studying how
+machine learning algorithms solve problems and represent spatial
+information. We introduce \texttt{maze-dataset}, a Python library for
+generating, processing, and visualizing datasets of mazes. This library
+supports a variety of maze generation algorithms providing both mazes
+with loops and ``perfect'' mazes without them. These generation
+algorithms can be configured with various parameters, and the resulting
+mazes can be filtered to satisfy desired properties. Also provided are
+tools for converting mazes to and from various formats suitable for a
+variety of neural network architectures, such as rasterized images and
+tokenized text sequences, as well as various visualization tools. As
+well as providing a simple interface for generating, storing, and
+loading these datasets, \texttt{maze-dataset} is extensively tested,
+type hinted, benchmarked, and documented.
 
 \begin{figure} 
-\begin{tikzpicture}[remember picture]
-  % Place a clickable blue box at the top left corner (offset down by 1cm)
-  \node[draw=blue, fill=blue!20, fill opacity=0.3,
-        hyperlink node url={https://github.com},
-        minimum width=2cm, minimum height=1cm,
-        anchor=north west]
-    at ($(current page.north west)+(0,-1cm)$) {};
-\end{tikzpicture}
-\caption{testinggggggggggggggggg}
+  \begin{minipage}{5in}
+    \input{docs/paper/diagram/diagram.tikz} 
+  \end{minipage}
+  \caption{Usage of maze-dataset.
+We create a `MazeDataset` from a `MazeDatasetConfig`.
+This contains `SolvedMaze` objects which can be converted to and from a variety of formats.
+Code in the image contains clickable links to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{documentation}.
+A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.}
+  \label{fig:diagram}
 \end{figure}
-
-\newpage
 
@@
 
-\hypertarget{features-and-implementation}{%
-\section{Features and
-Implementation}\label{features-and-implementation}}
+\hypertarget{features}{%
+\section{Features}\label{features}}
 
@@
 \subsection{Generation and Usage}\label{generation}}
-
-\textbf{TODO: JOSS says don't document the API itself. we should talk
-about the capabilities, but just rewrite this to link to the
-documentation.}
 
@@
 parameters. This object is passed to a \texttt{MazeDataset} class to
-create a dataset. Crucially, this \texttt{MazeDataset} inherits from a
-PyTorch (\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
-\texttt{Dataset}, and can thus be easily incorporated into existing data
-pre-processing and training pipelines, e.g., through the use of a
-\texttt{Dataloader} class.
+create a dataset. Crucially, this \texttt{MazeDataset} mimics the
+interface of a PyTorch (\protect\hyperlink{ref-pytorch}{Paszke et al.,
+2019}) \texttt{Dataset}, and can thus be easily incorporated into
+existing data pre-processing and training pipelines, e.g., through the
+use of a \texttt{Dataloader} class.
 
@@
 through the \texttt{from\_config()} factory method as necessary. Options
-include 1) whether to generate the dataset during runtime or load an
-existing dataset, 2) if and how to parallelize generation, and 3) where
-to store the generated dataset. Full documentation of configuration
-options is available in our repository
-(\protect\hyperlink{ref-maze-dataset-github}{Michael I. Ivanitskiy et
-al., 2023a}). Available maze generation algorithms are static methods of
-the \texttt{LatticeMazeGenerators} class and include the following:
+allow for saving/loading existing datasets instead of regenerating, and
+parallelization options for generation. Available maze generation
+algorithms are static methods of the \texttt{LatticeMazeGenerators}
+class and include the following:
 
@@
 \begin{figure}
+\hypertarget{fig:visualoutputs}{%
 \centering
-\includegraphics{figures/output-fmts.pdf}
+\includegraphics{docs/paper/figures/output-fmts.pdf}
 \caption{Various output formats. Top row (left to right): ASCII diagram,
-rasterized pixel grid, and advanced display.}
+rasterized pixel grid, and advanced display.}\label{fig:visualoutputs}
+}
 \end{figure}
@@
 class which returns rasterized pairs of (input, target) mazes as shown
-in Figure below.
+in \autoref{fig:e2hraster} below.
 
 \begin{figure}
+\hypertarget{fig:e2hraster}{%
 \centering
-\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
+\includegraphics[width=0.3\textwidth,height=\textheight]{docs/paper/figures/maze-raster-input-target.pdf}
 \caption{Input is the rasterized maze without the path marked (left),
@@
 Configuration options exist to adjust whether endpoints are included and
-if empty cells should be filled in.}
+if empty cells should be filled in.}\label{fig:e2hraster}
+}
 \end{figure}
@@
 \centering
-\includegraphics{figures/outputs-tokens-colored.tex}
+\includegraphics{docs/paper/figures/outputs-tokens-colored.tex}
 \caption{Example text output format with token regions highlighted.}
@@
 \centering
-\includegraphics{figures/TokenizerElement_structure.pdf}
+\includegraphics{docs/paper/figures/TokenizerElement_structure.pdf}
 \caption{Nested internal structure of \texttt{\_TokenizerElement}
@@
 \centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{figures/gridsize-vs-gentime.pdf}
+\includegraphics[width=0.95\textwidth,height=\textheight]{docs/paper/figures/gridsize-vs-gentime.pdf}
 \caption{Plots of maze generation time. Generation time scales

### 2025-03-29 docs/paper/paper.tex
wip paper

@@
     \centering
-\includegraphics{docs/paper/figures/TokenizerElement_structure.pdf}
-\caption{Nested internal structure of \texttt{\_TokenizerElement}
-objects inside a typical \texttt{MazeTokenizerModular} object.}
+    \input{docs/paper/figures/TokenizerElement_structure.tikz}
+    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular} object.}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{docs/paper/figures/gridsize-vs-gentime.pdf}
+\includegraphics[width=0.95\textwidth,height=\textheight]{docs/benchmarks/figures/gridsize-vs-gentime.pdf}
 \caption{Plots of maze generation time. Generation time scales
@@
 parallelization is disabled.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{docs/benchmarks/figures/n_mazes-vs-gentime.pdf}
+\caption{Maze size vs generation time}
 \end{figure}

### 2025-03-30 maze_dataset/tokenization/modular/__init__.py
ascii chart in docs

@@ -1,4 +1,40 @@
-"""implements `ModularMazeTokenizer` and related code"""
+"""implements `ModularMazeTokenizer` and related code
+
+the structure of a typical `MazeTokenizerModular` is something like this:
+```
++----------------------------------------------------+
+|                  MazeTokenizerModular              |
+|  +-----------------------------------------------+ |
+|  |                 _PromptSequencer              | |
+|  |         +-----------------------------+       | |
+|  |         |       _CoordTokenizer       |       | |
+|  |         +-----------------------------+       | |
+|  |     +------------------------------------+    | |
+|  |     |         _AdjListTokenizer          |    | |
+|  |     | +-----------+    +-------------+   |    | |
+|  |     | |_EdgeSubset|    |_EdgeGrouping|   |    | |
+|  |     | +-----------+    +-------------+   |    | |
+|  |     |          +-------------+           |    | |
+|  |     |          |_EdgePermuter|           |    | |
+|  |     |          +-------------+           |    | |
+|  |     +------------------------------------+    | |
+|  |         +-----------------------------+       | |
+|  |         |      _TargetTokenizer       |       | |
+|  |         +-----------------------------+       | |
+|  |  +------------------------------------------+ | |
+|  |  |              _PathTokenizer              | | |
+|  |  |  +---------------+   +----------------+  | | |
+|  |  |  |   _StepSize   |   | _StepTokenizer |  | | |
+|  |  |  +---------------+   +----------------+  | | |
+|  |  |                      | _StepTokenizer |  | | |
+|  |  |                      +----------------+  | | |
+|  |  |                             :            | | |
+|  |  +------------------------------------------+ | |
+|  +-----------------------------------------------+ |
++----------------------------------------------------+
+```
+
+"""
 
 __all__ = [
 	# modules

### 2025-03-30 docs/paper/paper.tex
paper stuff

--- 2025-03-29 docs/paper/paper.tex
@@
   \begin{minipage}{5in}
-    \input{docs/paper/diagram/diagram.tikz} 
+    \input{diagram/diagram.tikz} 
   \end{minipage}
@@
 \centering
-\includegraphics{docs/paper/figures/output-fmts.pdf}
+\includegraphics{figures/output-fmts.pdf}
 \caption{Various output formats. Top row (left to right): ASCII diagram,
@@
 \centering
-\includegraphics[width=0.3\textwidth,height=\textheight]{docs/paper/figures/maze-raster-input-target.pdf}
+\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
 \caption{Input is the rasterized maze without the path marked (left),
@@
 \centering
-\includegraphics{docs/paper/figures/outputs-tokens-colored.tex}
+\includegraphics{figures/outputs-tokens-colored.tex}
 \caption{Example text output format with token regions highlighted.}
@@
     \centering
-\includegraphics{docs/paper/figures/TokenizerElement_structure.pdf}
-\caption{Nested internal structure of \texttt{\_TokenizerElement}
-objects inside a typical \texttt{MazeTokenizerModular} object.}
+    \input{figures/TokenizerElement_structure.tikz}
+    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular} object.}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{docs/paper/figures/gridsize-vs-gentime.pdf}
+\includegraphics[width=0.95\textwidth,height=\textheight]{../benchmarks/figures/gridsize-vs-gentime.pdf}
 \caption{Plots of maze generation time. Generation time scales
@@
 parallelization is disabled.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../benchmarks/figures/n_mazes-vs-gentime.pdf}
+\caption{Maze size vs generation time}
 \end{figure}

### 2025-03-30 docs/benchmarks/benchmark_generation.py
benchmark generation action

@@ -13,6 +13,11 @@ BENCHMARK_KWARGS: dict[str, dict[str, Any]] = dict(
 		n_mazes_vals=list(range(1, 12, 2)),
 		trials=10,
 	),
+	large=dict(
+		grid_n_vals=[2, 3, 4, 5, 6, 7, 8, 11, 16, 22, 32, 45, 64],
+		n_mazes_vals=list(range(1, 12, 2)),
+		trials=10,
+	),
 )
 
 if __name__ == "__main__":

### 2025-03-30 docs/paper/paper.tex
wip

@@
 \usepackage{tikz}
-\usepackage[luatex, pdfversion=1.7]{hyperref}
 \usetikzlibrary{calc}
@@
 \centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{../benchmarks/figures/gridsize-vs-gentime.pdf}
+\includegraphics[width=0.95\textwidth,height=\textheight]{docs/benchmarks/figures/gridsize-vs-gentime.pdf}
 \caption{Plots of maze generation time. Generation time scales
@@
 \centering
-\includegraphics{../benchmarks/figures/n_mazes-vs-gentime.pdf}
+\includegraphics{docs/benchmarks/figures/n_mazes-vs-gentime.pdf}
 \caption{Maze size vs generation time}

### 2025-03-30 docs/paper/figures/outputs-ascii-colored.tex
paper wip

\begin{Verbatim}[commandchars=\\\{\}]
# # # # # # # # # # #
#         \textcolor{blue}{X X X} #   #
#   # # # \textcolor{blue}{X} # \textcolor{blue}{X} #   #
#       # \textcolor{blue}{X} # \textcolor{green}{S}     #
# # # # # \textcolor{blue}{X} # # # # #
# \textcolor{blue}{X X X X X} # \textcolor{red}{E} \textcolor{blue}{X X} #
# \textcolor{blue}{X} # # #   # # # \textcolor{blue}{X} #
# \textcolor{blue}{X} #           # \textcolor{blue}{X} #
# \textcolor{blue}{X} # # # # # # # \textcolor{blue}{X} #
# \textcolor{blue}{X X X X X X X X X} #
# # # # # # # # # # #
\end{Verbatim}

### 2025-03-30 docs/paper/paper.tex
paper wip

@@
   \end{minipage}
-  \caption{Usage of maze-dataset.
-We create a `MazeDataset` from a `MazeDatasetConfig`.
-This contains `SolvedMaze` objects which can be converted to and from a variety of formats.
+  \caption{
+    Usage of maze-dataset.
+We create a \texttt{MazeDataset} from a \texttt{MazeDatasetConfig}.
+This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats.
 Code in the image contains clickable links to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{documentation}.
-A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.}
+A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.
+  }
   \label{fig:diagram}
@@
 When initializing mazes, further configuration options can be specified
-through the \texttt{from\_config()} factory method as necessary. Options
-allow for saving/loading existing datasets instead of regenerating, and
-parallelization options for generation. Available maze generation
-algorithms are static methods of the \texttt{LatticeMazeGenerators}
-class and include the following:
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{gen\_dfs} \textbf{(randomized depth-first search)}: Parameters
-  can be passed to constrain the number of accessible cells, the number
-  of forks in the maze, and the maximum tree depth. Creates a spanning
-  tree by default or a partially spanning tree if constrained.
-\item
-  \texttt{gen\_wilson} \textbf{(Wilson's algorithm)}: Generates a random
-  spanning tree via loop-erased random walk
-  (\protect\hyperlink{ref-wilson}{Wilson, 1996}).
-\item
-  \texttt{gen\_percolation} \textbf{(percolation)}: Starting with no
-  connections, every possible lattice connection is set to either true
-  or false with some probability \(p\), independently of all other
-  connections. For the kinds of graphs that this process generates, we
-  refer to existing work
+through the
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
+factory method as necessary. Options allow for saving/loading existing
+datasets instead of regenerating, and parallelization options for
+generation. Available maze generation algorithms are static methods of
+the \texttt{LatticeMazeGenerators} class and include generation
+algorithms based on randomized depth-first search, Wilson's algorithm
+(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
 (\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
-  \protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam,
-  2004}).
-\item
-  \texttt{gen\_dfs\_percolation} \textbf{(randomized depth-first search
-  with percolation)}: A connection exists if it exists in a maze
-  generated via \texttt{gen\_dfs\ OR\ gen\_percolation}. Useful for
-  generating mazes that are not acyclic graphs.
-\end{itemize}
+\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
+Kruskal's algorithm
+(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956a}), and
+others.
 
 Furthermore, a dataset of mazes can be filtered to satisfy certain
-properties:
+properties. Custom filters can be specified, and some filters are
+included in
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
 
@@
 
-Custom filters can be specified, and several filters are included:
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{path\_length(min\_length:\ int)}: shortest length from the
-  origin to target should be at least \texttt{min\_length}.
-\item
-  \texttt{start\_end\_distance(min\_distance:\ int)}: Manhattan distance
-  between start and end should be at least \texttt{min\_distance},
-  ignoring walls.
-\item
-  \texttt{remove\_duplicates(...)}: remove mazes which are similar to
-  others in the dataset, measured via Hamming distance.
-\item
-  \texttt{remove\_duplicates\_fast()}: remove mazes which are exactly
-  identical to others in the dataset.
-\end{itemize}
-
 All implemented maze generation algorithms are stochastic by nature. For
@@
 
-Internally, mazes are \texttt{SolvedMaze} objects, which have path
-information, and a connection list optimized for storing sub-graphs of a
-lattice. These objects can be converted to and from several formats.
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3250}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3500}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3250}}@{}}
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-\texttt{as\_ascii()}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\texttt{as\_pixels()}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\texttt{MazePlot()}
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Simple text format for displaying mazes, useful for debugging in a
-terminal environment. & \texttt{numpy} array of \texttt{dtype=uint8} and
-shape \texttt{(height,\ width,\ 3)}. The last dimension is RGB color. &
-feature-rich plotting utility with support for multiple paths, heatmaps
-over positions, and more. \\
-\end{longtable}
-
-\begin{figure}
-\hypertarget{fig:visualoutputs}{%
+Internally, mazes are
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
+objects, which have path information, and a connection list optimized
+for storing sub-graphs of a lattice. These objects can be converted to
+and from several formats.
+
+\begin{figure}[H]
   \centering
-\includegraphics{figures/output-fmts.pdf}
-\caption{Various output formats. Top row (left to right): ASCII diagram,
-rasterized pixel grid, and advanced display.}\label{fig:visualoutputs}
-}
+  \begin{tabular}{p{2in} p{2in} p{2in}} 
+    \hline \\[.5em]
+    % algorithms
+    \texttt{as\_ascii()} & \texttt{as\_pixels()} & \texttt{MazePlot()} \\[.5em]
+    % descriptions
+      Simple text format for displaying mazes, useful for debugging in a terminal environment.
+      & \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
+      & feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
+    \hline \\
+    % examples
+      \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
+        \input{figures/outputs-ascii-colored.tex} 
+      \end{minipage}}
+      & \multicolumn{1}{c}{
+        \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
+      }
+      & \multicolumn{1}{c}{
+        \includegraphics[width=0.25\textwidth, trim={0 0 -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
+      } \\[1em]
+
+    \hline \\
+  \end{tabular}
+  \caption{Various output formats. Top row (left to right): ASCII diagram, rasterized pixel grid, and advanced display.}
+  \label{fig:output-fmts}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{docs/benchmarks/figures/n_mazes-vs-gentime.pdf}
-\caption{Maze size vs generation time}
-\end{figure}
-
 \hypertarget{implementation}{%
@@
 controlled via keyword arguments to the
-\texttt{MazeDataset.from\_config()} function.
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}
+function.
 
@@
 \protect\hyperlink{ref-prim}{Prim, 1957}) and Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal}{Kruskal, 1956}), among others
+(\protect\hyperlink{ref-kruskal}{Kruskal, 1956b}), among others
 (\protect\hyperlink{ref-mazegen_analysis}{Gabrovšek, 2019})), adding the
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kruskal1956shortest}{}}%
+Kruskal, J. B. (1956a). On the shortest spanning subtree of a graph and
+the traveling salesman problem. \emph{Proceedings of the American
+Mathematical Society}, \emph{7}(1), 48--50.
+
 \leavevmode\vadjust pre{\hypertarget{ref-kruskal}{}}%
-Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
+Kruskal, J. B. (1956b). On the shortest spanning subtree of a graph and
 the traveling salesman problem. \emph{Proceedings of the American

### 2025-03-30 docs/paper/paper.tex
wip

@@
 
-\hypertarget{training}{%
-\subsection{Visual Outputs for Training and Evaluation}\label{training}}
-
 In previous work, maze tasks have been used with Recurrent Convolutional
@@
 class which returns rasterized pairs of (input, target) mazes as shown
-in \autoref{fig:e2hraster} below.
+in \autoref{fig:e2h-raster} below.
 
 \begin{figure}
-\hypertarget{fig:e2hraster}{%
+\hypertarget{fig:e2h-raster}{%
 \centering
@@
 Configuration options exist to adjust whether endpoints are included and
-if empty cells should be filled in.}\label{fig:e2hraster}
+if empty cells should be filled in.}\label{fig:e2h-raster}
 }
@@
 \hypertarget{tokenized-output-formats}{%
-\section{Tokenized Output Formats}\label{tokenized-output-formats}}
+\subsection{Tokenized Output Formats}\label{tokenized-output-formats}}
 
@@
 
-\hypertarget{mtm}{%
-\subsection{\texorpdfstring{Stringification Options and
-\texttt{MazeTokenizerModular}}{Stringification Options and MazeTokenizerModular}}\label{mtm}}
-
 There are many algorithms by which one might tokenize a 2D maze into a
@@
 different internal representations, learned solution algorithms, and
-levels of performance. To explore how different maze tokenization
-algorithms affect these models, the \texttt{MazeTokenizerModular} class
-contains a rich set of options to customize how mazes are stringified.
-This class contains 19 discrete parameters, resulting in 5.9 million
-unique tokenizers. But wait, there's more! There are 6 additional
-parameters available in the library which are untested but further
-expand the the number of tokenizers by a factor of \(44/3\) to 86
-million.
+levels of performance. To allow exploration of how different maze
+tokenization algorithms affect these models, the
+\texttt{MazeTokenizerModular} class contains a rich set of options to
+customize how mazes are stringified. This class contains 19 discrete
+parameters, resulting in over 5.8 million unique tokenizers. But wait,
+there's more! There are 6 additional parameters available in the library
+which are untested but further expand the the number of tokenizers by a
+factor of \(44/3\) to 86 million.
 
@@
 different features of the maze. These regions are distinguished by color
-in Figure below.
+in \autoref{fig:token-regions}.
 
@@
 \begin{figure}
+\hypertarget{fig:token-regions}{%
 \centering
 \includegraphics{figures/outputs-tokens-colored.tex}
-\caption{Example text output format with token regions highlighted.}
+\caption{Example text output format with token regions
+highlighted.}\label{fig:token-regions}
+}
 \end{figure}

### 2025-03-30 maze_dataset/tokenization/modular/__init__.py

@@ -34,6 +34,46 @@ the structure of a typical `MazeTokenizerModular` is something like this:
 +----------------------------------------------------+
 ```
 
+Optional delimiter tokens may be added in many places in the output. Delimiter options are all configured using the parameters named `pre`, `intra`, and `post` in various `_TokenizerElement` classes. Each option controls a unique delimiter token.
+Here we describe each `_TokenizerElement` and the behaviors they support. We also discuss some of the model behaviors and properties that may be investigated using these options. 
+
+### Coordinates {#coordtokenizer}
+
+The `_CoordTokenizer` object controls how coordinates in the lattice are represented in across all token regions. Options include:
+
+- **Unique tokens**: Each coordinate is represented as a single unique token `"(i,j)"`
+- **Coordinate tuple tokens**: Each coordinate is represented as a sequence of 2 tokens, respectively encoding the row and column positions: `["i", ",", "j"]`
+
+### Adjacency List {#adjlisttokenizer}
+
+The `_AdjListTokenizer` object controls this token region. All tokenizations represent the maze connectivity as a sequence of connections or walls between pairs of adjacent coordinates in the lattice.
+
+- `_EdgeSubset`: Specifies the subset of lattice edges to be tokenized
+  - **All edges**: Every edge in the lattice
+  - **Connections**: Only edges which contain a connection
+  - **Walls**: Only edges which contain a wall
+- `_EdgePermuter`: Specifies how to sequence the two coordinates in each lattice edge
+  - **Random**
+  - **Sorted**: The smaller coordinate always comes first
+  - **Both permutations**: Each edge is represented twice, once with each permutation. This option attempts to represent connections in a more directionally symmetric manner. Including only one permutation of each edge may affect models' internal representations of edges, treating a path traversing the edge differently depending on if the coordinate sequence in the path matches the sequence in the adjacency list.
+- `shuffle_d0`: Whether to shuffle the edges randomly or sort them in the output by their first coordinate
+- `connection_token_ordinal`: Location in the sequence of the token representing whether the edge is a connection or a wall
+
+### Path {#pathtokenizer}
+
+The `_PathTokenizer` object controls this token region. Paths are all represented as a sequence of steps moving from the start to the end position. 
+
+- `_StepSize`: Specifies the size of each step
+  - **Singles**: Every coordinate traversed between start and end is directly represented
[…]

### 2025-03-30 docs/benchmarks/benchmark_generation.py
modify large run hparams

@@ -15,8 +15,8 @@ BENCHMARK_KWARGS: dict[str, dict[str, Any]] = dict(
 	),
 	large=dict(
 		grid_n_vals=[2, 3, 4, 5, 6, 7, 8, 11, 16, 22, 32, 45, 64],
-		n_mazes_vals=list(range(1, 12, 2)),
-		trials=10,
+		n_mazes_vals=[1, 2, 4, 8, 16],
+		trials=16,
 	),
 )
 

### 2025-03-30 maze_dataset/benchmark/config_sweep.py
pdf figures for success rate

@@ -442,6 +442,7 @@ def plot_grouped(  # noqa: C901
 	save_dir: Path | None = None,
 	show: bool = True,
 	logy: bool = False,
+	save_fmt: str = "svg",
 ) -> None:
 	"""Plot grouped sweep percolation value results for each distinct `endpoint_kwargs` in the configs
 
@@ -528,7 +529,9 @@ def plot_grouped(  # noqa: C901
 
 		# save and show
 		if save_dir:
-			save_path: Path = save_dir / f"ep_{endpoint_kwargs_to_name(ep_kw)}.svg"
+			save_path: Path = (
+				save_dir / f"ep_{endpoint_kwargs_to_name(ep_kw)}.{save_fmt}"
+			)
 			print(f"Saving plot to {save_path.as_posix()}")
 			save_path.parent.mkdir(exist_ok=True, parents=True)
 			plt.savefig(save_path)

### 2025-03-30 maze_dataset/tokenization/modular/__init__.py

@@ -35,7 +35,7 @@ the structure of a typical `MazeTokenizerModular` is something like this:
 ```
 
 Optional delimiter tokens may be added in many places in the output. Delimiter options are all configured using the parameters named `pre`, `intra`, and `post` in various `_TokenizerElement` classes. Each option controls a unique delimiter token.
-Here we describe each `_TokenizerElement` and the behaviors they support. We also discuss some of the model behaviors and properties that may be investigated using these options. 
+Here we describe each `_TokenizerElement` and the behaviors they support. We also discuss some of the model behaviors and properties that may be investigated using these options.
 
 ### Coordinates {#coordtokenizer}
 
@@ -61,7 +61,7 @@ The `_AdjListTokenizer` object controls this token region. All tokenizations rep
 
 ### Path {#pathtokenizer}
 
-The `_PathTokenizer` object controls this token region. Paths are all represented as a sequence of steps moving from the start to the end position. 
+The `_PathTokenizer` object controls this token region. Paths are all represented as a sequence of steps moving from the start to the end position.
 
 - `_StepSize`: Specifies the size of each step
   - **Singles**: Every coordinate traversed between start and end is directly represented

### 2025-03-30 maze_dataset/dataset/maze_dataset_config.py
some docs

@@ -86,7 +86,28 @@ EndpointKwargsType = dict[
 	],
 	bool | None | list[tuple[int, int]],
 ]
-"type hint for `MazeDatasetConfig.endpoint_kwargs`"
+"""type hint for `MazeDatasetConfig.endpoint_kwargs`
+
+- `except_on_no_valid_endpoint : bool` (default: `True`)
+	some of the conditions (dead ends if a maze is very open, no path between given start and end) can cause the maze generation to fail.
+	if `except_on_no_valid_endpoint` is `True`, then the maze generation will raise an error if it fails to generate a valid maze.
+	however, if `False`, then the maze generation will return a dataset with fewer mazes than requested.
+	If you are generating large datasets, consider using `MazeDatasetConfig.success_fraction_compensate()`
+	this uses a pysr-created function to roughly estimate the success fraction of the dataset.
+- `allowed_start : list[tuple[int, int]]` (default: `None`)
+	list of allowed starting position coordinates
+- `allowed_end : list[tuple[int, int]]` (default: `None`)
+	list of allowed ending position coordinates
+- `deadend_start : bool` (default: `False`)
+	if `True`, the starting position must be a dead end
+- `deadend_end : bool` (default: `False`)
+	if `True`, the ending position must be a dead end
+- `endpoints_not_equal : bool` (default: `True`)
+	if `True`, the starting and ending positions must be different
+
+
+
+"""
 
 
 def _load_endpoint_kwargs(data: dict) -> EndpointKwargsType:
@@ -106,8 +127,9 @@ def _load_endpoint_kwargs(data: dict) -> EndpointKwargsType:
 		}
 
 
+# not private because we need this to show up in docs
 @serializable_dataclass(kw_only=True, properties_to_serialize=["grid_shape"])
-class _MazeDatasetConfig_base(GPTDatasetConfig):  # noqa: N801
+class MazeDatasetConfig_base(GPTDatasetConfig):  # noqa: N801
 	"""base config -- we serialize, dump to json, and hash this to get the fname. all actual variables we want to be hashed are here"""
 
 	# NOTE: type: ignore[misc] is because it tells us non-default attributes aren't allowed after ones with defaults, but everything is kw_only
@@ -195,7 +217,7 @@ class _MazeDatasetConfig_base(GPTDatasetConfig):  # noqa: N801
 		but for comparing things, we do store it when serializing properly by setting
[…]

### 2025-03-30 maze_dataset/dataset/maze_dataset_config.py
format

@@ -276,7 +276,7 @@ class MazeDatasetConfig(MazeDatasetConfig_base):  # type: ignore[misc]
 		list of filters that have been applied to the dataset. We recommend applying filters to datasets directly,
 		but these are stored with the config in case you want to re-generate the dataset with the same filters.
 
-	"""	
+	"""
 
 	@property
 	def config_version(self) -> str:

### 2025-03-30 docs/paper/paper.tex
wip

@@
   that we do not believe is used in any existing maze generation
-  package.
+  package. This method is described in detail in
+  \autoref{implementation}.
 \item
@@
 \hypertarget{visual-output-formats}{%
-\section{Visual Output Formats}\label{visual-output-formats}}
+\subsection{Visual Output Formats}\label{visual-output-formats}}
 
@@
 
-Optional delimiter tokens may be added in many places in the output.
-Delimiter options are all configured using the parameters named
-\texttt{pre}, \texttt{intra}, and \texttt{post} in various
-\texttt{\_TokenizerElement} classes. Each option controls a unique
-delimiter token. Here we describe each \texttt{\_TokenizerElement} and
-the behaviors they support. We also discuss some of the model behaviors
-and properties that may be investigated using these options.
-
-\hypertarget{coordtokenizer}{%
-\subsubsection{Coordinates}\label{coordtokenizer}}
-
-The \texttt{\_CoordTokenizer} object controls how coordinates in the
-lattice are represented in across all token regions. Options include:
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Unique tokens}: Each coordinate is represented as a single
-  unique token \texttt{"(i,j)"}
-\item
-  \textbf{Coordinate tuple tokens}: Each coordinate is represented as a
-  sequence of 2 tokens, respectively encoding the row and column
-  positions: \texttt{{[}"i",\ ",",\ "j"{]}}
-\end{itemize}
-
-\hypertarget{adjlisttokenizer}{%
-\subsubsection{Adjacency List}\label{adjlisttokenizer}}
-
-The \texttt{\_AdjListTokenizer} object controls this token region. All
-tokenizations represent the maze connectivity as a sequence of
-connections or walls between pairs of adjacent coordinates in the
-lattice.
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{\_EdgeSubset}: Specifies the subset of lattice edges to be
-  tokenized
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \textbf{All edges}: Every edge in the lattice
-  \item
-    \textbf{Connections}: Only edges which contain a connection
-  \item
-    \textbf{Walls}: Only edges which contain a wall
-  \end{itemize}
-\item
-  \texttt{\_EdgePermuter}: Specifies how to sequence the two coordinates
-  in each lattice edge
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \textbf{Random}
-  \item
-    \textbf{Sorted}: The smaller coordinate always comes first
-  \item
-    \textbf{Both permutations}: Each edge is represented twice, once
-    with each permutation. This option attempts to represent connections
-    in a more directionally symmetric manner. Including only one
-    permutation of each edge may affect models' internal representations
-    of edges, treating a path traversing the edge differently depending
-    on if the coordinate sequence in the path matches the sequence in
-    the adjacency list.
-  \end{itemize}
-\item
-  \texttt{shuffle\_d0}: Whether to shuffle the edges randomly or sort
-  them in the output by their first coordinate
-\item
-  \texttt{connection\_token\_ordinal}: Location in the sequence of the
-  token representing whether the edge is a connection or a wall
-\end{itemize}
-
-\hypertarget{pathtokenizer}{%
-\subsubsection{Path}\label{pathtokenizer}}
-
-The \texttt{\_PathTokenizer} object controls this token region. Paths
-are all represented as a sequence of steps moving from the start to the
-end position.
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{\_StepSize}: Specifies the size of each step
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \textbf{Singles}: Every coordinate traversed between start and end
-    is directly represented
-  \item
-    \textbf{Forks}: Only coordinates at forking points in the maze are
-    represented. The paths between forking points are implicit. Using
-    this option might train models more directly to represent forking
-    points differently from coordinates where the maze connectivity
-    implies an obvious next step in the path.
-  \end{itemize}
-\item
-  \texttt{\_StepTokenizer}: Specifies how an individual step is
-  represented
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \textbf{Coordinate}: The coordinates of each step are directly
-    tokenized using a \texttt{\_CoordTokenizer}
-  \item
-    \textbf{Cardinal direction}: A single token corresponding to the
-    cardinal direction taken at the starting position of that step.
-    E.g., \texttt{NORTH}, \texttt{SOUTH}. If using a \texttt{\_StepSize}
-    other than \textbf{Singles}, this direction may not correspond to
-    the final direction traveled to arrive at the end position of the
-    step.
-  \item
-    \textbf{Relative direction}: A single token corresponding to the
-    first-person perspective relative direction taken at the starting
-    position of that step. E.g., \texttt{RIGHT}, \texttt{LEFT}.
-  \item
-    \textbf{Distance}: A single token corresponding to the number of
-    coordinate positions traversed in that step. E.g., using a
-    \texttt{\_StepSize} of \textbf{Singles}, the \textbf{Distance} token
-    would be the same for each step, corresponding to a distance of 1
-    coordinate. This option is only of interest in combination with a
-    \texttt{\_StepSize} other than \textbf{Singles}.
-  \end{itemize}
-\end{itemize}
-
-A \texttt{\_PathTokenizer} contains a sequence of one or more unique
-\texttt{\_StepTokenizer} objects. Different step representations may be
-mixed and permuted, allowing for investigation of model representations
-of multiple aspects of a maze solution at once.
-
-\hypertarget{token-training}{%
-\subsection{Tokenized Outputs for Training and
-Evaluation}\label{token-training}}
-
-During deployment we provide only the prompt up to the
-\texttt{\textless{}PATH\_START\textgreater{}} token.
-
-Examples of usage of this dataset to train autoregressive transformers
-can be found in our \texttt{maze-transformer} library
-(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
-et al., 2023b}). Other tokenization and vocabulary schemes are also
-included, such as representing each coordinate as a pair of \(i,j\)
-index tokens.
-
-\hypertarget{extensibility}{%
-\subsection{Extensibility}\label{extensibility}}
+\textbf{TODO: add fst stuff}
 
@@
 \hypertarget{benchmarks}{%
-\section{Benchmarks of Generation Speed}\label{benchmarks}}
+\subsection{Benchmarks of Generation Speed}\label{benchmarks}}
 
@@
 
+\hypertarget{success-rate-estimation}{%
+\subsection{Success Rate Estimation}\label{success-rate-estimation}}
+
+In order to replicate the exact dataset distribution of
+(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
+Bansal, et al., 2021}), we allow placing additional constraints in
+\texttt{MazeDatasetConfig.endpoint\_kwargs:\ EndpointKwargsType}, such
+as enforcing that the start or end point be in a ``dead end'' with only
+one accessible neighbor cell. However, combining this with cyclic mazes
+(such as those generated with percolation), as was required for the work
+in (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et
+al., 2024}), can lead to an absence of valid start and end points.
+Placing theoretical bounds on this success rate is difficult, as it
+depends on the exact maze generation algorithm and parameters used.
+However, our package provides a way to estimate the success rate of a
+given configuration using a symbolic regression model trained with PySR
+(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
+be found in
+\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
+
+Using the estimation simply requires the user to call
+\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}
+given their predefined \texttt{cfg}.
+
+\hypertarget{success-rate-estimation}{%
+\subsubsection{Success Rate Estimation
+Algorithm}\label{success-rate-estimation}}
+
+The base function learned by symbolic regression is not particularly
+insightful, and is potentially subject to change. It is defined as
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
+and takes a 5 dimensional float vector created by
+\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the 0)
+percolation value 1) grid size 2) endpoint deadend configuration 3)
+endpoint uniqueness 4) categorical generation function index.
+
+However, the outputs of this function are not directly usable due to
+minor divergences at the endpoints with respect to the percolation
+probability \(p\). Since we know that maze success is either guaranteed
+or impossible for \(p=0\) and \(p=1\), we define the \texttt{soft\_step}
+function to nudge the raw output of the symbolic regression. This
+function is defined with:
+
+\begin{itemize}
+\item
+  \textbf{Shifted Sigmoid}: Creates sharp transition at \(x=0.5\)
+  \[\sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}\]
+\item
+  \textbf{h-function}: Transition function combining polynomial and
+  sigmoid components
+  \[h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)\]
+\item
+  \textbf{Amplitude Scaling}: Weight-modulated polynomial
+  \[A(q,a,w) = w \cdot (1 - |2q-1|^a)\]
+\item
+  \textbf{Soft Step}: Identity-like for \(p \approx 0.5\), pushes \(x\)
+  to extremes otherwise
+  \[\text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))\]
+\end{itemize}
+
+Finally, we define
+
+\[
+  \text{cfg\_success\_predict\_fn}(\mathbf{x}) = \text{soft\_step}(\text{raw\_val}, x_0, 5, 10)
+\]
+
+where \texttt{raw\_val} is the output of the symbolic regression model.
+\(x_0\) is the percolation probability, while all other parameters from
+\texttt{\_to\_ps\_array()} only affect \texttt{raw\_val}.
+
+\begin{figure}
+\centering
+\includegraphics{figures/ep/ep_deadends_unique.pdf}
+\caption{An example of both empirical and predicted success rates as a
+function of the percolation probability \(p\) for various maze sizes,
+percolation with and without depth first search, and
+\texttt{endpoint\_kwargs} requiring that both the start and end be in
+unique dead ends.}
+\end{figure}
+
 \hypertarget{implementation}{%
@@
 an adjacency list to represent mazes would lead to a poor lookup time of
-whether any given connection exists, whilst using a dense adjacency
-matrix would waste memory by failing to exploit the structure (e.g.,
-only 4 of the diagonals would be filled in). Instead, we describe mazes
-with the following simple representation: for a \(d\)-dimensional
-lattice with \(r\) rows and \(c\) columns, we initialize a boolean array
+whether any given connection exists, while using an adjacency matrix
+would waste memory by failing to exploit the structure (e.g., only 4 of
+the diagonals would be filled in). Instead, we describe mazes with the
+following simple representation: for a \(d\)-dimensional lattice with
+\(r\) rows and \(c\) columns, we initialize a boolean array
 \(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
@@
 the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
-1968}) is applied to find the shortest path between them.
+1968}) is applied to find the shortest path between them. The endpoint
+selection can be affected by
+\texttt{MazeDatasetConfig.endpoint\_kwargs:\ EndpointKwargsType}, and
+complications caused by this are detailed in
+\autoref{success-rate-estimation}.
 
@@
 in the Python standard library, and parallel generation can be
-controlled via keyword arguments to the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}
-function.
-
-\hypertarget{limitations}{%
-\subsection{\texorpdfstring{Limitations of
-\texttt{maze-dataset}}{Limitations of maze-dataset}}\label{limitations}}
-
-For simplicity, the package primarily supports mazes that are sub-graphs
-of a 2-dimensional rectangular lattice. Some support for
-higher-dimensional lattices is present, but not all output formats are
-adapted for higher dimensional mazes.
-\protect\hyperlink{implementation-implementation}{Implementation}
-\protect\hyperlink{implementation-implementation}{Implementation}
+controlled via keyword arguments to
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pysr}{}}%
+Cranmer, M. (2023). Interpretable machine learning for science with PySR
+and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-dijkstra-prim}{}}%

### 2025-03-30 docs/paper/paper.tex
wip, mostly cleaning up

@@
 and the ability to easily transform between multiple representations of
-the mazes (Images, Textual, Tokenized) for training and testing models.
+the mazes (Raster, Textual, Tokenized) for training and testing models.
 
@@
   Most prior works provide mazes in some kind of image or raster format,
-  which is not suitable for training autoregressive text-based
-  transformer models -- a key usage case this work seeks to enable.
-  However, we still provide a variety of similar output formats:
+  and we provide a variety of similar output formats:
 
@@
   \item
-    we also include the \texttt{RasterizedMazeDataset} class, utilizing
-    \texttt{as\_pixels()}, in our codebase, which can exactly mimic the
-    outputs provided in
+    \texttt{RasterizedMazeDataset}, utilizing \texttt{as\_pixels()},
+    which can exactly mimic the outputs provided in
     \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
     Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
-    similar to the outputs of
-    (\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019}).
+    similar to the outputs of Németh
+    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
   \item
-    Our \texttt{as\_ascii()} method provides a format similar to that
-    used in Oppenheim
-    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{2018}).
+    \texttt{as\_ascii()} provides a format similar to that used in
+    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
+    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
   \item
-    Our \texttt{MazePlot} class provides a feature-rich plotting utility
-    with support for multiple paths, heatmaps over positions, and more.
-    This is similar to the outputs of Ehsan
-    (\protect\hyperlink{ref-gh_Ehsan_2022}{2022})
+    \texttt{MazePlot} provides a feature-rich plotting utility with
+    support for multiple paths, heatmaps over positions, and more. This
+    is similar to the outputs of
+    (\protect\hyperlink{ref-mazegenerator-net}{Alance AB, 2019};
+    \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
+    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
+    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
   \end{itemize}
@@
   2023}), but provides over 5.8 million unique formats for converting
-  mazes to a text stream. We maintain a single underlying format,
-  meaning that the same maze can be turned into a variety of text
-  streams to assess how the precise format of the text stream affects
-  the model.
+  mazes to a text stream, detailed in
+  \autoref{tokenized-output-formats}.
 \item
@@
   single human-readable file (\protect\hyperlink{ref-zanj}{M.
-  Ivanitskiy, n.d.}). This metadata is loaded when the dataset is
-  retrieved from disk and makes it simple to understand how exactly each
-  maze was generated. As far as we are aware, no existing packages do
+  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
   this reliably.
@@
   Storing mazes as images is not only difficult to work with, but also
-  inefficient. Directly storing adjacency matrices is also inefficient
-  as subgraphs of the lattice are sparse. Storing adjacency lists can be
-  efficient, but comes with a higher lookup cost and possible high
-  comparison cost. We use a simple, efficient representation of mazes
-  that is optimized for subgraphs of a \(d\)-dimensional finite lattice
-  that we do not believe is used in any existing maze generation
-  package. This method is described in detail in
+  inefficient. We use an extremely efficient method detailed in
   \autoref{implementation}.
@@
 Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956a}), and
+(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
 others.
@@
   \centering
-  \begin{tabular}{p{2in} p{2in} p{2in}} 
+  \begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
     \hline \\[.5em]
@@
       \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
+        \setlength{\baselineskip}{0.9em}
         \input{figures/outputs-ascii-colored.tex} 
@@
 All output sequences consist of four token regions representing
-different features of the maze. These regions are distinguished by color
-in \autoref{fig:token-regions}.
-
-\begin{itemize}
-\tightlist
-\item
-  {Adjacency list}: A text representation of the lattice graph
-\item
-  {Origin}: Starting coordinate
-\item
-  {Target}: Ending coordinate
-\item
-  {Path}: Maze solution sequence from the start to the end
-\end{itemize}
+different features of the maze of which we see an example in
+\autoref{fig:token-regions}.
 
 \begin{figure} 
-\hypertarget{fig:token-regions}{%
   \centering
-\includegraphics{figures/outputs-tokens-colored.tex}
-\caption{Example text output format with token regions
-highlighted.}\label{fig:token-regions}
-}
+  \begin{minipage}{5in}
+    \footnotesize
+    \input{figures/outputs-tokens-colored.tex}
+  \end{minipage}
+  \caption{
+    Example text output format with token regions highlighted.
+    \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the lattice graph,
+    \colorbox[RGB]{ 217,234,211 }{Origin}: starting coordinate,
+    \colorbox[RGB]{ 234,209,220 }{Target}: ending coordinate,
+    \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence
+  }
+  \label{fig:token-regions}
 \end{figure}
@@
 
-\hypertarget{success-rate-estimation}{%
+\hypertarget{success-rate-estimation-algorithm}{%
 \subsubsection{Success Rate Estimation
-Algorithm}\label{success-rate-estimation}}
+Algorithm}\label{success-rate-estimation-algorithm}}
 
@@
 
-\begin{itemize}
-\item
-  \textbf{Shifted Sigmoid}: Creates sharp transition at \(x=0.5\)
-  \[\sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}\]
-\item
-  \textbf{h-function}: Transition function combining polynomial and
-  sigmoid components
-  \[h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)\]
-\item
-  \textbf{Amplitude Scaling}: Weight-modulated polynomial
-  \[A(q,a,w) = w \cdot (1 - |2q-1|^a)\]
-\item
-  \textbf{Soft Step}: Identity-like for \(p \approx 0.5\), pushes \(x\)
-  to extremes otherwise
-  \[\text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))\]
-\end{itemize}
-
-Finally, we define
-
-\[
+A shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
+function: \[
+  \sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}
+  \qquad A(q,a,w) = w \cdot (1 - |2q-1|^a)
+\] \[
+  h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)
+\]
+
+We combine these to get the \texttt{soft\_step} function, which is
+identity-like for \(p \approx 0.5\), and pushes pushes \(x\) to extremes
+otherwise. \[
+  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
+\]
+
+Finally, we define \[
   \text{cfg\_success\_predict\_fn}(\mathbf{x}) = \text{soft\_step}(\text{raw\_val}, x_0, 5, 10)
@@
 \centering
-\includegraphics{figures/ep/ep_deadends_unique.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique.pdf}
 \caption{An example of both empirical and predicted success rates as a
@@
 et al., 2023b}) project, which aims to investigate spatial planning and
-world models in autoregressive transformer models trained on mazes Spies
-et al. (\protect\hyperlink{ref-spies2024causalworldmodels}{2024}). This
-project has also adapted itself to be useful for work on understanding
-the mechanisms by which recurrent convolutional and implicit networks
-(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
-given a rasterized view
+world models in autoregressive transformer models trained on mazes
+(\protect\hyperlink{ref-ivanitskiy2023structuredworldreps}{Michael
+Igorevich Ivanitskiy et al., 2023};
+\protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al., 2024}).
+This project has also adapted itself to be useful for work on
+understanding the mechanisms by which recurrent convolutional and
+implicit networks (\protect\hyperlink{ref-fung2022jfb}{Fung et al.,
+2022}) solve mazes given a rasterized view
 (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
@@
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-The
-\href{https://github.com/understanding-search/maze-dataset}{\texttt{maze-dataset}}
-library (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
-Ivanitskiy et al., 2023a}) introduced in this paper provides a flexible
-and extensible toolkit for generating, processing, and analyzing maze
-datasets. By supporting various procedural generation algorithms and
-conversion utilities, it enables the creation of mazes with customizable
-properties to suit diverse research needs. Planned improvements to the
-\texttt{maze-dataset} include adding more generation algorithms (such as
-Prim's algorithm (\protect\hyperlink{ref-dijkstra-prim}{Dijkstra, 1959};
-\protect\hyperlink{ref-jarnik-prim}{Jarnık, 1930};
-\protect\hyperlink{ref-prim}{Prim, 1957}) and Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal}{Kruskal, 1956b}), among others
-(\protect\hyperlink{ref-mazegen_analysis}{Gabrovšek, 2019})), adding the
-ability to augment a maze with an adjacency list to add ``shortcuts'' to
-the maze, and resolving certain limitations detailed in Section
-\protect\hyperlink{limitations}{Limitations}.
-
 \hypertarget{acknowledgements}{%
@@
 This work was partially supported by and many of the authors were
-brought together by AI Safety Camp and AI Safety Support. This work was
-partially funded by National Science Foundation awards DMS-2110745 and
-DMS-2309810. We are also grateful to LTFF and FAR Labs for hosting three
-of the authors for a Residency Visit, and to various members of FAR's
-technical staff for their advice. We thank the Mines Optimization and
-Deep Learning group (MODL) for fruitful discussions. We also thank
-Michael Rosenberg for recommending the usage of Finite State Transducers
-for storing tokenizer validation information.
+brought together by AI Safety Camp and AI Safety Support. We would like
+to thank our former collaborators at AI Safety Camp and other users and
+contributors to the \texttt{maze-dataset} package: Naveen Arunachalam,
+Benji Berczi, Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris
+Mathwin, Lucia Quirke, Can Rager, Adrians Skapars, Johannes Treutlein,
+and Dan Valentine.
+
+This work was partially funded by National Science Foundation awards
+DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
+for hosting three of the authors for a Residency Visit, and to various
+members of FAR's technical staff for their advice.
+
+We thank the Mines Optimization and Deep Learning group (MODL) for
+fruitful discussions. We also thank Michael Rosenberg for recommending
+the usage of Finite State Transducers for storing tokenizer validation
+information.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dijkstra-prim}{}}%
-Dijkstra, E. W. (1959). \emph{A note on two problems in connexion with
-graphs:({Numerische Mathematik}, 1 (1959), p 269-271)}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-percolation}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mazegen_analysis}{}}%
-Gabrovšek. (2019). Analysis of maze generating algorithms. \emph{{IPSI
-Transactions} on {Internet Research}}, \emph{15.1}, 23--30.
-\url{http://www.ipsitransactions.org/journals/papers/tir/2019jan/p5.pdf}
-
 \leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-jarnik-prim}{}}%
-Jarnık, V. (1930). About a certain minimal problem. \emph{Práce Moravské
-Prırodovedecké Spolecnosti}, \emph{6}, 57--63.
-
 \leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-kruskal1956shortest}{}}%
-Kruskal, J. B. (1956a). On the shortest spanning subtree of a graph and
+Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
 the traveling salesman problem. \emph{Proceedings of the American
 Mathematical Society}, \emph{7}(1), 48--50.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruskal}{}}%
-Kruskal, J. B. (1956b). On the shortest spanning subtree of a graph and
-the traveling salesman problem. \emph{Proceedings of the American
-Mathematical Society}, \emph{7}(1), 48--50.
-\url{https://doi.org/10.1090/S0002-9939-1956-0078686-7}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-prim}{}}%
-Prim, R. C. (1957). Shortest connection networks and some
-generalizations. \emph{The Bell System Technical Journal}, \emph{36}(6),
-1389--1401.
-
 \leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%

### 2025-03-30 docs/paper/paper.tex
wip

@@
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Rusheb Shah, Brandon Knutson, , Cecilia Diniz Behn, Samy Wu
-Fung},
+Spies, Rusheb Shah, Brandon Knutson, Cecilia Diniz Behn, Samy Wu Fung},
     pdfpublication={Journal of Open Source Software},
@@
     pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Rusheb Shah, Brandon Knutson, , Cecilia
+Aaron Sandoval, Alex F. Spies, Rusheb Shah, Brandon Knutson, Cecilia
 Diniz Behn, Samy Wu Fung},
@@
 }
-\author[%
+\author[1%
 %
 ]{Brandon Knutson%
-}
-\author[1%
-%
-]{%
   \,\orcidlink{0009-0004-8413-0239}\,%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
-Chen, X., Yang, F., \& Wang, C. (2024). iA \(\^{}*\): Imperative
-learning-based a \(\^{}*\) search for pathfinding. \emph{arXiv Preprint
-arXiv:2403.15870}.
+Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
+{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
 

### 2025-03-30 docs/benchmarks/benchmark_generation.py
format some scripts, add sys info printing to benchmark script

@@ -1,4 +1,10 @@
+"script for benchmarking generation performance"
+
+import platform
 from typing import Any
+
+import psutil
+
 from maze_dataset.benchmark.speed import run_benchmark
 from maze_dataset.generation.default_generators import DEFAULT_GENERATORS
 
@@ -20,7 +26,38 @@ BENCHMARK_KWARGS: dict[str, dict[str, Any]] = dict(
 	),
 )
 
+
+def get_system_info() -> dict[str, Any]:
+	"""Retrieve system hardware information"""
+	info: dict[str, Any] = {}
+
+	# Retrieve CPU frequency (in MHz)
+	cpu_freq: psutil._common.scpufreq = psutil.cpu_freq()  # type: ignore
+	info["cpu_current_freq"] = cpu_freq.current if cpu_freq else None
+	info["cpu_max_freq"] = cpu_freq.max if cpu_freq else None
+
+	# Retrieve CPU model information
+	cpu_model: str = platform.processor()
+	if not cpu_model:
+		# On some systems, platform.processor() returns an empty string.
+		cpu_model = platform.uname().processor
+	info["cpu_model"] = cpu_model
+
+	# Retrieve CPU core counts
+	info["cpu_logical_cores"] = psutil.cpu_count(logical=True)
+	info["cpu_physical_cores"] = psutil.cpu_count(logical=False)
+
+	# Retrieve total system RAM in bytes
+	virtual_mem: psutil._common.svmem = psutil.virtual_memory()  # type: ignore
+	info["total_ram"] = virtual_mem.total if virtual_mem else None
+
+	return info
+
+
 if __name__ == "__main__":
+
+	print(get_system_info())
+
 	import argparse
 
 	parser: argparse.ArgumentParser = argparse.ArgumentParser()

### 2025-03-30 docs/benchmarks/percolation_benchmarks.py

@@ -1,3 +1,5 @@
+"script for running percolation benchmarks"
+
 from pathlib import Path
 from typing import Any
 

### 2025-03-30 docs/examples/generate_plots.py

@@ -1,6 +1,4 @@
-#!/usr/bin/env python3
-"""
-Generate HTML examples of various maze configurations.
+"""Generate HTML examples of various maze configurations.
 
 This script generates a variety of maze examples with different configurations,
 saves them as SVG files, and creates an HTML page to display them with searchable
@@ -8,23 +6,19 @@ configuration details.
 """
 
 import json
-import os
 from pathlib import Path
 from typing import Any, Callable
 
-import matplotlib.pyplot as plt
-import numpy as np
-from matplotlib.figure import Figure
 import jinja2
+import matplotlib.pyplot as plt
+from matplotlib.figure import Figure
 from PIL import Image
 from tqdm import tqdm
 
-from maze_dataset import LatticeMaze, MazeDatasetConfig
+from maze_dataset import MazeDatasetConfig
 from maze_dataset.dataset.configs import _get_configs_for_examples
 from maze_dataset.dataset.maze_dataset import MazeDataset
-from maze_dataset.generation import LatticeMazeGenerators
 from maze_dataset.plotting import MazePlot
-
 
 plt.rcParams["svg.fonttype"] = "none"  # preserve text as text
 
@@ -122,7 +116,7 @@ def main() -> None:
 	maze_examples: list[dict] = []
 	all_tags: set[str] = set()
 
-	for i, cfg_src in tqdm(
+	for _i, cfg_src in tqdm(
 		enumerate(configs_sources),
 		desc="Generating maze examples",
 		total=len(configs_sources),

### 2025-03-30 docs/paper/paper.tex
benchmarks table

@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\usepackage{longtable,booktabs,array}
 
@@
 \fi
+\usepackage{makecell}
 \usepackage{graphicx}
@@
 \providecommand{\XeTeXLinkBox}[1]{#1}
+\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
 \ifLuaTeX
@@
     \input{figures/TokenizerElement_structure.tikz}
-    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular} object.}
+    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular}.}
 \end{figure}
@@
 We provide approximate benchmarks for relative generation time across
-various algorithms, parameter choices, maze sizes, and dataset sizes.
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1680}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2080}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1360}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1520}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2000}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1360}}@{}}
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Method \& Parameters
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Average time per maze (ms)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Generation algorithm & Generation parameters & all sizes & small
-(\(g \leq 10\)) & medium (\(10 < g \leq 32\)) & large (\(g > 32\)) \\
-:===============: & :===============: & :===============: &
-:===============: & :===============: & :===============: \\
-gen\_dfs & accessible\_cells=20 & 2.4 & 2.4 & 2.6 & 2.4 \\
-gen\_dfs & do\_forks=False & 3.0 & 2.4 & 3.7 & 3.8 \\
-gen\_dfs & max\_tree\_depth=0.5 & 4.5 & 2.2 & 4.9 & 11.6 \\
-gen\_dfs & -- & 31.1 & 2.8 & 28.0 & 136.5 \\
-gen\_dfs\_percolation & p=0.1 & 53.9 & 3.6 & 42.5 & 252.9 \\
-gen\_dfs\_percolation & p=0.4 & 58.8 & 3.7 & 44.7 & 280.2 \\
-gen\_percolation & -- & 59.1 & 3.3 & 43.6 & 285.2 \\
-gen\_wilson & -- & 767.9 & 10.1 & 212.9 & 4530.4 \\
-:===============: & :===============: & :===============: &
-:===============: & :===============: & :===============: \\
-\textbf{median (all runs)} & & 10.8 & 6.0 & 44.4 & 367.7 \\
-\textbf{mean (all runs)} & & 490.0 & 11.7 & 187.2 & 2769.6 \\
-\end{longtable}
+various algorithms, parameter choices, maze sizes, and dataset sizes in
+\autoref{tab:benchmarks} and \autoref{fig:benchmarks}. Experiments were
+performed on a
+\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners#standard-github-hosted-runners-for-public-repositories}{standard GitHub runner}
+without parallelism.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|ll|r|rrr|}
+  \hline
+  maze\_ctor
+          & keyword args           & all sizes 
+                                              & \makecell{small \\ $g \leq 10$} 
+                                                         & \makecell{medium \\ $g \in (10, 32]$} 
+                                                                    & \makecell{large \\ $g > 32$} \\
+  \hline\hline
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
+          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
+          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_kruskal}{kruskal}
+          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_percolation}{percolation}
+          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
+          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_wilson}{wilson}
+          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
+  \hline\hline
+  mean
+          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
+  median
+          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
+  \hline
+\end{tabular}
+\caption{Generation times for various algorithms and maze sizes.}
+\label{tab:benchmarks}
+\end{table}
 
 \begin{figure}
+\hypertarget{fig:benchmarks}{%
 \centering
@@
 averaged across all parameter sets for that algorithm, and
-parallelization is disabled.}
+parallelization is disabled.}\label{fig:benchmarks}
+}
 \end{figure}

### 2025-03-30 docs/paper/paper.tex
some changes, add links to tokenizer stuff

@@
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Rusheb Shah, Brandon Knutson, Cecilia Diniz Behn, Samy Wu Fung},
+Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
+Fung},
     pdfpublication={Journal of Open Source Software},
@@
     pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Rusheb Shah, Brandon Knutson, Cecilia
+Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
 Diniz Behn, Samy Wu Fung},
@@
 %
-]{Rusheb Shah%
-  \,\orcidlink{0009-0008-3058-0217}\,%
+]{Tilman Räuker%
+  \,\orcidlink{0009-0009-6321-4413}\,%
 }
@@
 for storing sub-graphs of a lattice. These objects can be converted to
-and from several formats.
+and from several formats to maximize their utility in different
+contexts.
 
@@
 
-To train autoregressive text models such as transformers, we convert
-mazes to token sequences in two steps. First, the maze is stringified
-using \texttt{as\_tokens()}. The \texttt{MazeTokenizerModular} class
-provides a powerful interface for configuring maze stringification
-behavior. Second, the sequence of strings is tokenized into integers
-using \texttt{encode()}. Tokenization uses a fixed vocabulary for
-simplicity. Mazes up to 50x50 are supported using unique tokens, and up
-to 128x128 when using coordinate tuple tokens.
+Autoregressive transformer models can be quite sensitive to the exact
+format of the input data, and may use delimiter tokens to perform
+reasoning steps (\protect\hyperlink{ref-pfau2024dotbydot}{Pfau et al.,
+2024}). To facilitate systematic investigation of this, we provide a
+variety of tokenized output formats.
+
+We convert mazes to token sequences in two steps. First, the maze is
+stringified using \texttt{as\_tokens()}. The
+\texttt{MazeTokenizerModular} class provides a powerful interface for
+configuring maze stringification behavior. Second, the sequence of
+strings is tokenized into integers using \texttt{encode()}. Tokenization
+uses a fixed vocabulary for simplicity. Mazes up to \(50 \times 50\) are
+supported using unique tokens, and up to \(128 \times 128\) when using
+coordinate tuple tokens.
 
@@
 \end{figure}
-
-\textbf{TODO: add fst stuff}
 
@@
 
-We refer to our GitHub repository
-(\protect\hyperlink{ref-maze-dataset-github}{Michael I. Ivanitskiy et
-al., 2023a}) for documentation and up-to-date implementation details.
+We refer to our
+\href{https://github.com/understanding-search/maze-dataset}{GitHub repository}
+and
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
+for documentation and up-to-date implementation details.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
+Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
+Hidden computation in transformer language models. \emph{arXiv Preprint
+arXiv:2404.15758}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%

### 2025-03-30 docs/paper/paper.tex
added a bunch of links

@@
   \item
-    \texttt{RasterizedMazeDataset}, utilizing \texttt{as\_pixels()},
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
+    utilizing
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
     which can exactly mimic the outputs provided in
@@
   \item
-    \texttt{as\_ascii()} provides a format similar to that used in
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
+    provides a format similar to that used in
     (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
@@
   \item
-    \texttt{MazePlot} provides a feature-rich plotting utility with
-    support for multiple paths, heatmaps over positions, and more. This
-    is similar to the outputs of
-    (\protect\hyperlink{ref-mazegenerator-net}{Alance AB, 2019};
-    \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
+    provides a feature-rich plotting utility with support for multiple
+    paths, heatmaps over positions, and more. This is similar to the
+    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
+    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
     \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
@@
 \item
-  The text format provided by \texttt{SolvedMaze(...).as\_tokens()} is
-  similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \& Wu,
-  2023}), but provides over 5.8 million unique formats for converting
-  mazes to a text stream, detailed in
+  The text format provided by
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
+  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
+  Wu, 2023}), but provides over 5.8 million unique formats for
+  converting mazes to a text stream, detailed in
   \autoref{tokenized-output-formats}.
@@
 
-To create a dataset, we first create a \texttt{MazeDatasetConfig}
+To create a dataset, we first create a
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
 configuration object, which specifies the seed, number, and size of
 mazes, as well as the generation algorithm and its corresponding
-parameters. This object is passed to a \texttt{MazeDataset} class to
-create a dataset. Crucially, this \texttt{MazeDataset} mimics the
-interface of a PyTorch (\protect\hyperlink{ref-pytorch}{Paszke et al.,
-2019}) \texttt{Dataset}, and can thus be easily incorporated into
-existing data pre-processing and training pipelines, e.g., through the
-use of a \texttt{Dataloader} class.
+parameters. This object is passed to a
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
+class to create a dataset. Crucially, this
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
+mimics the interface of a PyTorch
+(\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
+\href{https://pytorch.org/docs/stable/data.html}{\texttt{Dataset}}, and
+can thus be easily incorporated into existing data pre-processing and
+training pipelines, e.g., through the use of a \texttt{DataLoader}
+class.
 
@@
 generation. Available maze generation algorithms are static methods of
-the \texttt{LatticeMazeGenerators} class and include generation
-algorithms based on randomized depth-first search, Wilson's algorithm
-(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
-(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
-\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
-Kruskal's algorithm
+the
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
+class and include generation algorithms based on randomized depth-first
+search, Wilson's algorithm (\protect\hyperlink{ref-wilson}{Wilson,
+1996}), percolation (\protect\hyperlink{ref-percolation}{Duminil-Copin,
+2017}; \protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam,
+2004}), Kruskal's algorithm
 (\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
@@
 reproducibility, the \texttt{seed} parameter of
-\texttt{MazeDatasetConfig} may be set. In practice, we do not find that
-exact duplicates of mazes are generated with any meaningful frequency,
-even when generating large datasets.
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
+may be set. In practice, we do not find that exact duplicates of mazes
+are generated with any meaningful frequency, even when generating large
+datasets.
 
@@
     % algorithms
-    \texttt{as\_ascii()} & \texttt{as\_pixels()} & \texttt{MazePlot()} \\[.5em]
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
+    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
+    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html#MazePlot}{\texttt{MazePlot()}} \\[.5em]
     % descriptions
@@
 (\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) and provide the \texttt{RasterizedMazeDataset}
+Bansal, et al., 2021}) and provide the
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
 class which returns rasterized pairs of (input, target) mazes as shown
@@
 We convert mazes to token sequences in two steps. First, the maze is
-stringified using \texttt{as\_tokens()}. The
-\texttt{MazeTokenizerModular} class provides a powerful interface for
-configuring maze stringification behavior. Second, the sequence of
-strings is tokenized into integers using \texttt{encode()}. Tokenization
-uses a fixed vocabulary for simplicity. Mazes up to \(50 \times 50\) are
-supported using unique tokens, and up to \(128 \times 128\) when using
-coordinate tuple tokens.
+stringified using
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens()}}.
+The
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
+class provides a powerful interface for configuring maze stringification
+behavior. Second, the sequence of strings is tokenized into integers
+using \texttt{encode()}. Tokenization uses a fixed vocabulary for
+simplicity. Mazes up to \(50 \times 50\) are supported using unique
+tokens, and up to \(128 \times 128\) when using coordinate tuple tokens.
 
@@
 tokenization algorithms affect these models, the
-\texttt{MazeTokenizerModular} class contains a rich set of options to
-customize how mazes are stringified. This class contains 19 discrete
-parameters, resulting in over 5.8 million unique tokenizers. But wait,
-there's more! There are 6 additional parameters available in the library
-which are untested but further expand the the number of tokenizers by a
-factor of \(44/3\) to 86 million.
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
+class contains a rich set of options to customize how mazes are
+stringified. This class contains 19 discrete parameters, resulting in
+over 5.8 million unique tokenizers. But wait, there's more! There are 6
+additional parameters available in the library which are untested but
+further expand the the number of tokenizers by a factor of \(44/3\) to
+86 million.
 
@@
 
-Each \texttt{MazeTokenizerModular} is constructed from a set of several
-\texttt{\_TokenizerElement} objects, each of which specifies how
-different token regions or other elements of the stringification are
-produced.
+Each
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
+is constructed from a set of several
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+objects, each of which specifies how different token regions or other
+elements of the stringification are produced.
 
@@
 in integration tests. To create a new tokenizer, developers forking the
-library may simply create their own \texttt{\_TokenizerElement} subclass
-and implement the abstract methods. If the behavior change is
+library may simply create their own
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+subclass and implement the abstract methods. If the behavior change is
 sufficiently small, simply adding a parameter to an existing
-\texttt{\_TokenizerElement} subclass and updating its implementation
-will suffice. For small additions, simply adding new cases to existing
-unit tests will suffice.
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+subclass and updating its implementation will suffice. For small
+additions, simply adding new cases to existing unit tests will suffice.
 
@@
 integration tests. If any existing subclass contains features which
-aren't needed, a developer tool decorator is provided which can be
-applied to the unneeded \texttt{\_TokenizerElement} subclasses to prune
-those features and compact the available space of tokenizers.
+aren't needed, a developer tool decorator
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
+is provided which can be applied to the unneeded
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+subclasses to prune those features and compact the available space of
+tokenizers.
 
@@
 Bansal, et al., 2021}), we allow placing additional constraints in
-\texttt{MazeDatasetConfig.endpoint\_kwargs:\ EndpointKwargsType}, such
-as enforcing that the start or end point be in a ``dead end'' with only
-one accessible neighbor cell. However, combining this with cyclic mazes
-(such as those generated with percolation), as was required for the work
-in (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et
-al., 2024}), can lead to an absence of valid start and end points.
-Placing theoretical bounds on this success rate is difficult, as it
-depends on the exact maze generation algorithm and parameters used.
-However, our package provides a way to estimate the success rate of a
-given configuration using a symbolic regression model trained with PySR
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
+such as enforcing that the start or end point be in a ``dead end'' with
+only one accessible neighbor cell. However, combining this with cyclic
+mazes (such as those generated with percolation), as was required for
+the work in
+(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
+2024}), can lead to an absence of valid start and end points. Placing
+theoretical bounds on this success rate is difficult, as it depends on
+the exact maze generation algorithm and parameters used. However, our
+package provides a way to estimate the success rate of a given
+configuration using a symbolic regression model trained with PySR
 (\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
@@
 Using the estimation simply requires the user to call
-\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}
-given their predefined \texttt{cfg}.
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
+providing their initial \texttt{cfg} and then using the returned
+\texttt{cfg\_new} in its place.
 
@@
 probability \(p\). Since we know that maze success is either guaranteed
-or impossible for \(p=0\) and \(p=1\), we define the \texttt{soft\_step}
+or impossible for \(p=0\) and \(p=1\), we define the
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
 function to nudge the raw output of the symbolic regression. This
@@
 
-We combine these to get the \texttt{soft\_step} function, which is
-identity-like for \(p \approx 0.5\), and pushes pushes \(x\) to extremes
-otherwise. \[
+We combine these to get the
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
+function, which is identity-like for \(p \approx 0.5\), and pushes
+pushes \(x\) to extremes otherwise. \[
   \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
@@
 \(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
-a \texttt{connection\_list}. The value at \(A[0,i,j]\) determines
-whether a downward connection exists from node \([i,j]\) to
-\([i+1, j]\). Likewise, the value at \(A[1,i,j]\) determines whether a
-rightwards connection to \([i, j+1]\) exists. Thus, we avoid duplication
-of data about the existence of connections, at the cost of requiring
-additional care with indexing when looking for a connection upwards or
-to the left. Note that this setup allows for a periodic lattice.
+a
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
+The value at \(A[0,i,j]\) determines whether a downward connection
+exists from node \([i,j]\) to \([i+1, j]\). Likewise, the value at
+\(A[1,i,j]\) determines whether a rightwards connection to \([i, j+1]\)
+exists. Thus, we avoid duplication of data about the existence of
+connections, at the cost of requiring additional care with indexing when
+looking for a connection upwards or to the left. Note that this setup
+allows for a periodic lattice.
 
@@
 selection can be affected by
-\texttt{MazeDatasetConfig.endpoint\_kwargs:\ EndpointKwargsType}, and
-complications caused by this are detailed in
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
+and complications caused by this are detailed in
 \autoref{success-rate-estimation}.

### 2025-03-30 docs/paper/paper.tex
figure location and makecell package workaround for ci compilation

@@
 \fi
-\usepackage{makecell}
 \usepackage{graphicx}
@@
           & keyword args           & all sizes 
-                                              & \makecell{small \\ $g \leq 10$} 
-                                                         & \makecell{medium \\ $g \in (10, 32]$} 
-                                                                    & \makecell{large \\ $g > 32$} \\
+                                              & \shortstack{small \\ $g \leq 10$} 
+                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
+                                                                    & \shortstack{large \\ $g > 32$} \\
   \hline\hline
@@
 \centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{docs/benchmarks/figures/gridsize-vs-gentime.pdf}
+\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
 \caption{Plots of maze generation time. Generation time scales

### 2025-03-30 docs/paper/paper.tex
fixing some links for ci compilation

@@
     % algorithms
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html#MazePlot}{\texttt{MazePlot()}} \\[.5em]
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
+    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
+    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
     % descriptions
@@
   \hline\hline
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
           &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
           & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
           & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs}{dfs}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
           & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
           & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
           & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_kruskal}{kruskal}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
           &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_percolation}{percolation}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
           & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
           &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMazeGenerators.gen_wilson}{wilson}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
           &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\

### 2025-03-30 docs/paper/paper.tex
suggestions from @bknutson0

@@
 generating, processing, and visualizing datasets of mazes. This library
-supports a variety of maze generation algorithms providing both mazes
-with loops and ``perfect'' mazes without them. These generation
-algorithms can be configured with various parameters, and the resulting
-mazes can be filtered to satisfy desired properties. Also provided are
-tools for converting mazes to and from various formats suitable for a
-variety of neural network architectures, such as rasterized images and
-tokenized text sequences, as well as various visualization tools. As
-well as providing a simple interface for generating, storing, and
+supports a variety of maze generation algorithms providing mazes with or
+without loops, fully-connected or not, and many other variations. These
+generation algorithms can be configured with various parameters, and the
+resulting mazes can be filtered to satisfy desired properties. Also
+provided are tools for converting mazes to and from various formats
+suitable for a variety of neural network architectures, such as
+rasterized images, tokenized text sequences, and various visualizations.
+As well as providing a simple interface for generating, storing, and
 loading these datasets, \texttt{maze-dataset} is extensively tested,
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
+For example, we can require a minimum path length from origin to target
+of 3:
 
@@
 This package utilizes a simple, efficient representation of mazes. Using
-an adjacency list to represent mazes would lead to a poor lookup time of
-whether any given connection exists, while using an adjacency matrix
+an adjacency list to represent mazes would lead to a poor lookup time
+for whether any given connection exists, while using an adjacency matrix
 would waste memory by failing to exploit the structure (e.g., only 4 of
@@
   \texttt{maze-dataset} to study the effectiveness of imperative
-  learning
+  learning.
 \end{itemize}

### 2025-04-05 docs/paper/paper.tex
temp affiliation fix
https://github.com/understanding-search/maze-dataset/pull/55#discussion_r2026567100
changes from comments from CDB up to pg6, rebuild paper

@@
 
-\author[1%
+\author[1,3%
 %
@@
 }
-\author[%
+\author[2%
 %
@@
 }
-\author[2%
+\author[2,3%
 %
@@
 \affil[2]{Imperial College London%
+}
+\affil[3]{unsearch.org%
+}
+\affil[4]{Independent%
 }
@@
 machine learning algorithms solve problems and represent spatial
-information. We introduce \texttt{maze-dataset}, a Python library for
-generating, processing, and visualizing datasets of mazes. This library
-supports a variety of maze generation algorithms providing mazes with or
-without loops, fully-connected or not, and many other variations. These
-generation algorithms can be configured with various parameters, and the
-resulting mazes can be filtered to satisfy desired properties. Also
-provided are tools for converting mazes to and from various formats
-suitable for a variety of neural network architectures, such as
-rasterized images, tokenized text sequences, and various visualizations.
-As well as providing a simple interface for generating, storing, and
-loading these datasets, \texttt{maze-dataset} is extensively tested,
-type hinted, benchmarked, and documented.
+information. We introduce \texttt{maze-dataset}, a user-friendly Python
+library for generating, processing, and visualizing datasets of mazes.
+This library supports a variety of maze generation algorithms providing
+mazes with or without loops, mazes that are connected or not, and many
+other variations. These generation algorithms can be configured with
+various parameters, and the resulting mazes can be filtered to satisfy
+desired properties. Also provided are tools for converting mazes to and
+from various formats suitable for a variety of neural network
+architectures, such as rasterized images, tokenized text sequences, and
+various visualizations. As well as providing a simple interface for
+generating, storing, and loading these datasets, \texttt{maze-dataset}
+is extensively tested, type hinted, benchmarked, and documented.
 
@@
 
-The generation of mazes with a given algorithm is not inherently a
-complex task, but the ability to seamlessly switch out algorithms,
-modify algorithm parameters, or filter by desired properties all while
-preserving the ability to convert between different representations of
-the maze is not trivial. This library aims to greatly streamline the
-process of generating and working with datasets of mazes that can be
-described as subgraphs of an \(n \times n\) lattice with boolean
-connections and, optionally, start and end points that are nodes in the
-graph. Furthermore, we place emphasis on a wide variety of possible text
-output formats aimed at evaluating the spatial reasoning capabilities of
-Large Language Models and other text-based transformer models.
+While maze generation itself is straightforward, the architectural
+challenge comes from building a system supporting many algorithms with
+configurable parameters, property filtering, and representation
+transformation. This library aims to greatly streamline the process of
+generating and working with datasets of mazes that can be described as
+subgraphs of an \(n \times n\) lattice with boolean connections and,
+optionally, start and end points that are nodes in the graph.
+Furthermore, we place emphasis on a wide variety of possible text output
+formats aimed at evaluating the spatial reasoning capabilities of Large
+Language Models and other text-based transformer models.
 
@@
 (\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
-Although mazes are well suited for these investigations, we have found
-that existing maze generation packages
+Although mazes are well suited for these investigations, we found that
+existing maze generation packages
 (\protect\hyperlink{ref-cobbe2019procgen}{Cobbe et al., 2019};
@@
 \protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) do not support flexible maze generation
-algorithms that provide fine-grained control of generation parameters
-and the ability to easily transform between multiple representations of
-the mazes (Raster, Textual, Tokenized) for training and testing models.
+Bansal, et al., 2021}) lack support for transforming between multiple
+representations and provide limited control over the maze generation
+process.
 
@@
 \protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). However, nearly all of these packages generate
-and store mazes in a form that is not optimized for storage space or,
-more importantly, computer readability. The mazes produces by other
-packages are usually rasterized or in some form of image, rather than
-the underlying graph structure, and this makes it difficult to work with
-these datasets.
+Bansal, et al., 2021}). However, nearly all of these packages produce
+mazes represented as rasterized images or other visual formats rather
+than the underlying graph structure, and this makes it difficult to work
+with these datasets.
 
@@
 \item
-  Most prior works provide mazes in some kind of image or raster format,
-  and we provide a variety of similar output formats:
+  Most prior works provide mazes in visual or raster formats, and we
+  provide a variety of similar output formats:
 
@@
   Storing mazes as images is not only difficult to work with, but also
-  inefficient. We use an extremely efficient method detailed in
+  inefficient. We use a highly efficient method detailed in
   \autoref{implementation}.
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
-For example, we can require a minimum path length from origin to target
-of 3:
+For example, we can require a minimum path length of three steps from
+the origin to the target:
 
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-may be set. In practice, we do not find that exact duplicates of mazes
-are generated with any meaningful frequency, even when generating large
-datasets.
+may be set. In practice, using provided deduplication filters, we do
+find that exact duplicate mazes are generated very infrequently, even
+when generating very large datasets.
 
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
-objects, which have path information, and a connection list optimized
-for storing sub-graphs of a lattice. These objects can be converted to
-and from several formats to maximize their utility in different
-contexts.
+objects, which have path information and a tensor optimized for storing
+sub-graphs of a lattice. These objects can be converted to and from
+several formats to maximize their utility in different contexts.
 
@@
 \caption{Input is the rasterized maze without the path marked (left),
-and provide as a target the maze with all but the correct path removed.
-Configuration options exist to adjust whether endpoints are included and
-if empty cells should be filled in.}\label{fig:e2h-raster}
+and provide as a target the maze with all but the correct path removed
+(right). Configuration options exist to adjust whether endpoints are
+included and if empty cells should be filled in.}\label{fig:e2h-raster}
 }
@@
 using \texttt{encode()}. Tokenization uses a fixed vocabulary for
-simplicity. Mazes up to \(50 \times 50\) are supported using unique
-tokens, and up to \(128 \times 128\) when using coordinate tuple tokens.
+simplicity. Mazes up to \(50 \times 50\) are supported when using a
+unique token for each position, and up to \(128 \times 128\) are
+supported when positions in the maze are represented as a pair of
+coordinates.
 
@@
 stringified. This class contains 19 discrete parameters, resulting in
-over 5.8 million unique tokenizers. But wait, there's more! There are 6
-additional parameters available in the library which are untested but
+over 5.8 million unique tokenizers. There are 6 additional parameters
+available whose functionality is not verified via automated testing, but
 further expand the the number of tokenizers by a factor of \(44/3\) to
@@
 All output sequences consist of four token regions representing
-different features of the maze of which we see an example in
+different features of the maze; an example output sequence is shown in
 \autoref{fig:token-regions}.
@@
     Example text output format with token regions highlighted.
-    \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the lattice graph,
+    \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the graph,
     \colorbox[RGB]{ 217,234,211 }{Origin}: starting coordinate,
@@
 modular architecture and the automatic inclusion of any new tokenizers
-in integration tests. To create a new tokenizer, developers forking the
-library may simply create their own
+in integration tests. To create a new variety of tokenizer, developers
+forking the library may simply create their own
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and updating its implementation will suffice. For small
-additions, simply adding new cases to existing unit tests will suffice.
+subclass and updating its implementation will suffice.
 
@@
 \includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
-\caption{Plots of maze generation time. Generation time scales
-exponentially with maze size for all algorithms (left). Generation time
-does not depend on the number of mazes being generated, and there is
-minimal overhead to initializing the generation process for a small
-dataset (right). Wilson's algorithm is notably less efficient than
-others and has high variance. Note that for both plots, values are
-averaged across all parameter sets for that algorithm, and
-parallelization is disabled.}\label{fig:benchmarks}
+\caption{Plot of maze generation time. Generation time scales
+exponentially with maze size for all algorithms. Generation time per
+maze does not depend on the number of mazes being generated, and there
+is minimal overhead to initializing the generation process for a small
+dataset. Wilson's algorithm is notably less efficient than others and
+has high variance. Note that values are averaged across all parameter
+sets for that algorithm.}\label{fig:benchmarks}
 }
@@
 (\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}), we allow placing additional constraints in
+Bansal, et al., 2021}), we allow for additional constraints in
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
@@
 
-The base function learned by symbolic regression is not particularly
-insightful, and is potentially subject to change. It is defined as
+The base function learned by symbolic regression privdes limited insight
+and may be subject to change. It is defined as
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
@@
 \item
-  (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) use
-  \texttt{maze-dataset} to compare the effectiveness of transformers
-  trained with the MLM-\(\mathcal{U}\)
+  By (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) to
+  compare the effectiveness of transformers trained with the
+  MLM-\(\mathcal{U}\)
   (\protect\hyperlink{ref-MLMU-kitouni2024factorization}{Kitouni et al.,
@@
 \item
-  (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) use
-  \texttt{maze-dataset} to study the effectiveness of imperative
-  learning.
+  By (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
+  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) to
+  study the effectiveness of imperative learning.
+\item
+  By (\protect\hyperlink{ref-zhang2025tscend}{Zhang et al., 2025}) to
+  introduce a novel framework for reasoning diffusion models.
+\item
+  By (\protect\hyperlink{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to
+  improve spatial reasoning in LLMs with GRPO.
 \end{itemize}
@@
 DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
-for hosting three of the authors for a Residency Visit, and to various
+for hosting MII, AFS, and TR for a residency visit, and to various
 members of FAR's technical staff for their advice.
@@
 and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
+Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
+models' spatial intelligence via GRPO. \emph{arXiv Preprint
+arXiv:2502.14669}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
+Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
+scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
+arXiv:2502.01989}.
+
 \end{CSLReferences}

### 2025-04-08 docs/paper/paper.tex
minor paper edits, fix figure 7 success rate

@@
 
-\author[1,3%
+\author[1%
 %
@@
 }
-\author[2%
+\author[4%
 %
@@
 }
-\author[2,3%
+\author[2%
 %
@@
 }
-\author[2%
+\author[3%
 %
@@
 }
-\affil[3]{unsearch.org%
+\affil[3]{UnSearch.org%
 }
@@
 (\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}), we allow for additional constraints in
+Bansal, et al., 2021}), the parameter
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
-such as enforcing that the start or end point be in a ``dead end'' with
-only one accessible neighbor cell. However, combining this with cyclic
-mazes (such as those generated with percolation), as was required for
-the work in
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}}
+allows for additional constraints such as enforcing that the start or
+end point be in a ``dead end'' with only one accessible neighbor cell.
+However, combining these constraints with cyclic mazes (such as those
+generated with percolation), as was required for the work in
 (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
@@
 theoretical bounds on this success rate is difficult, as it depends on
-the exact maze generation algorithm and parameters used. However, our
-package provides a way to estimate the success rate of a given
+the exact maze generation algorithm and parameters used. To deal with
+this, our package provides a way to estimate the success rate of a given
 configuration using a symbolic regression model trained with PySR
@@
 \href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
-
-Using the estimation simply requires the user to call
+Using the estimation algorithm simply requires the user to call
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
@@
 function to nudge the raw output of the symbolic regression. This
-function is defined with:
-
-A shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
-function: \[
+function is defined with the following components:
+
+shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
+function given by \[
   \sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}
@@
 where \texttt{raw\_val} is the output of the symbolic regression model.
-\(x_0\) is the percolation probability, while all other parameters from
-\texttt{\_to\_ps\_array()} only affect \texttt{raw\_val}.
+The parameter \(x_0\) is the percolation probability, while all other
+parameters from \texttt{\_to\_ps\_array()} only affect
+\texttt{raw\_val}.
 
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
 \caption{An example of both empirical and predicted success rates as a
@@
 to thank our former collaborators at AI Safety Camp and other users and
-contributors to the \texttt{maze-dataset} package: Naveen Arunachalam,
-Benji Berczi, Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris
-Mathwin, Lucia Quirke, Can Rager, Adrians Skapars, Johannes Treutlein,
-and Dan Valentine.
+contributors to the \texttt{maze-dataset} package: Benji Berczi,
+Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris Mathwin, Lucia
+Quirke, Can Rager, Adrians Skapars, Rusheb Shah, Johannes Treutlein, and
+Dan Valentine.
 

### 2025-04-08 maze_dataset/benchmark/config_sweep.py

@@ -261,6 +261,8 @@ class SweepResult(SerializableDataclass, Generic[ParamType, SweepReturnType]):
 		plot_only: bool = False,
 		show: bool = True,
 		ax: plt.Axes | None = None,
+		minify_title: bool = False,
+		legend_kwargs: dict[str, Any]|None = None,
 	) -> plt.Axes:
 		"""Plot the results of percolation analysis"""
 		# set up figure
@@ -297,6 +299,12 @@ class SweepResult(SerializableDataclass, Generic[ParamType, SweepReturnType]):
 
 		# repr of config
 		cfg_shared: dict = self.configs_shared()
+		if minify_title:
+			cfg_shared["endpoint_kwargs"] = {
+				k : v
+				for k, v in cfg_shared["endpoint_kwargs"].items()
+				if k != "except_on_no_valid_endpoint"
+			}
 		cfg_repr: str = (
 			str(cfg_shared)
 			if cfg_keys is None
@@ -323,7 +331,12 @@ class SweepResult(SerializableDataclass, Generic[ParamType, SweepReturnType]):
 				f"{self.param_key} vs {self.analyze_func.__name__}\n{cfg_repr}",
 			)
 			ax_.grid(True)
-			ax_.legend(loc="center left")
+			# ax_.legend(loc="upper center", ncol=2, bbox_to_anchor=(0.5, -0.1))
+			legend_kwargs = {
+				**dict(loc="center left"),
+				**(legend_kwargs or dict()),
+			}
+			ax_.legend(**legend_kwargs)
 
 		# save and show
 		if save_path:
@@ -443,6 +456,9 @@ def plot_grouped(  # noqa: C901
 	show: bool = True,
 	logy: bool = False,
 	save_fmt: str = "svg",
+	figsize: tuple[int, int] = (22, 10),
+	minify_title: bool = False,
+	legend_kwargs: dict[str, Any] | None = None,
 ) -> None:
 	"""Plot grouped sweep percolation value results for each distinct `endpoint_kwargs` in the configs
 
@@ -485,7 +501,7 @@ def plot_grouped(  # noqa: C901
 		)
 		shared_keys: set[str] = set(results_epkw.configs_shared().keys())
 		cfg_keys: set[str] = shared_keys.intersection({"n_mazes", "endpoint_kwargs"})
-		fig, ax = plt.subplots(1, 1, figsize=(22, 10))
[…]

### 2025-04-08 docs/paper/paper.tex
wip, working on page flow

@@
 \hypertarget{generation}{%
-\subsection{Generation and Usage}\label{generation}}
+\subsection{Generation and Basic Usage}\label{generation}}
 
@@
 \begin{Highlighting}[]
-\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
+\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
+\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
+\NormalTok{)}
+\CommentTok{\# create a config}
 \NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
-\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }
-\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{, }
-\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }
-\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs,}
+\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique, filenames are automatically made to be}
+\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
+\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
+\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
+    \CommentTok{\# (optional) algorithm{-}specific parameters}
+\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
+    \CommentTok{\# (optional) many options for restricting start/end points}
+\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
 \NormalTok{)}
-\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(cfg)}
+\CommentTok{\# create a dataset}
+\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
+\NormalTok{  cfg, }\CommentTok{\# pass the config}
+\NormalTok{  ..., }\CommentTok{\# other options such as whether to load from disk,}
+       \CommentTok{\# generate in parallel, etc.}
+\NormalTok{)}
 \end{Highlighting}
@@
 Autoregressive transformer models can be quite sensitive to the exact
-format of the input data, and may use delimiter tokens to perform
+format of input data, and may even use delimiter tokens to perform
 reasoning steps (\protect\hyperlink{ref-pfau2024dotbydot}{Pfau et al.,
-2024}). To facilitate systematic investigation of this, we provide a
-variety of tokenized output formats.
+2024}; \protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al.,
+2024}). To facilitate systematic investigation of the effects of
+different representations of data on text model performance, we provide
+a variety of tokenized text output formats.
 
@@
 We refer to our
-\href{https://github.com/understanding-search/maze-dataset}{GitHub repository}
+\href{https://github.com/understanding-search/maze-dataset}{repository}
 and
@@
 the diagonals would be filled in). Instead, we describe mazes with the
-following simple representation: for a \(d\)-dimensional lattice with
-\(r\) rows and \(c\) columns, we initialize a boolean array
+following representation: for a \(d\)-dimensional lattice with \(r\)
+rows and \(c\) columns, we initialize a boolean array
 \(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
@@
 (\protect\hyperlink{ref-ivanitskiy2023structuredworldreps}{Michael
-Igorevich Ivanitskiy et al., 2023};
+Igorevich Ivanitskiy, Spies, et al., 2023};
+\protect\hyperlink{ref-maze-dataset-arxiv-2023}{Michael Igorevich
+Ivanitskiy, Shah, et al., 2023};
 \protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al., 2024}).
-This project has also adapted itself to be useful for work on
-understanding the mechanisms by which recurrent convolutional and
-implicit networks (\protect\hyperlink{ref-fung2022jfb}{Fung et al.,
-2022}) solve mazes given a rasterized view
+It was extended for work on understanding the mechanisms by which
+recurrent convolutional and implicit networks
+(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
+given a rasterized view
 (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), and for this we match the output format of
+2024}), which required matching the pixel-padded and endpoint
+constrained output format of
 (\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}).
+Bansal, et al., 2021}). Ongoing work using \texttt{maze-dataset} aims to
+investigate the effects of varying the tokenization format on the
+performance of pretrained Large Language Models (LLMs) on spatial
+reasoning.
 
@@
 
-This work was partially supported by and many of the authors were
-brought together by AI Safety Camp and AI Safety Support. We would like
+This work was partially funded by National Science Foundation awards
+DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
+for hosting authors MII, AFS, and TR for a residency visit, and to
+various members of FAR's technical staff for their advice.
+
+This work was partially supported by AI Safety Camp and AI Safety
+Support, which also brought many of the authors together. We would like
 to thank our former collaborators at AI Safety Camp and other users and
@@
 
-This work was partially funded by National Science Foundation awards
-DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
-for hosting MII, AFS, and TR for a residency visit, and to various
-members of FAR's technical staff for their advice.
-
 We thank the Mines Optimization and Deep Learning group (MODL) for
@@
 \url{https://github.com/understanding-search/maze-transformer}
+
+\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
+Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
+Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
+C. D., \& others. (2023). A configurable library for generating and
+manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
 

### 2025-04-08 docs/paper/paper.tex
messing with page layout

@@
 \NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
-\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique, filenames are automatically made to be}
+\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
 \NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
@@
 
-When initializing mazes, further configuration options can be specified
-through the
+When initializing a dataset, options which do not affect the mazes
+themselves can be specified through the
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
-factory method as necessary. Options allow for saving/loading existing
-datasets instead of regenerating, and parallelization options for
-generation. Available maze generation algorithms are static methods of
-the
+factory method as necessary. These options allow for saving/loading
+existing datasets instead of re-generating, parallelization options for
+generation, and more. Available maze generation algorithms are static
+methods of the
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
-class and include generation algorithms based on randomized depth-first
-search, Wilson's algorithm (\protect\hyperlink{ref-wilson}{Wilson,
-1996}), percolation (\protect\hyperlink{ref-percolation}{Duminil-Copin,
-2017}; \protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam,
-2004}), Kruskal's algorithm
+namespace class and include generation algorithms based on randomized
+depth-first search, Wilson's algorithm
+(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
+(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
+\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
+Kruskal's algorithm
 (\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
@@
 when generating very large datasets.
+
+For use cases where mazes of different sizes, generation algorithms, or
+other parameter variations are required, we provide the
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/collected_dataset.html\#MazeDatasetCollection}{\texttt{MazeDatasetCollection}}
+class, which allows for creating a single iterable dataset from multiple
+independent configurations.
 
@@
       & \multicolumn{1}{c}{
-        \includegraphics[width=0.25\textwidth, trim={0 0 -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
+        \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
       } \\[1em]
@@
   \end{tabular}
-  \caption{Various output formats. Top row (left to right): ASCII diagram, rasterized pixel grid, and advanced display.}
+  \caption{Various output formats. Top row (left to right): ASCII diagram, rasterized pixel grid, and advanced display tool.}
   \label{fig:output-fmts}

### 2025-04-08 docs/paper/paper.tex
more implementation details, page flow

@@
 
-This package utilizes a simple, efficient representation of mazes. Using
-an adjacency list to represent mazes would lead to a poor lookup time
-for whether any given connection exists, while using an adjacency matrix
-would waste memory by failing to exploit the structure (e.g., only 4 of
-the diagonals would be filled in). Instead, we describe mazes with the
-following representation: for a \(d\)-dimensional lattice with \(r\)
-rows and \(c\) columns, we initialize a boolean array
-\(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
-a
+This package utilizes a simple, efficient representation of mazes as
+subgraphs of a finite lattice, which we call a
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
+Using an adjacency matrix would be memory inefficient by failing to
+exploit the highly sparse structure -- for example, for a 2-dimsnionsal
+maze, only 4 of the diagonals would be filled in. On the other hand,
+using an adjacency list could lead to a poor lookup time for whether any
+given connection exists.
+
+Instead, we describe mazes with the following representation: for a
+\(d\)-dimensional lattice with \(r\) rows and \(c\) columns, we
+initialize a boolean array \[
+  A = \{0, 1\}^{d \times r \times c}
+\] which we refer to in the code as a
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
-The value at \(A[0,i,j]\) determines whether a downward connection
-exists from node \([i,j]\) to \([i+1, j]\). Likewise, the value at
-\(A[1,i,j]\) determines whether a rightwards connection to \([i, j+1]\)
-exists. Thus, we avoid duplication of data about the existence of
-connections, at the cost of requiring additional care with indexing when
-looking for a connection upwards or to the left. Note that this setup
-allows for a periodic lattice.
+In two dimensions, the value at \(A[0,i,j]\) determines whether a
+\emph{downward} connection exists from node \([i,j]\) to \([i+1, j]\).
+Likewise, the value at \(A[1,i,j]\) determines whether a
+\emph{rightward} connection to \([i, j+1]\) exists. Thus, we avoid
+duplication of data about the existence of connections, at the cost of
+requiring additional care with indexing when looking for a connection
+upwards or to the left. Note that this setup allows for a periodic
+lattice. This format generalizes to lattices of higher dimensions, but
+our package provides only limited support for generation and
+visualization of higher dimensional mazes. Generation of mazes is
+detailed in
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
 
@@
 1968}) is applied to find the shortest path between them. The endpoint
-selection can be affected by
+selection can be controlled via
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
@@
 
+\newpage
+
 \hypertarget{references}{%

### 2025-04-08 docs/paper/paper.tex
minor edits

@@
 formats aimed at evaluating the spatial reasoning capabilities of Large
-Language Models and other text-based transformer models.
+Language Models (LLMs) and other text-based transformer models.
 
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-may be set. In practice, using provided deduplication filters, we do
-find that exact duplicate mazes are generated very infrequently, even
-when generating very large datasets.
+may be set. In practice, using provided deduplication filters, we find
+that exact duplicate mazes are generated very infrequently, even when
+generating very large datasets.
 
@@
 investigate the effects of varying the tokenization format on the
-performance of pretrained Large Language Models (LLMs) on spatial
-reasoning.
+performance of pretrained LLMs on spatial reasoning.
 

### 2025-04-08 docs/paper/paper.tex
examples, docs, notebooks link

@@
     \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    provides a format similar to that used in
+    provides a format similar to
     (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
@@
 
+We direct readers to our
+\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples},
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs},
+and
+\href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
+for more information.
+
 \hypertarget{generation}{%
@@
 \NormalTok{  cfg, }\CommentTok{\# pass the config}
-\NormalTok{  ..., }\CommentTok{\# other options such as whether to load from disk,}
-       \CommentTok{\# generate in parallel, etc.}
+\NormalTok{  ..., }\CommentTok{\# other options for disk loading, parallelization, etc.}
 \NormalTok{)}

### 2025-04-08 docs/paper/paper.tex
provisional title

@@
     pdfaconformance=B,
-    pdftitle={maze-dataset},
+    pdftitle={maze-dataset: Algorithmic Variety and Representation
+Flexibility in Maze Generation},
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
@@
 \fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset. \emph{Journal of Open Source Software},
+et al. (1970). maze-dataset: Algorithmic Variety and Representation
+Flexibility in Maze Generation. \emph{Journal of Open Source Software},
 \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
@@
 
-\title{maze-dataset}
+\title{maze-dataset: Algorithmic Variety and Representation Flexibility
+in Maze Generation}
 

### 2025-04-09 docs/paper/paper.tex
fix figure again

--- 2025-04-08 docs/paper/paper.tex
@@
     pdfaconformance=B,
-    pdftitle={maze-dataset},
+    pdftitle={maze-dataset: User Friendly Maze Generation with
+Algorithmic Variety and Representation Flexibility},
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
@@
 \fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset. \emph{Journal of Open Source Software},
-\emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
+et al. (1970). maze-dataset: User Friendly Maze Generation with
+Algorithmic Variety and Representation Flexibility. \emph{Journal of
+Open Source Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE?
+\url{https://doi.org/N/A}}.}}
 
@@
 
-\title{maze-dataset}
+\title{maze-dataset: User Friendly Maze Generation with Algorithmic
+Variety and Representation Flexibility}
 
@@
 \texttt{endpoint\_kwargs} requiring that both the start and end be in
-unique dead ends.}
+unique dead ends. Empirical measures derived from a sample of 128
+mazes.}
 \end{figure}

### 2025-04-09 maze_dataset/benchmark/config_sweep.py

@@ -4,7 +4,7 @@ import functools
 import json
 import warnings
 from pathlib import Path
-from typing import Any, Callable, Generic, Sequence, TypeVar
+from typing import Any, Callable, Generic, Literal, Sequence, TypeVar
 
 import matplotlib.pyplot as plt
 import numpy as np
@@ -459,6 +459,7 @@ def plot_grouped(  # noqa: C901
 	figsize: tuple[int, int] = (22, 10),
 	minify_title: bool = False,
 	legend_kwargs: dict[str, Any] | None = None,
+	manual_titles: dict[Literal["x", "y", "title"], str] | None = None,
 ) -> None:
 	"""Plot grouped sweep percolation value results for each distinct `endpoint_kwargs` in the configs
 
@@ -544,6 +545,11 @@ def plot_grouped(  # noqa: C901
 
 					# Plot prediction as dashed line
 					ax.plot(p_dense, predictions, "--", color=color, alpha=0.8)
+
+			if manual_titles:
+				ax.set_xlabel(manual_titles["x"])
+				ax.set_ylabel(manual_titles["y"])
+				ax.set_title(manual_titles["title"])
 
 		# save and show
 		if save_dir:

### 2025-04-09 docs/benchmarks/benchmark_generation.py
make format, re-run nb

@@ -55,7 +55,6 @@ def get_system_info() -> dict[str, Any]:
 
 
 if __name__ == "__main__":
-
 	print(get_system_info())
 
 	import argparse

### 2025-04-09 maze_dataset/benchmark/config_sweep.py

@@ -262,7 +262,7 @@ class SweepResult(SerializableDataclass, Generic[ParamType, SweepReturnType]):
 		show: bool = True,
 		ax: plt.Axes | None = None,
 		minify_title: bool = False,
-		legend_kwargs: dict[str, Any]|None = None,
+		legend_kwargs: dict[str, Any] | None = None,
 	) -> plt.Axes:
 		"""Plot the results of percolation analysis"""
 		# set up figure
@@ -301,7 +301,7 @@ class SweepResult(SerializableDataclass, Generic[ParamType, SweepReturnType]):
 		cfg_shared: dict = self.configs_shared()
 		if minify_title:
 			cfg_shared["endpoint_kwargs"] = {
-				k : v
+				k: v
 				for k, v in cfg_shared["endpoint_kwargs"].items()
 				if k != "except_on_no_valid_endpoint"
 			}

### 2025-04-09 maze_dataset/generation/generators.py
add links to docs

@@ -51,7 +51,11 @@ def get_neighbors_in_bounds(
 
 
 class LatticeMazeGenerators:
-	"""namespace for lattice maze generation algorithms"""
+	"""namespace for lattice maze generation algorithms
+
+	examples of generated mazes can be found here:
+	https://understanding-search.github.io/maze-dataset/examples/maze_examples.html
+	"""
 
 	@staticmethod
 	def gen_dfs(

### 2025-04-09 docs/paper/paper.tex
add benchmarks links, change title, trying to fix section autorefs

--- 2025-04-08 docs/paper/paper.tex
@@
     pdfaconformance=B,
-    pdftitle={maze-dataset},
+    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
+Representational Flexibility},
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
@@
 \fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset. \emph{Journal of Open Source Software},
-\emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
+et al. (1970). maze-dataset: Maze Generation with Algorithmic Variety
+and Representational Flexibility. \emph{Journal of Open Source
+Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
 
@@
 
-\title{maze-dataset}
+\title{maze-dataset: Maze Generation with Algorithmic Variety and
+Representational Flexibility}
 
@@
   converting mazes to a text stream, detailed in
-  \autoref{tokenized-output-formats}.
+  \autoref{sec:tokenized-output-formats}.
 \item
@@
   inefficient. We use a highly efficient method detailed in
-  \autoref{implementation}.
+  \autoref{sec:implementation}.
 \item
@@
 
-\hypertarget{tokenized-output-formats}{%
-\subsection{Tokenized Output Formats}\label{tokenized-output-formats}}
+\hypertarget{sec:tokenized-output-formats}{%
+\subsection{Tokenized Output
+Formats}\label{sec:tokenized-output-formats}}
 
@@
 has high variance. Note that values are averaged across all parameter
-sets for that algorithm.}\label{fig:benchmarks}
+sets for that algorithm. More information can be found on the
+\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
+page}.}\label{fig:benchmarks}
 }
@@
 
-\hypertarget{success-rate-estimation}{%
-\subsection{Success Rate Estimation}\label{success-rate-estimation}}
+\hypertarget{sec:success-rate-estimation}{%
+\subsection{Success Rate Estimation}\label{sec:success-rate-estimation}}
 
@@
 \texttt{endpoint\_kwargs} requiring that both the start and end be in
-unique dead ends.}
+unique dead ends. Empirical measures derived from a sample of 128 mazes.
+More information can be found on the
+\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
+page}.}
 \end{figure}
 
-\hypertarget{implementation}{%
-\section{Implementation}\label{implementation}}
+\hypertarget{sec:implementation}{%
+\section{Implementation}\label{sec:implementation}}
 
@@
 and complications caused by this are detailed in
-\autoref{success-rate-estimation}.
+\autoref{sec:success-rate-estimation}.
 

### 2025-04-09 docs/paper/paper.tex
fix section references
minor changes to implementation section

@@
 \newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
+\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
+\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
 \ifLuaTeX
@@
   converting mazes to a text stream, detailed in
-  \autoref{sec:tokenized-output-formats}.
+  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
 \item
@@
   inefficient. We use a highly efficient method detailed in
-  \autoref{sec:implementation}.
+  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
 \item
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-Using an adjacency matrix would be memory inefficient by failing to
-exploit the highly sparse structure -- for example, for a 2-dimsnionsal
-maze, only 4 of the diagonals would be filled in. On the other hand,
-using an adjacency list could lead to a poor lookup time for whether any
-given connection exists.
+Using an adjacency matrix for storing mazes would be memory inefficient
+by failing to exploit the highly sparse structure -- for example, for a
+2-dimensional maze, only 4 of the diagonals would be filled in. On the
+other hand, using an adjacency list could lead to a poor lookup time for
+whether any given connection exists.
 
 Instead, we describe mazes with the following representation: for a
-\(d\)-dimensional lattice with \(r\) rows and \(c\) columns, we
+\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
 initialize a boolean array \[
-  A = \{0, 1\}^{d \times r \times c}
+  A = \{0, 1\}^{2 \times r \times c}
 \] which we refer to in the code as a
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
-In two dimensions, the value at \(A[0,i,j]\) determines whether a
-\emph{downward} connection exists from node \([i,j]\) to \([i+1, j]\).
-Likewise, the value at \(A[1,i,j]\) determines whether a
-\emph{rightward} connection to \([i, j+1]\) exists. Thus, we avoid
-duplication of data about the existence of connections, at the cost of
-requiring additional care with indexing when looking for a connection
-upwards or to the left. Note that this setup allows for a periodic
-lattice. This format generalizes to lattices of higher dimensions, but
-our package provides only limited support for generation and
-visualization of higher dimensional mazes. Generation of mazes is
-detailed in
+The value at \(A[0,i,j]\) determines whether a \emph{downward}
+connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
+value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
+to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
+existence of connections and facillitate fast lookup time, at the cost
+of requiring additional care with indexing. Note that this setup allows
+for a periodic lattice. Generation of mazes is detailed in
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
@@
 and complications caused by this are detailed in
-\autoref{sec:success-rate-estimation}.
+\hyperref[sec:success-rate-estimation]{section: \textit{\nameref{sec:success-rate-estimation}}}.
+A maze with a solution is denoted a
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}},
+which inherits from
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
 
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}.
+
+\newpage
 

### 2025-04-09 docs/paper/paper.tex
fix some typos

@@
 \end{tabular}
-\caption{Generation times for various algorithms and maze sizes.}
+\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
 \label{tab:benchmarks}
@@
 
-The base function learned by symbolic regression privdes limited insight
-and may be subject to change. It is defined as
+The base function learned by symbolic regression provides limited
+insight and may be subject to change. It is defined as
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
 and takes a 5 dimensional float vector created by
-\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the 0)
-percolation value 1) grid size 2) endpoint deadend configuration 3)
-endpoint uniqueness 4) categorical generation function index.
+\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the
+{[}percolation value, grid size, endpoint deadend configuration,
+endpoint uniqueness, categorical generation function index{]}.
 
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function, which is identity-like for \(p \approx 0.5\), and pushes
-pushes \(x\) to extremes otherwise. \[
+function, which is identity-like for \(p \approx 0.5\), and pushes \(x\)
+to extremes otherwise. \[
   \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
@@
 to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
-existence of connections and facillitate fast lookup time, at the cost
-of requiring additional care with indexing. Note that this setup allows
-for a periodic lattice. Generation of mazes is detailed in
+existence of connections and facilitate fast lookup time, at the cost of
+requiring additional care with indexing. Note that this setup allows for
+a periodic lattice. Generation of mazes is detailed in
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.

### 2025-04-09 docs/paper/paper.tex
minor fixes

@@
 by failing to exploit the highly sparse structure -- for example, for a
-2-dimensional maze, only 4 of the diagonals would be filled in. On the
-other hand, using an adjacency list could lead to a poor lookup time for
-whether any given connection exists.
+2-dimensional maze, only 4 off-diagonal bands would be have nonzero
+values. On the other hand, using an adjacency list could lead to a poor
+lookup time for whether any given connection exists.
 

### 2025-04-09 docs/paper/paper.tex
fix docs link
fix page breaks

@@
 \end{tabular}
-\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
+\caption{Generation times for various algorithms and maze sizes. More information can be found on the \href{https://understanding-search.github.io/maze-dataset/benchmarks}{benchmarks page}.}
 \label{tab:benchmarks}
@@
 \centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
+\includegraphics[width=0.9\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
 \caption{Plot of maze generation time. Generation time scales

### 2025-04-26 docs/paper/figures/tex/fig1_diagram.tex
move figure code to external files

--- 2025-04-09 docs/paper/paper.tex
@@
-\documentclass[10pt,a4paper,onecolumn]{article}
-\usepackage{marginnote}
-\usepackage{graphicx}
-\usepackage[rgb,svgnames]{xcolor}
-\usepackage{authblk,etoolbox}
-\usepackage{titlesec}
-\usepackage{calc}
-\usepackage{tikz}
-\usepackage[pdfa]{hyperref}
-\usepackage{hyperxmp}
-\hypersetup{%
-    unicode=true,
-    pdfapart=3,
-    pdfaconformance=B,
-    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility},
-    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
-Fung},
-    pdfpublication={Journal of Open Source Software},
-    pdfpublisher={Open Journals},
-    pdfissn={2475-9066},
-    pdfpubtype={journal},
-    pdfvolumenum={},
-    pdfissuenum={},
-    pdfdoi={N/A},
-    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
-Diniz Behn, Samy Wu Fung},
-    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
-    colorlinks=true,
-    linkcolor=[rgb]{0.0, 0.5, 1.0},
-    citecolor=Blue,
-    urlcolor=[rgb]{0.0, 0.5, 1.0},
-    breaklinks=true
-}
-% https://tex.stackexchange.com/a/535849
-% Create an OutputIntent in order to correctly specify colours
-\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
-\pdfcatalog{%
-  /OutputIntents [
-    <<
-      /Type /OutputIntent
-      /S /GTS_PDFA1
-      /DestOutputProfile \the\pdflastobj\space 0 R
-      /OutputConditionIdentifier (sRGB)
-      /Info (sRGB)
-    >>
-  ]
-}
-\pdfvariable omitcidset=1
-\usepackage{caption}
-\usepackage{orcidlink}
-\usepackage{tcolorbox}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{seqsplit}
-\usepackage{xstring}
-
-\usepackage{float}
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
-
-\usepackage{fixltx2e} % provides \textsubscript
-
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-% --- Page layout -------------------------------------------------------------
-\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
-            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}
-
-% --- Default font ------------------------------------------------------------
-\renewcommand\familydefault{\sfdefault}
-
-% --- Style -------------------------------------------------------------------
-\renewcommand{\captionfont}{\small\sffamily}
-\renewcommand{\captionlabelfont}{\bfseries}
-
-% --- Section/SubSection/SubSubSection ----------------------------------------
-\titleformat{\section}
-  {\normalfont\sffamily\Large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsection}
-  {\normalfont\sffamily\large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsubsection}
-  {\normalfont\sffamily\bfseries}
-  {}{0pt}{}
-\titleformat*{\paragraph}
-  {\sffamily\normalsize}
-
-% --- Header / Footer ---------------------------------------------------------
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-%\renewcommand{\headrulewidth}{0.50pt}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{joss/logo.png}}
-\fancyhead[C]{}
-\fancyhead[R]{}
-\renewcommand{\footrulewidth}{0.25pt}
-
-\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset: Maze Generation with Algorithmic Variety
-and Representational Flexibility. \emph{Journal of Open Source
-Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
-
-\fancyfoot[R]{\sffamily \thepage}
-\makeatletter
-\let\ps@plain\ps@fancy
-\fancyheadoffset[L]{4.5cm}
-\fancyfootoffset[L]{4.5cm}
-
-% --- Macros ---------
-
-\definecolor{linky}{rgb}{0.0, 0.5, 1.0}
-
-\newtcolorbox{repobox}
-   {colback=red, colframe=red!75!black,
-     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}
-
-\newcommand{\ExternalLink}{%
-   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
-       \begin{scope}[x=1ex, y=1ex]
-           \clip (-0.1,-0.1)
-               --++ (-0, 1.2)
-               --++ (0.6, 0)
-               --++ (0, -0.6)
-               --++ (0.6, 0)
-               --++ (0, -1);
-           \path[draw,
-               line width = 0.5,
-               rounded corners=0.5]
-               (0,0) rectangle (1,1);
-       \end{scope}
-       \path[draw, line width = 0.5] (0.5, 0.5)
-           -- (1, 1);
-       \path[draw, line width = 0.5] (0.6, 1)
-           -- (1, 1) -- (1, 0.6);
-       }
-   }
-
-\definecolor{c53baa1}{RGB}{83,186,161}
-\definecolor{c202826}{RGB}{32,40,38}
-\def \rorglobalscale {0.1}
-\newcommand{\rorlogo}{%
-\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
-  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
-    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364)..
-controls (3.9643, 1.4576) and (4.0508, 1.5081) ..
-(4.1098, 1.5878)..
-controls (4.169, 1.6674) and (4.1984, 1.7642) ..
-(4.1984, 1.8777)..
-controls (4.1984, 1.9719) and (4.182, 2.0503) ..
-(4.1495, 2.1132)..
-controls (4.1169, 2.1762) and (4.0727, 2.2262) ..
-(4.0174, 2.2635)..
-controls (3.9621, 2.3006) and (3.8976, 2.3273) ..
-(3.824, 2.3432)..
-controls (3.7505, 2.359) and (3.6727, 2.367) ..
-(3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688)..
-controls (2.9625, 1.8833) and (2.9572, 1.8976) ..
-(2.9514, 1.9119)..
-controls (2.9083, 2.0164) and (2.848, 2.1056) ..
-(2.7705, 2.1791)..
-controls (2.6929, 2.2527) and (2.6014, 2.3093) ..
-(2.495, 2.3487)..
-controls (2.3889, 2.3881) and (2.2728, 2.408) ..
-(2.1468, 2.408)..
-controls (2.0209, 2.408) and (1.905, 2.3881) ..
-(1.7986, 2.3487)..
-controls (1.6925, 2.3093) and (1.6007, 2.2527) ..
-(1.5232, 2.1791)..
-controls (1.4539, 2.1132) and (1.3983, 2.0346) ..
-(1.3565, 1.9436)..
-controls (1.3504, 2.009) and (1.3351, 2.0656) ..
-(1.3105, 2.1132)..
-controls (1.2779, 2.1762) and (1.2338, 2.2262) ..
-(1.1785, 2.2635)..
-controls (1.1232, 2.3006) and (1.0586, 2.3273) ..
-(0.985, 2.3432)..
-controls (0.9115, 2.359) and (0.8337, 2.367) ..
-(0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364)..
-controls (1.1255, 1.4576) and (1.2121, 1.5081) ..
-(1.2711, 1.5878)..
-controls (1.2737, 1.5915) and (1.2761, 1.5954) ..
-(1.2787, 1.5991)..
-controls (1.2782, 1.5867) and (1.2779, 1.5743) ..
-(1.2779, 1.5616)..
-controls (1.2779, 1.4327) and (1.2996, 1.3158) ..
-(1.3428, 1.2113)..
-controls (1.3859, 1.1068) and (1.4462, 1.0176) ..
-(1.5237, 0.944)..
-controls (1.601, 0.8705) and (1.6928, 0.8139) ..
-(1.7992, 0.7744)..
-controls (1.9053, 0.735) and (2.0214, 0.7152) ..
-(2.1474, 0.7152)..
-controls (2.2733, 0.7152) and (2.3892, 0.735) ..
-(2.4956, 0.7744)..
-controls (2.6016, 0.8139) and (2.6935, 0.8705) ..
-(2.771, 0.944)..
-controls (2.8482, 1.0176) and (2.9086, 1.1068) ..
-(2.952, 1.2113)..
-controls (2.9578, 1.2253) and (2.9631, 1.2398) ..
-(2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777)..
-controls (0.9438, 1.7534) and (0.92, 1.7357) ..
-(0.8911, 1.7243)..
-controls (0.8623, 1.7129) and (0.83, 1.706) ..
-(0.7945, 1.7039)..
-controls (0.7588, 1.7015) and (0.7252, 1.7005) ..
-(0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667)..
-controls (0.7477, 2.0667) and (0.7805, 2.0643) ..
-(0.8139, 2.0598)..
-controls (0.8472, 2.0553) and (0.8768, 2.0466) ..
-(0.9025, 2.0336)..
-controls (0.9282, 2.0206) and (0.9496, 2.0021) ..
-(0.9663, 1.9778)..
-controls (0.9829, 1.9534) and (0.9914, 1.9209) ..
-(0.9914, 1.8799)..
-controls (0.9914, 1.8362) and (0.9819, 1.8021) ..
-(0.9628, 1.7777) -- cycle(2.6125, 1.3533)..
-controls (2.5889, 1.2904) and (2.5553, 1.2359) ..
-(2.5112, 1.1896)..
-controls (2.4672, 1.1433) and (2.4146, 1.1073) ..
-(2.3529, 1.0814)..
-controls (2.2916, 1.0554) and (2.2228, 1.0427) ..
-(2.1471, 1.0427)..
-controls (2.0712, 1.0427) and (2.0026, 1.0557) ..
-(1.9412, 1.0814)..
-controls (1.8799, 1.107) and (1.8272, 1.1433) ..
-(1.783, 1.1896)..
-controls (1.7391, 1.2359) and (1.7052, 1.2904) ..
-(1.6817, 1.3533)..
-controls (1.6581, 1.4163) and (1.6465, 1.4856) ..
-(1.6465, 1.5616)..
-controls (1.6465, 1.6359) and (1.6581, 1.705) ..
-(1.6817, 1.7687)..
-controls (1.7052, 1.8325) and (1.7388, 1.8873) ..
-(1.783, 1.9336)..
-controls (1.8269, 1.9799) and (1.8796, 2.0159) ..
-(1.9412, 2.0418)..
-controls (2.0026, 2.0675) and (2.0712, 2.0804) ..
-(2.1471, 2.0804)..
-controls (2.223, 2.0804) and (2.2916, 2.0675) ..
-(2.3529, 2.0418)..
-controls (2.4143, 2.0161) and (2.467, 1.9799) ..
-(2.5112, 1.9336)..
-controls (2.5551, 1.8873) and (2.5889, 1.8322) ..
-(2.6125, 1.7687)..
-controls (2.636, 1.705) and (2.6477, 1.6359) ..
-(2.6477, 1.5616)..
-controls (2.6477, 1.4856) and (2.636, 1.4163) ..
-(2.6125, 1.3533) -- cycle(3.8015, 1.7777)..
-controls (3.7825, 1.7534) and (3.7587, 1.7357) ..
-(3.7298, 1.7243)..
-controls (3.701, 1.7129) and (3.6687, 1.706) ..
-(3.6333, 1.7039)..
-controls (3.5975, 1.7015) and (3.5639, 1.7005) ..
-(3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667)..
-controls (3.5864, 2.0667) and (3.6192, 2.0643) ..
-(3.6526, 2.0598)..
-controls (3.6859, 2.0553) and (3.7155, 2.0466) ..
-(3.7412, 2.0336)..
-controls (3.7669, 2.0206) and (3.7883, 2.0021) ..
-(3.805, 1.9778)..
-controls (3.8216, 1.9534) and (3.8301, 1.9209) ..
-(3.8301, 1.8799)..
-controls (3.8301, 1.8362) and (3.8206, 1.8021) ..
-(3.8015, 1.7777) -- cycle;
-  \end{scope}
-\end{tikzpicture}
-}
-
-% --- Title / Authors ---------------------------------------------------------
-% patch \maketitle so that it doesn't center
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-% patch \maketitle so that the font size for the title is normal
-\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
-% patch the patch by authblk so that the author block is flush left
-\def\maketitle{{%
-  \renewenvironment{tabular}[2][]
-    {\begin{flushleft}}
-    {\end{flushleft}}
-  \AB@maketitle}}
-\renewcommand\AB@affilsepx{ \protect\Affilfont}
-%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
-\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
-\renewcommand{\affil}[2][]%
-   {\newaffiltrue\let\AB@blk@and\AB@pand
-      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
-        \setcounter{Maxaffil}{0}\fi
-        \begingroup
-        \let\href=\href@Orig
-        \let\protect\@unexpandable@protect
-        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
-        \@temptokena=\expandafter{\AB@authors}%
-        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
-         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
-         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
-         \gdef\AB@las{}\gdef\AB@au@str{}%
-        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
-        \@temptokena=\expandafter{\AB@affillist}%
-        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
-          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
-      \endgroup
-       \let\AB@affilsep\AB@affilsepx
-}
-\makeatother
-\renewcommand\Authfont{\sffamily\bfseries}
-\renewcommand\Affilfont{\sffamily\small\mdseries}
-\setlength{\affilsep}{1em}
-
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{fontspec}
-
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\sffamily]{Ligatures=TeX}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-
-%% Font settings
-\usepackage{fontsetup} % Lazy way to get proper Greek lowercase glyphs
-
-% Use Hack https://sourcefoundry.org/hack/
-\setmonofont{Hack}
-
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\urlstyle{same}  % don't use monospace font for urls
-\ifLuaTeX
-\usepackage[bidi=basic]{babel}
-\else
-\usepackage[bidi=default]{babel}
-\fi
-\babelprovide[main,import]{american}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\tikzset{ % Define a TikZ style for an external hyperlink node.
-  hyperlink node url/.style={
-    alias=sourcenode,
-    append after command={
-      let \p1 = (sourcenode.north west),
-          \p2 = (sourcenode.south east),
-          \n1 = {\x2-\x1},
-          \n2 = {\y1-\y2} in
-      node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
-          {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
-    }
-  }
-}
-\providecommand{\XeTeXLinkBox}[1]{#1}
-\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
-\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
-\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-
-\title{maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility}
-
-\author[1%
-%
-\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
-  \,\orcidlink{0000-0002-4213-4993}\,%
-}
-\author[4%
-%
-]{Aaron Sandoval%
-  \,\orcidlink{0009-0002-8380-6140}\,%
-}
-\author[2%
-%
-]{Alex F. Spies%
-  \,\orcidlink{0000-0002-8708-1530}\,%
-}
-\author[3%
-%
-]{Tilman Räuker%
-  \,\orcidlink{0009-0009-6321-4413}\,%
-}
-\author[1%
-%
-]{Brandon Knutson%
-  \,\orcidlink{0009-0004-8413-0239}\,%
-}
-\author[1%
-%
-]{Cecilia Diniz Behn%
-  \,\orcidlink{0000-0002-8078-5105}\,%
-}
-\author[1%
-%
-]{Samy Wu Fung%
-  \,\orcidlink{0000-0002-2926-4582}\,%
-}
-
-\affil[1]{Colorado School of Mines, Department of Applied Mathematics
-and Statistics%
-}
-\affil[2]{Imperial College London%
-}
-\affil[3]{UnSearch.org%
-}
-\affil[4]{Independent%
-}
-\affil[$\mathparagraph$]{Corresponding author}
-\date{\vspace{-2.5ex}}
-
-\begin{document}
-\maketitle
-
-\marginpar{
-
-  \begin{flushleft}
-  %\hrule
-  \sffamily\small
-
-  {\bfseries DOI:} \href{https://doi.org/N/A}{\color{linky}{N/A}}
-
-  \vspace{2mm}
-    {\bfseries Software}
-  \begin{itemize}
-    \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{\color{linky}{Review}} \ExternalLink
-    \item \href{https://github.com/openjournals}{\color{linky}{Repository}} \ExternalLink
-    \item \href{https://doi.org/10.5281}{\color{linky}{Archive}} \ExternalLink
-  \end{itemize}
-
-  \vspace{2mm}
-
-    \par\noindent\hrulefill\par
-
-  \vspace{2mm}
-
-  {\bfseries Editor:} \href{https://joss.theoj.org}{Open
-Journals} \ExternalLink
-   \\
-  \vspace{1mm}
-    {\bfseries Reviewers:}
-  \begin{itemize}
-  \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{@openjournals}
-    \end{itemize}
-    \vspace{2mm}
-
-    {\bfseries Submitted:} 01 January 1970\\
-    {\bfseries Published:} 01 January 1970
-
-  \vspace{2mm}
-  {\bfseries License}\\
-  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{https://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).
-
-  \end{flushleft}
-}
-
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
-
-Solving mazes is a classic problem in computer science and artificial
-intelligence, and humans have been constructing mazes for thousands of
-years. Although finding the shortest path through a maze is a solved
-problem, this very fact makes it an excellent testbed for studying how
-machine learning algorithms solve problems and represent spatial
-information. We introduce \texttt{maze-dataset}, a user-friendly Python
-library for generating, processing, and visualizing datasets of mazes.
-This library supports a variety of maze generation algorithms providing
-mazes with or without loops, mazes that are connected or not, and many
-other variations. These generation algorithms can be configured with
-various parameters, and the resulting mazes can be filtered to satisfy
-desired properties. Also provided are tools for converting mazes to and
-from various formats suitable for a variety of neural network
-architectures, such as rasterized images, tokenized text sequences, and
-various visualizations. As well as providing a simple interface for
-generating, storing, and loading these datasets, \texttt{maze-dataset}
-is extensively tested, type hinted, benchmarked, and documented.
-
 \begin{figure} 
@@
 This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats.
-Code in the image contains clickable links to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{documentation}.
-A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.
+Code in the image contains clickable links to \docslink{maze_dataset.html}{documentation}.
+A variety of generated examples can be viewed \docslink{examples/maze_examples.html}{here}.
 	}
@@
   \end{figure}
-
-\hypertarget{statement-of-need}{%
-\section{Statement of Need}\label{statement-of-need}}
-
-While maze generation itself is straightforward, the architectural
-challenge comes from building a system supporting many algorithms with
-configurable parameters, property filtering, and representation
-transformation. This library aims to greatly streamline the process of
-generating and working with datasets of mazes that can be described as
-subgraphs of an \(n \times n\) lattice with boolean connections and,
-optionally, start and end points that are nodes in the graph.
-Furthermore, we place emphasis on a wide variety of possible text output
-formats aimed at evaluating the spatial reasoning capabilities of Large
-Language Models (LLMs) and other text-based transformer models.
-
-For interpretability and behavioral research, algorithmic tasks offer
-benefits by allowing systematic data generation and task decomposition,
-as well as simplifying the process of circuit discovery
-(\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
-Although mazes are well suited for these investigations, we found that
-existing maze generation packages
-(\protect\hyperlink{ref-cobbe2019procgen}{Cobbe et al., 2019};
-\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-harriesMazeExplorerCustomisable3D2019}{Harries et
-al., n.d.}; \protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) lack support for transforming between multiple
-representations and provide limited control over the maze generation
-process.
-
-\hypertarget{related-works}{%
-\subsection{Related Works}\label{related-works}}
-
-A multitude of public and open-source software packages exist for
-generating mazes (\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). However, nearly all of these packages produce
-mazes represented as rasterized images or other visual formats rather
-than the underlying graph structure, and this makes it difficult to work
-with these datasets.
-
-\begin{itemize}
-\item
-  Most prior works provide mazes in visual or raster formats, and we
-  provide a variety of similar output formats:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
-    utilizing
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
-    which can exactly mimic the outputs provided in
-    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
-    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
-    similar to the outputs of Németh
-    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    provides a format similar to
-    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
-    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
-    provides a feature-rich plotting utility with support for multiple
-    paths, heatmaps over positions, and more. This is similar to the
-    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
-    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
-    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
-  \end{itemize}
-\item
-  The text format provided by
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
-  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
-  Wu, 2023}), but provides over 5.8 million unique formats for
-  converting mazes to a text stream, detailed in
-  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
-\item
-  For rigorous investigations of the response of a model to various
-  distributional shifts, preserving metadata about the generation
-  algorithm with the dataset itself is essential. To this end, our
-  package efficiently stores the dataset along with its metadata in a
-  single human-readable file (\protect\hyperlink{ref-zanj}{M.
-  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
-  this reliably.
-\item
-  Storing mazes as images is not only difficult to work with, but also
-  inefficient. We use a highly efficient method detailed in
-  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
-\item
-  Our package is easily installable with source code freely available.
-  It is extensively tested, type hinted, benchmarked, and documented.
-  Many other maze generation packages lack this level of rigor and
-  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
-  2008}) appear to simply no longer be accessible.
-\end{itemize}
-
-\hypertarget{features}{%
-\section{Features}\label{features}}
-
-We direct readers to our
-\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples},
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs},
-and
-\href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
-for more information.
-
-\hypertarget{generation}{%
-\subsection{Generation and Basic Usage}\label{generation}}
-
-Our package can be installed from
-\href{https://pypi.org/project/maze-dataset/}{PyPi} via
-\texttt{pip\ install\ maze-dataset}, or directly from the
-\href{https://github.com/understanding-search/maze-dataset}{git
-repository} (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
-Ivanitskiy et al., 2023a}).
-
-To create a dataset, we first create a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-configuration object, which specifies the seed, number, and size of
-mazes, as well as the generation algorithm and its corresponding
-parameters. This object is passed to a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-class to create a dataset. Crucially, this
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-mimics the interface of a PyTorch
-(\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
-\href{https://pytorch.org/docs/stable/data.html}{\texttt{Dataset}}, and
-can thus be easily incorporated into existing data pre-processing and
-training pipelines, e.g., through the use of a \texttt{DataLoader}
-class.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
-\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
-\NormalTok{)}
-\CommentTok{\# create a config}
-\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
-\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
-\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
-\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
-\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
-    \CommentTok{\# (optional) algorithm{-}specific parameters}
-\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
-    \CommentTok{\# (optional) many options for restricting start/end points}
-\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
-\NormalTok{)}
-\CommentTok{\# create a dataset}
-\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
-\NormalTok{  cfg, }\CommentTok{\# pass the config}
-\NormalTok{  ..., }\CommentTok{\# other options for disk loading, parallelization, etc.}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-When initializing a dataset, options which do not affect the mazes
-themselves can be specified through the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
-factory method as necessary. These options allow for saving/loading
-existing datasets instead of re-generating, parallelization options for
-generation, and more. Available maze generation algorithms are static
-methods of the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
-namespace class and include generation algorithms based on randomized
-depth-first search, Wilson's algorithm
-(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
-(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
-\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
-Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
-others.
-
-Furthermore, a dataset of mazes can be filtered to satisfy certain
-properties. Custom filters can be specified, and some filters are
-included in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
-For example, we can require a minimum path length of three steps from
-the origin to the target:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dataset\_filtered: MazeDataset }\OperatorTok{=}\NormalTok{ dataset.filter\_by.path\_length(min\_length}\OperatorTok{=}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-All implemented maze generation algorithms are stochastic by nature. For
-reproducibility, the \texttt{seed} parameter of
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-may be set. In practice, using provided deduplication filters, we find
-that exact duplicate mazes are generated very infrequently, even when
-generating very large datasets.
-
-For use cases where mazes of different sizes, generation algorithms, or
-other parameter variations are required, we provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/collected_dataset.html\#MazeDatasetCollection}{\texttt{MazeDatasetCollection}}
-class, which allows for creating a single iterable dataset from multiple
-independent configurations.
-
-\hypertarget{visual-output-formats}{%
-\subsection{Visual Output Formats}\label{visual-output-formats}}
-
-Internally, mazes are
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
-objects, which have path information and a tensor optimized for storing
-sub-graphs of a lattice. These objects can be converted to and from
-several formats to maximize their utility in different contexts.
-
-\begin{figure}[H]
-  \centering
-  \begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
-    \hline \\[.5em]
-    % algorithms
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
-    % descriptions
-      Simple text format for displaying mazes, useful for debugging in a terminal environment.
-      & \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
-      & feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
-    \hline \\
-    % examples
-      \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
-        \setlength{\baselineskip}{0.9em}
-        \input{figures/outputs-ascii-colored.tex} 
-      \end{minipage}}
-      & \multicolumn{1}{c}{
-        \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
-      }
-      & \multicolumn{1}{c}{
-        \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
-      } \\[1em]
-
-    \hline \\
-  \end{tabular}
-  \caption{Various output formats. Top row (left to right): ASCII diagram, rasterized pixel grid, and advanced display tool.}
-  \label{fig:output-fmts}
-\end{figure}
-
-In previous work, maze tasks have been used with Recurrent Convolutional
-Neural Network (RCNN) derived architectures
-(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
-Huang, et al., 2021}). To facilitate the use of our package in this
-context, we replicate the format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) and provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
-class which returns rasterized pairs of (input, target) mazes as shown
-in \autoref{fig:e2h-raster} below.
-
-\begin{figure}
-\hypertarget{fig:e2h-raster}{%
-\centering
-\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
-\caption{Input is the rasterized maze without the path marked (left),
-and provide as a target the maze with all but the correct path removed
-(right). Configuration options exist to adjust whether endpoints are
-included and if empty cells should be filled in.}\label{fig:e2h-raster}
-}
-\end{figure}
-
-\hypertarget{sec:tokenized-output-formats}{%
-\subsection{Tokenized Output
-Formats}\label{sec:tokenized-output-formats}}
-
-Autoregressive transformer models can be quite sensitive to the exact
-format of input data, and may even use delimiter tokens to perform
-reasoning steps (\protect\hyperlink{ref-pfau2024dotbydot}{Pfau et al.,
-2024}; \protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al.,
-2024}). To facilitate systematic investigation of the effects of
-different representations of data on text model performance, we provide
-a variety of tokenized text output formats.
-
-We convert mazes to token sequences in two steps. First, the maze is
-stringified using
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens()}}.
-The
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class provides a powerful interface for configuring maze stringification
-behavior. Second, the sequence of strings is tokenized into integers
-using \texttt{encode()}. Tokenization uses a fixed vocabulary for
-simplicity. Mazes up to \(50 \times 50\) are supported when using a
-unique token for each position, and up to \(128 \times 128\) are
-supported when positions in the maze are represented as a pair of
-coordinates.
-
-There are many algorithms by which one might tokenize a 2D maze into a
-1D format usable by autoregressive text models. Training multiple models
-on the encodings output from each of these algorithms may produce very
-different internal representations, learned solution algorithms, and
-levels of performance. To allow exploration of how different maze
-tokenization algorithms affect these models, the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class contains a rich set of options to customize how mazes are
-stringified. This class contains 19 discrete parameters, resulting in
-over 5.8 million unique tokenizers. There are 6 additional parameters
-available whose functionality is not verified via automated testing, but
-further expand the the number of tokenizers by a factor of \(44/3\) to
-86 million.
-
-All output sequences consist of four token regions representing
-different features of the maze; an example output sequence is shown in
-\autoref{fig:token-regions}.
-
-\begin{figure} 
-  \centering
-  \begin{minipage}{5in}
-    \footnotesize
-    \input{figures/outputs-tokens-colored.tex}
-  \end{minipage}
-  \caption{
-    Example text output format with token regions highlighted.
-    \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the graph,
-    \colorbox[RGB]{ 217,234,211 }{Origin}: starting coordinate,
-    \colorbox[RGB]{ 234,209,220 }{Target}: ending coordinate,
-    \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence
-  }
-  \label{fig:token-regions}
-\end{figure}
-
-Each
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-is constructed from a set of several
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-objects, each of which specifies how different token regions or other
-elements of the stringification are produced.
-
-\begin{figure}
-    \centering
-    \input{figures/TokenizerElement_structure.tikz}
-    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular}.}
-\end{figure}
-
-The tokenizer architecture is purposefully designed such that adding and
-testing a wide variety of new tokenization algorithms is fast and
-minimizes disturbances to functioning code. This is enabled by the
-modular architecture and the automatic inclusion of any new tokenizers
-in integration tests. To create a new variety of tokenizer, developers
-forking the library may simply create their own
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and implement the abstract methods. If the behavior change is
-sufficiently small, simply adding a parameter to an existing
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and updating its implementation will suffice.
-
-The breadth of tokenizers is also easily scaled in the opposite
-direction. Due to the exponential scaling of parameter combinations,
-adding a small number of new features can significantly slow certain
-procedures which rely on constructing all possible tokenizers, such as
-integration tests. If any existing subclass contains features which
-aren't needed, a developer tool decorator
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
-is provided which can be applied to the unneeded
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclasses to prune those features and compact the available space of
-tokenizers.
-
-\hypertarget{benchmarks}{%
-\subsection{Benchmarks of Generation Speed}\label{benchmarks}}
-
-We provide approximate benchmarks for relative generation time across
-various algorithms, parameter choices, maze sizes, and dataset sizes in
-\autoref{tab:benchmarks} and \autoref{fig:benchmarks}. Experiments were
-performed on a
-\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners#standard-github-hosted-runners-for-public-repositories}{standard GitHub runner}
-without parallelism.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|ll|r|rrr|}
-  \hline
-  maze\_ctor
-          & keyword args           & all sizes 
-                                              & \shortstack{small \\ $g \leq 10$} 
-                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
-                                                                    & \shortstack{large \\ $g > 32$} \\
-  \hline\hline
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
-          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
-          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
-          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
-          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
-  \hline\hline
-  mean
-          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
-  median
-          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
-  \hline
-\end{tabular}
-\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
-\label{tab:benchmarks}
-\end{table}
-
-\begin{figure}
-\hypertarget{fig:benchmarks}{%
-\centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
-\caption{Plot of maze generation time. Generation time scales
-exponentially with maze size for all algorithms. Generation time per
-maze does not depend on the number of mazes being generated, and there
-is minimal overhead to initializing the generation process for a small
-dataset. Wilson's algorithm is notably less efficient than others and
-has high variance. Note that values are averaged across all parameter
-sets for that algorithm. More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}\label{fig:benchmarks}
-}
-\end{figure}
-
-\hypertarget{sec:success-rate-estimation}{%
-\subsection{Success Rate Estimation}\label{sec:success-rate-estimation}}
-
-In order to replicate the exact dataset distribution of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}), the parameter
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}}
-allows for additional constraints such as enforcing that the start or
-end point be in a ``dead end'' with only one accessible neighbor cell.
-However, combining these constraints with cyclic mazes (such as those
-generated with percolation), as was required for the work in
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), can lead to an absence of valid start and end points. Placing
-theoretical bounds on this success rate is difficult, as it depends on
-the exact maze generation algorithm and parameters used. To deal with
-this, our package provides a way to estimate the success rate of a given
-configuration using a symbolic regression model trained with PySR
-(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
-be found in
-\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
-Using the estimation algorithm simply requires the user to call
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
-providing their initial \texttt{cfg} and then using the returned
-\texttt{cfg\_new} in its place.
-
-\hypertarget{success-rate-estimation-algorithm}{%
-\subsubsection{Success Rate Estimation
-Algorithm}\label{success-rate-estimation-algorithm}}
-
-The base function learned by symbolic regression provides limited
-insight and may be subject to change. It is defined as
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
-and takes a 5 dimensional float vector created by
-\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the
-{[}percolation value, grid size, endpoint deadend configuration,
-endpoint uniqueness, categorical generation function index{]}.
-
-However, the outputs of this function are not directly usable due to
-minor divergences at the endpoints with respect to the percolation
-probability \(p\). Since we know that maze success is either guaranteed
-or impossible for \(p=0\) and \(p=1\), we define the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function to nudge the raw output of the symbolic regression. This
-function is defined with the following components:
-
-shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
-function given by \[
-  \sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}
-  \qquad A(q,a,w) = w \cdot (1 - |2q-1|^a)
-\] \[
-  h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)
-\]
-
-We combine these to get the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function, which is identity-like for \(p \approx 0.5\), and pushes \(x\)
-to extremes otherwise. \[
-  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
-\]
-
-Finally, we define \[
-  \text{cfg\_success\_predict\_fn}(\mathbf{x}) = \text{soft\_step}(\text{raw\_val}, x_0, 5, 10)
-\]
-
-where \texttt{raw\_val} is the output of the symbolic regression model.
-The parameter \(x_0\) is the percolation probability, while all other
-parameters from \texttt{\_to\_ps\_array()} only affect
-\texttt{raw\_val}.
-
-\begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
-\caption{An example of both empirical and predicted success rates as a
-function of the percolation probability \(p\) for various maze sizes,
-percolation with and without depth first search, and
-\texttt{endpoint\_kwargs} requiring that both the start and end be in
-unique dead ends. Empirical measures derived from a sample of 128 mazes.
-More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}
-\end{figure}
-
-\hypertarget{sec:implementation}{%
-\section{Implementation}\label{sec:implementation}}
-
-We refer to our
-\href{https://github.com/understanding-search/maze-dataset}{repository}
-and
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
-for documentation and up-to-date implementation details.
-
-This package utilizes a simple, efficient representation of mazes as
-subgraphs of a finite lattice, which we call a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-Using an adjacency matrix for storing mazes would be memory inefficient
-by failing to exploit the highly sparse structure -- for example, for a
-2-dimensional maze, only 4 of the diagonals would be filled in. On the
-other hand, using an adjacency list could lead to a poor lookup time for
-whether any given connection exists.
-
-Instead, we describe mazes with the following representation: for a
-\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
-initialize a boolean array \[
-  A = \{0, 1\}^{2 \times r \times c}
-\] which we refer to in the code as a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
-The value at \(A[0,i,j]\) determines whether a \emph{downward}
-connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
-value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
-to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
-existence of connections and facilitate fast lookup time, at the cost of
-requiring additional care with indexing. Note that this setup allows for
-a periodic lattice. Generation of mazes is detailed in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
-
-To produce solutions to mazes, two points are selected uniformly at
-random without replacement from the connected component of the maze, and
-the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
-1968}) is applied to find the shortest path between them. The endpoint
-selection can be controlled via
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
-and complications caused by this are detailed in
-\hyperref[sec:success-rate-estimation]{section: \textit{\nameref{sec:success-rate-estimation}}}.
-A maze with a solution is denoted a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}},
-which inherits from
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-
-Parallelization is implemented via the \texttt{multiprocessing} module
-in the Python standard library, and parallel generation can be
-controlled via keyword arguments to
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}.
-
-\newpage
-
-\hypertarget{usage-in-research}{%
-\section{Usage in Research}\label{usage-in-research}}
-
-This package was originally built for the needs of the
-(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
-et al., 2023b}) project, which aims to investigate spatial planning and
-world models in autoregressive transformer models trained on mazes
-(\protect\hyperlink{ref-ivanitskiy2023structuredworldreps}{Michael
-Igorevich Ivanitskiy, Spies, et al., 2023};
-\protect\hyperlink{ref-maze-dataset-arxiv-2023}{Michael Igorevich
-Ivanitskiy, Shah, et al., 2023};
-\protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al., 2024}).
-It was extended for work on understanding the mechanisms by which
-recurrent convolutional and implicit networks
-(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
-given a rasterized view
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), which required matching the pixel-padded and endpoint
-constrained output format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). Ongoing work using \texttt{maze-dataset} aims to
-investigate the effects of varying the tokenization format on the
-performance of pretrained LLMs on spatial reasoning.
-
-This package has also been utilized in work by other groups:
-
-\begin{itemize}
-\item
-  By (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) to
-  compare the effectiveness of transformers trained with the
-  MLM-\(\mathcal{U}\)
-  (\protect\hyperlink{ref-MLMU-kitouni2024factorization}{Kitouni et al.,
-  2024}) multistep prediction objective against standard autoregressive
-  training for multi-step planning on our maze task.
-\item
-  By (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) to
-  study the effectiveness of imperative learning.
-\item
-  By (\protect\hyperlink{ref-zhang2025tscend}{Zhang et al., 2025}) to
-  introduce a novel framework for reasoning diffusion models.
-\item
-  By (\protect\hyperlink{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to
-  improve spatial reasoning in LLMs with GRPO.
-\end{itemize}
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-This work was partially funded by National Science Foundation awards
-DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
-for hosting authors MII, AFS, and TR for a residency visit, and to
-various members of FAR's technical staff for their advice.
-
-This work was partially supported by AI Safety Camp and AI Safety
-Support, which also brought many of the authors together. We would like
-to thank our former collaborators at AI Safety Camp and other users and
-contributors to the \texttt{maze-dataset} package: Benji Berczi,
-Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris Mathwin, Lucia
-Quirke, Can Rager, Adrians Skapars, Rusheb Shah, Johannes Treutlein, and
-Dan Valentine.
-
-We thank the Mines Optimization and Deep Learning group (MODL) for
-fruitful discussions. We also thank Michael Rosenberg for recommending
-the usage of Finite State Transducers for storing tokenizer validation
-information.
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0.5}
-\leavevmode\vadjust pre{\hypertarget{ref-mazegenerator-net}{}}%
-Alance AB. (2019). \emph{Maze generator}.
-\url{http://www.mazegenerator.net}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
-Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
-1.0: A complete set of tools to prepare, present, and analyze
-navigational and spatial cognitive neuroscience experiments.
-\emph{Behavior Research Methods}, \emph{40}, 353--359.
-
-\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
-Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
-{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
-Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
-procedural generation to benchmark reinforcement learning. \emph{arXiv
-Preprint arXiv:1912.01588}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pysr}{}}%
-Cranmer, M. (2023). Interpretable machine learning for science with PySR
-and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
-Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
-models' spatial intelligence via GRPO. \emph{arXiv Preprint
-arXiv:2502.14669}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation}{}}%
-Duminil-Copin, H. (2017). \emph{Sixty years of percolation} (No.
-arXiv:1712.04651). {arXiv}. \url{http://arxiv.org/abs/1712.04651}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Ehsan_2022}{}}%
-Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation-clustersize}{}}%
-Fisher, M. E., \& Essam, J. W. (2004). Some {Cluster Size} and
-{Percolation Problems}. \emph{Journal of Mathematical Physics},
-\emph{2}(4), 609--619. \url{https://doi.org/10.1063/1.1703745}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
-Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
-(2022). Jfb: Jacobian-free backpropagation for implicit networks.
-\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
-\emph{36}, 6648--6656.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
-Guo, C., Barthelet, L., \& Morris, R. (2011). \emph{Maze generator and
-solver}. Wolfram Demonstrations Project,
-\url{https://demonstrations.wolfram.com/MazeGeneratorAndSolver/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
-Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
-{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
-Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
-{Games CoG}}, 1--4.
-
-\leavevmode\vadjust pre{\hypertarget{ref-A_star}{}}%
-Hart, P. E., Nilsson, N. J., \& Raphael, B. (1968). A {Formal Basis} for
-the {Heuristic Determination} of {Minimum Cost Paths}. \emph{IEEE
-Transactions on Systems Science and Cybernetics}, \emph{4}(2), 100--107.
-\url{https://doi.org/10.1109/TSSC.1968.300136}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
-Ivanitskiy, M. (n.d.). \emph{ZANJ}.
-\url{https://github.com/mivanit/ZANJ}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
-\emph{Maze dataset}.
-\url{https://github.com/understanding-search/maze-dataset}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
-\emph{Maze transformer interpretability}.
-\url{https://github.com/understanding-search/maze-transformer}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
-Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
-Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
-C. D., \& others. (2023). A configurable library for generating and
-manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
-Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
-Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
-D., \& others. (2023). Structured world representations in maze-solving
-transformers. \emph{arXiv Preprint arXiv:2312.02566}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
-Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
-Ibrahim, M. (2024). The factorization curse: Which tokens you predict
-underlie the reversal curse and more. \emph{Advances in Neural
-Information Processing Systems}, \emph{37}, 112329--112355.
-
-\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
-Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
-Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
-extrapolation for mazes with recurrent and implicit networks.
-\emph{arXiv Preprint arXiv:2410.03020}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruskal1956shortest}{}}%
-Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
-the traveling salesman problem. \emph{Proceedings of the American
-Mathematical Society}, \emph{7}(1), 48--50.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
-Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
-Performance insights and comparative analysis. \emph{arXiv Preprint
-arXiv:2308.11224}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
-Nag, A. (2020). MDL suite: A language, generator and compiler for
-describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
-1815.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Nemeth_2019}{}}%
-Németh, F. (2019). \emph{Maze-generation-algorithms}.
-\url{https://github.com/ferenc-nemeth/maze-generation-algorithms}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
-Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
-Transformers can navigate mazes with multi-step prediction. \emph{arXiv
-Preprint arXiv:2412.05117}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh-oppenheimj2018maze}{}}%
-Oppenheim, J. (2018). \emph{Maze-generator: Generate a random maze
-represented as a 2D array using depth-first search}.
-\url{https://github.com/oppenheimj/maze-generator/}; GitHub.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
-Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G.,
-Killeen, T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., Kopf,
-A., Yang, E., DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S.,
-Steiner, B., Fang, L., \ldots{} Chintala, S. (2019). {PyTorch: An
-Imperative Style, High-Performance Deep Learning Library}. In H.
-Wallach, H. Larochelle, A. Beygelzimer, F. d'Alché-Buc, E. Fox, \& R.
-Garnett (Eds.), \emph{Advances in neural information processing systems
-32} (pp. 8024--8035). Curran Associates, Inc.
-\url{http://papers.neurips.cc/paper/9015-pytorch-an-imperative-style-high-performance-deep-learning-library.pdf}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
-Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
-Hidden computation in transformer language models. \emph{arXiv Preprint
-arXiv:2404.15758}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
-Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
-transparent ai: A survey on interpreting the inner structures of deep
-neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
-Machine Learning (SaTML)}, 464--483.
-
-\leavevmode\vadjust pre{\hypertarget{ref-easy_to_hard}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Bansal, A., Emam, Z., Huang,
-F., Goldblum, M., \& Goldstein, T. (2021). \emph{Datasets for {Studying
-Generalization} from {Easy} to {Hard Examples}} (No. arXiv:2108.06011).
-{arXiv}. \url{https://doi.org/10.48550/arXiv.2108.06011}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
-Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
-Generalizing from easy to hard problems with recurrent networks.
-\emph{Advances in Neural Information Processing Systems}, \emph{34},
-6695--6706.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
-Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
-\emph{arXiv Preprint arXiv:2308.02522}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
-Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
-Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
-world models in maze-solving tasks. \emph{arXiv Preprint
-arXiv:2412.11867}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
-Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
-Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
-self-supervised neural-symbolic learning framework for robot autonomy.
-\emph{arXiv Preprint arXiv:2406.16087}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wilson}{}}%
-Wilson, D. B. (1996). Generating random spanning trees more quickly than
-the cover time. \emph{Proceedings of the Twenty-Eighth Annual {ACM}
-Symposium on {Theory} of Computing - {STOC} '96}, 296--303.
-\url{https://doi.org/10.1145/237814.237880}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
-Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
-scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
-arXiv:2502.01989}.
-
-\end{CSLReferences}
-
-\end{document}

### 2025-04-26 docs/paper/figures/tex/fig2_formats.tex
move figure code to external files

--- 2025-04-09 docs/paper/paper.tex
@@
-\documentclass[10pt,a4paper,onecolumn]{article}
-\usepackage{marginnote}
-\usepackage{graphicx}
-\usepackage[rgb,svgnames]{xcolor}
-\usepackage{authblk,etoolbox}
-\usepackage{titlesec}
-\usepackage{calc}
-\usepackage{tikz}
-\usepackage[pdfa]{hyperref}
-\usepackage{hyperxmp}
-\hypersetup{%
-    unicode=true,
-    pdfapart=3,
-    pdfaconformance=B,
-    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility},
-    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
-Fung},
-    pdfpublication={Journal of Open Source Software},
-    pdfpublisher={Open Journals},
-    pdfissn={2475-9066},
-    pdfpubtype={journal},
-    pdfvolumenum={},
-    pdfissuenum={},
-    pdfdoi={N/A},
-    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
-Diniz Behn, Samy Wu Fung},
-    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
-    colorlinks=true,
-    linkcolor=[rgb]{0.0, 0.5, 1.0},
-    citecolor=Blue,
-    urlcolor=[rgb]{0.0, 0.5, 1.0},
-    breaklinks=true
-}
-% https://tex.stackexchange.com/a/535849
-% Create an OutputIntent in order to correctly specify colours
-\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
-\pdfcatalog{%
-  /OutputIntents [
-    <<
-      /Type /OutputIntent
-      /S /GTS_PDFA1
-      /DestOutputProfile \the\pdflastobj\space 0 R
-      /OutputConditionIdentifier (sRGB)
-      /Info (sRGB)
-    >>
-  ]
-}
-\pdfvariable omitcidset=1
-\usepackage{caption}
-\usepackage{orcidlink}
-\usepackage{tcolorbox}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{seqsplit}
-\usepackage{xstring}
-
-\usepackage{float}
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
-
-\usepackage{fixltx2e} % provides \textsubscript
-
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-% --- Page layout -------------------------------------------------------------
-\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
-            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}
-
-% --- Default font ------------------------------------------------------------
-\renewcommand\familydefault{\sfdefault}
-
-% --- Style -------------------------------------------------------------------
-\renewcommand{\captionfont}{\small\sffamily}
-\renewcommand{\captionlabelfont}{\bfseries}
-
-% --- Section/SubSection/SubSubSection ----------------------------------------
-\titleformat{\section}
-  {\normalfont\sffamily\Large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsection}
-  {\normalfont\sffamily\large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsubsection}
-  {\normalfont\sffamily\bfseries}
-  {}{0pt}{}
-\titleformat*{\paragraph}
-  {\sffamily\normalsize}
-
-% --- Header / Footer ---------------------------------------------------------
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-%\renewcommand{\headrulewidth}{0.50pt}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{joss/logo.png}}
-\fancyhead[C]{}
-\fancyhead[R]{}
-\renewcommand{\footrulewidth}{0.25pt}
-
-\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset: Maze Generation with Algorithmic Variety
-and Representational Flexibility. \emph{Journal of Open Source
-Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
-
-\fancyfoot[R]{\sffamily \thepage}
-\makeatletter
-\let\ps@plain\ps@fancy
-\fancyheadoffset[L]{4.5cm}
-\fancyfootoffset[L]{4.5cm}
-
-% --- Macros ---------
-
-\definecolor{linky}{rgb}{0.0, 0.5, 1.0}
-
-\newtcolorbox{repobox}
-   {colback=red, colframe=red!75!black,
-     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}
-
-\newcommand{\ExternalLink}{%
-   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
-       \begin{scope}[x=1ex, y=1ex]
-           \clip (-0.1,-0.1)
-               --++ (-0, 1.2)
-               --++ (0.6, 0)
-               --++ (0, -0.6)
-               --++ (0.6, 0)
-               --++ (0, -1);
-           \path[draw,
-               line width = 0.5,
-               rounded corners=0.5]
-               (0,0) rectangle (1,1);
-       \end{scope}
-       \path[draw, line width = 0.5] (0.5, 0.5)
-           -- (1, 1);
-       \path[draw, line width = 0.5] (0.6, 1)
-           -- (1, 1) -- (1, 0.6);
-       }
-   }
-
-\definecolor{c53baa1}{RGB}{83,186,161}
-\definecolor{c202826}{RGB}{32,40,38}
-\def \rorglobalscale {0.1}
-\newcommand{\rorlogo}{%
-\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
-  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
-    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364)..
-controls (3.9643, 1.4576) and (4.0508, 1.5081) ..
-(4.1098, 1.5878)..
-controls (4.169, 1.6674) and (4.1984, 1.7642) ..
-(4.1984, 1.8777)..
-controls (4.1984, 1.9719) and (4.182, 2.0503) ..
-(4.1495, 2.1132)..
-controls (4.1169, 2.1762) and (4.0727, 2.2262) ..
-(4.0174, 2.2635)..
-controls (3.9621, 2.3006) and (3.8976, 2.3273) ..
-(3.824, 2.3432)..
-controls (3.7505, 2.359) and (3.6727, 2.367) ..
-(3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688)..
-controls (2.9625, 1.8833) and (2.9572, 1.8976) ..
-(2.9514, 1.9119)..
-controls (2.9083, 2.0164) and (2.848, 2.1056) ..
-(2.7705, 2.1791)..
-controls (2.6929, 2.2527) and (2.6014, 2.3093) ..
-(2.495, 2.3487)..
-controls (2.3889, 2.3881) and (2.2728, 2.408) ..
-(2.1468, 2.408)..
-controls (2.0209, 2.408) and (1.905, 2.3881) ..
-(1.7986, 2.3487)..
-controls (1.6925, 2.3093) and (1.6007, 2.2527) ..
-(1.5232, 2.1791)..
-controls (1.4539, 2.1132) and (1.3983, 2.0346) ..
-(1.3565, 1.9436)..
-controls (1.3504, 2.009) and (1.3351, 2.0656) ..
-(1.3105, 2.1132)..
-controls (1.2779, 2.1762) and (1.2338, 2.2262) ..
-(1.1785, 2.2635)..
-controls (1.1232, 2.3006) and (1.0586, 2.3273) ..
-(0.985, 2.3432)..
-controls (0.9115, 2.359) and (0.8337, 2.367) ..
-(0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364)..
-controls (1.1255, 1.4576) and (1.2121, 1.5081) ..
-(1.2711, 1.5878)..
-controls (1.2737, 1.5915) and (1.2761, 1.5954) ..
-(1.2787, 1.5991)..
-controls (1.2782, 1.5867) and (1.2779, 1.5743) ..
-(1.2779, 1.5616)..
-controls (1.2779, 1.4327) and (1.2996, 1.3158) ..
-(1.3428, 1.2113)..
-controls (1.3859, 1.1068) and (1.4462, 1.0176) ..
-(1.5237, 0.944)..
-controls (1.601, 0.8705) and (1.6928, 0.8139) ..
-(1.7992, 0.7744)..
-controls (1.9053, 0.735) and (2.0214, 0.7152) ..
-(2.1474, 0.7152)..
-controls (2.2733, 0.7152) and (2.3892, 0.735) ..
-(2.4956, 0.7744)..
-controls (2.6016, 0.8139) and (2.6935, 0.8705) ..
-(2.771, 0.944)..
-controls (2.8482, 1.0176) and (2.9086, 1.1068) ..
-(2.952, 1.2113)..
-controls (2.9578, 1.2253) and (2.9631, 1.2398) ..
-(2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777)..
-controls (0.9438, 1.7534) and (0.92, 1.7357) ..
-(0.8911, 1.7243)..
-controls (0.8623, 1.7129) and (0.83, 1.706) ..
-(0.7945, 1.7039)..
-controls (0.7588, 1.7015) and (0.7252, 1.7005) ..
-(0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667)..
-controls (0.7477, 2.0667) and (0.7805, 2.0643) ..
-(0.8139, 2.0598)..
-controls (0.8472, 2.0553) and (0.8768, 2.0466) ..
-(0.9025, 2.0336)..
-controls (0.9282, 2.0206) and (0.9496, 2.0021) ..
-(0.9663, 1.9778)..
-controls (0.9829, 1.9534) and (0.9914, 1.9209) ..
-(0.9914, 1.8799)..
-controls (0.9914, 1.8362) and (0.9819, 1.8021) ..
-(0.9628, 1.7777) -- cycle(2.6125, 1.3533)..
-controls (2.5889, 1.2904) and (2.5553, 1.2359) ..
-(2.5112, 1.1896)..
-controls (2.4672, 1.1433) and (2.4146, 1.1073) ..
-(2.3529, 1.0814)..
-controls (2.2916, 1.0554) and (2.2228, 1.0427) ..
-(2.1471, 1.0427)..
-controls (2.0712, 1.0427) and (2.0026, 1.0557) ..
-(1.9412, 1.0814)..
-controls (1.8799, 1.107) and (1.8272, 1.1433) ..
-(1.783, 1.1896)..
-controls (1.7391, 1.2359) and (1.7052, 1.2904) ..
-(1.6817, 1.3533)..
-controls (1.6581, 1.4163) and (1.6465, 1.4856) ..
-(1.6465, 1.5616)..
-controls (1.6465, 1.6359) and (1.6581, 1.705) ..
-(1.6817, 1.7687)..
-controls (1.7052, 1.8325) and (1.7388, 1.8873) ..
-(1.783, 1.9336)..
-controls (1.8269, 1.9799) and (1.8796, 2.0159) ..
-(1.9412, 2.0418)..
-controls (2.0026, 2.0675) and (2.0712, 2.0804) ..
-(2.1471, 2.0804)..
-controls (2.223, 2.0804) and (2.2916, 2.0675) ..
-(2.3529, 2.0418)..
-controls (2.4143, 2.0161) and (2.467, 1.9799) ..
-(2.5112, 1.9336)..
-controls (2.5551, 1.8873) and (2.5889, 1.8322) ..
-(2.6125, 1.7687)..
-controls (2.636, 1.705) and (2.6477, 1.6359) ..
-(2.6477, 1.5616)..
-controls (2.6477, 1.4856) and (2.636, 1.4163) ..
-(2.6125, 1.3533) -- cycle(3.8015, 1.7777)..
-controls (3.7825, 1.7534) and (3.7587, 1.7357) ..
-(3.7298, 1.7243)..
-controls (3.701, 1.7129) and (3.6687, 1.706) ..
-(3.6333, 1.7039)..
-controls (3.5975, 1.7015) and (3.5639, 1.7005) ..
-(3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667)..
-controls (3.5864, 2.0667) and (3.6192, 2.0643) ..
-(3.6526, 2.0598)..
-controls (3.6859, 2.0553) and (3.7155, 2.0466) ..
-(3.7412, 2.0336)..
-controls (3.7669, 2.0206) and (3.7883, 2.0021) ..
-(3.805, 1.9778)..
-controls (3.8216, 1.9534) and (3.8301, 1.9209) ..
-(3.8301, 1.8799)..
-controls (3.8301, 1.8362) and (3.8206, 1.8021) ..
-(3.8015, 1.7777) -- cycle;
-  \end{scope}
-\end{tikzpicture}
-}
-
-% --- Title / Authors ---------------------------------------------------------
-% patch \maketitle so that it doesn't center
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-% patch \maketitle so that the font size for the title is normal
-\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
-% patch the patch by authblk so that the author block is flush left
-\def\maketitle{{%
-  \renewenvironment{tabular}[2][]
-    {\begin{flushleft}}
-    {\end{flushleft}}
-  \AB@maketitle}}
-\renewcommand\AB@affilsepx{ \protect\Affilfont}
-%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
-\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
-\renewcommand{\affil}[2][]%
-   {\newaffiltrue\let\AB@blk@and\AB@pand
-      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
-        \setcounter{Maxaffil}{0}\fi
-        \begingroup
-        \let\href=\href@Orig
-        \let\protect\@unexpandable@protect
-        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
-        \@temptokena=\expandafter{\AB@authors}%
-        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
-         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
-         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
-         \gdef\AB@las{}\gdef\AB@au@str{}%
-        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
-        \@temptokena=\expandafter{\AB@affillist}%
-        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
-          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
-      \endgroup
-       \let\AB@affilsep\AB@affilsepx
-}
-\makeatother
-\renewcommand\Authfont{\sffamily\bfseries}
-\renewcommand\Affilfont{\sffamily\small\mdseries}
-\setlength{\affilsep}{1em}
-
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{fontspec}
-
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\sffamily]{Ligatures=TeX}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-
-%% Font settings
-\usepackage{fontsetup} % Lazy way to get proper Greek lowercase glyphs
-
-% Use Hack https://sourcefoundry.org/hack/
-\setmonofont{Hack}
-
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\urlstyle{same}  % don't use monospace font for urls
-\ifLuaTeX
-\usepackage[bidi=basic]{babel}
-\else
-\usepackage[bidi=default]{babel}
-\fi
-\babelprovide[main,import]{american}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\tikzset{ % Define a TikZ style for an external hyperlink node.
-  hyperlink node url/.style={
-    alias=sourcenode,
-    append after command={
-      let \p1 = (sourcenode.north west),
-          \p2 = (sourcenode.south east),
-          \n1 = {\x2-\x1},
-          \n2 = {\y1-\y2} in
-      node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
-          {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
-    }
-  }
-}
-\providecommand{\XeTeXLinkBox}[1]{#1}
-\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
-\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
-\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-
-\title{maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility}
-
-\author[1%
-%
-\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
-  \,\orcidlink{0000-0002-4213-4993}\,%
-}
-\author[4%
-%
-]{Aaron Sandoval%
-  \,\orcidlink{0009-0002-8380-6140}\,%
-}
-\author[2%
-%
-]{Alex F. Spies%
-  \,\orcidlink{0000-0002-8708-1530}\,%
-}
-\author[3%
-%
-]{Tilman Räuker%
-  \,\orcidlink{0009-0009-6321-4413}\,%
-}
-\author[1%
-%
-]{Brandon Knutson%
-  \,\orcidlink{0009-0004-8413-0239}\,%
-}
-\author[1%
-%
-]{Cecilia Diniz Behn%
-  \,\orcidlink{0000-0002-8078-5105}\,%
-}
-\author[1%
-%
-]{Samy Wu Fung%
-  \,\orcidlink{0000-0002-2926-4582}\,%
-}
-
-\affil[1]{Colorado School of Mines, Department of Applied Mathematics
-and Statistics%
-}
-\affil[2]{Imperial College London%
-}
-\affil[3]{UnSearch.org%
-}
-\affil[4]{Independent%
-}
-\affil[$\mathparagraph$]{Corresponding author}
-\date{\vspace{-2.5ex}}
-
-\begin{document}
-\maketitle
-
-\marginpar{
-
-  \begin{flushleft}
-  %\hrule
-  \sffamily\small
-
-  {\bfseries DOI:} \href{https://doi.org/N/A}{\color{linky}{N/A}}
-
-  \vspace{2mm}
-    {\bfseries Software}
-  \begin{itemize}
-    \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{\color{linky}{Review}} \ExternalLink
-    \item \href{https://github.com/openjournals}{\color{linky}{Repository}} \ExternalLink
-    \item \href{https://doi.org/10.5281}{\color{linky}{Archive}} \ExternalLink
-  \end{itemize}
-
-  \vspace{2mm}
-
-    \par\noindent\hrulefill\par
-
-  \vspace{2mm}
-
-  {\bfseries Editor:} \href{https://joss.theoj.org}{Open
-Journals} \ExternalLink
-   \\
-  \vspace{1mm}
-    {\bfseries Reviewers:}
-  \begin{itemize}
-  \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{@openjournals}
-    \end{itemize}
-    \vspace{2mm}
-
-    {\bfseries Submitted:} 01 January 1970\\
-    {\bfseries Published:} 01 January 1970
-
-  \vspace{2mm}
-  {\bfseries License}\\
-  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{https://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).
-
-  \end{flushleft}
-}
-
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
-
-Solving mazes is a classic problem in computer science and artificial
-intelligence, and humans have been constructing mazes for thousands of
-years. Although finding the shortest path through a maze is a solved
-problem, this very fact makes it an excellent testbed for studying how
-machine learning algorithms solve problems and represent spatial
-information. We introduce \texttt{maze-dataset}, a user-friendly Python
-library for generating, processing, and visualizing datasets of mazes.
-This library supports a variety of maze generation algorithms providing
-mazes with or without loops, mazes that are connected or not, and many
-other variations. These generation algorithms can be configured with
-various parameters, and the resulting mazes can be filtered to satisfy
-desired properties. Also provided are tools for converting mazes to and
-from various formats suitable for a variety of neural network
-architectures, such as rasterized images, tokenized text sequences, and
-various visualizations. As well as providing a simple interface for
-generating, storing, and loading these datasets, \texttt{maze-dataset}
-is extensively tested, type hinted, benchmarked, and documented.
-
-\begin{figure} 
-  \begin{minipage}{5in}
-    \input{diagram/diagram.tikz} 
-  \end{minipage}
-  \caption{
-    Usage of maze-dataset.
-We create a \texttt{MazeDataset} from a \texttt{MazeDatasetConfig}.
-This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats.
-Code in the image contains clickable links to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{documentation}.
-A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.
-  }
-  \label{fig:diagram}
-\end{figure}
-
-\hypertarget{statement-of-need}{%
-\section{Statement of Need}\label{statement-of-need}}
-
-While maze generation itself is straightforward, the architectural
-challenge comes from building a system supporting many algorithms with
-configurable parameters, property filtering, and representation
-transformation. This library aims to greatly streamline the process of
-generating and working with datasets of mazes that can be described as
-subgraphs of an \(n \times n\) lattice with boolean connections and,
-optionally, start and end points that are nodes in the graph.
-Furthermore, we place emphasis on a wide variety of possible text output
-formats aimed at evaluating the spatial reasoning capabilities of Large
-Language Models (LLMs) and other text-based transformer models.
-
-For interpretability and behavioral research, algorithmic tasks offer
-benefits by allowing systematic data generation and task decomposition,
-as well as simplifying the process of circuit discovery
-(\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
-Although mazes are well suited for these investigations, we found that
-existing maze generation packages
-(\protect\hyperlink{ref-cobbe2019procgen}{Cobbe et al., 2019};
-\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-harriesMazeExplorerCustomisable3D2019}{Harries et
-al., n.d.}; \protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) lack support for transforming between multiple
-representations and provide limited control over the maze generation
-process.
-
-\hypertarget{related-works}{%
-\subsection{Related Works}\label{related-works}}
-
-A multitude of public and open-source software packages exist for
-generating mazes (\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). However, nearly all of these packages produce
-mazes represented as rasterized images or other visual formats rather
-than the underlying graph structure, and this makes it difficult to work
-with these datasets.
-
-\begin{itemize}
-\item
-  Most prior works provide mazes in visual or raster formats, and we
-  provide a variety of similar output formats:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
-    utilizing
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
-    which can exactly mimic the outputs provided in
-    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
-    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
-    similar to the outputs of Németh
-    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    provides a format similar to
-    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
-    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
-    provides a feature-rich plotting utility with support for multiple
-    paths, heatmaps over positions, and more. This is similar to the
-    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
-    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
-    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
-  \end{itemize}
-\item
-  The text format provided by
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
-  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
-  Wu, 2023}), but provides over 5.8 million unique formats for
-  converting mazes to a text stream, detailed in
-  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
-\item
-  For rigorous investigations of the response of a model to various
-  distributional shifts, preserving metadata about the generation
-  algorithm with the dataset itself is essential. To this end, our
-  package efficiently stores the dataset along with its metadata in a
-  single human-readable file (\protect\hyperlink{ref-zanj}{M.
-  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
-  this reliably.
-\item
-  Storing mazes as images is not only difficult to work with, but also
-  inefficient. We use a highly efficient method detailed in
-  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
-\item
-  Our package is easily installable with source code freely available.
-  It is extensively tested, type hinted, benchmarked, and documented.
-  Many other maze generation packages lack this level of rigor and
-  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
-  2008}) appear to simply no longer be accessible.
-\end{itemize}
-
-\hypertarget{features}{%
-\section{Features}\label{features}}
-
-We direct readers to our
-\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples},
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs},
-and
-\href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
-for more information.
-
-\hypertarget{generation}{%
-\subsection{Generation and Basic Usage}\label{generation}}
-
-Our package can be installed from
-\href{https://pypi.org/project/maze-dataset/}{PyPi} via
-\texttt{pip\ install\ maze-dataset}, or directly from the
-\href{https://github.com/understanding-search/maze-dataset}{git
-repository} (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
-Ivanitskiy et al., 2023a}).
-
-To create a dataset, we first create a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-configuration object, which specifies the seed, number, and size of
-mazes, as well as the generation algorithm and its corresponding
-parameters. This object is passed to a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-class to create a dataset. Crucially, this
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-mimics the interface of a PyTorch
-(\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
-\href{https://pytorch.org/docs/stable/data.html}{\texttt{Dataset}}, and
-can thus be easily incorporated into existing data pre-processing and
-training pipelines, e.g., through the use of a \texttt{DataLoader}
-class.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
-\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
-\NormalTok{)}
-\CommentTok{\# create a config}
-\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
-\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
-\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
-\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
-\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
-    \CommentTok{\# (optional) algorithm{-}specific parameters}
-\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
-    \CommentTok{\# (optional) many options for restricting start/end points}
-\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
-\NormalTok{)}
-\CommentTok{\# create a dataset}
-\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
-\NormalTok{  cfg, }\CommentTok{\# pass the config}
-\NormalTok{  ..., }\CommentTok{\# other options for disk loading, parallelization, etc.}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-When initializing a dataset, options which do not affect the mazes
-themselves can be specified through the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
-factory method as necessary. These options allow for saving/loading
-existing datasets instead of re-generating, parallelization options for
-generation, and more. Available maze generation algorithms are static
-methods of the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
-namespace class and include generation algorithms based on randomized
-depth-first search, Wilson's algorithm
-(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
-(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
-\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
-Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
-others.
-
-Furthermore, a dataset of mazes can be filtered to satisfy certain
-properties. Custom filters can be specified, and some filters are
-included in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
-For example, we can require a minimum path length of three steps from
-the origin to the target:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dataset\_filtered: MazeDataset }\OperatorTok{=}\NormalTok{ dataset.filter\_by.path\_length(min\_length}\OperatorTok{=}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-All implemented maze generation algorithms are stochastic by nature. For
-reproducibility, the \texttt{seed} parameter of
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-may be set. In practice, using provided deduplication filters, we find
-that exact duplicate mazes are generated very infrequently, even when
-generating very large datasets.
-
-For use cases where mazes of different sizes, generation algorithms, or
-other parameter variations are required, we provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/collected_dataset.html\#MazeDatasetCollection}{\texttt{MazeDatasetCollection}}
-class, which allows for creating a single iterable dataset from multiple
-independent configurations.
-
-\hypertarget{visual-output-formats}{%
-\subsection{Visual Output Formats}\label{visual-output-formats}}
-
-Internally, mazes are
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
-objects, which have path information and a tensor optimized for storing
-sub-graphs of a lattice. These objects can be converted to and from
-several formats to maximize their utility in different contexts.
-
 \begin{figure}[H]
@@
 	  % algorithms
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
+	  \docslink{maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
+	  & \docslink{maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
+	  & \docslink{maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
 	  % descriptions
@@
 \end{figure}
-
-In previous work, maze tasks have been used with Recurrent Convolutional
-Neural Network (RCNN) derived architectures
-(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
-Huang, et al., 2021}). To facilitate the use of our package in this
-context, we replicate the format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) and provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
-class which returns rasterized pairs of (input, target) mazes as shown
-in \autoref{fig:e2h-raster} below.
-
-\begin{figure}
-\hypertarget{fig:e2h-raster}{%
-\centering
-\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
-\caption{Input is the rasterized maze without the path marked (left),
-and provide as a target the maze with all but the correct path removed
-(right). Configuration options exist to adjust whether endpoints are
-included and if empty cells should be filled in.}\label{fig:e2h-raster}
-}
-\end{figure}
-
-\hypertarget{sec:tokenized-output-formats}{%
-\subsection{Tokenized Output
-Formats}\label{sec:tokenized-output-formats}}
-
-Autoregressive transformer models can be quite sensitive to the exact
-format of input data, and may even use delimiter tokens to perform
-reasoning steps (\protect\hyperlink{ref-pfau2024dotbydot}{Pfau et al.,
-2024}; \protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al.,
-2024}). To facilitate systematic investigation of the effects of
-different representations of data on text model performance, we provide
-a variety of tokenized text output formats.
-
-We convert mazes to token sequences in two steps. First, the maze is
-stringified using
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens()}}.
-The
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class provides a powerful interface for configuring maze stringification
-behavior. Second, the sequence of strings is tokenized into integers
-using \texttt{encode()}. Tokenization uses a fixed vocabulary for
-simplicity. Mazes up to \(50 \times 50\) are supported when using a
-unique token for each position, and up to \(128 \times 128\) are
-supported when positions in the maze are represented as a pair of
-coordinates.
-
-There are many algorithms by which one might tokenize a 2D maze into a
-1D format usable by autoregressive text models. Training multiple models
-on the encodings output from each of these algorithms may produce very
-different internal representations, learned solution algorithms, and
-levels of performance. To allow exploration of how different maze
-tokenization algorithms affect these models, the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class contains a rich set of options to customize how mazes are
-stringified. This class contains 19 discrete parameters, resulting in
-over 5.8 million unique tokenizers. There are 6 additional parameters
-available whose functionality is not verified via automated testing, but
-further expand the the number of tokenizers by a factor of \(44/3\) to
-86 million.
-
-All output sequences consist of four token regions representing
-different features of the maze; an example output sequence is shown in
-\autoref{fig:token-regions}.
-
-\begin{figure} 
-  \centering
-  \begin{minipage}{5in}
-    \footnotesize
-    \input{figures/outputs-tokens-colored.tex}
-  \end{minipage}
-  \caption{
-    Example text output format with token regions highlighted.
-    \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the graph,
-    \colorbox[RGB]{ 217,234,211 }{Origin}: starting coordinate,
-    \colorbox[RGB]{ 234,209,220 }{Target}: ending coordinate,
-    \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence
-  }
-  \label{fig:token-regions}
-\end{figure}
-
-Each
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-is constructed from a set of several
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-objects, each of which specifies how different token regions or other
-elements of the stringification are produced.
-
-\begin{figure}
-    \centering
-    \input{figures/TokenizerElement_structure.tikz}
-    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular}.}
-\end{figure}
-
-The tokenizer architecture is purposefully designed such that adding and
-testing a wide variety of new tokenization algorithms is fast and
-minimizes disturbances to functioning code. This is enabled by the
-modular architecture and the automatic inclusion of any new tokenizers
-in integration tests. To create a new variety of tokenizer, developers
-forking the library may simply create their own
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and implement the abstract methods. If the behavior change is
-sufficiently small, simply adding a parameter to an existing
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and updating its implementation will suffice.
-
-The breadth of tokenizers is also easily scaled in the opposite
-direction. Due to the exponential scaling of parameter combinations,
-adding a small number of new features can significantly slow certain
-procedures which rely on constructing all possible tokenizers, such as
-integration tests. If any existing subclass contains features which
-aren't needed, a developer tool decorator
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
-is provided which can be applied to the unneeded
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclasses to prune those features and compact the available space of
-tokenizers.
-
-\hypertarget{benchmarks}{%
-\subsection{Benchmarks of Generation Speed}\label{benchmarks}}
-
-We provide approximate benchmarks for relative generation time across
-various algorithms, parameter choices, maze sizes, and dataset sizes in
-\autoref{tab:benchmarks} and \autoref{fig:benchmarks}. Experiments were
-performed on a
-\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners#standard-github-hosted-runners-for-public-repositories}{standard GitHub runner}
-without parallelism.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|ll|r|rrr|}
-  \hline
-  maze\_ctor
-          & keyword args           & all sizes 
-                                              & \shortstack{small \\ $g \leq 10$} 
-                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
-                                                                    & \shortstack{large \\ $g > 32$} \\
-  \hline\hline
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
-          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
-          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
-          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
-          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
-  \hline\hline
-  mean
-          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
-  median
-          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
-  \hline
-\end{tabular}
-\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
-\label{tab:benchmarks}
-\end{table}
-
-\begin{figure}
-\hypertarget{fig:benchmarks}{%
-\centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
-\caption{Plot of maze generation time. Generation time scales
-exponentially with maze size for all algorithms. Generation time per
-maze does not depend on the number of mazes being generated, and there
-is minimal overhead to initializing the generation process for a small
-dataset. Wilson's algorithm is notably less efficient than others and
-has high variance. Note that values are averaged across all parameter
-sets for that algorithm. More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}\label{fig:benchmarks}
-}
-\end{figure}
-
-\hypertarget{sec:success-rate-estimation}{%
-\subsection{Success Rate Estimation}\label{sec:success-rate-estimation}}
-
-In order to replicate the exact dataset distribution of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}), the parameter
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}}
-allows for additional constraints such as enforcing that the start or
-end point be in a ``dead end'' with only one accessible neighbor cell.
-However, combining these constraints with cyclic mazes (such as those
-generated with percolation), as was required for the work in
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), can lead to an absence of valid start and end points. Placing
-theoretical bounds on this success rate is difficult, as it depends on
-the exact maze generation algorithm and parameters used. To deal with
-this, our package provides a way to estimate the success rate of a given
-configuration using a symbolic regression model trained with PySR
-(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
-be found in
-\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
-Using the estimation algorithm simply requires the user to call
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
-providing their initial \texttt{cfg} and then using the returned
-\texttt{cfg\_new} in its place.
-
-\hypertarget{success-rate-estimation-algorithm}{%
-\subsubsection{Success Rate Estimation
-Algorithm}\label{success-rate-estimation-algorithm}}
-
-The base function learned by symbolic regression provides limited
-insight and may be subject to change. It is defined as
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
-and takes a 5 dimensional float vector created by
-\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the
-{[}percolation value, grid size, endpoint deadend configuration,
-endpoint uniqueness, categorical generation function index{]}.
-
-However, the outputs of this function are not directly usable due to
-minor divergences at the endpoints with respect to the percolation
-probability \(p\). Since we know that maze success is either guaranteed
-or impossible for \(p=0\) and \(p=1\), we define the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function to nudge the raw output of the symbolic regression. This
-function is defined with the following components:
-
-shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
-function given by \[
-  \sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}
-  \qquad A(q,a,w) = w \cdot (1 - |2q-1|^a)
-\] \[
-  h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)
-\]
-
-We combine these to get the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function, which is identity-like for \(p \approx 0.5\), and pushes \(x\)
-to extremes otherwise. \[
-  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
-\]
-
-Finally, we define \[
-  \text{cfg\_success\_predict\_fn}(\mathbf{x}) = \text{soft\_step}(\text{raw\_val}, x_0, 5, 10)
-\]
-
-where \texttt{raw\_val} is the output of the symbolic regression model.
-The parameter \(x_0\) is the percolation probability, while all other
-parameters from \texttt{\_to\_ps\_array()} only affect
-\texttt{raw\_val}.
-
-\begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
-\caption{An example of both empirical and predicted success rates as a
-function of the percolation probability \(p\) for various maze sizes,
-percolation with and without depth first search, and
-\texttt{endpoint\_kwargs} requiring that both the start and end be in
-unique dead ends. Empirical measures derived from a sample of 128 mazes.
-More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}
-\end{figure}
-
-\hypertarget{sec:implementation}{%
-\section{Implementation}\label{sec:implementation}}
-
-We refer to our
-\href{https://github.com/understanding-search/maze-dataset}{repository}
-and
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
-for documentation and up-to-date implementation details.
-
-This package utilizes a simple, efficient representation of mazes as
-subgraphs of a finite lattice, which we call a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-Using an adjacency matrix for storing mazes would be memory inefficient
-by failing to exploit the highly sparse structure -- for example, for a
-2-dimensional maze, only 4 of the diagonals would be filled in. On the
-other hand, using an adjacency list could lead to a poor lookup time for
-whether any given connection exists.
-
-Instead, we describe mazes with the following representation: for a
-\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
-initialize a boolean array \[
-  A = \{0, 1\}^{2 \times r \times c}
-\] which we refer to in the code as a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
-The value at \(A[0,i,j]\) determines whether a \emph{downward}
-connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
-value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
-to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
-existence of connections and facilitate fast lookup time, at the cost of
-requiring additional care with indexing. Note that this setup allows for
-a periodic lattice. Generation of mazes is detailed in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
-
-To produce solutions to mazes, two points are selected uniformly at
-random without replacement from the connected component of the maze, and
-the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
-1968}) is applied to find the shortest path between them. The endpoint
-selection can be controlled via
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
-and complications caused by this are detailed in
-\hyperref[sec:success-rate-estimation]{section: \textit{\nameref{sec:success-rate-estimation}}}.
-A maze with a solution is denoted a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}},
-which inherits from
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-
-Parallelization is implemented via the \texttt{multiprocessing} module
-in the Python standard library, and parallel generation can be
-controlled via keyword arguments to
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}.
-
-\newpage
-
-\hypertarget{usage-in-research}{%
-\section{Usage in Research}\label{usage-in-research}}
-
-This package was originally built for the needs of the
-(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
-et al., 2023b}) project, which aims to investigate spatial planning and
-world models in autoregressive transformer models trained on mazes
-(\protect\hyperlink{ref-ivanitskiy2023structuredworldreps}{Michael
-Igorevich Ivanitskiy, Spies, et al., 2023};
-\protect\hyperlink{ref-maze-dataset-arxiv-2023}{Michael Igorevich
-Ivanitskiy, Shah, et al., 2023};
-\protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al., 2024}).
-It was extended for work on understanding the mechanisms by which
-recurrent convolutional and implicit networks
-(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
-given a rasterized view
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), which required matching the pixel-padded and endpoint
-constrained output format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). Ongoing work using \texttt{maze-dataset} aims to
-investigate the effects of varying the tokenization format on the
-performance of pretrained LLMs on spatial reasoning.
-
-This package has also been utilized in work by other groups:
-
-\begin{itemize}
-\item
-  By (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) to
-  compare the effectiveness of transformers trained with the
-  MLM-\(\mathcal{U}\)
-  (\protect\hyperlink{ref-MLMU-kitouni2024factorization}{Kitouni et al.,
-  2024}) multistep prediction objective against standard autoregressive
-  training for multi-step planning on our maze task.
-\item
-  By (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) to
-  study the effectiveness of imperative learning.
-\item
-  By (\protect\hyperlink{ref-zhang2025tscend}{Zhang et al., 2025}) to
-  introduce a novel framework for reasoning diffusion models.
-\item
-  By (\protect\hyperlink{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to
-  improve spatial reasoning in LLMs with GRPO.
-\end{itemize}
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-This work was partially funded by National Science Foundation awards
-DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
-for hosting authors MII, AFS, and TR for a residency visit, and to
-various members of FAR's technical staff for their advice.
-
-This work was partially supported by AI Safety Camp and AI Safety
-Support, which also brought many of the authors together. We would like
-to thank our former collaborators at AI Safety Camp and other users and
-contributors to the \texttt{maze-dataset} package: Benji Berczi,
-Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris Mathwin, Lucia
-Quirke, Can Rager, Adrians Skapars, Rusheb Shah, Johannes Treutlein, and
-Dan Valentine.
-
-We thank the Mines Optimization and Deep Learning group (MODL) for
-fruitful discussions. We also thank Michael Rosenberg for recommending
-the usage of Finite State Transducers for storing tokenizer validation
-information.
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0.5}
-\leavevmode\vadjust pre{\hypertarget{ref-mazegenerator-net}{}}%
-Alance AB. (2019). \emph{Maze generator}.
-\url{http://www.mazegenerator.net}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
-Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
-1.0: A complete set of tools to prepare, present, and analyze
-navigational and spatial cognitive neuroscience experiments.
-\emph{Behavior Research Methods}, \emph{40}, 353--359.
-
-\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
-Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
-{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
-Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
-procedural generation to benchmark reinforcement learning. \emph{arXiv
-Preprint arXiv:1912.01588}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pysr}{}}%
-Cranmer, M. (2023). Interpretable machine learning for science with PySR
-and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
-Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
-models' spatial intelligence via GRPO. \emph{arXiv Preprint
-arXiv:2502.14669}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation}{}}%
-Duminil-Copin, H. (2017). \emph{Sixty years of percolation} (No.
-arXiv:1712.04651). {arXiv}. \url{http://arxiv.org/abs/1712.04651}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Ehsan_2022}{}}%
-Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation-clustersize}{}}%
-Fisher, M. E., \& Essam, J. W. (2004). Some {Cluster Size} and
-{Percolation Problems}. \emph{Journal of Mathematical Physics},
-\emph{2}(4), 609--619. \url{https://doi.org/10.1063/1.1703745}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
-Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
-(2022). Jfb: Jacobian-free backpropagation for implicit networks.
-\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
-\emph{36}, 6648--6656.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
-Guo, C., Barthelet, L., \& Morris, R. (2011). \emph{Maze generator and
-solver}. Wolfram Demonstrations Project,
-\url{https://demonstrations.wolfram.com/MazeGeneratorAndSolver/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
-Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
-{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
-Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
-{Games CoG}}, 1--4.
-
-\leavevmode\vadjust pre{\hypertarget{ref-A_star}{}}%
-Hart, P. E., Nilsson, N. J., \& Raphael, B. (1968). A {Formal Basis} for
-the {Heuristic Determination} of {Minimum Cost Paths}. \emph{IEEE
-Transactions on Systems Science and Cybernetics}, \emph{4}(2), 100--107.
-\url{https://doi.org/10.1109/TSSC.1968.300136}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
-Ivanitskiy, M. (n.d.). \emph{ZANJ}.
-\url{https://github.com/mivanit/ZANJ}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
-\emph{Maze dataset}.
-\url{https://github.com/understanding-search/maze-dataset}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
-\emph{Maze transformer interpretability}.
-\url{https://github.com/understanding-search/maze-transformer}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
-Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
-Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
-C. D., \& others. (2023). A configurable library for generating and
-manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
-Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
-Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
-D., \& others. (2023). Structured world representations in maze-solving
-transformers. \emph{arXiv Preprint arXiv:2312.02566}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
-Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
-Ibrahim, M. (2024). The factorization curse: Which tokens you predict
-underlie the reversal curse and more. \emph{Advances in Neural
-Information Processing Systems}, \emph{37}, 112329--112355.
-
-\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
-Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
-Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
-extrapolation for mazes with recurrent and implicit networks.
-\emph{arXiv Preprint arXiv:2410.03020}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruskal1956shortest}{}}%
-Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
-the traveling salesman problem. \emph{Proceedings of the American
-Mathematical Society}, \emph{7}(1), 48--50.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
-Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
-Performance insights and comparative analysis. \emph{arXiv Preprint
-arXiv:2308.11224}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
-Nag, A. (2020). MDL suite: A language, generator and compiler for
-describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
-1815.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Nemeth_2019}{}}%
-Németh, F. (2019). \emph{Maze-generation-algorithms}.
-\url{https://github.com/ferenc-nemeth/maze-generation-algorithms}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
-Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
-Transformers can navigate mazes with multi-step prediction. \emph{arXiv
-Preprint arXiv:2412.05117}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh-oppenheimj2018maze}{}}%
-Oppenheim, J. (2018). \emph{Maze-generator: Generate a random maze
-represented as a 2D array using depth-first search}.
-\url{https://github.com/oppenheimj/maze-generator/}; GitHub.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
-Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G.,
-Killeen, T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., Kopf,
-A., Yang, E., DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S.,
-Steiner, B., Fang, L., \ldots{} Chintala, S. (2019). {PyTorch: An
-Imperative Style, High-Performance Deep Learning Library}. In H.
-Wallach, H. Larochelle, A. Beygelzimer, F. d'Alché-Buc, E. Fox, \& R.
-Garnett (Eds.), \emph{Advances in neural information processing systems
-32} (pp. 8024--8035). Curran Associates, Inc.
-\url{http://papers.neurips.cc/paper/9015-pytorch-an-imperative-style-high-performance-deep-learning-library.pdf}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
-Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
-Hidden computation in transformer language models. \emph{arXiv Preprint
-arXiv:2404.15758}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
-Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
-transparent ai: A survey on interpreting the inner structures of deep
-neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
-Machine Learning (SaTML)}, 464--483.
-
-\leavevmode\vadjust pre{\hypertarget{ref-easy_to_hard}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Bansal, A., Emam, Z., Huang,
-F., Goldblum, M., \& Goldstein, T. (2021). \emph{Datasets for {Studying
-Generalization} from {Easy} to {Hard Examples}} (No. arXiv:2108.06011).
-{arXiv}. \url{https://doi.org/10.48550/arXiv.2108.06011}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
-Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
-Generalizing from easy to hard problems with recurrent networks.
-\emph{Advances in Neural Information Processing Systems}, \emph{34},
-6695--6706.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
-Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
-\emph{arXiv Preprint arXiv:2308.02522}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
-Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
-Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
-world models in maze-solving tasks. \emph{arXiv Preprint
-arXiv:2412.11867}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
-Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
-Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
-self-supervised neural-symbolic learning framework for robot autonomy.
-\emph{arXiv Preprint arXiv:2406.16087}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wilson}{}}%
-Wilson, D. B. (1996). Generating random spanning trees more quickly than
-the cover time. \emph{Proceedings of the Twenty-Eighth Annual {ACM}
-Symposium on {Theory} of Computing - {STOC} '96}, 296--303.
-\url{https://doi.org/10.1145/237814.237880}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
-Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
-scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
-arXiv:2502.01989}.
-
-\end{CSLReferences}
-
-\end{document}

### 2025-04-26 docs/paper/figures/tex/fig4_tokenfmt.tex
move figure code to external files

--- 2025-04-09 docs/paper/paper.tex
@@
-\documentclass[10pt,a4paper,onecolumn]{article}
-\usepackage{marginnote}
-\usepackage{graphicx}
-\usepackage[rgb,svgnames]{xcolor}
-\usepackage{authblk,etoolbox}
-\usepackage{titlesec}
-\usepackage{calc}
-\usepackage{tikz}
-\usepackage[pdfa]{hyperref}
-\usepackage{hyperxmp}
-\hypersetup{%
-    unicode=true,
-    pdfapart=3,
-    pdfaconformance=B,
-    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility},
-    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
-Fung},
-    pdfpublication={Journal of Open Source Software},
-    pdfpublisher={Open Journals},
-    pdfissn={2475-9066},
-    pdfpubtype={journal},
-    pdfvolumenum={},
-    pdfissuenum={},
-    pdfdoi={N/A},
-    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
-Diniz Behn, Samy Wu Fung},
-    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
-    colorlinks=true,
-    linkcolor=[rgb]{0.0, 0.5, 1.0},
-    citecolor=Blue,
-    urlcolor=[rgb]{0.0, 0.5, 1.0},
-    breaklinks=true
-}
-% https://tex.stackexchange.com/a/535849
-% Create an OutputIntent in order to correctly specify colours
-\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
-\pdfcatalog{%
-  /OutputIntents [
-    <<
-      /Type /OutputIntent
-      /S /GTS_PDFA1
-      /DestOutputProfile \the\pdflastobj\space 0 R
-      /OutputConditionIdentifier (sRGB)
-      /Info (sRGB)
-    >>
-  ]
-}
-\pdfvariable omitcidset=1
-\usepackage{caption}
-\usepackage{orcidlink}
-\usepackage{tcolorbox}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{seqsplit}
-\usepackage{xstring}
-
-\usepackage{float}
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
-
-\usepackage{fixltx2e} % provides \textsubscript
-
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-% --- Page layout -------------------------------------------------------------
-\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
-            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}
-
-% --- Default font ------------------------------------------------------------
-\renewcommand\familydefault{\sfdefault}
-
-% --- Style -------------------------------------------------------------------
-\renewcommand{\captionfont}{\small\sffamily}
-\renewcommand{\captionlabelfont}{\bfseries}
-
-% --- Section/SubSection/SubSubSection ----------------------------------------
-\titleformat{\section}
-  {\normalfont\sffamily\Large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsection}
-  {\normalfont\sffamily\large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsubsection}
-  {\normalfont\sffamily\bfseries}
-  {}{0pt}{}
-\titleformat*{\paragraph}
-  {\sffamily\normalsize}
-
-% --- Header / Footer ---------------------------------------------------------
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-%\renewcommand{\headrulewidth}{0.50pt}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{joss/logo.png}}
-\fancyhead[C]{}
-\fancyhead[R]{}
-\renewcommand{\footrulewidth}{0.25pt}
-
-\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset: Maze Generation with Algorithmic Variety
-and Representational Flexibility. \emph{Journal of Open Source
-Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
-
-\fancyfoot[R]{\sffamily \thepage}
-\makeatletter
-\let\ps@plain\ps@fancy
-\fancyheadoffset[L]{4.5cm}
-\fancyfootoffset[L]{4.5cm}
-
-% --- Macros ---------
-
-\definecolor{linky}{rgb}{0.0, 0.5, 1.0}
-
-\newtcolorbox{repobox}
-   {colback=red, colframe=red!75!black,
-     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}
-
-\newcommand{\ExternalLink}{%
-   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
-       \begin{scope}[x=1ex, y=1ex]
-           \clip (-0.1,-0.1)
-               --++ (-0, 1.2)
-               --++ (0.6, 0)
-               --++ (0, -0.6)
-               --++ (0.6, 0)
-               --++ (0, -1);
-           \path[draw,
-               line width = 0.5,
-               rounded corners=0.5]
-               (0,0) rectangle (1,1);
-       \end{scope}
-       \path[draw, line width = 0.5] (0.5, 0.5)
-           -- (1, 1);
-       \path[draw, line width = 0.5] (0.6, 1)
-           -- (1, 1) -- (1, 0.6);
-       }
-   }
-
-\definecolor{c53baa1}{RGB}{83,186,161}
-\definecolor{c202826}{RGB}{32,40,38}
-\def \rorglobalscale {0.1}
-\newcommand{\rorlogo}{%
-\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
-  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
-    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364)..
-controls (3.9643, 1.4576) and (4.0508, 1.5081) ..
-(4.1098, 1.5878)..
-controls (4.169, 1.6674) and (4.1984, 1.7642) ..
-(4.1984, 1.8777)..
-controls (4.1984, 1.9719) and (4.182, 2.0503) ..
-(4.1495, 2.1132)..
-controls (4.1169, 2.1762) and (4.0727, 2.2262) ..
-(4.0174, 2.2635)..
-controls (3.9621, 2.3006) and (3.8976, 2.3273) ..
-(3.824, 2.3432)..
-controls (3.7505, 2.359) and (3.6727, 2.367) ..
-(3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688)..
-controls (2.9625, 1.8833) and (2.9572, 1.8976) ..
-(2.9514, 1.9119)..
-controls (2.9083, 2.0164) and (2.848, 2.1056) ..
-(2.7705, 2.1791)..
-controls (2.6929, 2.2527) and (2.6014, 2.3093) ..
-(2.495, 2.3487)..
-controls (2.3889, 2.3881) and (2.2728, 2.408) ..
-(2.1468, 2.408)..
-controls (2.0209, 2.408) and (1.905, 2.3881) ..
-(1.7986, 2.3487)..
-controls (1.6925, 2.3093) and (1.6007, 2.2527) ..
-(1.5232, 2.1791)..
-controls (1.4539, 2.1132) and (1.3983, 2.0346) ..
-(1.3565, 1.9436)..
-controls (1.3504, 2.009) and (1.3351, 2.0656) ..
-(1.3105, 2.1132)..
-controls (1.2779, 2.1762) and (1.2338, 2.2262) ..
-(1.1785, 2.2635)..
-controls (1.1232, 2.3006) and (1.0586, 2.3273) ..
-(0.985, 2.3432)..
-controls (0.9115, 2.359) and (0.8337, 2.367) ..
-(0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364)..
-controls (1.1255, 1.4576) and (1.2121, 1.5081) ..
-(1.2711, 1.5878)..
-controls (1.2737, 1.5915) and (1.2761, 1.5954) ..
-(1.2787, 1.5991)..
-controls (1.2782, 1.5867) and (1.2779, 1.5743) ..
-(1.2779, 1.5616)..
-controls (1.2779, 1.4327) and (1.2996, 1.3158) ..
-(1.3428, 1.2113)..
-controls (1.3859, 1.1068) and (1.4462, 1.0176) ..
-(1.5237, 0.944)..
-controls (1.601, 0.8705) and (1.6928, 0.8139) ..
-(1.7992, 0.7744)..
-controls (1.9053, 0.735) and (2.0214, 0.7152) ..
-(2.1474, 0.7152)..
-controls (2.2733, 0.7152) and (2.3892, 0.735) ..
-(2.4956, 0.7744)..
-controls (2.6016, 0.8139) and (2.6935, 0.8705) ..
-(2.771, 0.944)..
-controls (2.8482, 1.0176) and (2.9086, 1.1068) ..
-(2.952, 1.2113)..
-controls (2.9578, 1.2253) and (2.9631, 1.2398) ..
-(2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777)..
-controls (0.9438, 1.7534) and (0.92, 1.7357) ..
-(0.8911, 1.7243)..
-controls (0.8623, 1.7129) and (0.83, 1.706) ..
-(0.7945, 1.7039)..
-controls (0.7588, 1.7015) and (0.7252, 1.7005) ..
-(0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667)..
-controls (0.7477, 2.0667) and (0.7805, 2.0643) ..
-(0.8139, 2.0598)..
-controls (0.8472, 2.0553) and (0.8768, 2.0466) ..
-(0.9025, 2.0336)..
-controls (0.9282, 2.0206) and (0.9496, 2.0021) ..
-(0.9663, 1.9778)..
-controls (0.9829, 1.9534) and (0.9914, 1.9209) ..
-(0.9914, 1.8799)..
-controls (0.9914, 1.8362) and (0.9819, 1.8021) ..
-(0.9628, 1.7777) -- cycle(2.6125, 1.3533)..
-controls (2.5889, 1.2904) and (2.5553, 1.2359) ..
-(2.5112, 1.1896)..
-controls (2.4672, 1.1433) and (2.4146, 1.1073) ..
-(2.3529, 1.0814)..
-controls (2.2916, 1.0554) and (2.2228, 1.0427) ..
-(2.1471, 1.0427)..
-controls (2.0712, 1.0427) and (2.0026, 1.0557) ..
-(1.9412, 1.0814)..
-controls (1.8799, 1.107) and (1.8272, 1.1433) ..
-(1.783, 1.1896)..
-controls (1.7391, 1.2359) and (1.7052, 1.2904) ..
-(1.6817, 1.3533)..
-controls (1.6581, 1.4163) and (1.6465, 1.4856) ..
-(1.6465, 1.5616)..
-controls (1.6465, 1.6359) and (1.6581, 1.705) ..
-(1.6817, 1.7687)..
-controls (1.7052, 1.8325) and (1.7388, 1.8873) ..
-(1.783, 1.9336)..
-controls (1.8269, 1.9799) and (1.8796, 2.0159) ..
-(1.9412, 2.0418)..
-controls (2.0026, 2.0675) and (2.0712, 2.0804) ..
-(2.1471, 2.0804)..
-controls (2.223, 2.0804) and (2.2916, 2.0675) ..
-(2.3529, 2.0418)..
-controls (2.4143, 2.0161) and (2.467, 1.9799) ..
-(2.5112, 1.9336)..
-controls (2.5551, 1.8873) and (2.5889, 1.8322) ..
-(2.6125, 1.7687)..
-controls (2.636, 1.705) and (2.6477, 1.6359) ..
-(2.6477, 1.5616)..
-controls (2.6477, 1.4856) and (2.636, 1.4163) ..
-(2.6125, 1.3533) -- cycle(3.8015, 1.7777)..
-controls (3.7825, 1.7534) and (3.7587, 1.7357) ..
-(3.7298, 1.7243)..
-controls (3.701, 1.7129) and (3.6687, 1.706) ..
-(3.6333, 1.7039)..
-controls (3.5975, 1.7015) and (3.5639, 1.7005) ..
-(3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667)..
-controls (3.5864, 2.0667) and (3.6192, 2.0643) ..
-(3.6526, 2.0598)..
-controls (3.6859, 2.0553) and (3.7155, 2.0466) ..
-(3.7412, 2.0336)..
-controls (3.7669, 2.0206) and (3.7883, 2.0021) ..
-(3.805, 1.9778)..
-controls (3.8216, 1.9534) and (3.8301, 1.9209) ..
-(3.8301, 1.8799)..
-controls (3.8301, 1.8362) and (3.8206, 1.8021) ..
-(3.8015, 1.7777) -- cycle;
-  \end{scope}
-\end{tikzpicture}
-}
-
-% --- Title / Authors ---------------------------------------------------------
-% patch \maketitle so that it doesn't center
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-% patch \maketitle so that the font size for the title is normal
-\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
-% patch the patch by authblk so that the author block is flush left
-\def\maketitle{{%
-  \renewenvironment{tabular}[2][]
-    {\begin{flushleft}}
-    {\end{flushleft}}
-  \AB@maketitle}}
-\renewcommand\AB@affilsepx{ \protect\Affilfont}
-%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
-\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
-\renewcommand{\affil}[2][]%
-   {\newaffiltrue\let\AB@blk@and\AB@pand
-      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
-        \setcounter{Maxaffil}{0}\fi
-        \begingroup
-        \let\href=\href@Orig
-        \let\protect\@unexpandable@protect
-        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
-        \@temptokena=\expandafter{\AB@authors}%
-        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
-         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
-         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
-         \gdef\AB@las{}\gdef\AB@au@str{}%
-        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
-        \@temptokena=\expandafter{\AB@affillist}%
-        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
-          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
-      \endgroup
-       \let\AB@affilsep\AB@affilsepx
-}
-\makeatother
-\renewcommand\Authfont{\sffamily\bfseries}
-\renewcommand\Affilfont{\sffamily\small\mdseries}
-\setlength{\affilsep}{1em}
-
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{fontspec}
-
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\sffamily]{Ligatures=TeX}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-
-%% Font settings
-\usepackage{fontsetup} % Lazy way to get proper Greek lowercase glyphs
-
-% Use Hack https://sourcefoundry.org/hack/
-\setmonofont{Hack}
-
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\urlstyle{same}  % don't use monospace font for urls
-\ifLuaTeX
-\usepackage[bidi=basic]{babel}
-\else
-\usepackage[bidi=default]{babel}
-\fi
-\babelprovide[main,import]{american}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\tikzset{ % Define a TikZ style for an external hyperlink node.
-  hyperlink node url/.style={
-    alias=sourcenode,
-    append after command={
-      let \p1 = (sourcenode.north west),
-          \p2 = (sourcenode.south east),
-          \n1 = {\x2-\x1},
-          \n2 = {\y1-\y2} in
-      node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
-          {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
-    }
-  }
-}
-\providecommand{\XeTeXLinkBox}[1]{#1}
-\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
-\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
-\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-
-\title{maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility}
-
-\author[1%
-%
-\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
-  \,\orcidlink{0000-0002-4213-4993}\,%
-}
-\author[4%
-%
-]{Aaron Sandoval%
-  \,\orcidlink{0009-0002-8380-6140}\,%
-}
-\author[2%
-%
-]{Alex F. Spies%
-  \,\orcidlink{0000-0002-8708-1530}\,%
-}
-\author[3%
-%
-]{Tilman Räuker%
-  \,\orcidlink{0009-0009-6321-4413}\,%
-}
-\author[1%
-%
-]{Brandon Knutson%
-  \,\orcidlink{0009-0004-8413-0239}\,%
-}
-\author[1%
-%
-]{Cecilia Diniz Behn%
-  \,\orcidlink{0000-0002-8078-5105}\,%
-}
-\author[1%
-%
-]{Samy Wu Fung%
-  \,\orcidlink{0000-0002-2926-4582}\,%
-}
-
-\affil[1]{Colorado School of Mines, Department of Applied Mathematics
-and Statistics%
-}
-\affil[2]{Imperial College London%
-}
-\affil[3]{UnSearch.org%
-}
-\affil[4]{Independent%
-}
-\affil[$\mathparagraph$]{Corresponding author}
-\date{\vspace{-2.5ex}}
-
-\begin{document}
-\maketitle
-
-\marginpar{
-
-  \begin{flushleft}
-  %\hrule
-  \sffamily\small
-
-  {\bfseries DOI:} \href{https://doi.org/N/A}{\color{linky}{N/A}}
-
-  \vspace{2mm}
-    {\bfseries Software}
-  \begin{itemize}
-    \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{\color{linky}{Review}} \ExternalLink
-    \item \href{https://github.com/openjournals}{\color{linky}{Repository}} \ExternalLink
-    \item \href{https://doi.org/10.5281}{\color{linky}{Archive}} \ExternalLink
-  \end{itemize}
-
-  \vspace{2mm}
-
-    \par\noindent\hrulefill\par
-
-  \vspace{2mm}
-
-  {\bfseries Editor:} \href{https://joss.theoj.org}{Open
-Journals} \ExternalLink
-   \\
-  \vspace{1mm}
-    {\bfseries Reviewers:}
-  \begin{itemize}
-  \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{@openjournals}
-    \end{itemize}
-    \vspace{2mm}
-
-    {\bfseries Submitted:} 01 January 1970\\
-    {\bfseries Published:} 01 January 1970
-
-  \vspace{2mm}
-  {\bfseries License}\\
-  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{https://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).
-
-  \end{flushleft}
-}
-
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
-
-Solving mazes is a classic problem in computer science and artificial
-intelligence, and humans have been constructing mazes for thousands of
-years. Although finding the shortest path through a maze is a solved
-problem, this very fact makes it an excellent testbed for studying how
-machine learning algorithms solve problems and represent spatial
-information. We introduce \texttt{maze-dataset}, a user-friendly Python
-library for generating, processing, and visualizing datasets of mazes.
-This library supports a variety of maze generation algorithms providing
-mazes with or without loops, mazes that are connected or not, and many
-other variations. These generation algorithms can be configured with
-various parameters, and the resulting mazes can be filtered to satisfy
-desired properties. Also provided are tools for converting mazes to and
-from various formats suitable for a variety of neural network
-architectures, such as rasterized images, tokenized text sequences, and
-various visualizations. As well as providing a simple interface for
-generating, storing, and loading these datasets, \texttt{maze-dataset}
-is extensively tested, type hinted, benchmarked, and documented.
-
-\begin{figure} 
-  \begin{minipage}{5in}
-    \input{diagram/diagram.tikz} 
-  \end{minipage}
-  \caption{
-    Usage of maze-dataset.
-We create a \texttt{MazeDataset} from a \texttt{MazeDatasetConfig}.
-This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats.
-Code in the image contains clickable links to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{documentation}.
-A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.
-  }
-  \label{fig:diagram}
-\end{figure}
-
-\hypertarget{statement-of-need}{%
-\section{Statement of Need}\label{statement-of-need}}
-
-While maze generation itself is straightforward, the architectural
-challenge comes from building a system supporting many algorithms with
-configurable parameters, property filtering, and representation
-transformation. This library aims to greatly streamline the process of
-generating and working with datasets of mazes that can be described as
-subgraphs of an \(n \times n\) lattice with boolean connections and,
-optionally, start and end points that are nodes in the graph.
-Furthermore, we place emphasis on a wide variety of possible text output
-formats aimed at evaluating the spatial reasoning capabilities of Large
-Language Models (LLMs) and other text-based transformer models.
-
-For interpretability and behavioral research, algorithmic tasks offer
-benefits by allowing systematic data generation and task decomposition,
-as well as simplifying the process of circuit discovery
-(\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
-Although mazes are well suited for these investigations, we found that
-existing maze generation packages
-(\protect\hyperlink{ref-cobbe2019procgen}{Cobbe et al., 2019};
-\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-harriesMazeExplorerCustomisable3D2019}{Harries et
-al., n.d.}; \protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) lack support for transforming between multiple
-representations and provide limited control over the maze generation
-process.
-
-\hypertarget{related-works}{%
-\subsection{Related Works}\label{related-works}}
-
-A multitude of public and open-source software packages exist for
-generating mazes (\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). However, nearly all of these packages produce
-mazes represented as rasterized images or other visual formats rather
-than the underlying graph structure, and this makes it difficult to work
-with these datasets.
-
-\begin{itemize}
-\item
-  Most prior works provide mazes in visual or raster formats, and we
-  provide a variety of similar output formats:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
-    utilizing
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
-    which can exactly mimic the outputs provided in
-    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
-    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
-    similar to the outputs of Németh
-    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    provides a format similar to
-    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
-    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
-    provides a feature-rich plotting utility with support for multiple
-    paths, heatmaps over positions, and more. This is similar to the
-    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
-    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
-    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
-  \end{itemize}
-\item
-  The text format provided by
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
-  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
-  Wu, 2023}), but provides over 5.8 million unique formats for
-  converting mazes to a text stream, detailed in
-  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
-\item
-  For rigorous investigations of the response of a model to various
-  distributional shifts, preserving metadata about the generation
-  algorithm with the dataset itself is essential. To this end, our
-  package efficiently stores the dataset along with its metadata in a
-  single human-readable file (\protect\hyperlink{ref-zanj}{M.
-  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
-  this reliably.
-\item
-  Storing mazes as images is not only difficult to work with, but also
-  inefficient. We use a highly efficient method detailed in
-  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
-\item
-  Our package is easily installable with source code freely available.
-  It is extensively tested, type hinted, benchmarked, and documented.
-  Many other maze generation packages lack this level of rigor and
-  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
-  2008}) appear to simply no longer be accessible.
-\end{itemize}
-
-\hypertarget{features}{%
-\section{Features}\label{features}}
-
-We direct readers to our
-\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples},
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs},
-and
-\href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
-for more information.
-
-\hypertarget{generation}{%
-\subsection{Generation and Basic Usage}\label{generation}}
-
-Our package can be installed from
-\href{https://pypi.org/project/maze-dataset/}{PyPi} via
-\texttt{pip\ install\ maze-dataset}, or directly from the
-\href{https://github.com/understanding-search/maze-dataset}{git
-repository} (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
-Ivanitskiy et al., 2023a}).
-
-To create a dataset, we first create a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-configuration object, which specifies the seed, number, and size of
-mazes, as well as the generation algorithm and its corresponding
-parameters. This object is passed to a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-class to create a dataset. Crucially, this
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-mimics the interface of a PyTorch
-(\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
-\href{https://pytorch.org/docs/stable/data.html}{\texttt{Dataset}}, and
-can thus be easily incorporated into existing data pre-processing and
-training pipelines, e.g., through the use of a \texttt{DataLoader}
-class.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
-\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
-\NormalTok{)}
-\CommentTok{\# create a config}
-\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
-\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
-\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
-\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
-\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
-    \CommentTok{\# (optional) algorithm{-}specific parameters}
-\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
-    \CommentTok{\# (optional) many options for restricting start/end points}
-\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
-\NormalTok{)}
-\CommentTok{\# create a dataset}
-\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
-\NormalTok{  cfg, }\CommentTok{\# pass the config}
-\NormalTok{  ..., }\CommentTok{\# other options for disk loading, parallelization, etc.}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-When initializing a dataset, options which do not affect the mazes
-themselves can be specified through the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
-factory method as necessary. These options allow for saving/loading
-existing datasets instead of re-generating, parallelization options for
-generation, and more. Available maze generation algorithms are static
-methods of the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
-namespace class and include generation algorithms based on randomized
-depth-first search, Wilson's algorithm
-(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
-(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
-\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
-Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
-others.
-
-Furthermore, a dataset of mazes can be filtered to satisfy certain
-properties. Custom filters can be specified, and some filters are
-included in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
-For example, we can require a minimum path length of three steps from
-the origin to the target:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dataset\_filtered: MazeDataset }\OperatorTok{=}\NormalTok{ dataset.filter\_by.path\_length(min\_length}\OperatorTok{=}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-All implemented maze generation algorithms are stochastic by nature. For
-reproducibility, the \texttt{seed} parameter of
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-may be set. In practice, using provided deduplication filters, we find
-that exact duplicate mazes are generated very infrequently, even when
-generating very large datasets.
-
-For use cases where mazes of different sizes, generation algorithms, or
-other parameter variations are required, we provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/collected_dataset.html\#MazeDatasetCollection}{\texttt{MazeDatasetCollection}}
-class, which allows for creating a single iterable dataset from multiple
-independent configurations.
-
-\hypertarget{visual-output-formats}{%
-\subsection{Visual Output Formats}\label{visual-output-formats}}
-
-Internally, mazes are
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
-objects, which have path information and a tensor optimized for storing
-sub-graphs of a lattice. These objects can be converted to and from
-several formats to maximize their utility in different contexts.
-
-\begin{figure}[H]
-  \centering
-  \begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
-    \hline \\[.5em]
-    % algorithms
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
-    % descriptions
-      Simple text format for displaying mazes, useful for debugging in a terminal environment.
-      & \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
-      & feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
-    \hline \\
-    % examples
-      \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
-        \setlength{\baselineskip}{0.9em}
-        \input{figures/outputs-ascii-colored.tex} 
-      \end{minipage}}
-      & \multicolumn{1}{c}{
-        \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
-      }
-      & \multicolumn{1}{c}{
-        \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
-      } \\[1em]
-
-    \hline \\
-  \end{tabular}
-  \caption{Various output formats. Top row (left to right): ASCII diagram, rasterized pixel grid, and advanced display tool.}
-  \label{fig:output-fmts}
-\end{figure}
-
-In previous work, maze tasks have been used with Recurrent Convolutional
-Neural Network (RCNN) derived architectures
-(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
-Huang, et al., 2021}). To facilitate the use of our package in this
-context, we replicate the format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) and provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
-class which returns rasterized pairs of (input, target) mazes as shown
-in \autoref{fig:e2h-raster} below.
-
-\begin{figure}
-\hypertarget{fig:e2h-raster}{%
-\centering
-\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
-\caption{Input is the rasterized maze without the path marked (left),
-and provide as a target the maze with all but the correct path removed
-(right). Configuration options exist to adjust whether endpoints are
-included and if empty cells should be filled in.}\label{fig:e2h-raster}
-}
-\end{figure}
-
-\hypertarget{sec:tokenized-output-formats}{%
-\subsection{Tokenized Output
-Formats}\label{sec:tokenized-output-formats}}
-
-Autoregressive transformer models can be quite sensitive to the exact
-format of input data, and may even use delimiter tokens to perform
-reasoning steps (\protect\hyperlink{ref-pfau2024dotbydot}{Pfau et al.,
-2024}; \protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al.,
-2024}). To facilitate systematic investigation of the effects of
-different representations of data on text model performance, we provide
-a variety of tokenized text output formats.
-
-We convert mazes to token sequences in two steps. First, the maze is
-stringified using
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens()}}.
-The
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class provides a powerful interface for configuring maze stringification
-behavior. Second, the sequence of strings is tokenized into integers
-using \texttt{encode()}. Tokenization uses a fixed vocabulary for
-simplicity. Mazes up to \(50 \times 50\) are supported when using a
-unique token for each position, and up to \(128 \times 128\) are
-supported when positions in the maze are represented as a pair of
-coordinates.
-
-There are many algorithms by which one might tokenize a 2D maze into a
-1D format usable by autoregressive text models. Training multiple models
-on the encodings output from each of these algorithms may produce very
-different internal representations, learned solution algorithms, and
-levels of performance. To allow exploration of how different maze
-tokenization algorithms affect these models, the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class contains a rich set of options to customize how mazes are
-stringified. This class contains 19 discrete parameters, resulting in
-over 5.8 million unique tokenizers. There are 6 additional parameters
-available whose functionality is not verified via automated testing, but
-further expand the the number of tokenizers by a factor of \(44/3\) to
-86 million.
-
-All output sequences consist of four token regions representing
-different features of the maze; an example output sequence is shown in
-\autoref{fig:token-regions}.
-
 \begin{figure} 
@@
 \end{figure}
-
-Each
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-is constructed from a set of several
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-objects, each of which specifies how different token regions or other
-elements of the stringification are produced.
-
-\begin{figure}
-    \centering
-    \input{figures/TokenizerElement_structure.tikz}
-    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular}.}
-\end{figure}
-
-The tokenizer architecture is purposefully designed such that adding and
-testing a wide variety of new tokenization algorithms is fast and
-minimizes disturbances to functioning code. This is enabled by the
-modular architecture and the automatic inclusion of any new tokenizers
-in integration tests. To create a new variety of tokenizer, developers
-forking the library may simply create their own
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and implement the abstract methods. If the behavior change is
-sufficiently small, simply adding a parameter to an existing
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and updating its implementation will suffice.
-
-The breadth of tokenizers is also easily scaled in the opposite
-direction. Due to the exponential scaling of parameter combinations,
-adding a small number of new features can significantly slow certain
-procedures which rely on constructing all possible tokenizers, such as
-integration tests. If any existing subclass contains features which
-aren't needed, a developer tool decorator
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
-is provided which can be applied to the unneeded
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclasses to prune those features and compact the available space of
-tokenizers.
-
-\hypertarget{benchmarks}{%
-\subsection{Benchmarks of Generation Speed}\label{benchmarks}}
-
-We provide approximate benchmarks for relative generation time across
-various algorithms, parameter choices, maze sizes, and dataset sizes in
-\autoref{tab:benchmarks} and \autoref{fig:benchmarks}. Experiments were
-performed on a
-\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners#standard-github-hosted-runners-for-public-repositories}{standard GitHub runner}
-without parallelism.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|ll|r|rrr|}
-  \hline
-  maze\_ctor
-          & keyword args           & all sizes 
-                                              & \shortstack{small \\ $g \leq 10$} 
-                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
-                                                                    & \shortstack{large \\ $g > 32$} \\
-  \hline\hline
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
-          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
-          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
-          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
-          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
-  \hline\hline
-  mean
-          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
-  median
-          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
-  \hline
-\end{tabular}
-\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
-\label{tab:benchmarks}
-\end{table}
-
-\begin{figure}
-\hypertarget{fig:benchmarks}{%
-\centering
-\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
-\caption{Plot of maze generation time. Generation time scales
-exponentially with maze size for all algorithms. Generation time per
-maze does not depend on the number of mazes being generated, and there
-is minimal overhead to initializing the generation process for a small
-dataset. Wilson's algorithm is notably less efficient than others and
-has high variance. Note that values are averaged across all parameter
-sets for that algorithm. More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}\label{fig:benchmarks}
-}
-\end{figure}
-
-\hypertarget{sec:success-rate-estimation}{%
-\subsection{Success Rate Estimation}\label{sec:success-rate-estimation}}
-
-In order to replicate the exact dataset distribution of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}), the parameter
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}}
-allows for additional constraints such as enforcing that the start or
-end point be in a ``dead end'' with only one accessible neighbor cell.
-However, combining these constraints with cyclic mazes (such as those
-generated with percolation), as was required for the work in
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), can lead to an absence of valid start and end points. Placing
-theoretical bounds on this success rate is difficult, as it depends on
-the exact maze generation algorithm and parameters used. To deal with
-this, our package provides a way to estimate the success rate of a given
-configuration using a symbolic regression model trained with PySR
-(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
-be found in
-\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
-Using the estimation algorithm simply requires the user to call
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
-providing their initial \texttt{cfg} and then using the returned
-\texttt{cfg\_new} in its place.
-
-\hypertarget{success-rate-estimation-algorithm}{%
-\subsubsection{Success Rate Estimation
-Algorithm}\label{success-rate-estimation-algorithm}}
-
-The base function learned by symbolic regression provides limited
-insight and may be subject to change. It is defined as
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
-and takes a 5 dimensional float vector created by
-\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the
-{[}percolation value, grid size, endpoint deadend configuration,
-endpoint uniqueness, categorical generation function index{]}.
-
-However, the outputs of this function are not directly usable due to
-minor divergences at the endpoints with respect to the percolation
-probability \(p\). Since we know that maze success is either guaranteed
-or impossible for \(p=0\) and \(p=1\), we define the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function to nudge the raw output of the symbolic regression. This
-function is defined with the following components:
-
-shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
-function given by \[
-  \sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}
-  \qquad A(q,a,w) = w \cdot (1 - |2q-1|^a)
-\] \[
-  h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)
-\]
-
-We combine these to get the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function, which is identity-like for \(p \approx 0.5\), and pushes \(x\)
-to extremes otherwise. \[
-  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
-\]
-
-Finally, we define \[
-  \text{cfg\_success\_predict\_fn}(\mathbf{x}) = \text{soft\_step}(\text{raw\_val}, x_0, 5, 10)
-\]
-
-where \texttt{raw\_val} is the output of the symbolic regression model.
-The parameter \(x_0\) is the percolation probability, while all other
-parameters from \texttt{\_to\_ps\_array()} only affect
-\texttt{raw\_val}.
-
-\begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
-\caption{An example of both empirical and predicted success rates as a
-function of the percolation probability \(p\) for various maze sizes,
-percolation with and without depth first search, and
-\texttt{endpoint\_kwargs} requiring that both the start and end be in
-unique dead ends. Empirical measures derived from a sample of 128 mazes.
-More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}
-\end{figure}
-
-\hypertarget{sec:implementation}{%
-\section{Implementation}\label{sec:implementation}}
-
-We refer to our
-\href{https://github.com/understanding-search/maze-dataset}{repository}
-and
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
-for documentation and up-to-date implementation details.
-
-This package utilizes a simple, efficient representation of mazes as
-subgraphs of a finite lattice, which we call a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-Using an adjacency matrix for storing mazes would be memory inefficient
-by failing to exploit the highly sparse structure -- for example, for a
-2-dimensional maze, only 4 of the diagonals would be filled in. On the
-other hand, using an adjacency list could lead to a poor lookup time for
-whether any given connection exists.
-
-Instead, we describe mazes with the following representation: for a
-\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
-initialize a boolean array \[
-  A = \{0, 1\}^{2 \times r \times c}
-\] which we refer to in the code as a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
-The value at \(A[0,i,j]\) determines whether a \emph{downward}
-connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
-value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
-to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
-existence of connections and facilitate fast lookup time, at the cost of
-requiring additional care with indexing. Note that this setup allows for
-a periodic lattice. Generation of mazes is detailed in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
-
-To produce solutions to mazes, two points are selected uniformly at
-random without replacement from the connected component of the maze, and
-the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
-1968}) is applied to find the shortest path between them. The endpoint
-selection can be controlled via
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
-and complications caused by this are detailed in
-\hyperref[sec:success-rate-estimation]{section: \textit{\nameref{sec:success-rate-estimation}}}.
-A maze with a solution is denoted a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}},
-which inherits from
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-
-Parallelization is implemented via the \texttt{multiprocessing} module
-in the Python standard library, and parallel generation can be
-controlled via keyword arguments to
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}.
-
-\newpage
-
-\hypertarget{usage-in-research}{%
-\section{Usage in Research}\label{usage-in-research}}
-
-This package was originally built for the needs of the
-(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
-et al., 2023b}) project, which aims to investigate spatial planning and
-world models in autoregressive transformer models trained on mazes
-(\protect\hyperlink{ref-ivanitskiy2023structuredworldreps}{Michael
-Igorevich Ivanitskiy, Spies, et al., 2023};
-\protect\hyperlink{ref-maze-dataset-arxiv-2023}{Michael Igorevich
-Ivanitskiy, Shah, et al., 2023};
-\protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al., 2024}).
-It was extended for work on understanding the mechanisms by which
-recurrent convolutional and implicit networks
-(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
-given a rasterized view
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), which required matching the pixel-padded and endpoint
-constrained output format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). Ongoing work using \texttt{maze-dataset} aims to
-investigate the effects of varying the tokenization format on the
-performance of pretrained LLMs on spatial reasoning.
-
-This package has also been utilized in work by other groups:
-
-\begin{itemize}
-\item
-  By (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) to
-  compare the effectiveness of transformers trained with the
-  MLM-\(\mathcal{U}\)
-  (\protect\hyperlink{ref-MLMU-kitouni2024factorization}{Kitouni et al.,
-  2024}) multistep prediction objective against standard autoregressive
-  training for multi-step planning on our maze task.
-\item
-  By (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) to
-  study the effectiveness of imperative learning.
-\item
-  By (\protect\hyperlink{ref-zhang2025tscend}{Zhang et al., 2025}) to
-  introduce a novel framework for reasoning diffusion models.
-\item
-  By (\protect\hyperlink{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to
-  improve spatial reasoning in LLMs with GRPO.
-\end{itemize}
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-This work was partially funded by National Science Foundation awards
-DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
-for hosting authors MII, AFS, and TR for a residency visit, and to
-various members of FAR's technical staff for their advice.
-
-This work was partially supported by AI Safety Camp and AI Safety
-Support, which also brought many of the authors together. We would like
-to thank our former collaborators at AI Safety Camp and other users and
-contributors to the \texttt{maze-dataset} package: Benji Berczi,
-Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris Mathwin, Lucia
-Quirke, Can Rager, Adrians Skapars, Rusheb Shah, Johannes Treutlein, and
-Dan Valentine.
-
-We thank the Mines Optimization and Deep Learning group (MODL) for
-fruitful discussions. We also thank Michael Rosenberg for recommending
-the usage of Finite State Transducers for storing tokenizer validation
-information.
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0.5}
-\leavevmode\vadjust pre{\hypertarget{ref-mazegenerator-net}{}}%
-Alance AB. (2019). \emph{Maze generator}.
-\url{http://www.mazegenerator.net}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
-Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
-1.0: A complete set of tools to prepare, present, and analyze
-navigational and spatial cognitive neuroscience experiments.
-\emph{Behavior Research Methods}, \emph{40}, 353--359.
-
-\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
-Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
-{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
-Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
-procedural generation to benchmark reinforcement learning. \emph{arXiv
-Preprint arXiv:1912.01588}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pysr}{}}%
-Cranmer, M. (2023). Interpretable machine learning for science with PySR
-and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
-Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
-models' spatial intelligence via GRPO. \emph{arXiv Preprint
-arXiv:2502.14669}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation}{}}%
-Duminil-Copin, H. (2017). \emph{Sixty years of percolation} (No.
-arXiv:1712.04651). {arXiv}. \url{http://arxiv.org/abs/1712.04651}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Ehsan_2022}{}}%
-Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation-clustersize}{}}%
-Fisher, M. E., \& Essam, J. W. (2004). Some {Cluster Size} and
-{Percolation Problems}. \emph{Journal of Mathematical Physics},
-\emph{2}(4), 609--619. \url{https://doi.org/10.1063/1.1703745}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
-Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
-(2022). Jfb: Jacobian-free backpropagation for implicit networks.
-\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
-\emph{36}, 6648--6656.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
-Guo, C., Barthelet, L., \& Morris, R. (2011). \emph{Maze generator and
-solver}. Wolfram Demonstrations Project,
-\url{https://demonstrations.wolfram.com/MazeGeneratorAndSolver/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
-Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
-{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
-Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
-{Games CoG}}, 1--4.
-
-\leavevmode\vadjust pre{\hypertarget{ref-A_star}{}}%
-Hart, P. E., Nilsson, N. J., \& Raphael, B. (1968). A {Formal Basis} for
-the {Heuristic Determination} of {Minimum Cost Paths}. \emph{IEEE
-Transactions on Systems Science and Cybernetics}, \emph{4}(2), 100--107.
-\url{https://doi.org/10.1109/TSSC.1968.300136}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
-Ivanitskiy, M. (n.d.). \emph{ZANJ}.
-\url{https://github.com/mivanit/ZANJ}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
-\emph{Maze dataset}.
-\url{https://github.com/understanding-search/maze-dataset}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
-\emph{Maze transformer interpretability}.
-\url{https://github.com/understanding-search/maze-transformer}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
-Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
-Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
-C. D., \& others. (2023). A configurable library for generating and
-manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
-Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
-Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
-D., \& others. (2023). Structured world representations in maze-solving
-transformers. \emph{arXiv Preprint arXiv:2312.02566}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
-Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
-Ibrahim, M. (2024). The factorization curse: Which tokens you predict
-underlie the reversal curse and more. \emph{Advances in Neural
-Information Processing Systems}, \emph{37}, 112329--112355.
-
-\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
-Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
-Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
-extrapolation for mazes with recurrent and implicit networks.
-\emph{arXiv Preprint arXiv:2410.03020}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruskal1956shortest}{}}%
-Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
-the traveling salesman problem. \emph{Proceedings of the American
-Mathematical Society}, \emph{7}(1), 48--50.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
-Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
-Performance insights and comparative analysis. \emph{arXiv Preprint
-arXiv:2308.11224}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
-Nag, A. (2020). MDL suite: A language, generator and compiler for
-describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
-1815.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Nemeth_2019}{}}%
-Németh, F. (2019). \emph{Maze-generation-algorithms}.
-\url{https://github.com/ferenc-nemeth/maze-generation-algorithms}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
-Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
-Transformers can navigate mazes with multi-step prediction. \emph{arXiv
-Preprint arXiv:2412.05117}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh-oppenheimj2018maze}{}}%
-Oppenheim, J. (2018). \emph{Maze-generator: Generate a random maze
-represented as a 2D array using depth-first search}.
-\url{https://github.com/oppenheimj/maze-generator/}; GitHub.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
-Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G.,
-Killeen, T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., Kopf,
-A., Yang, E., DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S.,
-Steiner, B., Fang, L., \ldots{} Chintala, S. (2019). {PyTorch: An
-Imperative Style, High-Performance Deep Learning Library}. In H.
-Wallach, H. Larochelle, A. Beygelzimer, F. d'Alché-Buc, E. Fox, \& R.
-Garnett (Eds.), \emph{Advances in neural information processing systems
-32} (pp. 8024--8035). Curran Associates, Inc.
-\url{http://papers.neurips.cc/paper/9015-pytorch-an-imperative-style-high-performance-deep-learning-library.pdf}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
-Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
-Hidden computation in transformer language models. \emph{arXiv Preprint
-arXiv:2404.15758}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
-Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
-transparent ai: A survey on interpreting the inner structures of deep
-neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
-Machine Learning (SaTML)}, 464--483.
-
-\leavevmode\vadjust pre{\hypertarget{ref-easy_to_hard}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Bansal, A., Emam, Z., Huang,
-F., Goldblum, M., \& Goldstein, T. (2021). \emph{Datasets for {Studying
-Generalization} from {Easy} to {Hard Examples}} (No. arXiv:2108.06011).
-{arXiv}. \url{https://doi.org/10.48550/arXiv.2108.06011}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
-Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
-Generalizing from easy to hard problems with recurrent networks.
-\emph{Advances in Neural Information Processing Systems}, \emph{34},
-6695--6706.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
-Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
-\emph{arXiv Preprint arXiv:2308.02522}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
-Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
-Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
-world models in maze-solving tasks. \emph{arXiv Preprint
-arXiv:2412.11867}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
-Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
-Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
-self-supervised neural-symbolic learning framework for robot autonomy.
-\emph{arXiv Preprint arXiv:2406.16087}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wilson}{}}%
-Wilson, D. B. (1996). Generating random spanning trees more quickly than
-the cover time. \emph{Proceedings of the Twenty-Eighth Annual {ACM}
-Symposium on {Theory} of Computing - {STOC} '96}, 296--303.
-\url{https://doi.org/10.1145/237814.237880}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
-Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
-scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
-arXiv:2502.01989}.
-
-\end{CSLReferences}
-
-\end{document}

### 2025-04-26 docs/paper/figures/tex/tab1_benchmarks.tex
move figure code to external files

\begin{table}[H]
\centering
\begin{tabular}{|ll|r|rrr|}
	\hline
	maze\_ctor
			& keyword args           & all sizes 
												& \shortstack{small \\ $g \leq 10$} 
															& \shortstack{medium \\ $g \in (10, 32]$} 
																	& \shortstack{large \\ $g > 32$} \\
	\hline\hline
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
			&                        &   28.0   &    2.8   &   20.3   &  131.8   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
			& accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
			& do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
			& max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
			& p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
			& p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
			&                        &   12.8   &    1.9   &   10.3   &   55.8   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
			& p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
			&                        &   10.2   &    1.7   &    8.9   &   42.1   \\
	\docslink{maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
			&                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
	\hline\hline
	mean
			&                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
	median
			&                        &   11.1   &    6.5   &   32.9   &  302.7   \\
	\hline
\end{tabular}
\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docslink{benchmarks}{benchmarks page}.}
\label{tab:benchmarks}
\end{table}

### 2025-04-26 docs/paper/paper.tex
move figure code to external files

--- 2025-04-09 docs/paper/paper.tex
@@
 
-\begin{figure} 
-  \begin{minipage}{5in}
-    \input{diagram/diagram.tikz} 
-  \end{minipage}
-  \caption{
-    Usage of maze-dataset.
-We create a \texttt{MazeDataset} from a \texttt{MazeDatasetConfig}.
-This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats.
-Code in the image contains clickable links to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{documentation}.
-A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.
-  }
-  \label{fig:diagram}
-\end{figure}
+\input{figures/tex/fig1_diagram.tex}
 
@@
 
-\begin{figure}[H]
-  \centering
-  \begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
-    \hline \\[.5em]
-    % algorithms
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
-    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
-    % descriptions
-      Simple text format for displaying mazes, useful for debugging in a terminal environment.
-      & \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
-      & feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
-    \hline \\
-    % examples
-      \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
-        \setlength{\baselineskip}{0.9em}
-        \input{figures/outputs-ascii-colored.tex} 
-      \end{minipage}}
-      & \multicolumn{1}{c}{
-        \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
-      }
-      & \multicolumn{1}{c}{
-        \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
-      } \\[1em]
-
-    \hline \\
-  \end{tabular}
-  \caption{Various output formats. Top row (left to right): ASCII diagram, rasterized pixel grid, and advanced display tool.}
-  \label{fig:output-fmts}
-\end{figure}
+\input{figures/tex/fig2_formats.tex}
 
@@
 
-\begin{figure} 
-  \centering
-  \begin{minipage}{5in}
-    \footnotesize
-    \input{figures/outputs-tokens-colored.tex}
-  \end{minipage}
-  \caption{
-    Example text output format with token regions highlighted.
-    \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the graph,
-    \colorbox[RGB]{ 217,234,211 }{Origin}: starting coordinate,
-    \colorbox[RGB]{ 234,209,220 }{Target}: ending coordinate,
-    \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence
-  }
-  \label{fig:token-regions}
-\end{figure}
+\input{figures/tex/fig4_tokenfmt.tex}
 
@@
 
-\begin{figure}
-    \centering
-    \input{figures/TokenizerElement_structure.tikz}
-    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular}.}
-\end{figure}
+\input{figures/tex/fig5_mmt.tex}
 
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{|ll|r|rrr|}
-  \hline
-  maze\_ctor
-          & keyword args           & all sizes 
-                                              & \shortstack{small \\ $g \leq 10$} 
-                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
-                                                                    & \shortstack{large \\ $g > 32$} \\
-  \hline\hline
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
-          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
-          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
-          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
-          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
-          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
-          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
-  \hline\hline
-  mean
-          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
-  median
-          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
-  \hline
-\end{tabular}
-\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
-\label{tab:benchmarks}
-\end{table}
+\input{figures/tex/tab1_benchmarks.tex}
 

### 2025-04-26 docs/paper/paper.tex
move acknowledgements

@@
 is constructed from a set of several
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
 objects, each of which specifies how different token regions or other
@@
 forking the library may simply create their own
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
 subclass and implement the abstract methods. If the behavior change is
 sufficiently small, simply adding a parameter to an existing
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
 subclass and updating its implementation will suffice.
@@
 is provided which can be applied to the unneeded
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
 subclasses to prune those features and compact the available space of
@@
 
-This work was partially funded by National Science Foundation awards
-DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
-for hosting authors MII, AFS, and TR for a residency visit, and to
-various members of FAR's technical staff for their advice.
-
-This work was partially supported by AI Safety Camp and AI Safety
-Support, which also brought many of the authors together. We would like
-to thank our former collaborators at AI Safety Camp and other users and
-contributors to the \texttt{maze-dataset} package: Benji Berczi,
-Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris Mathwin, Lucia
-Quirke, Can Rager, Adrians Skapars, Rusheb Shah, Johannes Treutlein, and
-Dan Valentine.
-
-We thank the Mines Optimization and Deep Learning group (MODL) for
-fruitful discussions. We also thank Michael Rosenberg for recommending
-the usage of Finite State Transducers for storing tokenizer validation
-information.
+\input{figures/tex/acknowledgements.tex}
 

### 2025-04-26 docs/paper/figures/tex/fig7_sre.tex
cut down impl and sre sections, better links

--- 2025-04-26 docs/paper/paper.tex
@@
-\documentclass[10pt,a4paper,onecolumn]{article}
-\usepackage{marginnote}
-\usepackage{graphicx}
-\usepackage[rgb,svgnames]{xcolor}
-\usepackage{authblk,etoolbox}
-\usepackage{titlesec}
-\usepackage{calc}
-\usepackage{tikz}
-\usepackage[pdfa]{hyperref}
-\usepackage{hyperxmp}
-\hypersetup{%
-    unicode=true,
-    pdfapart=3,
-    pdfaconformance=B,
-    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility},
-    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
-Fung},
-    pdfpublication={Journal of Open Source Software},
-    pdfpublisher={Open Journals},
-    pdfissn={2475-9066},
-    pdfpubtype={journal},
-    pdfvolumenum={},
-    pdfissuenum={},
-    pdfdoi={N/A},
-    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
-Diniz Behn, Samy Wu Fung},
-    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
-    colorlinks=true,
-    linkcolor=[rgb]{0.0, 0.5, 1.0},
-    citecolor=Blue,
-    urlcolor=[rgb]{0.0, 0.5, 1.0},
-    breaklinks=true
-}
-% https://tex.stackexchange.com/a/535849
-% Create an OutputIntent in order to correctly specify colours
-\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
-\pdfcatalog{%
-  /OutputIntents [
-    <<
-      /Type /OutputIntent
-      /S /GTS_PDFA1
-      /DestOutputProfile \the\pdflastobj\space 0 R
-      /OutputConditionIdentifier (sRGB)
-      /Info (sRGB)
-    >>
-  ]
-}
-\pdfvariable omitcidset=1
-\usepackage{caption}
-\usepackage{orcidlink}
-\usepackage{tcolorbox}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{seqsplit}
-\usepackage{xstring}
-
-\usepackage{float}
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
-
-\usepackage{fixltx2e} % provides \textsubscript
-
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-% --- Page layout -------------------------------------------------------------
-\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
-            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}
-
-% --- Default font ------------------------------------------------------------
-\renewcommand\familydefault{\sfdefault}
-
-% --- Style -------------------------------------------------------------------
-\renewcommand{\captionfont}{\small\sffamily}
-\renewcommand{\captionlabelfont}{\bfseries}
-
-% --- Section/SubSection/SubSubSection ----------------------------------------
-\titleformat{\section}
-  {\normalfont\sffamily\Large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsection}
-  {\normalfont\sffamily\large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsubsection}
-  {\normalfont\sffamily\bfseries}
-  {}{0pt}{}
-\titleformat*{\paragraph}
-  {\sffamily\normalsize}
-
-% --- Header / Footer ---------------------------------------------------------
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-%\renewcommand{\headrulewidth}{0.50pt}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{joss/logo.png}}
-\fancyhead[C]{}
-\fancyhead[R]{}
-\renewcommand{\footrulewidth}{0.25pt}
-
-\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset: Maze Generation with Algorithmic Variety
-and Representational Flexibility. \emph{Journal of Open Source
-Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
-
-\fancyfoot[R]{\sffamily \thepage}
-\makeatletter
-\let\ps@plain\ps@fancy
-\fancyheadoffset[L]{4.5cm}
-\fancyfootoffset[L]{4.5cm}
-
-% --- Macros ---------
-
-\definecolor{linky}{rgb}{0.0, 0.5, 1.0}
-
-\newtcolorbox{repobox}
-   {colback=red, colframe=red!75!black,
-     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}
-
-\newcommand{\ExternalLink}{%
-   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
-       \begin{scope}[x=1ex, y=1ex]
-           \clip (-0.1,-0.1)
-               --++ (-0, 1.2)
-               --++ (0.6, 0)
-               --++ (0, -0.6)
-               --++ (0.6, 0)
-               --++ (0, -1);
-           \path[draw,
-               line width = 0.5,
-               rounded corners=0.5]
-               (0,0) rectangle (1,1);
-       \end{scope}
-       \path[draw, line width = 0.5] (0.5, 0.5)
-           -- (1, 1);
-       \path[draw, line width = 0.5] (0.6, 1)
-           -- (1, 1) -- (1, 0.6);
-       }
-   }
-
-\definecolor{c53baa1}{RGB}{83,186,161}
-\definecolor{c202826}{RGB}{32,40,38}
-\def \rorglobalscale {0.1}
-\newcommand{\rorlogo}{%
-\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
-  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
-    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
-    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364)..
-controls (3.9643, 1.4576) and (4.0508, 1.5081) ..
-(4.1098, 1.5878)..
-controls (4.169, 1.6674) and (4.1984, 1.7642) ..
-(4.1984, 1.8777)..
-controls (4.1984, 1.9719) and (4.182, 2.0503) ..
-(4.1495, 2.1132)..
-controls (4.1169, 2.1762) and (4.0727, 2.2262) ..
-(4.0174, 2.2635)..
-controls (3.9621, 2.3006) and (3.8976, 2.3273) ..
-(3.824, 2.3432)..
-controls (3.7505, 2.359) and (3.6727, 2.367) ..
-(3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688)..
-controls (2.9625, 1.8833) and (2.9572, 1.8976) ..
-(2.9514, 1.9119)..
-controls (2.9083, 2.0164) and (2.848, 2.1056) ..
-(2.7705, 2.1791)..
-controls (2.6929, 2.2527) and (2.6014, 2.3093) ..
-(2.495, 2.3487)..
-controls (2.3889, 2.3881) and (2.2728, 2.408) ..
-(2.1468, 2.408)..
-controls (2.0209, 2.408) and (1.905, 2.3881) ..
-(1.7986, 2.3487)..
-controls (1.6925, 2.3093) and (1.6007, 2.2527) ..
-(1.5232, 2.1791)..
-controls (1.4539, 2.1132) and (1.3983, 2.0346) ..
-(1.3565, 1.9436)..
-controls (1.3504, 2.009) and (1.3351, 2.0656) ..
-(1.3105, 2.1132)..
-controls (1.2779, 2.1762) and (1.2338, 2.2262) ..
-(1.1785, 2.2635)..
-controls (1.1232, 2.3006) and (1.0586, 2.3273) ..
-(0.985, 2.3432)..
-controls (0.9115, 2.359) and (0.8337, 2.367) ..
-(0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364)..
-controls (1.1255, 1.4576) and (1.2121, 1.5081) ..
-(1.2711, 1.5878)..
-controls (1.2737, 1.5915) and (1.2761, 1.5954) ..
-(1.2787, 1.5991)..
-controls (1.2782, 1.5867) and (1.2779, 1.5743) ..
-(1.2779, 1.5616)..
-controls (1.2779, 1.4327) and (1.2996, 1.3158) ..
-(1.3428, 1.2113)..
-controls (1.3859, 1.1068) and (1.4462, 1.0176) ..
-(1.5237, 0.944)..
-controls (1.601, 0.8705) and (1.6928, 0.8139) ..
-(1.7992, 0.7744)..
-controls (1.9053, 0.735) and (2.0214, 0.7152) ..
-(2.1474, 0.7152)..
-controls (2.2733, 0.7152) and (2.3892, 0.735) ..
-(2.4956, 0.7744)..
-controls (2.6016, 0.8139) and (2.6935, 0.8705) ..
-(2.771, 0.944)..
-controls (2.8482, 1.0176) and (2.9086, 1.1068) ..
-(2.952, 1.2113)..
-controls (2.9578, 1.2253) and (2.9631, 1.2398) ..
-(2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777)..
-controls (0.9438, 1.7534) and (0.92, 1.7357) ..
-(0.8911, 1.7243)..
-controls (0.8623, 1.7129) and (0.83, 1.706) ..
-(0.7945, 1.7039)..
-controls (0.7588, 1.7015) and (0.7252, 1.7005) ..
-(0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667)..
-controls (0.7477, 2.0667) and (0.7805, 2.0643) ..
-(0.8139, 2.0598)..
-controls (0.8472, 2.0553) and (0.8768, 2.0466) ..
-(0.9025, 2.0336)..
-controls (0.9282, 2.0206) and (0.9496, 2.0021) ..
-(0.9663, 1.9778)..
-controls (0.9829, 1.9534) and (0.9914, 1.9209) ..
-(0.9914, 1.8799)..
-controls (0.9914, 1.8362) and (0.9819, 1.8021) ..
-(0.9628, 1.7777) -- cycle(2.6125, 1.3533)..
-controls (2.5889, 1.2904) and (2.5553, 1.2359) ..
-(2.5112, 1.1896)..
-controls (2.4672, 1.1433) and (2.4146, 1.1073) ..
-(2.3529, 1.0814)..
-controls (2.2916, 1.0554) and (2.2228, 1.0427) ..
-(2.1471, 1.0427)..
-controls (2.0712, 1.0427) and (2.0026, 1.0557) ..
-(1.9412, 1.0814)..
-controls (1.8799, 1.107) and (1.8272, 1.1433) ..
-(1.783, 1.1896)..
-controls (1.7391, 1.2359) and (1.7052, 1.2904) ..
-(1.6817, 1.3533)..
-controls (1.6581, 1.4163) and (1.6465, 1.4856) ..
-(1.6465, 1.5616)..
-controls (1.6465, 1.6359) and (1.6581, 1.705) ..
-(1.6817, 1.7687)..
-controls (1.7052, 1.8325) and (1.7388, 1.8873) ..
-(1.783, 1.9336)..
-controls (1.8269, 1.9799) and (1.8796, 2.0159) ..
-(1.9412, 2.0418)..
-controls (2.0026, 2.0675) and (2.0712, 2.0804) ..
-(2.1471, 2.0804)..
-controls (2.223, 2.0804) and (2.2916, 2.0675) ..
-(2.3529, 2.0418)..
-controls (2.4143, 2.0161) and (2.467, 1.9799) ..
-(2.5112, 1.9336)..
-controls (2.5551, 1.8873) and (2.5889, 1.8322) ..
-(2.6125, 1.7687)..
-controls (2.636, 1.705) and (2.6477, 1.6359) ..
-(2.6477, 1.5616)..
-controls (2.6477, 1.4856) and (2.636, 1.4163) ..
-(2.6125, 1.3533) -- cycle(3.8015, 1.7777)..
-controls (3.7825, 1.7534) and (3.7587, 1.7357) ..
-(3.7298, 1.7243)..
-controls (3.701, 1.7129) and (3.6687, 1.706) ..
-(3.6333, 1.7039)..
-controls (3.5975, 1.7015) and (3.5639, 1.7005) ..
-(3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667)..
-controls (3.5864, 2.0667) and (3.6192, 2.0643) ..
-(3.6526, 2.0598)..
-controls (3.6859, 2.0553) and (3.7155, 2.0466) ..
-(3.7412, 2.0336)..
-controls (3.7669, 2.0206) and (3.7883, 2.0021) ..
-(3.805, 1.9778)..
-controls (3.8216, 1.9534) and (3.8301, 1.9209) ..
-(3.8301, 1.8799)..
-controls (3.8301, 1.8362) and (3.8206, 1.8021) ..
-(3.8015, 1.7777) -- cycle;
-  \end{scope}
-\end{tikzpicture}
-}
-
-% --- Title / Authors ---------------------------------------------------------
-% patch \maketitle so that it doesn't center
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-% patch \maketitle so that the font size for the title is normal
-\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
-% patch the patch by authblk so that the author block is flush left
-\def\maketitle{{%
-  \renewenvironment{tabular}[2][]
-    {\begin{flushleft}}
-    {\end{flushleft}}
-  \AB@maketitle}}
-\renewcommand\AB@affilsepx{ \protect\Affilfont}
-%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
-\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
-\renewcommand{\affil}[2][]%
-   {\newaffiltrue\let\AB@blk@and\AB@pand
-      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
-        \setcounter{Maxaffil}{0}\fi
-        \begingroup
-        \let\href=\href@Orig
-        \let\protect\@unexpandable@protect
-        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
-        \@temptokena=\expandafter{\AB@authors}%
-        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
-         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
-         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
-         \gdef\AB@las{}\gdef\AB@au@str{}%
-        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
-        \@temptokena=\expandafter{\AB@affillist}%
-        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
-          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
-      \endgroup
-       \let\AB@affilsep\AB@affilsepx
-}
-\makeatother
-\renewcommand\Authfont{\sffamily\bfseries}
-\renewcommand\Affilfont{\sffamily\small\mdseries}
-\setlength{\affilsep}{1em}
-
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{fontspec}
-
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\sffamily]{Ligatures=TeX}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-
-%% Font settings
-\usepackage{fontsetup} % Lazy way to get proper Greek lowercase glyphs
-
-% Use Hack https://sourcefoundry.org/hack/
-\setmonofont{Hack}
-
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\urlstyle{same}  % don't use monospace font for urls
-\ifLuaTeX
-\usepackage[bidi=basic]{babel}
-\else
-\usepackage[bidi=default]{babel}
-\fi
-\babelprovide[main,import]{american}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\tikzset{ % Define a TikZ style for an external hyperlink node.
-  hyperlink node url/.style={
-    alias=sourcenode,
-    append after command={
-      let \p1 = (sourcenode.north west),
-          \p2 = (sourcenode.south east),
-          \n1 = {\x2-\x1},
-          \n2 = {\y1-\y2} in
-      node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
-          {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
-    }
-  }
-}
-\providecommand{\XeTeXLinkBox}[1]{#1}
-\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
-\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
-\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-
-\title{maze-dataset: Maze Generation with Algorithmic Variety and
-Representational Flexibility}
-
-\author[1%
-%
-\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
-  \,\orcidlink{0000-0002-4213-4993}\,%
-}
-\author[4%
-%
-]{Aaron Sandoval%
-  \,\orcidlink{0009-0002-8380-6140}\,%
-}
-\author[2%
-%
-]{Alex F. Spies%
-  \,\orcidlink{0000-0002-8708-1530}\,%
-}
-\author[3%
-%
-]{Tilman Räuker%
-  \,\orcidlink{0009-0009-6321-4413}\,%
-}
-\author[1%
-%
-]{Brandon Knutson%
-  \,\orcidlink{0009-0004-8413-0239}\,%
-}
-\author[1%
-%
-]{Cecilia Diniz Behn%
-  \,\orcidlink{0000-0002-8078-5105}\,%
-}
-\author[1%
-%
-]{Samy Wu Fung%
-  \,\orcidlink{0000-0002-2926-4582}\,%
-}
-
-\affil[1]{Colorado School of Mines, Department of Applied Mathematics
-and Statistics%
-}
-\affil[2]{Imperial College London%
-}
-\affil[3]{UnSearch.org%
-}
-\affil[4]{Independent%
-}
-\affil[$\mathparagraph$]{Corresponding author}
-\date{\vspace{-2.5ex}}
-
-\begin{document}
-\maketitle
-
-\marginpar{
-
-  \begin{flushleft}
-  %\hrule
-  \sffamily\small
-
-  {\bfseries DOI:} \href{https://doi.org/N/A}{\color{linky}{N/A}}
-
-  \vspace{2mm}
-    {\bfseries Software}
-  \begin{itemize}
-    \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{\color{linky}{Review}} \ExternalLink
-    \item \href{https://github.com/openjournals}{\color{linky}{Repository}} \ExternalLink
-    \item \href{https://doi.org/10.5281}{\color{linky}{Archive}} \ExternalLink
-  \end{itemize}
-
-  \vspace{2mm}
-
-    \par\noindent\hrulefill\par
-
-  \vspace{2mm}
-
-  {\bfseries Editor:} \href{https://joss.theoj.org}{Open
-Journals} \ExternalLink
-   \\
-  \vspace{1mm}
-    {\bfseries Reviewers:}
-  \begin{itemize}
-  \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals}{@openjournals}
-    \end{itemize}
-    \vspace{2mm}
-
-    {\bfseries Submitted:} 01 January 1970\\
-    {\bfseries Published:} 01 January 1970
-
-  \vspace{2mm}
-  {\bfseries License}\\
-  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{https://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).
-
-  \end{flushleft}
-}
-
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
-
-Solving mazes is a classic problem in computer science and artificial
-intelligence, and humans have been constructing mazes for thousands of
-years. Although finding the shortest path through a maze is a solved
-problem, this very fact makes it an excellent testbed for studying how
-machine learning algorithms solve problems and represent spatial
-information. We introduce \texttt{maze-dataset}, a user-friendly Python
-library for generating, processing, and visualizing datasets of mazes.
-This library supports a variety of maze generation algorithms providing
-mazes with or without loops, mazes that are connected or not, and many
-other variations. These generation algorithms can be configured with
-various parameters, and the resulting mazes can be filtered to satisfy
-desired properties. Also provided are tools for converting mazes to and
-from various formats suitable for a variety of neural network
-architectures, such as rasterized images, tokenized text sequences, and
-various visualizations. As well as providing a simple interface for
-generating, storing, and loading these datasets, \texttt{maze-dataset}
-is extensively tested, type hinted, benchmarked, and documented.
-
-\input{figures/tex/fig1_diagram.tex}
-
-\hypertarget{statement-of-need}{%
-\section{Statement of Need}\label{statement-of-need}}
-
-While maze generation itself is straightforward, the architectural
-challenge comes from building a system supporting many algorithms with
-configurable parameters, property filtering, and representation
-transformation. This library aims to greatly streamline the process of
-generating and working with datasets of mazes that can be described as
-subgraphs of an \(n \times n\) lattice with boolean connections and,
-optionally, start and end points that are nodes in the graph.
-Furthermore, we place emphasis on a wide variety of possible text output
-formats aimed at evaluating the spatial reasoning capabilities of Large
-Language Models (LLMs) and other text-based transformer models.
-
-For interpretability and behavioral research, algorithmic tasks offer
-benefits by allowing systematic data generation and task decomposition,
-as well as simplifying the process of circuit discovery
-(\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
-Although mazes are well suited for these investigations, we found that
-existing maze generation packages
-(\protect\hyperlink{ref-cobbe2019procgen}{Cobbe et al., 2019};
-\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-harriesMazeExplorerCustomisable3D2019}{Harries et
-al., n.d.}; \protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) lack support for transforming between multiple
-representations and provide limited control over the maze generation
-process.
-
-\hypertarget{related-works}{%
-\subsection{Related Works}\label{related-works}}
-
-A multitude of public and open-source software packages exist for
-generating mazes (\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). However, nearly all of these packages produce
-mazes represented as rasterized images or other visual formats rather
-than the underlying graph structure, and this makes it difficult to work
-with these datasets.
-
-\begin{itemize}
-\item
-  Most prior works provide mazes in visual or raster formats, and we
-  provide a variety of similar output formats:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
-    utilizing
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
-    which can exactly mimic the outputs provided in
-    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
-    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
-    similar to the outputs of Németh
-    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    provides a format similar to
-    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
-    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
-  \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
-    provides a feature-rich plotting utility with support for multiple
-    paths, heatmaps over positions, and more. This is similar to the
-    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
-    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
-    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
-  \end{itemize}
-\item
-  The text format provided by
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
-  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
-  Wu, 2023}), but provides over 5.8 million unique formats for
-  converting mazes to a text stream, detailed in
-  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
-\item
-  For rigorous investigations of the response of a model to various
-  distributional shifts, preserving metadata about the generation
-  algorithm with the dataset itself is essential. To this end, our
-  package efficiently stores the dataset along with its metadata in a
-  single human-readable file (\protect\hyperlink{ref-zanj}{M.
-  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
-  this reliably.
-\item
-  Storing mazes as images is not only difficult to work with, but also
-  inefficient. We use a highly efficient method detailed in
-  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
-\item
-  Our package is easily installable with source code freely available.
-  It is extensively tested, type hinted, benchmarked, and documented.
-  Many other maze generation packages lack this level of rigor and
-  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
-  2008}) appear to simply no longer be accessible.
-\end{itemize}
-
-\hypertarget{features}{%
-\section{Features}\label{features}}
-
-We direct readers to our
-\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples},
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs},
-and
-\href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
-for more information.
-
-\hypertarget{generation}{%
-\subsection{Generation and Basic Usage}\label{generation}}
-
-Our package can be installed from
-\href{https://pypi.org/project/maze-dataset/}{PyPi} via
-\texttt{pip\ install\ maze-dataset}, or directly from the
-\href{https://github.com/understanding-search/maze-dataset}{git
-repository} (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
-Ivanitskiy et al., 2023a}).
-
-To create a dataset, we first create a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-configuration object, which specifies the seed, number, and size of
-mazes, as well as the generation algorithm and its corresponding
-parameters. This object is passed to a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-class to create a dataset. Crucially, this
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-mimics the interface of a PyTorch
-(\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
-\href{https://pytorch.org/docs/stable/data.html}{\texttt{Dataset}}, and
-can thus be easily incorporated into existing data pre-processing and
-training pipelines, e.g., through the use of a \texttt{DataLoader}
-class.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
-\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
-\NormalTok{)}
-\CommentTok{\# create a config}
-\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
-\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
-\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
-\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
-\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
-    \CommentTok{\# (optional) algorithm{-}specific parameters}
-\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
-    \CommentTok{\# (optional) many options for restricting start/end points}
-\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
-\NormalTok{)}
-\CommentTok{\# create a dataset}
-\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
-\NormalTok{  cfg, }\CommentTok{\# pass the config}
-\NormalTok{  ..., }\CommentTok{\# other options for disk loading, parallelization, etc.}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-When initializing a dataset, options which do not affect the mazes
-themselves can be specified through the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
-factory method as necessary. These options allow for saving/loading
-existing datasets instead of re-generating, parallelization options for
-generation, and more. Available maze generation algorithms are static
-methods of the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
-namespace class and include generation algorithms based on randomized
-depth-first search, Wilson's algorithm
-(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
-(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
-\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
-Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
-others.
-
-Furthermore, a dataset of mazes can be filtered to satisfy certain
-properties. Custom filters can be specified, and some filters are
-included in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
-For example, we can require a minimum path length of three steps from
-the origin to the target:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dataset\_filtered: MazeDataset }\OperatorTok{=}\NormalTok{ dataset.filter\_by.path\_length(min\_length}\OperatorTok{=}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-All implemented maze generation algorithms are stochastic by nature. For
-reproducibility, the \texttt{seed} parameter of
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-may be set. In practice, using provided deduplication filters, we find
-that exact duplicate mazes are generated very infrequently, even when
-generating very large datasets.
-
-For use cases where mazes of different sizes, generation algorithms, or
-other parameter variations are required, we provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/collected_dataset.html\#MazeDatasetCollection}{\texttt{MazeDatasetCollection}}
-class, which allows for creating a single iterable dataset from multiple
-independent configurations.
-
-\hypertarget{visual-output-formats}{%
-\subsection{Visual Output Formats}\label{visual-output-formats}}
-
-Internally, mazes are
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
-objects, which have path information and a tensor optimized for storing
-sub-graphs of a lattice. These objects can be converted to and from
-several formats to maximize their utility in different contexts.
-
-\input{figures/tex/fig2_formats.tex}
-
-In previous work, maze tasks have been used with Recurrent Convolutional
-Neural Network (RCNN) derived architectures
-(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
-Huang, et al., 2021}). To facilitate the use of our package in this
-context, we replicate the format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) and provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
-class which returns rasterized pairs of (input, target) mazes as shown
-in \autoref{fig:e2h-raster} below.
-
-\begin{figure}
-\hypertarget{fig:e2h-raster}{%
-\centering
-\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
-\caption{Input is the rasterized maze without the path marked (left),
-and provide as a target the maze with all but the correct path removed
-(right). Configuration options exist to adjust whether endpoints are
-included and if empty cells should be filled in.}\label{fig:e2h-raster}
-}
-\end{figure}
-
-\hypertarget{sec:tokenized-output-formats}{%
-\subsection{Tokenized Output
-Formats}\label{sec:tokenized-output-formats}}
-
-Autoregressive transformer models can be quite sensitive to the exact
-format of input data, and may even use delimiter tokens to perform
-reasoning steps (\protect\hyperlink{ref-pfau2024dotbydot}{Pfau et al.,
-2024}; \protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al.,
-2024}). To facilitate systematic investigation of the effects of
-different representations of data on text model performance, we provide
-a variety of tokenized text output formats.
-
-We convert mazes to token sequences in two steps. First, the maze is
-stringified using
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens()}}.
-The
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class provides a powerful interface for configuring maze stringification
-behavior. Second, the sequence of strings is tokenized into integers
-using \texttt{encode()}. Tokenization uses a fixed vocabulary for
-simplicity. Mazes up to \(50 \times 50\) are supported when using a
-unique token for each position, and up to \(128 \times 128\) are
-supported when positions in the maze are represented as a pair of
-coordinates.
-
-There are many algorithms by which one might tokenize a 2D maze into a
-1D format usable by autoregressive text models. Training multiple models
-on the encodings output from each of these algorithms may produce very
-different internal representations, learned solution algorithms, and
-levels of performance. To allow exploration of how different maze
-tokenization algorithms affect these models, the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class contains a rich set of options to customize how mazes are
-stringified. This class contains 19 discrete parameters, resulting in
-over 5.8 million unique tokenizers. There are 6 additional parameters
-available whose functionality is not verified via automated testing, but
-further expand the the number of tokenizers by a factor of \(44/3\) to
-86 million.
-
-All output sequences consist of four token regions representing
-different features of the maze; an example output sequence is shown in
-\autoref{fig:token-regions}.
-
-\input{figures/tex/fig4_tokenfmt.tex}
-
-Each
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-is constructed from a set of several
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-objects, each of which specifies how different token regions or other
-elements of the stringification are produced.
-
-\input{figures/tex/fig5_mmt.tex}
-
-The tokenizer architecture is purposefully designed such that adding and
-testing a wide variety of new tokenization algorithms is fast and
-minimizes disturbances to functioning code. This is enabled by the
-modular architecture and the automatic inclusion of any new tokenizers
-in integration tests. To create a new variety of tokenizer, developers
-forking the library may simply create their own
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and implement the abstract methods. If the behavior change is
-sufficiently small, simply adding a parameter to an existing
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and updating its implementation will suffice.
-
-The breadth of tokenizers is also easily scaled in the opposite
-direction. Due to the exponential scaling of parameter combinations,
-adding a small number of new features can significantly slow certain
-procedures which rely on constructing all possible tokenizers, such as
-integration tests. If any existing subclass contains features which
-aren't needed, a developer tool decorator
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
-is provided which can be applied to the unneeded
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclasses to prune those features and compact the available space of
-tokenizers.
-
-\hypertarget{benchmarks}{%
-\subsection{Benchmarks of Generation Speed}\label{benchmarks}}
-
-We provide approximate benchmarks for relative generation time across
-various algorithms, parameter choices, maze sizes, and dataset sizes in
-\autoref{tab:benchmarks} and \autoref{fig:benchmarks}. Experiments were
-performed on a
-\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners#standard-github-hosted-runners-for-public-repositories}{standard GitHub runner}
-without parallelism.
-
-\input{figures/tex/tab1_benchmarks.tex}
-
-\begin{figure}
-\hypertarget{fig:benchmarks}{%
-\centering
-\includegraphics[width=0.9\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
-\caption{Plot of maze generation time. Generation time scales
-exponentially with maze size for all algorithms. Generation time per
-maze does not depend on the number of mazes being generated, and there
-is minimal overhead to initializing the generation process for a small
-dataset. Wilson's algorithm is notably less efficient than others and
-has high variance. Note that values are averaged across all parameter
-sets for that algorithm. More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}\label{fig:benchmarks}
-}
-\end{figure}
-
-\hypertarget{sec:success-rate-estimation}{%
-\subsection{Success Rate Estimation}\label{sec:success-rate-estimation}}
-
-In order to replicate the exact dataset distribution of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}), the parameter
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}}
-allows for additional constraints such as enforcing that the start or
-end point be in a ``dead end'' with only one accessible neighbor cell.
-However, combining these constraints with cyclic mazes (such as those
-generated with percolation), as was required for the work in
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), can lead to an absence of valid start and end points. Placing
-theoretical bounds on this success rate is difficult, as it depends on
-the exact maze generation algorithm and parameters used. To deal with
-this, our package provides a way to estimate the success rate of a given
-configuration using a symbolic regression model trained with PySR
-(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
-be found in
-\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
-Using the estimation algorithm simply requires the user to call
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
-providing their initial \texttt{cfg} and then using the returned
-\texttt{cfg\_new} in its place.
-
-\hypertarget{success-rate-estimation-algorithm}{%
-\subsubsection{Success Rate Estimation
-Algorithm}\label{success-rate-estimation-algorithm}}
-
-The base function learned by symbolic regression provides limited
-insight and may be subject to change. It is defined as
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
-and takes a 5 dimensional float vector created by
-\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the
-{[}percolation value, grid size, endpoint deadend configuration,
-endpoint uniqueness, categorical generation function index{]}.
-
-However, the outputs of this function are not directly usable due to
-minor divergences at the endpoints with respect to the percolation
-probability \(p\). Since we know that maze success is either guaranteed
-or impossible for \(p=0\) and \(p=1\), we define the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function to nudge the raw output of the symbolic regression. This
-function is defined with the following components:
-
-shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
-function given by \[
-  \sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}
-  \qquad A(q,a,w) = w \cdot (1 - |2q-1|^a)
-\] \[
-  h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)
-\]
-
-We combine these to get the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function, which is identity-like for \(p \approx 0.5\), and pushes \(x\)
-to extremes otherwise. \[
-  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
-\]
-
-Finally, we define \[
-  \text{cfg\_success\_predict\_fn}(\mathbf{x}) = \text{soft\_step}(\text{raw\_val}, x_0, 5, 10)
-\]
-
-where \texttt{raw\_val} is the output of the symbolic regression model.
-The parameter \(x_0\) is the percolation probability, while all other
-parameters from \texttt{\_to\_ps\_array()} only affect
-\texttt{raw\_val}.
-
 \begin{figure}
@@
 \end{figure}
-
-\hypertarget{sec:implementation}{%
-\section{Implementation}\label{sec:implementation}}
-
-We refer to our
-\href{https://github.com/understanding-search/maze-dataset}{repository}
-and
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
-for documentation and up-to-date implementation details.
-
-This package utilizes a simple, efficient representation of mazes as
-subgraphs of a finite lattice, which we call a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-Using an adjacency matrix for storing mazes would be memory inefficient
-by failing to exploit the highly sparse structure -- for example, for a
-2-dimensional maze, only 4 off-diagonal bands would be have nonzero
-values. On the other hand, using an adjacency list could lead to a poor
-lookup time for whether any given connection exists.
-
-Instead, we describe mazes with the following representation: for a
-\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
-initialize a boolean array \[
-  A = \{0, 1\}^{2 \times r \times c}
-\] which we refer to in the code as a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
-The value at \(A[0,i,j]\) determines whether a \emph{downward}
-connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
-value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
-to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
-existence of connections and facilitate fast lookup time, at the cost of
-requiring additional care with indexing. Note that this setup allows for
-a periodic lattice. Generation of mazes is detailed in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
-
-To produce solutions to mazes, two points are selected uniformly at
-random without replacement from the connected component of the maze, and
-the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
-1968}) is applied to find the shortest path between them. The endpoint
-selection can be controlled via
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
-and complications caused by this are detailed in
-\hyperref[sec:success-rate-estimation]{section: \textit{\nameref{sec:success-rate-estimation}}}.
-A maze with a solution is denoted a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}},
-which inherits from
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-
-Parallelization is implemented via the \texttt{multiprocessing} module
-in the Python standard library, and parallel generation can be
-controlled via keyword arguments to
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}.
-
-\newpage
-
-\hypertarget{usage-in-research}{%
-\section{Usage in Research}\label{usage-in-research}}
-
-This package was originally built for the needs of the
-(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
-et al., 2023b}) project, which aims to investigate spatial planning and
-world models in autoregressive transformer models trained on mazes
-(\protect\hyperlink{ref-ivanitskiy2023structuredworldreps}{Michael
-Igorevich Ivanitskiy, Spies, et al., 2023};
-\protect\hyperlink{ref-maze-dataset-arxiv-2023}{Michael Igorevich
-Ivanitskiy, Shah, et al., 2023};
-\protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al., 2024}).
-It was extended for work on understanding the mechanisms by which
-recurrent convolutional and implicit networks
-(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
-given a rasterized view
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), which required matching the pixel-padded and endpoint
-constrained output format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). Ongoing work using \texttt{maze-dataset} aims to
-investigate the effects of varying the tokenization format on the
-performance of pretrained LLMs on spatial reasoning.
-
-This package has also been utilized in work by other groups:
-
-\begin{itemize}
-\item
-  By (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) to
-  compare the effectiveness of transformers trained with the
-  MLM-\(\mathcal{U}\)
-  (\protect\hyperlink{ref-MLMU-kitouni2024factorization}{Kitouni et al.,
-  2024}) multistep prediction objective against standard autoregressive
-  training for multi-step planning on our maze task.
-\item
-  By (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) to
-  study the effectiveness of imperative learning.
-\item
-  By (\protect\hyperlink{ref-zhang2025tscend}{Zhang et al., 2025}) to
-  introduce a novel framework for reasoning diffusion models.
-\item
-  By (\protect\hyperlink{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to
-  improve spatial reasoning in LLMs with GRPO.
-\end{itemize}
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-\input{figures/tex/acknowledgements.tex}
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0.5}
-\leavevmode\vadjust pre{\hypertarget{ref-mazegenerator-net}{}}%
-Alance AB. (2019). \emph{Maze generator}.
-\url{http://www.mazegenerator.net}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
-Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
-1.0: A complete set of tools to prepare, present, and analyze
-navigational and spatial cognitive neuroscience experiments.
-\emph{Behavior Research Methods}, \emph{40}, 353--359.
-
-\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
-Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
-{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
-Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
-procedural generation to benchmark reinforcement learning. \emph{arXiv
-Preprint arXiv:1912.01588}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pysr}{}}%
-Cranmer, M. (2023). Interpretable machine learning for science with PySR
-and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
-Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
-models' spatial intelligence via GRPO. \emph{arXiv Preprint
-arXiv:2502.14669}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation}{}}%
-Duminil-Copin, H. (2017). \emph{Sixty years of percolation} (No.
-arXiv:1712.04651). {arXiv}. \url{http://arxiv.org/abs/1712.04651}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Ehsan_2022}{}}%
-Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation-clustersize}{}}%
-Fisher, M. E., \& Essam, J. W. (2004). Some {Cluster Size} and
-{Percolation Problems}. \emph{Journal of Mathematical Physics},
-\emph{2}(4), 609--619. \url{https://doi.org/10.1063/1.1703745}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
-Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
-(2022). Jfb: Jacobian-free backpropagation for implicit networks.
-\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
-\emph{36}, 6648--6656.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
-Guo, C., Barthelet, L., \& Morris, R. (2011). \emph{Maze generator and
-solver}. Wolfram Demonstrations Project,
-\url{https://demonstrations.wolfram.com/MazeGeneratorAndSolver/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
-Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
-{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
-Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
-{Games CoG}}, 1--4.
-
-\leavevmode\vadjust pre{\hypertarget{ref-A_star}{}}%
-Hart, P. E., Nilsson, N. J., \& Raphael, B. (1968). A {Formal Basis} for
-the {Heuristic Determination} of {Minimum Cost Paths}. \emph{IEEE
-Transactions on Systems Science and Cybernetics}, \emph{4}(2), 100--107.
-\url{https://doi.org/10.1109/TSSC.1968.300136}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
-Ivanitskiy, M. (n.d.). \emph{ZANJ}.
-\url{https://github.com/mivanit/ZANJ}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
-\emph{Maze dataset}.
-\url{https://github.com/understanding-search/maze-dataset}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
-Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
-D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
-\emph{Maze transformer interpretability}.
-\url{https://github.com/understanding-search/maze-transformer}
-
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
-Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
-Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
-C. D., \& others. (2023). A configurable library for generating and
-manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
-Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
-Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
-D., \& others. (2023). Structured world representations in maze-solving
-transformers. \emph{arXiv Preprint arXiv:2312.02566}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
-Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
-Ibrahim, M. (2024). The factorization curse: Which tokens you predict
-underlie the reversal curse and more. \emph{Advances in Neural
-Information Processing Systems}, \emph{37}, 112329--112355.
-
-\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
-Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
-Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
-extrapolation for mazes with recurrent and implicit networks.
-\emph{arXiv Preprint arXiv:2410.03020}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruskal1956shortest}{}}%
-Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
-the traveling salesman problem. \emph{Proceedings of the American
-Mathematical Society}, \emph{7}(1), 48--50.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
-Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
-Performance insights and comparative analysis. \emph{arXiv Preprint
-arXiv:2308.11224}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
-Nag, A. (2020). MDL suite: A language, generator and compiler for
-describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
-1815.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Nemeth_2019}{}}%
-Németh, F. (2019). \emph{Maze-generation-algorithms}.
-\url{https://github.com/ferenc-nemeth/maze-generation-algorithms}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
-Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
-Transformers can navigate mazes with multi-step prediction. \emph{arXiv
-Preprint arXiv:2412.05117}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gh-oppenheimj2018maze}{}}%
-Oppenheim, J. (2018). \emph{Maze-generator: Generate a random maze
-represented as a 2D array using depth-first search}.
-\url{https://github.com/oppenheimj/maze-generator/}; GitHub.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
-Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G.,
-Killeen, T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., Kopf,
-A., Yang, E., DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S.,
-Steiner, B., Fang, L., \ldots{} Chintala, S. (2019). {PyTorch: An
-Imperative Style, High-Performance Deep Learning Library}. In H.
-Wallach, H. Larochelle, A. Beygelzimer, F. d'Alché-Buc, E. Fox, \& R.
-Garnett (Eds.), \emph{Advances in neural information processing systems
-32} (pp. 8024--8035). Curran Associates, Inc.
-\url{http://papers.neurips.cc/paper/9015-pytorch-an-imperative-style-high-performance-deep-learning-library.pdf}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
-Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
-Hidden computation in transformer language models. \emph{arXiv Preprint
-arXiv:2404.15758}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
-Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
-transparent ai: A survey on interpreting the inner structures of deep
-neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
-Machine Learning (SaTML)}, 464--483.
-
-\leavevmode\vadjust pre{\hypertarget{ref-easy_to_hard}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Bansal, A., Emam, Z., Huang,
-F., Goldblum, M., \& Goldstein, T. (2021). \emph{Datasets for {Studying
-Generalization} from {Easy} to {Hard Examples}} (No. arXiv:2108.06011).
-{arXiv}. \url{https://doi.org/10.48550/arXiv.2108.06011}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
-Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
-Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
-Generalizing from easy to hard problems with recurrent networks.
-\emph{Advances in Neural Information Processing Systems}, \emph{34},
-6695--6706.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
-Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
-\emph{arXiv Preprint arXiv:2308.02522}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
-Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
-Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
-world models in maze-solving tasks. \emph{arXiv Preprint
-arXiv:2412.11867}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
-Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
-Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
-self-supervised neural-symbolic learning framework for robot autonomy.
-\emph{arXiv Preprint arXiv:2406.16087}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wilson}{}}%
-Wilson, D. B. (1996). Generating random spanning trees more quickly than
-the cover time. \emph{Proceedings of the Twenty-Eighth Annual {ACM}
-Symposium on {Theory} of Computing - {STOC} '96}, 296--303.
-\url{https://doi.org/10.1145/237814.237880}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
-Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
-scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
-arXiv:2502.01989}.
-
-\end{CSLReferences}
-
-\end{document}

### 2025-04-26 docs/paper/paper.tex
cut down impl and sre sections, better links

@@
 \newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
+\newcommand{\docslinkcodemain}[1]{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\##1}{\texttt{#1}}}
 \newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
@@
 
-\begin{figure}
-\hypertarget{fig:benchmarks}{%
-\centering
-\includegraphics[width=0.9\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
-\caption{Plot of maze generation time. Generation time scales
-exponentially with maze size for all algorithms. Generation time per
-maze does not depend on the number of mazes being generated, and there
-is minimal overhead to initializing the generation process for a small
-dataset. Wilson's algorithm is notably less efficient than others and
-has high variance. Note that values are averaged across all parameter
-sets for that algorithm. More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}\label{fig:benchmarks}
-}
-\end{figure}
+\input{figures/tex/fig6_benchmarks.tex}
 
@@
 \href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
-Using the estimation algorithm simply requires the user to call
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
-providing their initial \texttt{cfg} and then using the returned
-\texttt{cfg\_new} in its place.
-
-\hypertarget{success-rate-estimation-algorithm}{%
-\subsubsection{Success Rate Estimation
-Algorithm}\label{success-rate-estimation-algorithm}}
-
-The base function learned by symbolic regression provides limited
-insight and may be subject to change. It is defined as
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
-and takes a 5 dimensional float vector created by
-\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the
-{[}percolation value, grid size, endpoint deadend configuration,
-endpoint uniqueness, categorical generation function index{]}.
-
-However, the outputs of this function are not directly usable due to
-minor divergences at the endpoints with respect to the percolation
-probability \(p\). Since we know that maze success is either guaranteed
-or impossible for \(p=0\) and \(p=1\), we define the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function to nudge the raw output of the symbolic regression. This
-function is defined with the following components:
-
-shifted sigmoid \(\sigma_s\), amplitude scaling \(A\), and \(h\)
-function given by \[
-  \sigma_s(x) = (1 + e^{-10^3 \cdot (x-0.5)})^{-1}
-  \qquad A(q,a,w) = w \cdot (1 - |2q-1|^a)
-\] \[
-  h(q,a) = q \cdot (1 - |2q-1|^a) \cdot (1-\sigma_s(q)) + (1-(1-q) \cdot (1 - |2(1-q)-1|^a)) \cdot \sigma_s(q)
-\]
-
-We combine these to get the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
-function, which is identity-like for \(p \approx 0.5\), and pushes \(x\)
-to extremes otherwise. \[
-  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
-\]
-
-Finally, we define \[
-  \text{cfg\_success\_predict\_fn}(\mathbf{x}) = \text{soft\_step}(\text{raw\_val}, x_0, 5, 10)
-\]
-
-where \texttt{raw\_val} is the output of the symbolic regression model.
-The parameter \(x_0\) is the percolation probability, while all other
-parameters from \texttt{\_to\_ps\_array()} only affect
-\texttt{raw\_val}.
-
-\begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
-\caption{An example of both empirical and predicted success rates as a
-function of the percolation probability \(p\) for various maze sizes,
-percolation with and without depth first search, and
-\texttt{endpoint\_kwargs} requiring that both the start and end be in
-unique dead ends. Empirical measures derived from a sample of 128 mazes.
-More information can be found on the
-\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-page}.}
-\end{figure}
+
+\input{figures/tex/fig7_sre.tex}
 
@@
 
-We refer to our
-\href{https://github.com/understanding-search/maze-dataset}{repository}
-and
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs}
-for documentation and up-to-date implementation details.
-
-This package utilizes a simple, efficient representation of mazes as
-subgraphs of a finite lattice, which we call a
+Using an adjacency matrix for storing mazes would be memory inefficient
+by failing to exploit the highly sparse structure, while using an
+adjacency list could lead to a poor lookup time. This package utilizes a
+simple, efficient representation of mazes as subgraphs of a finite
+lattice, which we call a
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-Using an adjacency matrix for storing mazes would be memory inefficient
-by failing to exploit the highly sparse structure -- for example, for a
-2-dimensional maze, only 4 off-diagonal bands would be have nonzero
-values. On the other hand, using an adjacency list could lead to a poor
-lookup time for whether any given connection exists.
-
-Instead, we describe mazes with the following representation: for a
+
+We describe mazes with the following representation: for a
 \(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
-
-To produce solutions to mazes, two points are selected uniformly at
-random without replacement from the connected component of the maze, and
-the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
-1968}) is applied to find the shortest path between them. The endpoint
-selection can be controlled via
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
-and complications caused by this are detailed in
-\hyperref[sec:success-rate-estimation]{section: \textit{\nameref{sec:success-rate-estimation}}}.
-A maze with a solution is denoted a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}},
-which inherits from
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-
-Parallelization is implemented via the \texttt{multiprocessing} module
-in the Python standard library, and parallel generation can be
-controlled via keyword arguments to
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{MazeDataset.from\_config()}}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-A_star}{}}%
-Hart, P. E., Nilsson, N. J., \& Raphael, B. (1968). A {Formal Basis} for
-the {Heuristic Determination} of {Minimum Cost Paths}. \emph{IEEE
-Transactions on Systems Science and Cybernetics}, \emph{4}(2), 100--107.
-\url{https://doi.org/10.1109/TSSC.1968.300136}
-
 \leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%

### 2025-04-26 docs/paper/paper.tex
recompile paper

@@
 
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \IfFileExists{parskip.sty}{%
@@
   is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
-  Wu, 2023}), but provides over 5.8 million unique formats for
-  converting mazes to a text stream, detailed in
+  Wu, 2023}) but with many more options, detailed
   \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
 \item
-  For rigorous investigations of the response of a model to various
-  distributional shifts, preserving metadata about the generation
-  algorithm with the dataset itself is essential. To this end, our
-  package efficiently stores the dataset along with its metadata in a
-  single human-readable file (\protect\hyperlink{ref-zanj}{M.
+  Preserving metadata about the generation algorithm with the dataset
+  itself is essential for studying the effects of distributional shifts.
+  Our package efficiently stores the dataset along with its metadata in
+  a single human-readable file (\protect\hyperlink{ref-zanj}{M.
   Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
@@
 \item
-  Storing mazes as images is not only difficult to work with, but also
-  inefficient. We use a highly efficient method detailed in
+  Storing mazes as images or adjacency matrices is not only difficult to
+  work with, but also inefficient. We use a highly efficient method
+  detailed in
   \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
@@
 
-\begin{figure}
-\hypertarget{fig:e2h-raster}{%
-\centering
-\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
-\caption{Input is the rasterized maze without the path marked (left),
-and provide as a target the maze with all but the correct path removed
-(right). Configuration options exist to adjust whether endpoints are
-included and if empty cells should be filled in.}\label{fig:e2h-raster}
-}
-\end{figure}
+\input{figures/tex/fig3_raster.tex}
 
@@
 different representations of data on text model performance, we provide
-a variety of tokenized text output formats.
-
-We convert mazes to token sequences in two steps. First, the maze is
-stringified using
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens()}}.
-The
+a variety of text output formats. By passing an instance of
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class provides a powerful interface for configuring maze stringification
-behavior. Second, the sequence of strings is tokenized into integers
-using \texttt{encode()}. Tokenization uses a fixed vocabulary for
-simplicity. Mazes up to \(50 \times 50\) are supported when using a
-unique token for each position, and up to \(128 \times 128\) are
-supported when positions in the maze are represented as a pair of
-coordinates.
-
-There are many algorithms by which one might tokenize a 2D maze into a
-1D format usable by autoregressive text models. Training multiple models
-on the encodings output from each of these algorithms may produce very
-different internal representations, learned solution algorithms, and
-levels of performance. To allow exploration of how different maze
-tokenization algorithms affect these models, the
+to
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens(...)}},
+a maze can be converted to a text sequence. The
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class contains a rich set of options to customize how mazes are
-stringified. This class contains 19 discrete parameters, resulting in
-over 5.8 million unique tokenizers. There are 6 additional parameters
-available whose functionality is not verified via automated testing, but
-further expand the the number of tokenizers by a factor of \(44/3\) to
-86 million.
+class contains a rich set of options with class contains 19 discrete
+parameters, resulting in over 5.8 million unique possible tokenizers.
 
@@
 
-The tokenizer architecture is purposefully designed such that adding and
-testing a wide variety of new tokenization algorithms is fast and
-minimizes disturbances to functioning code. This is enabled by the
-modular architecture and the automatic inclusion of any new tokenizers
-in integration tests. To create a new variety of tokenizer, developers
-forking the library may simply create their own
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and implement the abstract methods. If the behavior change is
-sufficiently small, simply adding a parameter to an existing
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclass and updating its implementation will suffice.
-
-The breadth of tokenizers is also easily scaled in the opposite
-direction. Due to the exponential scaling of parameter combinations,
-adding a small number of new features can significantly slow certain
-procedures which rely on constructing all possible tokenizers, such as
-integration tests. If any existing subclass contains features which
-aren't needed, a developer tool decorator
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
-is provided which can be applied to the unneeded
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-subclasses to prune those features and compact the available space of
-tokenizers.
-
 \hypertarget{benchmarks}{%
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}}
-allows for additional constraints such as enforcing that the start or
+allows for additional constraints, such as enforcing that the start or
 end point be in a ``dead end'' with only one accessible neighbor cell.
-However, combining these constraints with cyclic mazes (such as those
-generated with percolation), as was required for the work in
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), can lead to an absence of valid start and end points. Placing
-theoretical bounds on this success rate is difficult, as it depends on
-the exact maze generation algorithm and parameters used. To deal with
-this, our package provides a way to estimate the success rate of a given
-configuration using a symbolic regression model trained with PySR
-(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
-be found in
+However, combining these constraints with cyclic mazes\footnote{Such as
+  those generated with percolation, as was required for the work in
+  (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et
+  al., 2024}).} can lead to an absence of valid start and end points. To
+deal with this, our package provides a way to estimate the success rate
+of a given configuration using a symbolic regression model trained with
+PySR (\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this
+can be found in
 \href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
@@
 existence of connections and facilitate fast lookup time, at the cost of
-requiring additional care with indexing. Note that this setup allows for
-a periodic lattice. Generation of mazes is detailed in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
+requiring additional care with indexing.
 

### 2025-05-28 docs/paper/figures/tex/fig8_impl.tex
fix some figures, cut some sections

\begin{figure}
\centering
\begin{tabular}{ccc}
	Maze 
	& $A[0]$ (downward)
	& $A[1]$ (rightward) \\
	\raisebox{-0.5\height}{
		\includegraphics[width=5em]{figures/tex/maze-impl.pdf}
	}
	& \raisebox{-0.5em}{
		\large $\begin{bmatrix}
			1 & 1 & 1 \\
			0 & 0 & 1 \\
			0 & 0 & 0
		\end{bmatrix}$
	}
	& \raisebox{-0.5em}{
		\large $\begin{bmatrix}
			1 & 0 & 0 \\
			0 & 1 & 0 \\
			1 & 1 & 0
		\end{bmatrix}$
	}
\end{tabular}
\caption{
	We describe mazes with the following representation: for a $2$-dimensional lattice with $r$ rows and $c$ columns, we initialize a boolean array $A = \{0, 1\}^{2 \times r \times c}$ which we refer to in the code as a \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.connection_list}{\texttt{connection\_list}}. The value at $A[0,i,j]$ determines whether a \emph{downward} connection exists from node $[i,j]$ to $[i+1, j]$. Likewise, the value at $A[1,i,j]$ determines whether a \emph{rightward} connection to $[i, j+1]$ exists. Thus, we avoid duplication of data about the existence of connections and facilitate fast lookup time, at the cost of requiring additional care with indexing.
}
\end{figure}

### 2025-05-28 docs/paper/paper.tex
fix some figures, cut some sections

@@
 \def\languageshorthands#1{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
 
@@
 \fi
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\tikzset{ % Define a TikZ style for an external hyperlink node.
-  hyperlink node url/.style={
-    alias=sourcenode,
-    append after command={
-      let \p1 = (sourcenode.north west),
-          \p2 = (sourcenode.south east),
-          \n1 = {\x2-\x1},
-          \n2 = {\y1-\y2} in
-      node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
-          {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
-    }
-  }
-}
-\providecommand{\XeTeXLinkBox}[1]{#1}
-\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
-\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
-\newcommand{\docslinkcodemain}[1]{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\##1}{\texttt{#1}}}
-\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
+\input{preamble.tex}
 \ifLuaTeX
@@
   Wu, 2023}) but with many more options, detailed
-  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
+  \secref{sec:tokenized-output-formats}.
 \item
@@
   work with, but also inefficient. We use a highly efficient method
-  detailed in
-  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
+  detailed in \secref{sec:implementation}.
 \item
@@
 \href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
-for more information.
-
-\hypertarget{generation}{%
-\subsection{Generation and Basic Usage}\label{generation}}
-
-Our package can be installed from
+for more information. Our package can be installed from
 \href{https://pypi.org/project/maze-dataset/}{PyPi} via
@@
 
-To create a dataset, we first create a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-configuration object, which specifies the seed, number, and size of
-mazes, as well as the generation algorithm and its corresponding
-parameters. This object is passed to a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-class to create a dataset. Crucially, this
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
-mimics the interface of a PyTorch
-(\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
-\href{https://pytorch.org/docs/stable/data.html}{\texttt{Dataset}}, and
-can thus be easily incorporated into existing data pre-processing and
-training pipelines, e.g., through the use of a \texttt{DataLoader}
-class.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
-\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
-\NormalTok{)}
-\CommentTok{\# create a config}
-\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
-\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
-\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
-\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
-\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
-    \CommentTok{\# (optional) algorithm{-}specific parameters}
-\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
-    \CommentTok{\# (optional) many options for restricting start/end points}
-\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
-\NormalTok{)}
-\CommentTok{\# create a dataset}
-\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
-\NormalTok{  cfg, }\CommentTok{\# pass the config}
-\NormalTok{  ..., }\CommentTok{\# other options for disk loading, parallelization, etc.}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-When initializing a dataset, options which do not affect the mazes
-themselves can be specified through the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
-factory method as necessary. These options allow for saving/loading
-existing datasets instead of re-generating, parallelization options for
-generation, and more. Available maze generation algorithms are static
-methods of the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
-namespace class and include generation algorithms based on randomized
-depth-first search, Wilson's algorithm
-(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
-(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
-\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
-Kruskal's algorithm
-(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
-others.
-
-Furthermore, a dataset of mazes can be filtered to satisfy certain
-properties. Custom filters can be specified, and some filters are
-included in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
-For example, we can require a minimum path length of three steps from
-the origin to the target:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dataset\_filtered: MazeDataset }\OperatorTok{=}\NormalTok{ dataset.filter\_by.path\_length(min\_length}\OperatorTok{=}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-All implemented maze generation algorithms are stochastic by nature. For
-reproducibility, the \texttt{seed} parameter of
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
-may be set. In practice, using provided deduplication filters, we find
-that exact duplicate mazes are generated very infrequently, even when
-generating very large datasets.
-
-For use cases where mazes of different sizes, generation algorithms, or
-other parameter variations are required, we provide the
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/collected_dataset.html\#MazeDatasetCollection}{\texttt{MazeDatasetCollection}}
-class, which allows for creating a single iterable dataset from multiple
-independent configurations.
+Datasets of mazes are created from a
+\docslinkcodemain{MazeDatasetConfig} configuration object, which allows
+specifying the number of mazes, their size, the generation algorithm,
+and various parameters for the generation algorithm. Datasets can also
+be filtered after generation to satisfy certain properties. Custom
+filters can be specified, and some filters are included in
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}
 
@@
 
-Each
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-is constructed from a set of several
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
-objects, each of which specifies how different token regions or other
-elements of the stringification are produced.
-
-\input{figures/tex/fig5_mmt.tex}
-
 \hypertarget{benchmarks}{%
@@
 
-We provide approximate benchmarks for relative generation time across
-various algorithms, parameter choices, maze sizes, and dataset sizes in
+We benchmarks for generation time across various configurations in
 \autoref{tab:benchmarks} and \autoref{fig:benchmarks}. Experiments were
 performed on a
-\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners#standard-github-hosted-runners-for-public-repositories}{standard GitHub runner}
-without parallelism.
+\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners\#standard-github-hosted-runners-for-public-repositories}{standard
+GitHub runner} without parallelism.
 
@@
 
-In order to replicate the exact dataset distribution of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}), the parameter
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}}
-allows for additional constraints, such as enforcing that the start or
-end point be in a ``dead end'' with only one accessible neighbor cell.
-However, combining these constraints with cyclic mazes\footnote{Such as
-  those generated with percolation, as was required for the work in
-  (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et
-  al., 2024}).} can lead to an absence of valid start and end points. To
-deal with this, our package provides a way to estimate the success rate
-of a given configuration using a symbolic regression model trained with
-PySR (\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this
-can be found in
-\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
+Maze generation under certain constraints may not always be successful,
+and for this we provide a way to estimate the success rate of a given
+configuration, described in \autoref{fig:sre}.
 
@@
 
-We describe mazes with the following representation: for a
-\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
-initialize a boolean array \[
-  A = \{0, 1\}^{2 \times r \times c}
-\] which we refer to in the code as a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
-The value at \(A[0,i,j]\) determines whether a \emph{downward}
-connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
-value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
-to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
-existence of connections and facilitate fast lookup time, at the cost of
-requiring additional care with indexing.
+\input{figures/tex/fig8_impl.tex}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pysr}{}}%
-Cranmer, M. (2023). Interpretable machine learning for science with PySR
-and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-percolation}{}}%
-Duminil-Copin, H. (2017). \emph{Sixty years of percolation} (No.
-arXiv:1712.04651). {arXiv}. \url{http://arxiv.org/abs/1712.04651}
-
 \leavevmode\vadjust pre{\hypertarget{ref-gh_Ehsan_2022}{}}%
 Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}
-
-\leavevmode\vadjust pre{\hypertarget{ref-percolation-clustersize}{}}%
-Fisher, M. E., \& Essam, J. W. (2004). Some {Cluster Size} and
-{Percolation Problems}. \emph{Journal of Mathematical Physics},
-\emph{2}(4), 609--619. \url{https://doi.org/10.1063/1.1703745}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kruskal1956shortest}{}}%
-Kruskal, J. B. (1956). On the shortest spanning subtree of a graph and
-the traveling salesman problem. \emph{Proceedings of the American
-Mathematical Society}, \emph{7}(1), 48--50.
-
 \leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
@@
 \url{https://github.com/oppenheimj/maze-generator/}; GitHub.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
-Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G.,
-Killeen, T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., Kopf,
-A., Yang, E., DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S.,
-Steiner, B., Fang, L., \ldots{} Chintala, S. (2019). {PyTorch: An
-Imperative Style, High-Performance Deep Learning Library}. In H.
-Wallach, H. Larochelle, A. Beygelzimer, F. d'Alché-Buc, E. Fox, \& R.
-Garnett (Eds.), \emph{Advances in neural information processing systems
-32} (pp. 8024--8035). Curran Associates, Inc.
-\url{http://papers.neurips.cc/paper/9015-pytorch-an-imperative-style-high-performance-deep-learning-library.pdf}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wilson}{}}%
-Wilson, D. B. (1996). Generating random spanning trees more quickly than
-the cover time. \emph{Proceedings of the Twenty-Eighth Annual {ACM}
-Symposium on {Theory} of Computing - {STOC} '96}, 296--303.
-\url{https://doi.org/10.1145/237814.237880}
-
 \leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%

### 2025-05-28 docs/paper/preamble.tex
fix some figures, cut some sections

\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{calc}
\tikzset{ % Define a TikZ style for an external hyperlink node.
  hyperlink node url/.style={
	alias=sourcenode,
	append after command={
	  let \p1 = (sourcenode.north west),
		  \p2 = (sourcenode.south east),
		  \n1 = {\x2-\x1},
		  \n2 = {\y1-\y2} in
	  node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
		  {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
	}
  }
}
\providecommand{\XeTeXLinkBox}[1]{#1}
\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
\newcommand{\docslinkcodemain}[1]{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\##1}{\texttt{#1}}}
\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>0.8\textheight 0.8\textheight\else\Gin@nat@height\fi}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\makeatother

### 2025-05-28 docs/paper/figures/outputs-ascii-colored.tex
wip

@@
-\begin{Verbatim}[commandchars=\\\{\}]
-# # # # # # # # # # #
-#         \textcolor{blue}{X X X} #   #
-#   # # # \textcolor{blue}{X} # \textcolor{blue}{X} #   #
-#       # \textcolor{blue}{X} # \textcolor{green}{S}     #
-# # # # # \textcolor{blue}{X} # # # # #
-# \textcolor{blue}{X X X X X} # \textcolor{red}{E} \textcolor{blue}{X X} #
-# \textcolor{blue}{X} # # #   # # # \textcolor{blue}{X} #
-# \textcolor{blue}{X} #           # \textcolor{blue}{X} #
-# \textcolor{blue}{X} # # # # # # # \textcolor{blue}{X} #
-# \textcolor{blue}{X X X X X X X X X} #
-# # # # # # # # # # #
-\end{Verbatim}
+{\ttfamily
+\setlength{\tabcolsep}{0.5em}
+\setlength{\arraystretch}{0.8}
+\begin{tabular}{ccccccccccc}
+\# & \# & \# & \# & \# & \# & \# & \# & \# & \# & \# \\
+\# &   &   &   &   & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# &  & \# \\
+\# &   & \# & \# & \# & \textcolor{blue}{X} & \# & \textcolor{blue}{X} & \# &   & \# \\
+\# &   &   &   & \# & \textcolor{blue}{X} & \# & \textcolor{green}{S} &   &   & \# \\
+\# & \# & \# & \# & \# & \textcolor{blue}{X} & \# & \# & \# & \# & \# \\
+\# & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# & \textcolor{red}{E} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# \\
+\# & \textcolor{blue}{X} & \# & \# & \# &   & \# & \# & \# & \textcolor{blue}{X} & \# \\
+\# & \textcolor{blue}{X} & \# &   &   &   &   &   & \# & \textcolor{blue}{X} & \# \\
+\# & \textcolor{blue}{X} & \# & \# & \# & \# & \# & \# & \# & \textcolor{blue}{X} & \# \\
+\# & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# \\
+\# & \# & \# & \# & \# & \# & \# & \# & \# & \# & \#
+\end{tabular}
+}

### 2025-05-28 docs/paper/paper.tex
wip

@@
 
+\newpage
+
 \hypertarget{features}{%
@@
 sub-graphs of a lattice. These objects can be converted to and from
-several formats to maximize their utility in different contexts.
-
-\input{figures/tex/fig2_formats.tex}
+several formats, shown in \autoref{fig:output-fmts}, to maximize their
+utility in different contexts.
 
@@
 class which returns rasterized pairs of (input, target) mazes as shown
-in \autoref{fig:e2h-raster} below.
+in \autoref{fig:e2h-raster}.
+
+\input{figures/tex/fig2_formats.tex}
 
 \input{figures/tex/fig3_raster.tex}
+
+\newpage
 
@@
 parameters, resulting in over 5.8 million unique possible tokenizers.
-
 All output sequences consist of four token regions representing
@@
 \hypertarget{benchmarks}{%
-\subsection{Benchmarks of Generation Speed}\label{benchmarks}}
+\subsection{Benchmarks}\label{benchmarks}}
 
@@
 \href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners\#standard-github-hosted-runners-for-public-repositories}{standard
-GitHub runner} without parallelism.
+GitHub runner} without parallelism. Additionally, maze generation under
+certain constraints may not always be successful, and for this we
+provide a way to estimate the success rate of a given configuration,
+described in \autoref{fig:sre}.
 
@@
 \input{figures/tex/fig6_benchmarks.tex}
-
-\hypertarget{sec:success-rate-estimation}{%
-\subsection{Success Rate Estimation}\label{sec:success-rate-estimation}}
-
-Maze generation under certain constraints may not always be successful,
-and for this we provide a way to estimate the success rate of a given
-configuration, described in \autoref{fig:sre}.
 
@@
 simple, efficient representation of mazes as subgraphs of a finite
-lattice, which we call a
+lattice, detailed in \autoref{fig:maze-impl}, which we call a
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
@@
 \input{figures/tex/fig8_impl.tex}
-
-\newpage
 

### 2025-05-28 docs/paper/paper.tex
wip

@@
 different representations of data on text model performance, we provide
-a variety of text output formats. By passing an instance of
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-to
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens(...)}},
-a maze can be converted to a text sequence. The
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
-class contains a rich set of options with class contains 19 discrete
-parameters, resulting in over 5.8 million unique possible tokenizers.
-All output sequences consist of four token regions representing
-different features of the maze; an example output sequence is shown in
+a variety of text output formats, with an example given in
 \autoref{fig:token-regions}.

### 2025-07-27 docs/paper/paper.tex
add some DOIs
more DOIs
fix umlaut in paper?

@@
     pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
+Spies, Tilman R"auker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
 Fung},
@@
     pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
+Aaron Sandoval, Alex F. Spies, Tilman R"auker, Brandon Knutson, Cecilia
 Diniz Behn, Samy Wu Fung},
@@
 %
-]{Tilman Räuker%
+]{Tilman R"auker%
   \,\orcidlink{0009-0009-6321-4413}\,%
@@
 \emph{Behavior Research Methods}, \emph{40}, 353--359.
+\url{https://doi.org/10.3758/brm.40.1.353}
 
@@
 {A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
+\url{https://doi.org/10.48550/arXiv.2403.15870}
 
@@
 Preprint arXiv:1912.01588}.
+\url{https://doi.org/10.48550/arXiv.1912.01588}
 
@@
 models' spatial intelligence via GRPO. \emph{arXiv Preprint
-arXiv:2502.14669}.
+arXiv:2502.14669}. \url{https://doi.org/10.48550/arXiv.2502.14669}
 
@@
 \emph{Proceedings of the AAAI Conference on Artificial Intelligence},
-\emph{36}, 6648--6656.
+\emph{36}, 6648--6656. \url{https://doi.org/10.1609/aaai.v36i6.20619}
 
@@
 Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
-{Games CoG}}, 1--4.
+{Games CoG}}, 1--4. \url{https://doi.org/10.1109/cig.2019.8848048}
 
@@
 Ivanitskiy, M. (n.d.). \emph{ZANJ}.
-\url{https://github.com/mivanit/ZANJ}
+\url{https://doi.org/10.5281/zenodo.15540393}
 
@@
 D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
-\emph{Maze dataset}.
-\url{https://github.com/understanding-search/maze-dataset}
+\emph{Maze dataset}. \url{https://doi.org/10.48550/arXiv.2309.10498}
 
@@
 \emph{Maze transformer interpretability}.
-\url{https://github.com/understanding-search/maze-transformer}
+\url{https://doi.org/10.48550/arXiv.2312.02566}
 
@@
 manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
+\url{https://doi.org/10.48550/arXiv.2309.10498}
 
@@
 transformers. \emph{arXiv Preprint arXiv:2312.02566}.
+\url{https://doi.org/10.48550/arXiv.2312.02566}
 
@@
 Information Processing Systems}, \emph{37}, 112329--112355.
+\url{https://doi.org/10.48550/arXiv.2406.05183}
 
@@
 \emph{arXiv Preprint arXiv:2410.03020}.
+\url{https://doi.org/10.48550/arXiv.2410.03020}
 
@@
 Performance insights and comparative analysis. \emph{arXiv Preprint
-arXiv:2308.11224}.
+arXiv:2308.11224}. \url{https://doi.org/10.48550/arXiv.2308.11224}
 
@@
 describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
-1815.
+1815. \url{https://doi.org/10.21105/joss.01815}
 
@@
 Preprint arXiv:2412.05117}.
+\url{https://doi.org/10.48550/arXiv.2412.05117}
 
@@
 Hidden computation in transformer language models. \emph{arXiv Preprint
-arXiv:2404.15758}.
+arXiv:2404.15758}. \url{https://doi.org/10.48550/arXiv.2404.15758}
 
@@
 Machine Learning (SaTML)}, 464--483.
+\url{https://doi.org/10.1109/satml54575.2023.00039}
 
@@
 \emph{Advances in Neural Information Processing Systems}, \emph{34},
-6695--6706.
+6695--6706. \url{https://doi.org/10.48550/arXiv.2106.04537}
 
@@
 \emph{arXiv Preprint arXiv:2308.02522}.
+\url{https://doi.org/10.48550/arXiv.2308.02522}
 
@@
 world models in maze-solving tasks. \emph{arXiv Preprint
-arXiv:2412.11867}.
+arXiv:2412.11867}. \url{https://doi.org/10.48550/arXiv.2412.11867}
 
@@
 \emph{arXiv Preprint arXiv:2406.16087}.
+\url{https://doi.org/10.48550/arXiv.2406.16087}
 
@@
 scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
-arXiv:2502.01989}.
+arXiv:2502.01989}. \url{https://doi.org/10.48550/arXiv.2502.01989}
 

### 2025-07-27 docs/paper/paper.tex
Fix problematic links?

--- 2025-05-28 docs/paper/paper.tex
@@
 as well as simplifying the process of circuit discovery
-(\protect\hyperlink{ref-interpretability-survery}{Räuker et al., 2023}).
+(\protect\hyperlink{ref-interpretability-survey}{Räuker et al., 2023}).
 Although mazes are well suited for these investigations, we found that
@@
   \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
+    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}}{RasterizedMazeDataset},
     utilizing
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
+    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}}{as\_pixels()},
     which can exactly mimic the outputs provided in
@@
   \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
+    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}}{as\_ascii()}
     provides a format similar to
@@
   \item
-    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
+    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html#MazePlot}{\texttt{MazePlot}}}{MazePlot}
     provides a feature-rich plotting utility with support for multiple
@@
   The text format provided by
-  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
+  \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}}{SolvedMaze(...).as\_tokens()}
   is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
@@
 Hidden computation in transformer language models. \emph{arXiv Preprint
-arXiv:2404.15758}.
+arXiv:2404.15758}. \url{https://doi.org/10.48550/arXiv.2404.15758}
 
-\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survey}{}}%
 Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward

### 2025-07-27 docs/paper/paper.tex
try again to fix LaTeX error

@@
   \item
-    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}}{RasterizedMazeDataset},
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
     utilizing
-    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}}{as\_pixels()},
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
     which can exactly mimic the outputs provided in
-    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
-    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
-    similar to the outputs of Németh
-    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
+    \texttt{easy-to-hard-data}
+    (\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
+    Bansal, et al., 2021}) and can be configured to be similar to the
+    outputs of Németh (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
   \item
-    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}}{as\_ascii()}
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
     provides a format similar to
@@
   \item
-    \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html#MazePlot}{\texttt{MazePlot}}}{MazePlot}
-    provides a feature-rich plotting utility with support for multiple
+    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
+    provides a feature‑rich plotting utility with support for multiple
     paths, heatmaps over positions, and more. This is similar to the
@@
   The text format provided by
-  \texorpdfstring{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}}{SolvedMaze(...).as\_tokens()}
+  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
   is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&

### 2025-07-28 docs/paper/figures/tex/fig4_tokenfmt.tex

@@
 	\centering
+	\hspace{-1em}
 	\begin{minipage}{5in}
@@
 	  \colorbox[RGB]{ 234,209,220 }{Target}: ending coordinate,
-	  \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence
+	  \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence.
+	  By passing an instance of \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#MazeTokenizerModular}{\texttt{MazeTokenizerModular}} to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#MazeDataset.as_tokens}{\texttt{as\_tokens(...)}} a maze can be converted to a text sequence.
+The \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#MazeTokenizerModular}{\texttt{MazeTokenizerModular}} class contains a rich set of options with 19 discrete parameters, resulting in over 5.8 million unique possible tokenizers.
 	}

### 2025-07-28 docs/paper/figures/tex/fig6_benchmarks.tex

\begin{figure}
	\hypertarget{fig:benchmarks}{%
	\centering
	\includegraphics[width=0.9\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
	\caption{Plot of maze generation time. Generation time scales
	exponentially with maze size for all algorithms. Generation time per
	maze does not depend on the number of mazes being generated, and there
	is minimal overhead to initializing the generation process for a small
	dataset. Wilson's algorithm is notably less efficient than others and
	has high variance. Note that values are averaged across all parameter
	sets for that algorithm. More information can be found on the
	\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
	page}.}\label{fig:benchmarks}
	}
\end{figure}

### 2025-07-28 docs/paper/figures/tex/fig7_sre.tex

@@
 	\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
-	\caption{An example of both empirical and predicted success rates as a
-	function of the percolation probability \(p\) for various maze sizes,
-	percolation with and without depth first search, and
-	\texttt{endpoint\_kwargs} requiring that both the start and end be in
-	unique dead ends. Empirical measures derived from a sample of 128 mazes.
-	More information can be found on the
-	\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-	page}.}
+	\caption{
+		In order to replicate the exact dataset distribution of \cite{easy_to_hard}, the parameter \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs}} allows for additional constraints, such as enforcing that the start or end point be in a ``dead end'' with only one accessible neighbor cell.
+However, combining these constraints with cyclic mazes can lead to an absence of valid start and end points.
+To deal with this, our package provides a way to estimate the success rate of a given configuration using a symbolic regression model trained with PySR \cite{pysr}.
+		An example of both empirical and predicted success rates as a function of the percolation probability \(p\) for various maze sizes, percolation with and without depth first search, and \texttt{endpoint\_kwargs} requiring that both the start and end be in unique dead ends.
+Empirical measures derived from a sample of 128 mazes.
+More information can be found on the \href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks page} and in the notebook \href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
+	}
+	\label{fig:sre}
 \end{figure}

### 2025-07-28 docs/paper/paper.tex

@@
     pdfissuenum={},
-    pdfdoi={N/A},
+    pdfdoi={10.xxxxxx/draft},
     pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
@@
 \fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
-et al. (1970). maze-dataset: Maze Generation with Algorithmic Variety
+et al. (2025). maze-dataset: Maze Generation with Algorithmic Variety
 and Representational Flexibility. \emph{Journal of Open Source
-Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}
+Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE?
+\url{https://doi.org/10.xxxxxx/draft}}.}}
 
@@
 \def\languageshorthands#1{}
+\usepackage{lineno}
+\linenumbers
+\usepackage{draftwatermark}
 
@@
 
-  {\bfseries DOI:} \href{https://doi.org/N/A}{\color{linky}{N/A}}
+  {\bfseries DOI:} \href{https://doi.org/10.xxxxxx/draft}{\color{linky}{10.xxxxxx/draft}}
 
@@
     {\bfseries Submitted:} 01 January 1970\\
-    {\bfseries Published:} 01 January 1970
+    {\bfseries Published:} unpublished
 
@@
 
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
+\section{Summary}\label{summary}
 
@@
 
-\hypertarget{statement-of-need}{%
-\section{Statement of Need}\label{statement-of-need}}
+\section{Statement of Need}\label{statement-of-need}
 
@@
 as well as simplifying the process of circuit discovery
-(\protect\hyperlink{ref-interpretability-survey}{Räuker et al., 2023}).
-Although mazes are well suited for these investigations, we found that
-existing maze generation packages
-(\protect\hyperlink{ref-cobbe2019procgen}{Cobbe et al., 2019};
-\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-harriesMazeExplorerCustomisable3D2019}{Harries et
-al., n.d.}; \protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) lack support for transforming between multiple
+(\citeproc{ref-interpretability-survey}{Räuker et al., 2023}). Although
+mazes are well suited for these investigations, we found that existing
+maze generation packages (\citeproc{ref-cobbe2019procgen}{Cobbe et al.,
+2019}; \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
+\citeproc{ref-harriesMazeExplorerCustomisable3D2019}{Harries et al.,
+n.d.}; \citeproc{ref-gh_Nemeth_2019}{Németh, 2019};
+\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal, et
+al., 2021}) lack support for transforming between multiple
 representations and provide limited control over the maze generation
@@
 
-\hypertarget{related-works}{%
-\subsection{Related Works}\label{related-works}}
+\subsection{Related Works}\label{related-works}
 
 A multitude of public and open-source software packages exist for
-generating mazes (\protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-\protect\hyperlink{ref-gh_Nemeth_2019}{Németh, 2019};
-\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). However, nearly all of these packages produce
-mazes represented as rasterized images or other visual formats rather
-than the underlying graph structure, and this makes it difficult to work
-with these datasets.
+generating mazes (\citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
+\citeproc{ref-gh_Nemeth_2019}{Németh, 2019};
+\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal, et
+al., 2021}). However, nearly all of these packages produce mazes
+represented as rasterized images or other visual formats rather than the
+underlying graph structure, and this makes it difficult to work with
+these datasets.
 
@@
     \texttt{easy-to-hard-data}
-    (\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-    Bansal, et al., 2021}) and can be configured to be similar to the
-    outputs of Németh (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
+    (\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal,
+    et al., 2021}) and can be configured to be similar to the outputs of
+    Németh (\citeproc{ref-gh_Nemeth_2019}{2019})
   \item
@@
     provides a format similar to
-    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
-    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
+    (\citeproc{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
+    \citeproc{ref-eval-gpt-visual}{Singla, 2023})
   \item
@@
     paths, heatmaps over positions, and more. This is similar to the
-    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
-    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
-    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
-    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
+    outputs of (\citeproc{ref-mazegenerator-net}{Alance AB, 2019};
+    \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
+    \citeproc{ref-mathematica-maze}{Guo et al., 2011};
+    \citeproc{ref-mdl-suite}{Nag, 2020})
   \end{itemize}
@@
   \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
-  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
-  Wu, 2023}) but with many more options, detailed
-  \secref{sec:tokenized-output-formats}.
+  is similar to that of (\citeproc{ref-eval-LLM-graphs}{Liu \& Wu,
+  2023}) but with many more options, detailed in
+  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
 \item
@@
   Our package efficiently stores the dataset along with its metadata in
-  a single human-readable file (\protect\hyperlink{ref-zanj}{M.
-  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
-  this reliably.
+  a single human-readable file (\citeproc{ref-zanj}{M. Ivanitskiy,
+  n.d.}). As far as we are aware, no existing packages do this reliably.
 \item
@@
   work with, but also inefficient. We use a highly efficient method
-  detailed in \secref{sec:implementation}.
+  detailed in
+  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
 \item
@@
   Many other maze generation packages lack this level of rigor and
-  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
-  2008}) appear to simply no longer be accessible.
+  scope, and some (\citeproc{ref-ayaz2008maze}{Ayaz et al., 2008})
+  appear to simply no longer be accessible.
 \end{itemize}
@@
 
-\hypertarget{features}{%
-\section{Features}\label{features}}
+\section{Features}\label{features}
 
@@
 \href{https://github.com/understanding-search/maze-dataset}{git
-repository} (\protect\hyperlink{ref-maze-dataset-github}{Michael I.
-Ivanitskiy et al., 2023a}).
+repository} (\citeproc{ref-maze-dataset-github}{Michael I. Ivanitskiy et
+al., 2023a}).
 
 Datasets of mazes are created from a
-\docslinkcodemain{MazeDatasetConfig} configuration object, which allows
-specifying the number of mazes, their size, the generation algorithm,
-and various parameters for the generation algorithm. Datasets can also
-be filtered after generation to satisfy certain properties. Custom
-filters can be specified, and some filters are included in
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
+configuration object, which allows specifying the number of mazes, their
+size, the generation algorithm, and various parameters for the
+generation algorithm. Datasets can also be filtered after generation to
+satisfy certain properties. Custom filters can be specified, and some
+filters are included in
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}
 
-\hypertarget{visual-output-formats}{%
-\subsection{Visual Output Formats}\label{visual-output-formats}}
+\subsection{Visual Output Formats}\label{visual-output-formats}
 
@@
 Neural Network (RCNN) derived architectures
-(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
-Huang, et al., 2021}). To facilitate the use of our package in this
-context, we replicate the format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}) and provide the
+(\citeproc{ref-deepthinking}{Schwarzschild, Borgnia, Gupta, Huang, et
+al., 2021}). To facilitate the use of our package in this context, we
+replicate the format of (\citeproc{ref-easy_to_hard}{Schwarzschild,
+Borgnia, Gupta, Bansal, et al., 2021}) and provide the
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
@@
 
-\hypertarget{sec:tokenized-output-formats}{%
 \subsection{Tokenized Output
-Formats}\label{sec:tokenized-output-formats}}
+Formats}\label{sec:tokenized-output-formats}
 
@@
 format of input data, and may even use delimiter tokens to perform
-reasoning steps (\protect\hyperlink{ref-pfau2024dotbydot}{Pfau et al.,
-2024}; \protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al.,
-2024}). To facilitate systematic investigation of the effects of
-different representations of data on text model performance, we provide
-a variety of text output formats, with an example given in
+reasoning steps (\citeproc{ref-pfau2024dotbydot}{Pfau et al., 2024};
+\citeproc{ref-spies2024causalworldmodels}{Spies et al., 2024}). To
+facilitate systematic investigation of the effects of different
+representations of data on text model performance, we provide a variety
+of text output formats, with an example given in
 \autoref{fig:token-regions}.
@@
 
-\hypertarget{benchmarks}{%
-\subsection{Benchmarks}\label{benchmarks}}
+\subsection{Benchmarks}\label{benchmarks}
 
@@
 
-\hypertarget{sec:implementation}{%
-\section{Implementation}\label{sec:implementation}}
+\section{Implementation}\label{sec:implementation}
 
@@
 
-\hypertarget{usage-in-research}{%
-\section{Usage in Research}\label{usage-in-research}}
+\section{Usage in Research}\label{usage-in-research}
 
 This package was originally built for the needs of the
-(\protect\hyperlink{ref-maze-transformer-github}{Michael I. Ivanitskiy
-et al., 2023b}) project, which aims to investigate spatial planning and
-world models in autoregressive transformer models trained on mazes
-(\protect\hyperlink{ref-ivanitskiy2023structuredworldreps}{Michael
-Igorevich Ivanitskiy, Spies, et al., 2023};
-\protect\hyperlink{ref-maze-dataset-arxiv-2023}{Michael Igorevich
-Ivanitskiy, Shah, et al., 2023};
-\protect\hyperlink{ref-spies2024causalworldmodels}{Spies et al., 2024}).
-It was extended for work on understanding the mechanisms by which
-recurrent convolutional and implicit networks
-(\protect\hyperlink{ref-fung2022jfb}{Fung et al., 2022}) solve mazes
-given a rasterized view
-(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
-2024}), which required matching the pixel-padded and endpoint
-constrained output format of
-(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
-Bansal, et al., 2021}). Ongoing work using \texttt{maze-dataset} aims to
-investigate the effects of varying the tokenization format on the
-performance of pretrained LLMs on spatial reasoning.
+(\citeproc{ref-maze-transformer-github}{Michael I. Ivanitskiy et al.,
+2023b}) project, which aims to investigate spatial planning and world
+models in autoregressive transformer models trained on mazes
+(\citeproc{ref-ivanitskiy2023structuredworldreps}{Michael Igorevich
+Ivanitskiy, Spies, et al., 2023};
+\citeproc{ref-maze-dataset-arxiv-2023}{Michael Igorevich Ivanitskiy,
+Shah, et al., 2023}; \citeproc{ref-spies2024causalworldmodels}{Spies et
+al., 2024}). It was extended for work on understanding the mechanisms by
+which recurrent convolutional and implicit networks
+(\citeproc{ref-fung2022jfb}{Fung et al., 2022}) solve mazes given a
+rasterized view (\citeproc{ref-knutson2024logicalextrapolation}{Knutson
+et al., 2024}), which required matching the pixel-padded and endpoint
+constrained output format of (\citeproc{ref-easy_to_hard}{Schwarzschild,
+Borgnia, Gupta, Bansal, et al., 2021}). Ongoing work using
+\texttt{maze-dataset} aims to investigate the effects of varying the
+tokenization format on the performance of pretrained LLMs on spatial
+reasoning.
 
@@
 \item
-  By (\protect\hyperlink{ref-nolte2024multistep}{Nolte et al., 2024}) to
-  compare the effectiveness of transformers trained with the
-  MLM-\(\mathcal{U}\)
-  (\protect\hyperlink{ref-MLMU-kitouni2024factorization}{Kitouni et al.,
-  2024}) multistep prediction objective against standard autoregressive
+  By (\citeproc{ref-nolte2024multistep}{Nolte et al., 2024}) to compare
+  the effectiveness of transformers trained with the MLM-\(\mathcal{U}\)
+  (\citeproc{ref-MLMU-kitouni2024factorization}{Kitouni et al., 2024})
+  multistep prediction objective against standard autoregressive
   training for multi-step planning on our maze task.
 \item
-  By (\protect\hyperlink{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\protect\hyperlink{ref-chen2024iaimperative}{Chen et al., 2024}) to
-  study the effectiveness of imperative learning.
+  By (\citeproc{ref-wang2024imperative}{Wang et al., 2024}) and
+  (\citeproc{ref-chen2024iaimperative}{Chen et al., 2024}) to study the
+  effectiveness of imperative learning.
 \item
-  By (\protect\hyperlink{ref-zhang2025tscend}{Zhang et al., 2025}) to
-  introduce a novel framework for reasoning diffusion models.
+  By (\citeproc{ref-zhang2025tscend}{Zhang et al., 2025}) to introduce a
+  novel framework for reasoning diffusion models.
 \item
-  By (\protect\hyperlink{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to
-  improve spatial reasoning in LLMs with GRPO.
+  By (\citeproc{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to improve
+  spatial reasoning in LLMs with GRPO.
 \end{itemize}
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0.5}
-\leavevmode\vadjust pre{\hypertarget{ref-mazegenerator-net}{}}%
+\bibitem[\citeproctext]{ref-mazegenerator-net}
 Alance AB. (2019). \emph{Maze generator}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
+\bibitem[\citeproctext]{ref-ayaz2008maze}
 Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
+\bibitem[\citeproctext]{ref-chen2024iaimperative}
 Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
+\bibitem[\citeproctext]{ref-cobbe2019procgen}
 Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
+\bibitem[\citeproctext]{ref-dao2025alphamaze}
 Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Ehsan_2022}{}}%
+\bibitem[\citeproctext]{ref-gh_Ehsan_2022}
 Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}
 
-\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
+\bibitem[\citeproctext]{ref-fung2022jfb}
 Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
+\bibitem[\citeproctext]{ref-mathematica-maze}
 Guo, C., Barthelet, L., \& Morris, R. (2011). \emph{Maze generator and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
+\bibitem[\citeproctext]{ref-harriesMazeExplorerCustomisable3D2019}
 Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
+\bibitem[\citeproctext]{ref-zanj}
 Ivanitskiy, M. (n.d.). \emph{ZANJ}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
+\bibitem[\citeproctext]{ref-maze-dataset-github}
 Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
+\bibitem[\citeproctext]{ref-maze-transformer-github}
 Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
+\bibitem[\citeproctext]{ref-maze-dataset-arxiv-2023}
 Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
+\bibitem[\citeproctext]{ref-ivanitskiy2023structuredworldreps}
 Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
+\bibitem[\citeproctext]{ref-MLMU-kitouni2024factorization}
 Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
+\bibitem[\citeproctext]{ref-knutson2024logicalextrapolation}
 Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
+\bibitem[\citeproctext]{ref-eval-LLM-graphs}
 Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
+\bibitem[\citeproctext]{ref-mdl-suite}
 Nag, A. (2020). MDL suite: A language, generator and compiler for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gh_Nemeth_2019}{}}%
+\bibitem[\citeproctext]{ref-gh_Nemeth_2019}
 Németh, F. (2019). \emph{Maze-generation-algorithms}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
+\bibitem[\citeproctext]{ref-nolte2024multistep}
 Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gh-oppenheimj2018maze}{}}%
+\bibitem[\citeproctext]{ref-gh-oppenheimj2018maze}
 Oppenheim, J. (2018). \emph{Maze-generator: Generate a random maze
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
+\bibitem[\citeproctext]{ref-pfau2024dotbydot}
 Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survey}{}}%
+\bibitem[\citeproctext]{ref-interpretability-survey}
 Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-easy_to_hard}{}}%
+\bibitem[\citeproctext]{ref-easy_to_hard}
 Schwarzschild, A., Borgnia, E., Gupta, A., Bansal, A., Emam, Z., Huang,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
+\bibitem[\citeproctext]{ref-deepthinking}
 Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
+\bibitem[\citeproctext]{ref-eval-gpt-visual}
 Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
+\bibitem[\citeproctext]{ref-spies2024causalworldmodels}
 Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
+\bibitem[\citeproctext]{ref-wang2024imperative}
 Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
+\bibitem[\citeproctext]{ref-zhang2025tscend}
 Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time

### 2025-07-28 maze_dataset/dataset/maze_dataset.py
make format, fix some issues

@@ -10,7 +10,7 @@ import multiprocessing
 import typing
 import warnings
 from pathlib import Path
-from typing import Literal, Optional, cast, overload
+from typing import Literal, cast, overload
 
 import numpy as np
 import tqdm
@@ -41,7 +41,7 @@ from maze_dataset.maze import LatticeMaze, SolvedMaze
 _GLOBAL_WORKER_CONFIG: MazeDatasetConfig
 
 
-def _generate_maze_helper(index: int) -> Optional[SolvedMaze]:  # noqa: ARG001
+def _generate_maze_helper(index: int) -> SolvedMaze | None:  # noqa: ARG001
 	"""Helper function for generating mazes in parallel.
 
 	> [!CAUTION]
@@ -59,7 +59,7 @@ def _generate_maze_helper(index: int) -> Optional[SolvedMaze]:  # noqa: ARG001
 	# Generate the solution
 	# mypy doesnt realize EndpointKwargsType has only string keys: `Keywords must be strings  [misc]`
 	# TYPING: error: No overload variant of "generate_random_path" of "LatticeMaze" matches argument type "dict[Literal['allowed_start', 'allowed_end', 'deadend_start', 'deadend_end', 'endpoints_not_equal', 'except_on_no_valid_endpoint'], bool | list[tuple[int, int]] | None]"  [call-overload]
-	solution: Optional[CoordArray] = maze.generate_random_path(**endpoint_kwargs)  # type: ignore[misc, call-overload]
+	solution: CoordArray | None = maze.generate_random_path(**endpoint_kwargs)  # type: ignore[misc, call-overload]
 
 	# Validate the solution
 	if (
@@ -150,7 +150,7 @@ class MazeDataset(GPTDataset[MazeDatasetConfig]):
 
 		"""
 		return cast(
-			MazeDataset,
+			"MazeDataset",
 			super().from_config(
 				cfg=cfg,
 				do_generate=do_generate,

### 2025-07-28 maze_dataset/maze/lattice_maze.py

@@ -413,7 +413,7 @@ class LatticeMaze(SerializableDataclass):
 		deadend_end: bool = False,
 		endpoints_not_equal: bool = False,
 		except_on_no_valid_endpoint: typing.Literal[False] = False,
-	) -> typing.Optional[CoordArray]: ...
+	) -> CoordArray | None: ...
 	def generate_random_path(  # noqa: C901
 		self,
 		allowed_start: CoordList | None = None,
@@ -422,7 +422,7 @@ class LatticeMaze(SerializableDataclass):
 		deadend_end: bool = False,
 		endpoints_not_equal: bool = False,
 		except_on_no_valid_endpoint: bool = True,
-	) -> typing.Optional[CoordArray]:
+	) -> CoordArray | None:
 		"""return a path between randomly chosen start and end nodes within the connected component
 
 		Note that setting special conditions on start and end positions might cause the same position to be selected as both start and end.

### 2025-07-28 maze_dataset/dataset/maze_dataset.py
fix more ruff issues

@@ -109,7 +109,8 @@ def _maze_gen_init_worker(config: MazeDatasetConfig) -> None:
 		)
 
 
-class MazeDataset(GPTDataset[MazeDatasetConfig]):
+# TODO: we probably don't need to hash datasets, right?
+class MazeDataset(GPTDataset[MazeDatasetConfig]):  # noqa: PLW1641
 	"""a maze dataset class. This is a collection of solved mazes, and should be initialized via `MazeDataset.from_config`"""
 
 	def __init__(

### 2025-07-28 maze_dataset/maze/lattice_maze.py

@@ -1215,6 +1215,16 @@ class TargetedLatticeMaze(LatticeMaze):  # type: ignore[misc]
 		"check equality, calls parent class equality check"
 		return super().__eq__(other)
 
+	def __hash__(self) -> int:
+		"hash the `TargetedLatticeMaze` by hashing a tuple of the connection list and start and end positions as bytes"
+		return hash(
+			(
+				self.connection_list.tobytes(),
+				self.start_pos.tobytes(),
+				self.end_pos.tobytes(),
+			)
+		)
+
 	def _get_start_pos_tokens(self) -> list[str | CoordTup]:
 		return [
 			SPECIAL_TOKENS.ORIGIN_START,

### 2025-07-30 docs/paper/paper.tex

--- 2025-07-27 docs/paper/paper.tex
@@
   is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
-  Wu, 2023}) but with many more options, detailed
-  \secref{sec:tokenized-output-formats}.
+  Wu, 2023}) but with many more options, detailed in
+  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
 \item
@@
 Datasets of mazes are created from a
-\docslinkcodemain{MazeDatasetConfig} configuration object, which allows
-specifying the number of mazes, their size, the generation algorithm,
-and various parameters for the generation algorithm. Datasets can also
-be filtered after generation to satisfy certain properties. Custom
-filters can be specified, and some filters are included in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
+configuration object, which allows specifying the number of mazes, their
+size, the generation algorithm, and various parameters for the
+generation algorithm. Datasets can also be filtered after generation to
+satisfy certain properties. Custom filters can be specified, and some
+filters are included in
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
 
@@
 a variety of text output formats, with an example given in
-\autoref{fig:token-regions}.
+\autoref{fig:token-regions}. We utilize Finite State Transducers
+(\protect\hyperlink{ref-Gallant2015Transducers}{Gallant, 2015}) for
+efficiently storing valid tokenizers.
 
@@
 \input{figures/tex/fig8_impl.tex}
+
+Our package is implemented in
+Python(\protect\hyperlink{ref-python}{Rossum, 1995}), and makes use of
+the extensive scientific computing ecosystem, including NumPy
+(\protect\hyperlink{ref-numpy}{Harris et al., 2020}) for array
+manipulation, plotting tools Waskom
+(\protect\hyperlink{ref-seaborn}{2021}), Jupyter notebooks
+(\protect\hyperlink{ref-jupyter}{Kluyver et al., 2016}), and PySR
+(\protect\hyperlink{ref-pysr}{Cranmer, 2023}) for symbolic regression.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pysr}{}}%
+Cranmer, M. (2023). Interpretable machine learning for science with PySR
+and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
+\url{https://doi.org/10.48550/arXiv.2305.01582}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Gallant2015Transducers}{}}%
+Gallant, A. (2015). \emph{Index 1,600,000,000 keys with automata and
+rust}. \url{https://burntsushi.net/transducers/}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-mathematica-maze}{}}%
@@
 {Games CoG}}, 1--4. \url{https://doi.org/10.1109/cig.2019.8848048}
+
+\leavevmode\vadjust pre{\hypertarget{ref-numpy}{}}%
+Harris, C. R., Millman, K. J., Walt, S. J. van der, Gommers, R.,
+Virtanen, P., \& al., et. (2020). Array programming with NumPy.
+\emph{Nature}, \emph{585}, 357--362.
+\url{https://doi.org/10.1038/s41586-020-2649-2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-matplotlib}{}}%
+Hunter, J. D. (2007). Matplotlib: A 2D graphics environment.
+\emph{Computing in Science and Engineering}, \emph{9}(3), 90--95.
+\url{https://doi.org/10.1109/MCSE.2007.55}
 
@@
 \url{https://doi.org/10.48550/arXiv.2406.05183}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jupyter}{}}%
+Kluyver, T., Ragan-Kelley, B., Perez, F., Granger, B., Bussonnier, M.,
+Frederic, J., Kelley, K., Hamrick, J., Grout, J., Corlay, S., Ivanov,
+P., Avila, D., Abdalla, S., \& Willing, C. (2016). Jupyter notebooks - a
+publishing format for reproducible computational workflows.
+\emph{Proceedings of the 20th International Conference on Electronic
+Publishing}, 87--90. \url{https://doi.org/10.3233/978-1-61499-649-1-87}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-python}{}}%
+Rossum, G. van. (1995). \emph{Python reference manual} (CS-R9525).
+Centrum voor Wiskunde; Informatica (CWI).
+\url{https://ir.cwi.nl/pub/5008/05008D.pdf}
+
 \leavevmode\vadjust pre{\hypertarget{ref-easy_to_hard}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-seaborn}{}}%
+Waskom, M. L. (2021). {seaborn}: Statistical data visualization.
+\emph{Journal of Open Source Software}, \emph{6}(60), 3021.
+\url{https://doi.org/10.21105/joss.03021}
+
 \leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%

### 2025-08-04 maze_dataset/generation/generators.py
rename numpy rng global

@@ -12,7 +12,7 @@ from maze_dataset.generation.seed import GLOBAL_SEED
 from maze_dataset.maze import ConnectionList, Coord, LatticeMaze, SolvedMaze
 from maze_dataset.maze.lattice_maze import NEIGHBORS_MASK, _fill_edges_with_walls
 
-numpy_rng = np.random.default_rng(GLOBAL_SEED)
+_NUMPY_RNG: np.random.Generator = np.random.default_rng(GLOBAL_SEED)
 random.seed(GLOBAL_SEED)
 
 

### 2025-08-04 maze_dataset/tokenization/modular/elements.py

@@ -25,7 +25,7 @@ from maze_dataset.constants import (
 	Coord,
 	CoordTup,
 )
-from maze_dataset.generation import numpy_rng
+from maze_dataset.generation import _NUMPY_RNG
 from maze_dataset.maze.lattice_maze import LatticeMaze, SolvedMaze
 from maze_dataset.token_utils import (
 	connection_list_to_adj_list,
@@ -206,7 +206,7 @@ class EdgeGroupings(__TokenizerElementNamespace):
 				np.unique(sorted_edges[:, 0, :], return_index=True, axis=0)[1][1:],
 			)
 			if self.shuffle_group:
-				[numpy_rng.shuffle(g, axis=0) for g in groups]
+				[_NUMPY_RNG.shuffle(g, axis=0) for g in groups]
 			return groups
 
 
@@ -268,7 +268,7 @@ class EdgePermuters(__TokenizerElementNamespace):
 
 		@staticmethod
 		def _permute(lattice_edges: ConnectionArray) -> ConnectionArray:
-			numpy_rng.permuted(lattice_edges, axis=1, out=lattice_edges)
+			_NUMPY_RNG.permuted(lattice_edges, axis=1, out=lattice_edges)
 			return lattice_edges
 
 	@serializable_dataclass(frozen=True, kw_only=True)
@@ -510,7 +510,7 @@ class AdjListTokenizers(__TokenizerElementNamespace):
 			# shuffle the groups if specified
 			if self.shuffle_d0:
 				if isinstance(groups, np.ndarray):
-					numpy_rng.shuffle(groups, axis=0)
+					_NUMPY_RNG.shuffle(groups, axis=0)
 				elif isinstance(groups, list):
 					random.shuffle(groups)
 				else:

### 2025-08-04 tests/unit/tokenization/test_token_utils.py

@@ -8,7 +8,7 @@ from jaxtyping import Int
 
 from maze_dataset import LatticeMaze
 from maze_dataset.constants import VOCAB, Connection, ConnectionArray
-from maze_dataset.generation import numpy_rng
+from maze_dataset.generation import _NUMPY_RNG
 from maze_dataset.testing_utils import GRID_N, MAZE_DATASET
 from maze_dataset.token_utils import (
 	_coord_to_strings_UT,
@@ -1692,12 +1692,12 @@ def test_lattice_connection_arrray(n):
 					lambda: lattice_connection_array(GRID_N),
 					lambda: np.flip(lattice_connection_array(GRID_N), axis=1),
 					lambda: lattice_connection_array(GRID_N - 1),
-					lambda: numpy_rng.choice(
+					lambda: _NUMPY_RNG.choice(
 						lattice_connection_array(GRID_N),
 						2 * GRID_N,
 						axis=0,
 					),
-					lambda: numpy_rng.choice(
+					lambda: _NUMPY_RNG.choice(
 						lattice_connection_array(GRID_N),
 						1,
 						axis=0,

### 2025-08-04 maze_dataset/generation/__init__.py
add a way to register new generator functions

@@ -3,28 +3,28 @@
 `DEFAULT_GENERATORS` is a list of generator name, generator kwargs pairs used in tests and demos
 
 you can add your own maze generators by:
-- adding a static method implementing your generation function to `LatticeMazeGenerators`, with the signature `(grid_shape: Coord, **kwargs) -> LatticeMaze`
-- adding the pair to `GENERATORS_MAP`
+- adding a static method implementing your generation function to `LatticeMazeGenerators`, with the signature `(grid_shape: Coord, **kwargs) -> LatticeMaze` and adding the `(name, func)` pair to `GENERATORS_MAP`
+- using the `@register_maze_generator` decorator on your generation function. However, this is only for testing purposes where modifying the original package is not possible.
 
 If you implement a new generation function, please make a pull request!
 https://github.com/understanding-search/maze-dataset/pulls
 """
 
 from maze_dataset.generation.generators import (
+	_NUMPY_RNG,
 	GENERATORS_MAP,
 	LatticeMazeGenerators,
 	get_maze_with_solution,
-	numpy_rng,
 )
 
 __all__ = [
 	# submodules
 	"default_generators",
 	"generators",
-	"seed",
+	"registrationseed",
 	# imports
 	"LatticeMazeGenerators",
 	"GENERATORS_MAP",
 	"get_maze_with_solution",
-	"numpy_rng",
+	"_NUMPY_RNG",
 ]

### 2025-08-04 maze_dataset/generation/registration.py

@@ -1,0 +1,173 @@
+"""Registration system for custom maze generators"""
+
+import inspect
+from typing import Callable, Union, get_args, get_origin
+
+from maze_dataset.constants import Coord, CoordTup
+from maze_dataset.generation.generators import GENERATORS_MAP, LatticeMazeGenerators
+from maze_dataset.maze import LatticeMaze
+
+# Type for maze generator functions
+MazeGeneratorFunc = Callable[[Union[Coord, CoordTup], ...], LatticeMaze]
+
+
+class MazeGeneratorRegistrationError(TypeError):
+	"""error for maze generator registration issues"""
+
+	pass
+
+
+def _check_grid_shape_annotation(
+	annotation: Union[type, str],
+	func_name: str,
+) -> None:
+	"""Check if the annotation for grid_shape is valid
+
+	# Parameters:
+	- `annotation` : Union[type, str] the type annotation of the grid_shape parameter
+
+	# Raises:
+	- `MazeGeneratorRegistrationError` if the annotation is not compatible
+	"""
+	annotation_str: str = str(annotation)
+
+	if not any(
+		[
+			annotation == Coord,
+			annotation == CoordTup,
+			annotation_str.lower() == "tuple[int, int]",
+			# TODO: these are pretty loose checks, would be better to make this more robust
+			"ndarray" in annotation_str,
+			"Int8" in annotation_str,
+		]
+	):
+		err_msg = (
+			f"Maze generator function '{func_name}' first parameter 'grid_shape' "
+			f"must be typed as 'Coord | CoordTup' or compatible type, "
+			f"got {annotation = }, {annotation_str = }"
+		)
+		raise MazeGeneratorRegistrationError(err_msg)
+
+
+def validate_MazeGeneratorFunc(
+	func: MazeGeneratorFunc,
+) -> None:
+	"""validate the signature of a maze generator function
+
+	return `None` if valid, otherwise raises `MazeGeneratorRegistrationError`
+	(which is a subclass of `TypeError`)
+
+	# Parameters:
+	- `func : MazeGeneratorFunc` function to validate
+
+	# Raises:
+	- `MazeGeneratorRegistrationError` : type error describing the issue with the function signature
+	"""
+	func_name: str = func.__name__
+	sig: inspect.Signature = inspect.signature(func)
+	params: list[str] = list(sig.parameters.keys())
+
+	if not params or params[0] != "grid_shape":
+		err_msg = (
+			f"Maze generator function '{func_name}' must have 'grid_shape' "
[…]

### 2025-08-04 tests/unit/generation/test_custom_generator.py

@@ -1,0 +1,299 @@
+"""Tests for custom maze generator registration system"""
+
+from pathlib import Path
+
+import numpy as np
+import pytest
+from zanj import ZANJ
+
+from maze_dataset import LatticeMaze, MazeDataset, MazeDatasetConfig
+from maze_dataset.constants import Coord, CoordTup
+from maze_dataset.generation import (
+	GENERATORS_MAP,
+	LatticeMazeGenerators,
+	get_maze_with_solution,
+)
+from maze_dataset.generation.registration import (
+	MazeGeneratorRegistrationError,
+	register_maze_generator,
+)
+
+# Temp directory for file operations
+TEMP_PATH = Path("tests/_temp/test_custom_generator/")
+
+
+def test_register_valid_function():
+	"""Test that a valid function can be registered successfully"""
+
+	@register_maze_generator
+	def gen_test_valid(
+		grid_shape: Coord | CoordTup,
+		lattice_dim: int = 2,
+	) -> LatticeMaze:
+		"""Simple test generator - fully connected grid"""
+		grid_shape_: Coord = np.array(grid_shape)
+		connection_list: np.ndarray = np.zeros(
+			(lattice_dim, *grid_shape_), dtype=np.bool_
+		)
+
+		# Create fully connected grid
+		if grid_shape_[1] > 1:
+			connection_list[1, :, : grid_shape_[1] - 1] = True
+		if grid_shape_[0] > 1:
+			connection_list[0, : grid_shape_[0] - 1, :] = True
+
+		return LatticeMaze(
+			connection_list=connection_list,
+			generation_meta=dict(
+				func_name="gen_test_valid",
+				grid_shape=grid_shape_,
+				fully_connected=True,
+			),
+		)
+
+	# Test registration worked
+	assert "gen_test_valid" in GENERATORS_MAP
+	assert hasattr(LatticeMazeGenerators, "gen_test_valid")
+
+	# Test function works
+	maze = get_maze_with_solution("gen_test_valid", (5, 5))
+	assert maze.maze.grid_shape == (5, 5)
+
+	# Test via LatticeMazeGenerators
+	maze2 = LatticeMazeGenerators.gen_test_valid((4, 4))
+	assert maze2.grid_shape == (4, 4)
+
+
+def test_maze_dataset_config_with_custom_generator():
+	"""Test creating, saving, and loading MazeDatasetConfig with custom generator"""
+
+	@register_maze_generator
+	def gen_test_config(
[…]

### 2025-08-04 maze_dataset/generation/generators.py
add a TODO

@@ -636,6 +636,7 @@ this variable is primarily used in `MazeDatasetConfig._to_ps_array` and `MazeDat
 """
 
 
+# TODO: we should deprecate this, always get a dataset when you want a maze with a solution
 def get_maze_with_solution(
 	gen_name: str,
 	grid_shape: Coord | CoordTup,

### 2025-08-04 tests/unit/generation/test_custom_generator.py
fix some warnings

@@ -57,7 +57,7 @@ def test_register_valid_function():
 
 	# Test function works
 	maze = get_maze_with_solution("gen_test_valid", (5, 5))
-	assert maze.maze.grid_shape == (5, 5)
+	assert maze.grid_shape == (5, 5)
 
 	# Test via LatticeMazeGenerators
 	maze2 = LatticeMazeGenerators.gen_test_valid((4, 4))
@@ -165,7 +165,7 @@ def test_maze_dataset_with_custom_generator():
 	# Test dataset properties
 	assert len(dataset) == 2
 	for maze in dataset:
-		assert maze.maze.grid_shape == (4, 4)
+		assert maze.grid_shape == (4, 4)
 
 	# Test save/load dataset
 	TEMP_PATH.mkdir(parents=True, exist_ok=True)
@@ -177,10 +177,8 @@ def test_maze_dataset_with_custom_generator():
 	assert len(loaded_dataset) == len(dataset)
 	assert loaded_dataset.cfg.name == dataset.cfg.name
 	for original, loaded in zip(dataset, loaded_dataset, strict=True):
-		assert original.maze.grid_shape == loaded.maze.grid_shape
-		assert np.array_equal(
-			original.maze.connection_list, loaded.maze.connection_list
-		)
+		assert original.grid_shape == loaded.grid_shape
+		assert np.array_equal(original.connection_list, loaded.connection_list)
 
 
 def test_registration_error_missing_grid_shape():

### 2025-08-04 maze_dataset/generation/generators.py
typing fixes

@@ -2,7 +2,7 @@
 
 import random
 import warnings
-from typing import Any, Callable
+from typing import Callable, Concatenate, ParamSpec
 
 import numpy as np
 from jaxtyping import Bool
@@ -60,6 +60,7 @@ class LatticeMazeGenerators:
 	@staticmethod
 	def gen_dfs(
 		grid_shape: Coord | CoordTup,
+		*,
 		lattice_dim: int = 2,
 		accessible_cells: float | None = None,
 		max_tree_depth: float | None = None,
@@ -609,8 +610,15 @@ class LatticeMazeGenerators:
 		)
 
 
+P_GeneratorKwargs = ParamSpec("P_GeneratorKwargs")
+MazeGeneratorFunc = Callable[
+	Concatenate[Coord | CoordTup, P_GeneratorKwargs],
+	LatticeMaze,
+]
+
+
 # cant automatically populate this because it messes with pickling :(
-GENERATORS_MAP: dict[str, Callable[[Coord | CoordTup, Any], "LatticeMaze"]] = {
+GENERATORS_MAP: dict[str, MazeGeneratorFunc] = {
 	"gen_dfs": LatticeMazeGenerators.gen_dfs,
 	# TYPING: error: Dict entry 1 has incompatible type
 	# "str": "Callable[[ndarray[Any, Any] | tuple[int, int], KwArg(Any)], LatticeMaze]";

### 2025-08-04 maze_dataset/generation/registration.py

@@ -1,14 +1,17 @@
 """Registration system for custom maze generators"""
 
 import inspect
-from typing import Callable, Union, get_args, get_origin
+from typing import TypeVar, Union, get_args, get_origin
 
 from maze_dataset.constants import Coord, CoordTup
-from maze_dataset.generation.generators import GENERATORS_MAP, LatticeMazeGenerators
+from maze_dataset.generation.generators import (
+	GENERATORS_MAP,
+	LatticeMazeGenerators,
+	MazeGeneratorFunc,
+)
 from maze_dataset.maze import LatticeMaze
 
-# Type for maze generator functions
-MazeGeneratorFunc = Callable[[Union[Coord, CoordTup], ...], LatticeMaze]
+F_MazeGeneratorFunc = TypeVar("F_MazeGeneratorFunc", bound=MazeGeneratorFunc)
 
 
 class MazeGeneratorRegistrationError(TypeError):
@@ -50,7 +53,7 @@ def _check_grid_shape_annotation(
 
 
 def validate_MazeGeneratorFunc(
-	func: MazeGeneratorFunc,
+	func: F_MazeGeneratorFunc,
 ) -> None:
 	"""validate the signature of a maze generator function
 
@@ -110,7 +113,7 @@ def validate_MazeGeneratorFunc(
 		raise MazeGeneratorRegistrationError(err_msg)
 
 
-def register_maze_generator(func: MazeGeneratorFunc) -> MazeGeneratorFunc:
+def register_maze_generator(func: F_MazeGeneratorFunc) -> F_MazeGeneratorFunc:
 	"""Decorator to register a custom maze generator function.
 
 	This decorator allows users to register their own maze generation functions

### 2025-08-04 tests/unit/generation/test_custom_generator.py

@@ -181,6 +181,10 @@ def test_maze_dataset_with_custom_generator():
 		assert np.array_equal(original.connection_list, loaded.connection_list)
 
 
+# bunch of type ignores here, because we are testing to make sure that
+# the registration system raises errors for invalid function signatures
+
+
 def test_registration_error_missing_grid_shape():
 	"""Test error when function is missing grid_shape parameter"""
 
@@ -192,7 +196,7 @@ def test_registration_error_missing_grid_shape():
 		MazeGeneratorRegistrationError,
 		match="must have 'grid_shape' as its first parameter",
 	):
-		register_maze_generator(invalid_missing_grid_shape)
+		register_maze_generator(invalid_missing_grid_shape)  # type: ignore[type-var]
 
 
 def test_registration_error_wrong_param_name():
@@ -206,7 +210,7 @@ def test_registration_error_wrong_param_name():
 		MazeGeneratorRegistrationError,
 		match="must have 'grid_shape' as its first parameter",
 	):
-		register_maze_generator(invalid_wrong_param_name)
+		register_maze_generator(invalid_wrong_param_name)  # type: ignore[type-var]
 
 
 def test_registration_error_missing_type_annotation():
@@ -220,7 +224,7 @@ def test_registration_error_missing_type_annotation():
 		MazeGeneratorRegistrationError,
 		match=r"must be typed as 'Coord \| CoordTup' or compatible type",
 	):
-		register_maze_generator(invalid_missing_type_annotation)
+		register_maze_generator(invalid_missing_type_annotation)  # type: ignore[type-var]
 
 
 def test_registration_error_missing_return_annotation():
@@ -234,7 +238,7 @@ def test_registration_error_missing_return_annotation():
 		MazeGeneratorRegistrationError,
 		match="must have a return type annotation of LatticeMaze",
 	):
-		register_maze_generator(invalid_missing_return_annotation)
+		register_maze_generator(invalid_missing_return_annotation)  # type: ignore[type-var]
 
 
 def test_registration_error_wrong_return_type():
@@ -245,7 +249,7 @@ def test_registration_error_wrong_return_type():
 		return "wrong"
 
 	with pytest.raises(MazeGeneratorRegistrationError, match="must return LatticeMaze"):
-		register_maze_generator(invalid_wrong_return_type)
+		register_maze_generator(invalid_wrong_return_type)  # type: ignore[type-var]
 
 
[…]

### 2025-08-04 docs/paper/paper.tex
narrowed it down to figs 4, 6, 7, 8 and tab 1 are the issue in paper compilation

--- 2025-07-28 docs/paper/paper.tex
@@
 filters are included in
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.
 
@@
 of text output formats, with an example given in
-\autoref{fig:token-regions}.
-
-\input{figures/tex/fig4_tokenfmt.tex}
+\autoref{fig:token-regions}. We utilize Finite State Transducers
+(\citeproc{ref-Gallant2015Transducers}{Gallant, 2015}) for efficiently
+storing valid tokenizers.
 
@@
 
-\input{figures/tex/tab1_benchmarks.tex}
-
-\input{figures/tex/fig6_benchmarks.tex}
-
-\input{figures/tex/fig7_sre.tex}
-
 \section{Implementation}\label{sec:implementation}
@@
 
-\input{figures/tex/fig8_impl.tex}
+Our package is implemented in Python(\citeproc{ref-python}{Rossum,
+1995}), and makes use of the extensive scientific computing ecosystem,
+including NumPy (\citeproc{ref-numpy}{Harris et al., 2020}) for array
+manipulation, plotting tools (\citeproc{ref-matplotlib}{Hunter, 2007};
+\citeproc{ref-seaborn}{Waskom, 2021}), Jupyter notebooks
+(\citeproc{ref-jupyter}{Kluyver et al., 2016}), and PySR
+(\citeproc{ref-pysr}{Cranmer, 2023}) for symbolic regression.
 
@@
 
+\bibitem[\citeproctext]{ref-pysr}
+Cranmer, M. (2023). Interpretable machine learning for science with PySR
+and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
+\url{https://doi.org/10.48550/arXiv.2305.01582}
+
 \bibitem[\citeproctext]{ref-dao2025alphamaze}
@@
 
+\bibitem[\citeproctext]{ref-Gallant2015Transducers}
+Gallant, A. (2015). \emph{Index 1,600,000,000 keys with automata and
+rust}. \url{https://burntsushi.net/transducers/}.
+
 \bibitem[\citeproctext]{ref-mathematica-maze}
@@
 {Games CoG}}, 1--4. \url{https://doi.org/10.1109/cig.2019.8848048}
+
+\bibitem[\citeproctext]{ref-numpy}
+Harris, C. R., Millman, K. J., Walt, S. J. van der, Gommers, R.,
+Virtanen, P., \& al., et. (2020). Array programming with NumPy.
+\emph{Nature}, \emph{585}, 357--362.
+\url{https://doi.org/10.1038/s41586-020-2649-2}
+
+\bibitem[\citeproctext]{ref-matplotlib}
+Hunter, J. D. (2007). Matplotlib: A 2D graphics environment.
+\emph{Computing in Science and Engineering}, \emph{9}(3), 90--95.
+\url{https://doi.org/10.1109/MCSE.2007.55}
 
@@
 \url{https://doi.org/10.48550/arXiv.2406.05183}
+
+\bibitem[\citeproctext]{ref-jupyter}
+Kluyver, T., Ragan-Kelley, B., Perez, F., Granger, B., Bussonnier, M.,
+Frederic, J., Kelley, K., Hamrick, J., Grout, J., Corlay, S., Ivanov,
+P., Avila, D., Abdalla, S., \& Willing, C. (2016). Jupyter notebooks - a
+publishing format for reproducible computational workflows.
+\emph{Proceedings of the 20th International Conference on Electronic
+Publishing}, 87--90. \url{https://doi.org/10.3233/978-1-61499-649-1-87}
 
@@
 
+\bibitem[\citeproctext]{ref-python}
+Rossum, G. van. (1995). \emph{Python reference manual} (CS-R9525).
+Centrum voor Wiskunde; Informatica (CWI).
+\url{https://ir.cwi.nl/pub/5008/05008D.pdf}
+
 \bibitem[\citeproctext]{ref-easy_to_hard}
@@
 
+\bibitem[\citeproctext]{ref-seaborn}
+Waskom, M. L. (2021). {seaborn}: Statistical data visualization.
+\emph{Journal of Open Source Software}, \emph{6}(60), 3021.
+\url{https://doi.org/10.21105/joss.03021}
+
 \bibitem[\citeproctext]{ref-zhang2025tscend}

### 2025-08-04 docs/paper/figures/tex/fig7_sre.tex
fix figure 7

@@
 	\caption{
-		In order to replicate the exact dataset distribution of \cite{easy_to_hard}, the parameter \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs}} allows for additional constraints, such as enforcing that the start or end point be in a ``dead end'' with only one accessible neighbor cell.
+		In order to replicate the exact dataset distribution of \cite{easy_to_hard}, the parameter 
+		\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs}}
+		allows for additional constraints, such as enforcing that the start or end point be in a ``dead end'' with only one accessible neighbor cell.
 However, combining these constraints with cyclic mazes can lead to an absence of valid start and end points.
@@
 Empirical measures derived from a sample of 128 mazes.
-More information can be found on the \href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks page} and in the notebook \href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
+More information can be found on the \href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks page} and in the notebook
+		\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.
 	}

### 2025-08-04 docs/paper/figures/tex/fig7_sre.tex
fix citations, full paper compiling but fig heights messed up

@@
 	\caption{
-		In order to replicate the exact dataset distribution of \cite{easy_to_hard}, the parameter 
+		In order to replicate the exact dataset distribution of (\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal, et
+		al., 2021}), % custom cite here due to weird interactions with inara
+		the parameter 
 		\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs}}
@@
 However, combining these constraints with cyclic mazes can lead to an absence of valid start and end points.
-To deal with this, our package provides a way to estimate the success rate of a given configuration using a symbolic regression model trained with PySR \cite{pysr}.
+To deal with this, our package provides a way to estimate the success rate of a given configuration using a symbolic regression model trained with PySR (\citeproc{ref-pysr}{Cranmer, 2023}).
+% custom cite here due to weird interactions with inara
 		An example of both empirical and predicted success rates as a function of the percolation probability \(p\) for various maze sizes, percolation with and without depth first search, and \texttt{endpoint\_kwargs} requiring that both the start and end be in unique dead ends.

### 2025-08-04 docs/paper/figures/tex/fig6_benchmarks.tex
figure 6 fixed?
fix some figure sizing, all except diagram

@@
 		\centering
-	\includegraphics[width=0.9\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
-	\caption{Plot of maze generation time. Generation time scales
+		\includegraphics[width=0.9\textwidth]{figures/benchmarks/gridsize-vs-gentime.pdf}
+		\caption{
+			Plot of maze generation time. Generation time scales
 			exponentially with maze size for all algorithms. Generation time per
@@
 			sets for that algorithm. More information can be found on the
-	\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
-	page}.}\label{fig:benchmarks}
+			\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks page}.
+		}
+		\label{fig:benchmarks}
 	}

### 2025-08-04 docs/paper/paper.tex
figure 6 fixed?
fix figure 4
table 1 working
fix figure 7
fix citations, full paper compiling but fig heights messed up
figure sizing, formatting edits

@@
 library for generating, processing, and visualizing datasets of mazes.
-This library supports a variety of maze generation algorithms providing
-mazes with or without loops, mazes that are connected or not, and many
-other variations. These generation algorithms can be configured with
-various parameters, and the resulting mazes can be filtered to satisfy
-desired properties. Also provided are tools for converting mazes to and
-from various formats suitable for a variety of neural network
-architectures, such as rasterized images, tokenized text sequences, and
-various visualizations. As well as providing a simple interface for
-generating, storing, and loading these datasets, \texttt{maze-dataset}
-is extensively tested, type hinted, benchmarked, and documented.
+This library supports a variety of maze generation algorithms which can
+be configured with various parameters, and the resulting mazes can be
+filtered to satisfy desired properties. Also provided are tools for
+converting mazes to and from various formats suitable for a variety of
+neural network architectures, such as rasterized images, tokenized text
+sequences, and various visualizations. As well as providing a simple
+interface for generating, storing, and loading these datasets,
+\texttt{maze-dataset} is extensively tested, type hinted, benchmarked,
+and documented.
 
@@
 
+\input{figures/tex/fig4_tokenfmt.tex}
+
 \subsection{Benchmarks}\label{benchmarks}
@@
 
+\input{figures/tex/tab1_benchmarks.tex}
+
+\input{figures/tex/fig6_benchmarks.tex}
+
+\input{figures/tex/fig7_sre.tex}
+
 \section{Implementation}\label{sec:implementation}
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
+
+\input{figures/tex/fig8_impl.tex}
 

### 2025-08-21 docs/paper/figures/tex/fig1_diagram.tex
fix minor typos

@@
 \begin{figure} 
-	\begin{minipage}{5in}
+	\resizebox{\linewidth}{!}{
     	\input{diagram/diagram.tikz}
-	\end{minipage}
+	}
 	\caption{
@@
 This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats.
-Code in the image contains clickable links to \docslink{maze_dataset.html}{documentation}.
-A variety of generated examples can be viewed \docslink{examples/maze_examples.html}{here}.
+A variety of generated examples can be viewed on the \docslink{examples/maze_examples.html}{examples page}, and more information can be found in the \docslink{maze_dataset.html}{documentation}.
 	}

### 2025-08-21 docs/paper/figures/tex/fig6_benchmarks.tex
fix minor typos

@@
 		\centering
-		\includegraphics[width=0.9\textwidth]{figures/benchmarks/gridsize-vs-gentime.pdf}
+		\includegraphics[width=\textwidth]{figures/benchmarks/gridsize-vs-gentime.pdf}
 		\caption{
 			Plot of maze generation time. Generation time scales
-			exponentially with maze size for all algorithms. Generation time per
+			roughly exponentially with maze size for all algorithms. Generation time per
 			maze does not depend on the number of mazes being generated, and there

### 2025-08-21 docs/paper/paper.tex
fix minor typos

@@
 challenge comes from building a system supporting many algorithms with
-configurable parameters, property filtering, and representation
-transformation. This library aims to greatly streamline the process of
-generating and working with datasets of mazes that can be described as
-subgraphs of an \(n \times n\) lattice with boolean connections and,
-optionally, start and end points that are nodes in the graph.
-Furthermore, we place emphasis on a wide variety of possible text output
-formats aimed at evaluating the spatial reasoning capabilities of Large
-Language Models (LLMs) and other text-based transformer models.
+configurable parameters, property filtering, representation
+transformation, and reproducibility. This library aims to greatly
+streamline the process of generating and working with datasets of mazes
+that can be described as subgraphs of an \(n \times n\) lattice with
+boolean connections and, optionally, start and end points that are nodes
+in the graph. Furthermore, we place emphasis on a wide variety of
+possible text output formats aimed at evaluating the spatial reasoning
+capabilities of Large Language Models (LLMs) and other text-based
+transformer models.
 
@@
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
-objects, which have path information and a tensor optimized for storing
+objects, which have path information and an array optimized for storing
 sub-graphs of a lattice. These objects can be converted to and from
@@
 
+\newpage
+
 \section{Implementation}\label{sec:implementation}
@@
 This package was originally built for the needs of the
+\texttt{maze-transformer} project
 (\citeproc{ref-maze-transformer-github}{Michael I. Ivanitskiy et al.,
-2023b}) project, which aims to investigate spatial planning and world
-models in autoregressive transformer models trained on mazes
+2023b}), which aims to investigate spatial planning and world models in
+autoregressive transformer models trained on mazes
 (\citeproc{ref-ivanitskiy2023structuredworldreps}{Michael Igorevich
@@
 
-This package has also been utilized in work by other groups:
+At the time of writing, this software package has been actively used in
+work by other groups:
 
@@
   By (\citeproc{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\citeproc{ref-chen2024iaimperative}{Chen et al., 2024}) to study the
-  effectiveness of imperative learning.
-\item
-  By (\citeproc{ref-zhang2025tscend}{Zhang et al., 2025}) to introduce a
-  novel framework for reasoning diffusion models.
+  (\citeproc{ref-chen2024iaimperative}{Chen et al., 2024}) to study
+  imperative learning.
+\item
+  By (\citeproc{ref-zhang2025tscend}{Zhang et al., 2025a}) to introduce
+  a novel framework for reasoning diffusion models.
 \item
@@
   spatial reasoning in LLMs with GRPO.
+\item
+  By (\citeproc{ref-cai2025morse}{Cai et al., 2025}) to create a
+  multimodal reasoning benchmark, via mazes in videos.
+\item
+  By (\citeproc{ref-xu2025visual}{Xu et al., 2025}) to study visual
+  planning in LLMs.
+\item
+  By (\citeproc{ref-lee2025adaptive}{Lee et al., 2025}) to evaluate
+  adaptive inference-time scaling with diffusion models on maze
+  navigation tasks.
+\item
+  By (\citeproc{ref-zhang2025vfscale}{Zhang et al., 2025b}) to test
+  verifier-free diffusion models.
 \end{itemize}
 
+\newpage
+
 \section{Acknowledgements}\label{acknowledgements}
@@
 \input{figures/tex/acknowledgements.tex}
-
-\newpage
 
@@
 \url{https://doi.org/10.3758/brm.40.1.353}
+
+\bibitem[\citeproctext]{ref-cai2025morse}
+Cai, Z., Wang, A., Satheesh, A., Nakhawa, A., Jae, H., Powell, K., Liu,
+M., Jay, N., Oh, S., Wang, X., \& others. (2025). MORSE-500: A
+programmatically controllable video benchmark to stress-test multimodal
+reasoning. \emph{arXiv Preprint arXiv:2506.05523}.
+\url{https://doi.org/10.48550/arXiv.2506.05523}
 
@@
 
+\bibitem[\citeproctext]{ref-lee2025adaptive}
+Lee, G., Bao, T. N. N., Yoon, J., Lee, D., Kim, M., Bengio, Y., \& Ahn,
+S. (2025). Adaptive cyclic diffusion for inference scaling. \emph{arXiv
+Preprint arXiv:2505.14036}.
+\url{https://doi.org/10.48550/arXiv.2505.14036}
+
 \bibitem[\citeproctext]{ref-eval-LLM-graphs}
@@
 
+\bibitem[\citeproctext]{ref-xu2025visual}
+Xu, Y., Li, C., Zhou, H., Wan, X., Zhang, C., Korhonen, A., \& Vulić, I.
+(2025). Visual planning: Let's think only with images. \emph{arXiv
+Preprint arXiv:2505.11409}.
+\url{https://doi.org/10.48550/arXiv.2505.11409}
+
 \bibitem[\citeproctext]{ref-zhang2025tscend}
-Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
+Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025a). T-SCEND: Test-time
 scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
@@
 
+\bibitem[\citeproctext]{ref-zhang2025vfscale}
+Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025b). VFScale: Intrinsic
+reasoning through verifier-free test-time scalable diffusion model.
+\emph{arXiv Preprint arXiv:2502.01989}.
+\url{https://doi.org/10.48550/arXiv.2502.01989}
+
 \end{CSLReferences}

### 2025-08-21 docs/paper/figures/tex/fig8_impl.tex
minor edits to implementation figure (fig 5)

@@
 \begin{tabular}{ccc}
-	Maze 
+	Maze \hspace{1em} 
 	& $A[0]$ (downward)
@@
 	\raisebox{-0.5\height}{
-		\includegraphics[width=5em]{figures/tex/maze-impl.pdf}
-	}
+		\includegraphics[width=7em]{figures/tex/maze-impl.pdf}
+	} \hspace{1em}
 	& \raisebox{-0.5em}{
-		\large $\begin{bmatrix}
+		\huge $\begin{bmatrix}
 			1 & 1 & 1 \\
@@
 	& \raisebox{-0.5em}{
-		\large $\begin{bmatrix}
+		\huge $\begin{bmatrix}
 			1 & 0 & 0 \\
@@
 \caption{
-	We describe mazes with the following representation: for a $2$-dimensional lattice with $r$ rows and $c$ columns, we initialize a boolean array $A = \{0, 1\}^{2 \times r \times c}$ which we refer to in the code as a \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.connection_list}{\texttt{connection\_list}}.
+	We describe mazes with the following representation: for a $2$-dimensional lattice with $r$ rows and $c$ columns, we initialize a boolean array $A = \{0, 1\}^{2 \times r \times c}$ which we refer to in the code as a
+	\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}. 
 	The value at $A[0,i,j]$ determines whether a \emph{downward} connection exists from node $[i,j]$ to $[i+1, j]$.
@@
 }
+\label{fig:maze-impl}
 \end{figure}

### 2025-10-12 docs/paper/figures/tex/fig2_formats.tex
JOSS paper fixes
see https://github.com/openjournals/joss-reviews/issues/8633#issuecomment-3270514168

- `numpy` -> NumPy in figure 2
- proper capitalization of "feature-rich" in figure 2
- fixed legend in figure 7
- author name consistency: "Alex F. Spies" -> "Alexander F. Spies"

@@
 		Simple text format for displaying mazes, useful for debugging in a terminal environment.
-		& \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
-		& feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
+		& NumPy array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
+		& Feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
 	  \hline \\
 	  % examples
-		\multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
-		  \setlength{\baselineskip}{0.9em}
+		\multicolumn{1}{l}{
+		  \hspace{-1.5em}\raisebox{0.85\height}{
 			  \input{figures/outputs-ascii-colored.tex}
-		\end{minipage}}
+		  }
+		}
+		% 	\begin{minipage}[b]{1.6in}
+		%   \setlength{\baselineskip}{0.9em}
+		% 	\end{minipage}
 		& \multicolumn{1}{c}{
@@
 		}
-		& \multicolumn{1}{c}{
+		& \multicolumn{1}{r}{
 		  \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}

### 2025-10-12 docs/paper/paper.tex
reorder sections, implementation before benchmarks
JOSS paper fixes
see https://github.com/openjournals/joss-reviews/issues/8633#issuecomment-3270514168

- `numpy` -> NumPy in figure 2
- proper capitalization of "feature-rich" in figure 2
- fixed legend in figure 7
- author name consistency: "Alex F. Spies" -> "Alexander F. Spies"

@@
 Representational Flexibility},
-    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
-Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
+    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alexander
+F. Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
 Fung},
@@
     pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
-Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
-Diniz Behn, Samy Wu Fung},
+Aaron Sandoval, Alexander F. Spies, Tilman Räuker, Brandon Knutson,
+Cecilia Diniz Behn, Samy Wu Fung},
     pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
@@
 %
-]{Alex F. Spies%
+]{Alexander F. Spies%
   \,\orcidlink{0000-0002-8708-1530}\,%
@@
     \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
-    provides a feature‑rich plotting utility with support for multiple
+    provides a feature-rich plotting utility with support for multiple
     paths, heatmaps over positions, and more. This is similar to the
@@
 
+\section{Implementation}\label{sec:implementation}
+
+Using an adjacency matrix for storing mazes would be memory inefficient
+by failing to exploit the highly sparse structure, while using an
+adjacency list could lead to a poor lookup time. This package utilizes a
+simple, efficient representation of mazes as subgraphs of a finite
+lattice, detailed in \autoref{fig:maze-impl}, which we call a
+\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
+
+\input{figures/tex/fig8_impl.tex}
+
+Our package is implemented in Python(\citeproc{ref-python}{Rossum,
+1995}), and makes use of the extensive scientific computing ecosystem,
+including NumPy (\citeproc{ref-numpy}{Harris et al., 2020}) for array
+manipulation, plotting tools (\citeproc{ref-matplotlib}{Hunter, 2007};
+\citeproc{ref-seaborn}{Waskom, 2021}), Jupyter notebooks
+(\citeproc{ref-jupyter}{Kluyver et al., 2016}), and PySR
+(\citeproc{ref-pysr}{Cranmer, 2023}) for symbolic regression.
+
 \subsection{Benchmarks}\label{benchmarks}
@@
 \input{figures/tex/fig7_sre.tex}
-
-\newpage
-
-\section{Implementation}\label{sec:implementation}
-
-Using an adjacency matrix for storing mazes would be memory inefficient
-by failing to exploit the highly sparse structure, while using an
-adjacency list could lead to a poor lookup time. This package utilizes a
-simple, efficient representation of mazes as subgraphs of a finite
-lattice, detailed in \autoref{fig:maze-impl}, which we call a
-\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
-
-\input{figures/tex/fig8_impl.tex}
-
-Our package is implemented in Python(\citeproc{ref-python}{Rossum,
-1995}), and makes use of the extensive scientific computing ecosystem,
-including NumPy (\citeproc{ref-numpy}{Harris et al., 2020}) for array
-manipulation, plotting tools (\citeproc{ref-matplotlib}{Hunter, 2007};
-\citeproc{ref-seaborn}{Waskom, 2021}), Jupyter notebooks
-(\citeproc{ref-jupyter}{Kluyver et al., 2016}), and PySR
-(\citeproc{ref-pysr}{Cranmer, 2023}) for symbolic regression.
 

### 2025-10-12 maze_dataset/benchmark/config_sweep.py

@@ -558,7 +558,7 @@ def plot_grouped(  # noqa: C901
 			)
 			print(f"Saving plot to {save_path.as_posix()}")
 			save_path.parent.mkdir(exist_ok=True, parents=True)
-			plt.savefig(save_path)
+			plt.savefig(save_path, bbox_inches="tight")
 
 		if show:
 			plt.show()

### 2025-10-12 docs/paper/figures/tex/fig2_formats.tex
[joss paper] minor figure 2 edits

@@
 	  % algorithms
-	  \docslink{maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-	  & \docslink{maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
+	  \docslink{maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
+	  & \docslink{maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
 	  & \docslink{maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
 	  % descriptions
-		Simple text format for displaying mazes, useful for debugging in a terminal environment.
-		& NumPy array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
-		& Feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
+	  	{ \footnotesize NumPy array of \texttt{dtype=uint8} and shape \texttt{(height*2+1, width*2+1, 3)}. The last dimension is RGB color. }
+		& { \footnotesize Simple text format for displaying mazes, useful for debugging in a terminal environment. Text representation of \texttt{as\_pixels()}. }
+		& { \footnotesize Feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. } \\[1em]
 	  \hline \\
 	  % examples
-		\multicolumn{1}{l}{
+		\multicolumn{1}{c}{
+			\includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
+		}
+		& \multicolumn{1}{l}{
 		  \hspace{-1.5em}\raisebox{0.85\height}{
@@
 		  }
-		}
-		% 	\begin{minipage}[b]{1.6in}
-		%   \setlength{\baselineskip}{0.9em}
-		% 	\end{minipage}
-		& \multicolumn{1}{c}{
-		  \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
 		}

### 2025-10-20 docs/paper/paper.tex
recompile paper

@@
     \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
-    provides a format similar to
-    (\citeproc{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
-    \citeproc{ref-eval-gpt-visual}{Singla, 2023})
+    provides a format similar to Singla
+    (\citeproc{ref-eval-gpt-visual}{2023}) and Oppenheim
+    (\citeproc{ref-gh-oppenheimj2018maze}{2018})
   \item
@@
     paths, heatmaps over positions, and more. This is similar to the
-    outputs of (\citeproc{ref-mazegenerator-net}{Alance AB, 2019};
-    \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
+    outputs of many packages (\citeproc{ref-mazegenerator-net}{Alance
+    AB, 2019}; \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
     \citeproc{ref-mathematica-maze}{Guo et al., 2011};
-    \citeproc{ref-mdl-suite}{Nag, 2020})
+    \citeproc{ref-mdl-suite}{Nag, 2020}).
   \end{itemize}
@@
   \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
-  is similar to that of (\citeproc{ref-eval-LLM-graphs}{Liu \& Wu,
-  2023}) but with many more options, detailed in
+  is similar to that of Liu \& Wu (\citeproc{ref-eval-LLM-graphs}{2023})
+  but with many more options, detailed in
   \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
@@
 al., 2021}). To facilitate the use of our package in this context, we
-replicate the format of (\citeproc{ref-easy_to_hard}{Schwarzschild,
-Borgnia, Gupta, Bansal, et al., 2021}) and provide the
+replicate the format of Schwarzschild, Borgnia, Gupta, Bansal, et al.
+(\citeproc{ref-easy_to_hard}{2021}) and provide the
 \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
@@
 
-Our package is implemented in Python(\citeproc{ref-python}{Rossum,
+Our package is implemented in Python (\citeproc{ref-python}{Rossum,
 1995}), and makes use of the extensive scientific computing ecosystem,
@@
 et al., 2024}), which required matching the pixel-padded and endpoint
-constrained output format of (\citeproc{ref-easy_to_hard}{Schwarzschild,
-Borgnia, Gupta, Bansal, et al., 2021}). Ongoing work using
+constrained output format of Schwarzschild, Borgnia, Gupta, Bansal, et
+al. (\citeproc{ref-easy_to_hard}{2021}). Ongoing work using
 \texttt{maze-dataset} aims to investigate the effects of varying the
@@
 \item
-  By (\citeproc{ref-nolte2024multistep}{Nolte et al., 2024}) to compare
+  By Nolte et al. (\citeproc{ref-nolte2024multistep}{2024}) to compare
   the effectiveness of transformers trained with the MLM-\(\mathcal{U}\)
@@
 \item
-  By (\citeproc{ref-wang2024imperative}{Wang et al., 2024}) and
-  (\citeproc{ref-chen2024iaimperative}{Chen et al., 2024}) to study
-  imperative learning.
-\item
-  By (\citeproc{ref-zhang2025tscend}{Zhang et al., 2025a}) to introduce
-  a novel framework for reasoning diffusion models.
-\item
-  By (\citeproc{ref-dao2025alphamaze}{Dao \& Vu, 2025}) to improve
+  By Wang et al. (\citeproc{ref-wang2024imperative}{2024}) and Chen et
+  al. (\citeproc{ref-chen2024iaimperative}{2024}) to study imperative
+  learning.
+\item
+  By Zhang et al. (\citeproc{ref-zhang2025tscend}{2025a}) to introduce a
+  novel framework for reasoning diffusion models.
+\item
+  By Dao \& Vu (\citeproc{ref-dao2025alphamaze}{2025}) to improve
   spatial reasoning in LLMs with GRPO.
 \item
-  By (\citeproc{ref-cai2025morse}{Cai et al., 2025}) to create a
+  By Cai et al. (\citeproc{ref-cai2025morse}{2025}) to create a
   multimodal reasoning benchmark, via mazes in videos.
 \item
-  By (\citeproc{ref-xu2025visual}{Xu et al., 2025}) to study visual
+  By Xu et al. (\citeproc{ref-xu2025visual}{2025}) to study visual
   planning in LLMs.
 \item
-  By (\citeproc{ref-lee2025adaptive}{Lee et al., 2025}) to evaluate
+  By Lee et al. (\citeproc{ref-lee2025adaptive}{2025}) to evaluate
   adaptive inference-time scaling with diffusion models on maze
@@
 \item
-  By (\citeproc{ref-zhang2025vfscale}{Zhang et al., 2025b}) to test
+  By Zhang et al. (\citeproc{ref-zhang2025vfscale}{2025b}) to test
   verifier-free diffusion models.
@@
 Cranmer, M. (2023). Interpretable machine learning for science with PySR
-and SymbolicRegression. jl. \emph{arXiv Preprint arXiv:2305.01582}.
+and SymbolicRegression.jl. \emph{arXiv Preprint arXiv:2305.01582}.
 \url{https://doi.org/10.48550/arXiv.2305.01582}
